### 2013-05-03 thesis.tex
Inducted my thesis into git

\documentclass{article}

%% \usepackage{sdrt}
\usepackage{gb4e}
\usepackage{bussproofs}

%% Linear implication from Alessio Guglielmi
%% https://groups.google.com/forum/?fromgroups=#!topic/comp.text.tex/0B4C3F_BsVI
\def\limp {\mathbin{{-}\mkern-3.5mu{\circ}}}

\begin{document}

\author{Ji\v{r}\'{i} Mar\v{s}\'{i}k}
\title{Master Thesis}
\maketitle

\section{Preliminaries}

\subsection{Motivation}

To develop applications capable of understanding natural language, we
have to analyze the discourse as a complex structure. The structure of
the discourse and the rhetorical relations between its constituent
propositions have shown to have an important effect on anaphora and on
the semantic content of the discourse.

Let us consider these examples from \cite{asher2003logics}.

\begin{exe}
  \ex \label{narxexp-nar} \begin{xlist}
    \ex \label{narxexp-nar-a} Max fell.
    \ex \label{narxexp-nar-b} John helped him up.
  \end{xlist}
  \ex \label{narxexp-exp} \begin{xlist}
    \ex \label{narxexp-exp-a} Max fell.
    \ex \label{narxexp-exp-b} John pushed him.
  \end{xlist}
\end{exe}

In (\ref{narxexp-nar}), we intuitively recognize that the second
proposition (\ref{narxexp-nar-b}) serves as a narrative continuation of
(\ref{narxexp-nar-a}), whereas in (\ref{narxexp-exp}), the proposition
(\ref{narxexp-exp-b}) serves as an explanation to the event described in
(\ref{narxexp-exp-a}). This distinction has interesting consequences as
to what we can infer from these two discourse excerpts. In the case of
(\ref{narxexp-nar}), we can infer that the event described in
(\ref{narxexp-nar-b}) occurred after the event described in
(\ref{narxexp-nar-a}). In the latter case (\ref{narxexp-exp}), we can
conversely infer that the events happened in the opposite order. We feel
that a complete understanding of the two examples above entails
inferring the correct temporal order and we would welcome a principled
way to make this kind of decisions.

\begin{exe}
  \ex \label{salmon} \begin{xlist}
    \ex Max had a lovely evening last night.
    \ex He had a great meal.
    \ex \label{salmon-here} He ate salmon.
    \ex He devoured lots of cheese.
    \ex He then won a dancing competition.
  \end{xlist}
\end{exe}

%% \sdrtree{
%%  &                      & \LAB{Max had a lovely evening} \\
%%           & \LAB{He had a great meal} & & \LAB{He won a dancing competition} \\
%% \LAB{He ate salmon} &     & \LAB{He devoured cheese}
%% }

The discourse in (\ref{salmon}) and its structure in (TODO) demonstrate
another feature of discourse structures. First off, seeing the
hiearchical structure of the discourse gives us important information
about the granularity of the descriptions given in the individual
propositions, information that could be useful for performing tasks such
as text summarization.

Furthermore, the discourse structure has grammatical consequences. It is
thanks to our knowledge of the discourse structure that we can predict
that a proposition like ``It was a beautiful pink.''  could not
coherently follow our excerpt. SDRT, \cite{asher2003logics}, the theory
of discourse structure that we adhere to, would not license a discourse
structure in which the new proposition connects to (\ref{salmon-here})
as its \emph{Elaboration}. As a consequence, it states that the pronoun
``it'' cannot have the salmon as its antecedent, which is a constraint
that is not enforced by prior theories of discourse such as DRT.

\begin{exe}
  \ex \label{diaimp} \begin{xlist}
    \ex A: Smith doesn't seem to have a girlfriend.
    \ex \label{diaimp-b} B: He's been paying lots of visits to New York lately.
  \end{xlist}
\end{exe}

In the example dialogue (\ref{diaimp}), based on the prosody of
proposition (\ref{diaimp-b}) the discourse structure would link the two
propositions with an \emph{Evidence} or \emph{Counter-evidence}
relation. Knowing this structure would allow us to either infer that B
believes that Smith has a girlfriend in New York or that Smith doesn't
have a girlfriend because he is too busy in New York.

Based on the findings in \cite{asher2003logics}, we surmise that a clear
picture of the discourse structure is essential to capturing the
intended meaning of a discourse.

To support our approach, we will stand on the shoulders of many giants,
the first of them being Richard Montague. As in his approach
\cite{montague1973proper}, we will assign functions and values as
denotations of wordforms and use function application to compose them
together to yield the denotations of phrases, propositions and
discourses. To account for the discourse-level phenomena, we will defer
to the theories of DRT \cite{kamp1993discourse} and SDRT
\cite{asher2003logics}.

The grammatical formalism of our choice for this task will be the
Abstract Categorial Grammars (ACG) \cite{de2001towards}. This framework
lets use lambda calculus to express the syntax-semantics interface in a
fashion similar to Montague's. Furthermore, elegant techniques for using
ACGs to implemenent DRT and SDRT using continuations have been
discovered \cite{de2006towards} \cite{asher2011sdrt}
\cite{asher2011montagovian}.

Finally, our work will be supported by the existence of a corpus
annotated with the rhetorical relations of SDRT, the Annodis corpus
\cite{afantenos2012empirical}.

\subsection{Abstract Categorial Grammars}

We present the grammatical framework in which we will develop our
system. Abstract categorial grammars are built upon two mathematical
structures, \emph{(higher-order) signatures} and \emph{lexicons}.

A \textbf{higher-order signature} is a set of elements that we call
\emph{constants}, each of which is associated with a type. Formally, it
is defined as a triple $\Sigma = \mathopen{<}A, C, \tau\mathclose{>}$,
where:
\begin{itemize}
  \item $C$ is the (finite) set of constants
  \item $A$ is a (finite) set of atomic types
  \item $\tau$ is the type associating mapping from $C$ to $\mathcal{T}(A)$,
    the set of types built over $A$
\end{itemize}

In our case, $\mathcal{T}(A)$ is the implicative fragment of linear and
intuitionistic logic with $A$ being the atomic propositions. This means
that $\mathcal{T}(A)$ contains all the $a \in A$ and all the $\alpha \limp
\beta$ and $\alpha \to \beta$ for $\alpha, \beta \in \mathcal{T}(A)$.

A signature $\Sigma = \mathopen{<}A, C, \tau\mathclose{>}$, by itself,
already lets us define an interesting set of structures, that is the set
$\Lambda(\Sigma)$ of \emph{well-typed lambda terms} built upon the
signature $\Sigma$. The set of well-typed terms and their types are
established through the judgment schemas in Figure
\ref{fig:type-judgments}.

\begin{figure}
  \begin{prooftree}
    \AxiomC{$\emptyset, \Gamma_i \vdash_\Sigma c : \tau(c)$ (cons)}
  \end{prooftree}
  \begin{prooftree}
    \AxiomC{$(x : \alpha), \Gamma_i \vdash_\Sigma x : \alpha$ (l-var)}
  \end{prooftree}
  \begin{prooftree}
    \AxiomC{$\emptyset, (\Gamma_i, x : \alpha) \vdash_\Sigma x : \alpha$ (i-var)}
  \end{prooftree}
  \begin{prooftree}
    \AxiomC{$(\Gamma_l, x : \alpha), \Gamma_i \vdash_\Sigma t : \beta$}
    \RightLabel{(l-abs)}
    \UnaryInfC{$\Gamma_l, \Gamma_i \vdash_\Sigma \lambda^{\circ} x. t : \alpha \limp \beta$}
  \end{prooftree}
  \begin{prooftree}
    \AxiomC{$\Gamma_l, (\Gamma_i, x : \alpha) \vdash_\Sigma t : \beta$}
    \RightLabel{(i-abs)}
    \UnaryInfC{$\Gamma_l, \Gamma_i \vdash_\Sigma \lambda x. t : \alpha \to \beta$}
  \end{prooftree}
  \begin{prooftree}
    \AxiomC{$\Gamma_l, \Gamma_i \vdash_\Sigma t : \alpha \limp \beta$}
    \AxiomC{$\Delta_l, \Delta_i \vdash_\Sigma u : \alpha$}
    \RightLabel{(l-app)}
    \BinaryInfC{$(\Gamma_l, \Delta_l), (\Gamma_i, \Delta_i) \vdash_\Sigma (t\ u) : \beta$}
  \end{prooftree}
  \begin{prooftree}
    \AxiomC{$\Gamma_l, \Gamma_i \vdash_\Sigma t : \alpha \to \beta$}
    \AxiomC{$\emptyset, \Delta_i \vdash_\Sigma u : \alpha$}
    \RightLabel{(i-app)}
    \BinaryInfC{$\Gamma_l, (\Gamma_i, \Delta_i) \vdash_\Sigma (t\ u) : \beta$}
  \end{prooftree}
  \caption{\label{fig:type-judgments} Type judgment schemas of the
    well-typed lambda terms $\Lambda(\Sigma)$ for a signature $\Sigma =
    \mathopen{<}A, C, \tau\mathclose{>}$}
\end{figure}

The definition of a well-typed lambda term already gives us an
interesting combinatorial structure. To make this structure even more
useful, we often focus ourselves only on terms that have a specific
\emph{distinguished type}. Using this notion of a signature of typed
constants and some distinguished type, we can describe languages of,
e.g. tree-like (and by extension string-like), lambda terms.

The idea of a signature is coupled with that of a \textbf{lexicon},
which is a mapping between two different signatures (mapping the
constants of one into well-typed terms of the other). Formally speaking,
a lexicon $\mathcal{L}$ from a signature $\Sigma_1 = \mathopen{<}A_1, C_1,
\tau_1\mathclose{>}$ (which we call the abstract signature) to a
signature $\Sigma_2 = \mathopen{<}A_2, C_2, \tau_2\mathclose{>}$ (which
we call the object signature) is a pair $\mathopen{<}F, G\mathclose{>}$
such that:
\begin{itemize}
\item $G$ is a mapping from $C_1$ to $\Lambda(\Sigma_2)$ assigning to
  every constant of the abstract signature a term in the object
  signature, which can be understood as its
  interpretation/implementation/realization.
\item $F$ is a mapping from $A_1$ to $\mathcal{T}(A_2)$ which links the
  abstract-level types with the object-level types that they realized
  in.
\item $F$ and $G$ are compatible, meaning that for any $c \in C_1$, we
  have $\vdash_{\Sigma_2} G(c) : \hat{F}(\tau_1(c))$ (we will be using
  $\hat{F}$ and $\hat{G}$ to refer to the homomorphic extensions of $F$
  and $G$ to $\mathcal{T}(A_1)$ and $\Lambda(\Sigma_1)$ respectively).
\end{itemize}

An abstract categorial grammar for us will then be just a collection of
signatures with their distinguished types and lexicons connecting these
signatures. A common pattern will have us using two signatures for the
surface forms of utterances (strings) and their logical forms (logical
propositions) and an abstract signature which is connected to both of
the object signatures via lexicons. Parsing is then just a matter of
inverting the surface lexicon to get the abstract term and then applying
to it the logical lexicon. Generation is symmetric, we simply invert the
logical lexicon and apply the surface lexicon.

\subsubsection{Example ACG}

We will look at an example of an ACG from

$\Sigma_{Synt}$ will be our abstract signature with $A_{Synt} = \{np, n,
s\}$, $C_{Synt} = \{C_{every}, C_{some}, C_{love}, C_{man}, C_{woman}\}$
and $\tau_{Synt}$ as follows:
\begin{itemize}
\item
\end{itemize}

...

\subsection{Interaction Grammars}

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}

### 2013-05-06 thesis.tex
Inline example of a signature (string)

@@
 
+Let us take a look at an example from
+\cite{pogodalla2007generalizing}. We will start with a very simple
+signature of string expressions with the concatenation operator,
+$\Sigma_{String} = \mathopen{<}A_{String}, C_{String},
+\tau_{String}\mathclose{>}$. Our signature will need only one atomic
+type, $A_{String} = \{\textsc{String}\}$. Our constants will be the
+wordforms of our example fragment ($every$, $some$, $man$, $woman$,
+$loves$), the empty string $\epsilon$ and the string concatention
+operator, $+$. $\tau_{String}$ assigns the type $\textsc{String}$ to all
+the wordforms and to $\epsilon$ and the type $\textsc{String} \limp
+\textsc{String} \limp \textsc{String}$\footnote{As a convention,
+  whenever we omit parentheses in a type, we presume the $\limp$ and
+  $\to$ type constructors always bind from the right, meaning that $a
+  \to b \to c$ should be read as $a \to (b \to c)$.}
+to the concatenation operator $+$.
+
+Now we can look at some well-typed terms from the string domain
+(i.e. elements of $\Lambda(\Sigma_{String})$). The term $t_1 =
+some\ +\ woman$\footnote{Here we introduce another notational
+  convention. The unadorned way of writing this term would be $t_1 =
+  (+\ some)\ woman$. Firstly, we will allow ourselves to drop the
+  parentheses, meaning that any string of expressions $f\ x\ y$ should
+  be read as $(f\ x)\ y$. Secondly, we will admit a specific infix
+  notation for some constants which denote functions of two
+  arguments. This will then let us write $x\ +\ y$ instead of
+  $+\ x\ y$.} denotes the concatenation of $some$ and $woman$, which is
+the phrase $some\ woman$; its type is $\textsc{String}$. A term such as
+$t_2 = \lambda^{\circ} x. x\ +\ man$ denotes a function that appends the
+string $man$ to its argument; its type is $\textsc{String} \limp
+\textsc{String}$. If we restrict the set $\Lambda(\Sigma_{String})$ only
+to terms having the type $\textsc{String}$, we get the set of
+expressions which denote strings.
+
 The idea of a signature is coupled with that of a \textbf{lexicon},

### 2013-05-06 thesis.tex
Example lexicon (Syntax)

@@
 
+\subsubsection{Higher-order signatures}
+
 A \textbf{higher-order signature} is a set of elements that we call
@@
 
+\subsubsection{Example signature}
+\label{sssec:example-sig}
+
 Let us take a look at an example from
@@
 to terms having the type $\textsc{String}$, we get the set of
-expressions which denote strings.
+expressions which denote strings (a set that contains terms like $t_1$
+but not $t_2$).
+
+\subsubsection{Lexicons}
 
@@
 
-\subsubsection{Example ACG}
+\subsubsection{Example lexicon}
 
-We will look at an example of an ACG from
+\newcommand{\synt}[1]{C_{\textrm{#1}}}
 
-$\Sigma_{Synt}$ will be our abstract signature with $A_{Synt} = \{np, n,
-s\}$, $C_{Synt} = \{C_{every}, C_{some}, C_{love}, C_{man}, C_{woman}\}$
-and $\tau_{Synt}$ as follows:
+To illustrate the ideas of a lexicon and an abstract categorial grammar,
+we will expand our example from \ref{sssec:example-sig}. We will
+consider another signature, $\Sigma_{Synt}$, which will describe the
+syntax of quantified noun phrases in the style of Montague
+\cite{montague1973proper}. The atomic types $A_{Synt}$ will consist of
+the types $np$, $n$ and $s$. Our constants, $C_{Synt} = \{\synt{every},
+\synt{some}, \synt{love}, \synt{man}, \synt{woman}\}$, will have the
+following types, as predicted by $\tau_{Synt}$:
 \begin{itemize}
-\item
+\item $\tau_{Synt}(\synt{every}) = \tau_{Synt}(\synt{some}) = n \limp
+  ((np \limp s) \limp s)$
+\item $\tau_{Synt}(\synt{love}) = np \limp np \limp s$
+\item $\tau_{Synt}(\synt{man}) = \tau_{Synt}(\synt{woman}) = n$
 \end{itemize}
 
-...
+Let us explore some terms from $\Lambda(\Sigma_{Synt})$. $t_3 =
+\synt{some}\ \synt{woman}$ has type $(np \limp s) \limp s$, which is the
+type that serves to describe quantified noun phrases in our
+example. Terms of this type expect a verb phrase (or some other
+predicate) of type $np \limp s$, and can yield a sentence of type $s$.
 
-\subsection{Interaction Grammars}
+Now we will see how this approach can handle transitive verbs. In $t_4 =
+\lambda^{\circ} x.\ (\synt{some}\ \synt{woman})\ (\lambda^{\circ}
+y.\ \synt{love}\ x\ y)$, which is a term of type $np \limp s$
+representing the verb phrase \emph{loves some woman}, we first take the
+expression $\synt{love}\ x\ y$ (\emph{$x$ loves $y$}, type $s$) and we
+abstract over $y$ to get the predicate $\lambda^{\circ}
+y.\ \synt{love}\ x\ y$ (\emph{is loved by $x$}, type $np \limp s$). By
+applying the quantified noun phrase $t_3$ to this term, we get
+$(\synt{some}\ \synt{woman})\ (\lambda^{\circ} y.\ \synt{love}\ x\ y)$
+(\emph{$x$ loves some woman}, type $s$). Finally, by abstracting over
+$x$, we get the verb phrase $t_4$.
+
+$t_5 = (\synt{every}\ \synt{man})\ (\lambda^{\circ}
+x.\ (\synt{some}\ \synt{woman})\ (\lambda^{\circ}
+y.\ \synt{love}\ x\ y))$ is the result of applying the quantified noun
+phrase $\synt{every}\ \synt{man}$ (of type $(np \limp s) \limp s$) to
+$t_4$. What we get is a term of type $s$ which describes the sentence
+\emph{every man loves some woman}. Note that as before, we can restrict
+the set $\Lambda(\Sigma_{Synt})$ to terms having the distinguished type
+$s$ and we will arrive at the set of terms which describe sentences
+(this set will include $t_5$, but not $t_3$ or $t_4$).
+
+In the example above, I have been presuming some kind of implied
+connection between the terms of $\Lambda(\Sigma_{Synt})$ and phrases of
+some natural language, strings. Indeed, while we have defined a system
+of structures which describe the quantification and predicate structures
+of a microscopic fragment of English and a system for talking about
+strings in concatenation, we have not given any explicit link between
+the two. It is at this moment that we will introduce a lexicon to link
+these two levels of description.
+
+Our lexicon $\mathcal{L}_{syntax}$ will map the constants of
+$\Sigma_{Synt}$, our abstract signature, into terms from
+$\Lambda(\Sigma_{String})$, our object signature. If we view terms of
+$\Lambda(\Sigma_{Synt})$ as abstract computations, the lexicon will
+instantiate this abstraction by providing an implementation for the
+constants of $\Sigma_{Synt}$. This way we can map an abstract
+computation representing a phrase into a more specific object
+computation that computes the string representing the phrase.
+
+So, how does our lexicon $\mathcal{L}_{syntax}$ look like? It will map
+all the atomic types of $\Sigma_{Synt}$ into $\textsc{String}$, which we
+can write as $\mathcal{L}_{syntax}(n) = \mathcal{L}_{syntax}(np) =
+\mathcal{L}_{syntax}(s)$\footnote{Whenever we talk about some lexicon,
+  we will use the lexicon itself, which is formally a pair of mappings
+  $\mathopen{<}F, G\mathclose{>}$, to mean either $F$, $G$ or their
+  homomorphic extensions $\hat{F}$ and $\hat{G}$ depending on whether
+  the argument is an atomic type, a constant, a complex type or a term
+  respectively.}. From that, we can use the homomorphic extension of our
+type mapping to arrive at the object-level interpretations of complex
+types, such as the type of predicates, $np \limp s$, which becomes a
+unary string function, $\textsc{String} \limp \textsc{String}$, or the
+type of quantified noun phrases, $(np \limp s) \limp s$, which becomes a
+higher-order string function, $(\textsc{String} \limp \textsc{String})
+\limp \textsc{String}$. With the types out of the way, we can give the
+interpretation of the individual constants of $\Sigma_{Synt}$:
+\begin{itemize}
+\item $\mathcal{L}_{syntax}(\synt{every}) = \lambda^{\circ} x
+  R.\ R\ (every\ +\ x)$
+\item $\mathcal{L}_{syntax}(\synt{some}) = \lambda^{\circ} x
+  R.\ R\ (some\ +\ x)$
+\item $\mathcal{L}_{syntax}(\synt{love}) = \lambda^{\circ} x
+  y.\ x\ +\ loves\ +\ y$
+\item $\mathcal{L}_{syntax}(\synt{man}) = man$
+\item $\mathcal{L}_{syntax}(\synt{woman}) = woman$
+\end{itemize}
+
+We can now go back to our examples from $\Lambda(\Sigma_{Synt})$ and
+look at what they look like after applying the lexicon to them.
+\begin{itemize}
+\item $t_3 = \synt{some}\ \synt{woman}$ \\ Mapping this term using the
+  lexicon and $\beta$-reducing gives us $\mathcal{L}_{syntax}(t_3)
+  =_{\beta} \lambda^{\circ} R.\ R\ (some\ +\ woman)$, which is a
+  type-raised version of the string \emph{some woman}.
+\item $t_4 = \lambda^{\circ}
+  x.\ (\synt{some}\ \synt{woman})\ (\lambda^{\circ}
+  y.\ \synt{love}\ x\ y)$ \\ Our verb phrase ends up being mapped onto
+  a function which appends the phrase \emph{loves some woman} to its
+  argument, $\mathcal{L}_{syntax}(t_4) =_{\beta} \lambda^{\circ}
+  x.\ x\ +\ loves\ +\ some\ +\ woman$
+\item $t_5 = (\synt{every}\ \synt{man})\ (\lambda^{\circ}
+  x.\ (\synt{some}\ \synt{woman})\ (\lambda^{\circ}
+  y.\ \synt{love}\ x\ y))$ \\ Finally, the sentence (type $s$) is mapped
+  into a simple $\textsc{String}$, $\mathcal{L}_{syntax}(t_5) =_{\beta}
+  every\ +\ man\ +\ loves\ +\ some\ +\ woman$.
+\end{itemize}
+
+Now we have enough machinery in play to define the set of strings which
+form our fragment of English. We will consider the \emph{abstract
+  language} generated by our signature $\Sigma_{Synt}$ and the
+distinguished type $s$, that is, elements of $\Lambda(\Sigma_{Synt})$
+having type $s$. Then we can define the \emph{object language}, which is
+the image of the abstract language when transformed using the
+lexicon. The object language contains strings, terms from
+$\Lambda(\Sigma_{String})$ having the type $\textsc{String}$. However,
+the object language does not contain all such terms, but only those
+object terms, for which there exists a term in the abstract language
+such that its image given by the lexicon yields the same object term.
 

### 2013-05-06 thesis.tex
Example lexicon (semantics)

@@
 \subsubsection{Example lexicon}
+\label{sssec:example-lex}
 
@@
 
+\subsubsection{Example semantic lexicon}
+
+One perk of working with abstract categorial grammars is that one
+abstract term can be interpreted multiple ways, using different
+lexicons. In the previous section, we considered the abstract terms of
+the $\Sigma_{Synt}$ signature and interpreted them as computations on
+strings. Now we will take the same abstract terms, but try to intepret
+them as computations on entities and truth values.
+
+We will introduce a new signature, $\Sigma_{Sem}$, for our semantic
+computations. The atomic types $A_{Sem}$ will consist of a type for
+entities, $e$, and a type for truth values, $t$. The constants $C_{Sem}$
+will contain atomic predicates from our fragment, $\textbf{man}$,
+$\textbf{woman}$ and $\textbf{love}$, logical connectives, $\Rightarrow$
+and $\land$, and quantifiers, $\forall$ and $\exists$. Their types as
+given by $\tau_{Sem}$ are:
+\begin{itemize}
+\item $\tau_{Sem}(\textbf{man}) = \tau_{Sem}(\textbf{woman}) = e \limp
+  t$
+\item $\tau_{Sem}(\textbf{love}) = e \limp e \limp t$
+\item $\tau_{Sem}(\Rightarrow) = \tau_{Sem}(\land) = t \limp t \limp t$
+\item $\tau_{Sem}(\forall) = \tau_{Sem}(\exists) = (e \to t) \limp t$
+\end{itemize}
+
+Now to see how this signature connects to our abstract signature
+$\Sigma_{Synt}$, we will give a lexicon from the latter to the
+former. First, we give the type mappings. $\mathcal{L}_{sem}(n) = e
+\limp t$ meaning that nouns will yield boolean functions of one argument
+(unary \emph{predicates} in functional programming
+terminology). $\mathcal{L}_{sem}(np) = e$, simple noun phrases will be
+mapped into terms which yield some entity. Finally,
+$\mathcal{L}_{sem}(s) = t$, sentences will be assigned terms which
+compute their truthiness. Let us see how this plan is achieved in the
+$\mathcal{L}_{sem}$ lexicon.
+\begin{itemize}
+\item $\mathcal{L}_{sem}(\synt{every}) = \lambda^{\circ} P Q.\ \forall
+  x.\ ((P\ x) \Rightarrow (Q\ x))$\footnote{We will be using $\forall
+    x.\ M$ as a shortcut for $\forall\ (\lambda x.\ M)$ and $\exists
+    x.\ M$ for $\exists\ (\lambda x.\ M)$.}
+\item $\mathcal{L}_{sem}(\synt{some}) = \lambda^{\circ} P Q.\ \exists
+  x.\ ((P\ x) \land (Q\ x))$
+\item $\mathcal{L}_{sem}(\synt{love}) = \textbf{love}$
+\item $\mathcal{L}_{sem}(\synt{man}) = \textbf{man}$
+\item $\mathcal{L}_{sem}(\synt{woman}) = \textbf{woman}$
+\end{itemize}
+
+Let us now consider the example terms of the $\Sigma_{Synt}$ signature
+from \ref{sssec:example-lex}. $t_3 = \synt{some}\ \synt{woman}$ will be
+interpreted as $\mathcal{L}_{sem}(t_3) =_{\beta} \lambda^{\circ}
+Q.\ \exists x.\ ((\textbf{woman}\ x) \land (Q\ x))$ having type $(e \limp
+t) \limp t$. It is a function that expects some predicate function and
+tests whether that predicate holds at least for some woman (some entity
+for which the $\textbf{woman}$ predicate holds).
+
+In $t_4 = \lambda^{\circ}
+x.\ (\synt{some}\ \synt{woman})\ (\lambda^{\circ}
+y.\ \synt{love}\ x\ y)$, we get $\mathcal{L}_{sem}(t_4) =_{\beta}
+\lambda^{\circ} x.\ \exists y.\ ((\textbf{woman}\ y) \land
+(\textbf{love}\ x\ y))$ of type $e \limp t$. What we have here is a
+predicate function testing whether the argument entity loves some woman.
+
+Finally, in $t_5 = (\synt{every}\ \synt{man})\ (\lambda^{\circ}
+x.\ (\synt{some}\ \synt{woman})\ (\lambda^{\circ}
+y.\ \synt{love}\ x\ y))$, we have $\mathcal{L}_{sem}(t_5) =_{\beta}
+\forall x.\ ((\textbf{man}\ x) \Rightarrow (\exists
+y.\ ((\textbf{woman}\ y) \land (\textbf{love}\ x\ y))))$ of type
+$t$. This computation yields a truth value, which tells us whether the
+proposition that every man loves some woman is true given some universe
+that the quantifiers range over and some values of the $\textbf{man}$,
+$\textbf{woman}$ and $\textbf{love}$ predicates. The term itself, in its
+$\beta$-normal form, can serve as our semantic representation for the
+proposition \emph{every man loves some woman}.
+
+We will finish our exposition of abstract categorial grammars by
+highlighting a prominent feature of the example grammar we have just
+described and that is its treatment of scope ambiguity. You might have
+noticed that $t_5 = (\synt{every}\ \synt{man})\ (\lambda^{\circ}
+x.\ (\synt{some}\ \synt{woman})\ (\lambda^{\circ}
+y.\ \synt{love}\ x\ y))$ is not the only term in
+$\Lambda(\Sigma_{Synt})$ for which $\mathcal{L}_{syntax}(t_5) =_{\beta}
+every\ +\ man\ +\ loves\ +\ some\ +\ woman$. We could just as well take $t_6
+= (\synt{some}\ \synt{woman})\ (\lambda^{\circ}
+y.\ (\synt{every}\ \synt{man})\ (\lambda^{\circ}
+x.\ \synt{love}\ x\ y))$ which is not $\beta$-equivalent to $t_5$. When
+we apply our syntactic lexicon to $t_6$, we see that we end up with an
+expression denoting the same string, $\mathcal{L}_{syntax}(t_6)
+=_{\beta} every\ +\ man\ +\ loves\ +\ some\ +\ woman$.
+
+However, when we consider a different interpretation for the abstract
+constants in $\Sigma_{Synt}$, for example our semantic lexicon
+$\mathcal{L}_{sem}$, we can see the differences rise to surface.
+
+$$
+\mathcal{L}_{sem}(t_5) =_{\beta} \forall x.\ ((\textbf{man}\ x)
+\Rightarrow (\exists y.\ ((\textbf{woman}\ y) \land
+(\textbf{love}\ x\ y))))
+$$
+$$
+\mathcal{L}_{sem}(t_6) =_{\beta} \exists y.\ ((\textbf{woman}\ y)
+\land (\forall x.\ ((\textbf{man}\ x) \Rightarrow
+(\textbf{love}\ x\ y))))
+$$
+
+When we try to parse some sentence using abstract categorial grammars,
+we are trying to find the abstract terms which upon transformation by
+the lexicon and $\beta$-reduction yield the sentence encoded in some
+object term. This is basically trying to invert the lexicon function,
+modulo $\beta$-equivalence. It is therefore no surprise then that we can
+find multiple abstract terms which all map to the same string term but
+which can be mapped to distinct terms using the semantic lexicon. It is
+this mechanism that enables abstract categorial grammars to handle
+ambiguity. When we reverse the scenario and go from semantic
+representations to strings, the same situation can occur (more than one
+abstract term having the same semantics but different surface strings)
+and this is what enables paraphrases.
+
 \bibliographystyle{plain}

### 2013-05-06 acg.tex
Moved acg stuff into acg.tex

--- 2013-05-06 thesis.tex
@@
-\documentclass{article}
-
-%% \usepackage{sdrt}
-\usepackage{gb4e}
-\usepackage{bussproofs}
-
-%% Linear implication from Alessio Guglielmi
-%% https://groups.google.com/forum/?fromgroups=#!topic/comp.text.tex/0B4C3F_BsVI
-\def\limp {\mathbin{{-}\mkern-3.5mu{\circ}}}
-
-\begin{document}
-
-\author{Ji\v{r}\'{i} Mar\v{s}\'{i}k}
-\title{Master Thesis}
-\maketitle
-
-\section{Preliminaries}
-
-\subsection{Motivation}
-
-To develop applications capable of understanding natural language, we
-have to analyze the discourse as a complex structure. The structure of
-the discourse and the rhetorical relations between its constituent
-propositions have shown to have an important effect on anaphora and on
-the semantic content of the discourse.
-
-Let us consider these examples from \cite{asher2003logics}.
-
-\begin{exe}
-  \ex \label{narxexp-nar} \begin{xlist}
-    \ex \label{narxexp-nar-a} Max fell.
-    \ex \label{narxexp-nar-b} John helped him up.
-  \end{xlist}
-  \ex \label{narxexp-exp} \begin{xlist}
-    \ex \label{narxexp-exp-a} Max fell.
-    \ex \label{narxexp-exp-b} John pushed him.
-  \end{xlist}
-\end{exe}
-
-In (\ref{narxexp-nar}), we intuitively recognize that the second
-proposition (\ref{narxexp-nar-b}) serves as a narrative continuation of
-(\ref{narxexp-nar-a}), whereas in (\ref{narxexp-exp}), the proposition
-(\ref{narxexp-exp-b}) serves as an explanation to the event described in
-(\ref{narxexp-exp-a}). This distinction has interesting consequences as
-to what we can infer from these two discourse excerpts. In the case of
-(\ref{narxexp-nar}), we can infer that the event described in
-(\ref{narxexp-nar-b}) occurred after the event described in
-(\ref{narxexp-nar-a}). In the latter case (\ref{narxexp-exp}), we can
-conversely infer that the events happened in the opposite order. We feel
-that a complete understanding of the two examples above entails
-inferring the correct temporal order and we would welcome a principled
-way to make this kind of decisions.
-
-\begin{exe}
-  \ex \label{salmon} \begin{xlist}
-    \ex Max had a lovely evening last night.
-    \ex He had a great meal.
-    \ex \label{salmon-here} He ate salmon.
-    \ex He devoured lots of cheese.
-    \ex He then won a dancing competition.
-  \end{xlist}
-\end{exe}
-
-%% \sdrtree{
-%%  &                      & \LAB{Max had a lovely evening} \\
-%%           & \LAB{He had a great meal} & & \LAB{He won a dancing competition} \\
-%% \LAB{He ate salmon} &     & \LAB{He devoured cheese}
-%% }
-
-The discourse in (\ref{salmon}) and its structure in (TODO) demonstrate
-another feature of discourse structures. First off, seeing the
-hiearchical structure of the discourse gives us important information
-about the granularity of the descriptions given in the individual
-propositions, information that could be useful for performing tasks such
-as text summarization.
-
-Furthermore, the discourse structure has grammatical consequences. It is
-thanks to our knowledge of the discourse structure that we can predict
-that a proposition like ``It was a beautiful pink.''  could not
-coherently follow our excerpt. SDRT, \cite{asher2003logics}, the theory
-of discourse structure that we adhere to, would not license a discourse
-structure in which the new proposition connects to (\ref{salmon-here})
-as its \emph{Elaboration}. As a consequence, it states that the pronoun
-``it'' cannot have the salmon as its antecedent, which is a constraint
-that is not enforced by prior theories of discourse such as DRT.
-
-\begin{exe}
-  \ex \label{diaimp} \begin{xlist}
-    \ex A: Smith doesn't seem to have a girlfriend.
-    \ex \label{diaimp-b} B: He's been paying lots of visits to New York lately.
-  \end{xlist}
-\end{exe}
-
-In the example dialogue (\ref{diaimp}), based on the prosody of
-proposition (\ref{diaimp-b}) the discourse structure would link the two
-propositions with an \emph{Evidence} or \emph{Counter-evidence}
-relation. Knowing this structure would allow us to either infer that B
-believes that Smith has a girlfriend in New York or that Smith doesn't
-have a girlfriend because he is too busy in New York.
-
-Based on the findings in \cite{asher2003logics}, we surmise that a clear
-picture of the discourse structure is essential to capturing the
-intended meaning of a discourse.
-
-To support our approach, we will stand on the shoulders of many giants,
-the first of them being Richard Montague. As in his approach
-\cite{montague1973proper}, we will assign functions and values as
-denotations of wordforms and use function application to compose them
-together to yield the denotations of phrases, propositions and
-discourses. To account for the discourse-level phenomena, we will defer
-to the theories of DRT \cite{kamp1993discourse} and SDRT
-\cite{asher2003logics}.
-
-The grammatical formalism of our choice for this task will be the
-Abstract Categorial Grammars (ACG) \cite{de2001towards}. This framework
-lets use lambda calculus to express the syntax-semantics interface in a
-fashion similar to Montague's. Furthermore, elegant techniques for using
-ACGs to implemenent DRT and SDRT using continuations have been
-discovered \cite{de2006towards} \cite{asher2011sdrt}
-\cite{asher2011montagovian}.
-
-Finally, our work will be supported by the existence of a corpus
-annotated with the rhetorical relations of SDRT, the Annodis corpus
-\cite{afantenos2012empirical}.
-
 \subsection{Abstract Categorial Grammars}
@@
 and this is what enables paraphrases.
-
-\bibliographystyle{plain}
-\bibliography{biblio}
-
-\end{document}

### 2013-05-06 thesis.tex
Moved acg stuff into acg.tex

--- 2013-05-06 thesis.tex
@@
 
-\subsection{Abstract Categorial Grammars}
-
-We present the grammatical framework in which we will develop our
-system. Abstract categorial grammars are built upon two mathematical
-structures, \emph{(higher-order) signatures} and \emph{lexicons}.
-
-A \textbf{higher-order signature} is a set of elements that we call
-\emph{constants}, each of which is associated with a type. Formally, it
-is defined as a triple $\Sigma = \mathopen{<}A, C, \tau\mathclose{>}$,
-where:
-\begin{itemize}
-  \item $C$ is the (finite) set of constants
-  \item $A$ is a (finite) set of atomic types
-  \item $\tau$ is the type associating mapping from $C$ to $\mathcal{T}(A)$,
-    the set of types built over $A$
-\end{itemize}
-
-In our case, $\mathcal{T}(A)$ is the implicative fragment of linear and
-intuitionistic logic with $A$ being the atomic propositions. This means
-that $\mathcal{T}(A)$ contains all the $a \in A$ and all the $\alpha \limp
-\beta$ and $\alpha \to \beta$ for $\alpha, \beta \in \mathcal{T}(A)$.
-
-A signature $\Sigma = \mathopen{<}A, C, \tau\mathclose{>}$, by itself,
-already lets us define an interesting set of structures, that is the set
-$\Lambda(\Sigma)$ of \emph{well-typed lambda terms} built upon the
-signature $\Sigma$. The set of well-typed terms and their types are
-established through the judgment schemas in Figure
-\ref{fig:type-judgments}.
-
-\begin{figure}
-  \begin{prooftree}
-    \AxiomC{$\emptyset, \Gamma_i \vdash_\Sigma c : \tau(c)$ (cons)}
-  \end{prooftree}
-  \begin{prooftree}
-    \AxiomC{$(x : \alpha), \Gamma_i \vdash_\Sigma x : \alpha$ (l-var)}
-  \end{prooftree}
-  \begin{prooftree}
-    \AxiomC{$\emptyset, (\Gamma_i, x : \alpha) \vdash_\Sigma x : \alpha$ (i-var)}
-  \end{prooftree}
-  \begin{prooftree}
-    \AxiomC{$(\Gamma_l, x : \alpha), \Gamma_i \vdash_\Sigma t : \beta$}
-    \RightLabel{(l-abs)}
-    \UnaryInfC{$\Gamma_l, \Gamma_i \vdash_\Sigma \lambda^{\circ} x. t : \alpha \limp \beta$}
-  \end{prooftree}
-  \begin{prooftree}
-    \AxiomC{$\Gamma_l, (\Gamma_i, x : \alpha) \vdash_\Sigma t : \beta$}
-    \RightLabel{(i-abs)}
-    \UnaryInfC{$\Gamma_l, \Gamma_i \vdash_\Sigma \lambda x. t : \alpha \to \beta$}
-  \end{prooftree}
-  \begin{prooftree}
-    \AxiomC{$\Gamma_l, \Gamma_i \vdash_\Sigma t : \alpha \limp \beta$}
-    \AxiomC{$\Delta_l, \Delta_i \vdash_\Sigma u : \alpha$}
-    \RightLabel{(l-app)}
-    \BinaryInfC{$(\Gamma_l, \Delta_l), (\Gamma_i, \Delta_i) \vdash_\Sigma (t\ u) : \beta$}
-  \end{prooftree}
-  \begin{prooftree}
-    \AxiomC{$\Gamma_l, \Gamma_i \vdash_\Sigma t : \alpha \to \beta$}
-    \AxiomC{$\emptyset, \Delta_i \vdash_\Sigma u : \alpha$}
-    \RightLabel{(i-app)}
-    \BinaryInfC{$\Gamma_l, (\Gamma_i, \Delta_i) \vdash_\Sigma (t\ u) : \beta$}
-  \end{prooftree}
-  \caption{\label{fig:type-judgments} Type judgment schemas of the
-    well-typed lambda terms $\Lambda(\Sigma)$ for a signature $\Sigma =
-    \mathopen{<}A, C, \tau\mathclose{>}$}
-\end{figure}
-
-The definition of a well-typed lambda term already gives us an
-interesting combinatorial structure. To make this structure even more
-useful, we often focus ourselves only on terms that have a specific
-\emph{distinguished type}. Using this notion of a signature of typed
-constants and some distinguished type, we can describe languages of,
-e.g. tree-like (and by extension string-like), lambda terms.
-
-Let us take a look at an example from
-\cite{pogodalla2007generalizing}. We will start with a very simple
-signature of string expressions with the concatenation operator,
-$\Sigma_{String} = \mathopen{<}A_{String}, C_{String},
-\tau_{String}\mathclose{>}$. Our signature will need only one atomic
-type, $A_{String} = \{\textsc{String}\}$. Our constants will be the
-wordforms of our example fragment ($every$, $some$, $man$, $woman$,
-$loves$), the empty string $\epsilon$ and the string concatention
-operator, $+$. $\tau_{String}$ assigns the type $\textsc{String}$ to all
-the wordforms and to $\epsilon$ and the type $\textsc{String} \limp
-\textsc{String} \limp \textsc{String}$\footnote{As a convention,
-  whenever we omit parentheses in a type, we presume the $\limp$ and
-  $\to$ type constructors always bind from the right, meaning that $a
-  \to b \to c$ should be read as $a \to (b \to c)$.}
-to the concatenation operator $+$.
-
-Now we can look at some well-typed terms from the string domain
-(i.e. elements of $\Lambda(\Sigma_{String})$). The term $t_1 =
-some\ +\ woman$\footnote{Here we introduce another notational
-  convention. The unadorned way of writing this term would be $t_1 =
-  (+\ some)\ woman$. Firstly, we will allow ourselves to drop the
-  parentheses, meaning that any string of expressions $f\ x\ y$ should
-  be read as $(f\ x)\ y$. Secondly, we will admit a specific infix
-  notation for some constants which denote functions of two
-  arguments. This will then let us write $x\ +\ y$ instead of
-  $+\ x\ y$.} denotes the concatenation of $some$ and $woman$, which is
-the phrase $some\ woman$; its type is $\textsc{String}$. A term such as
-$t_2 = \lambda^{\circ} x. x\ +\ man$ denotes a function that appends the
-string $man$ to its argument; its type is $\textsc{String} \limp
-\textsc{String}$. If we restrict the set $\Lambda(\Sigma_{String})$ only
-to terms having the type $\textsc{String}$, we get the set of
-expressions which denote strings.
-
-The idea of a signature is coupled with that of a \textbf{lexicon},
-which is a mapping between two different signatures (mapping the
-constants of one into well-typed terms of the other). Formally speaking,
-a lexicon $\mathcal{L}$ from a signature $\Sigma_1 = \mathopen{<}A_1, C_1,
-\tau_1\mathclose{>}$ (which we call the abstract signature) to a
-signature $\Sigma_2 = \mathopen{<}A_2, C_2, \tau_2\mathclose{>}$ (which
-we call the object signature) is a pair $\mathopen{<}F, G\mathclose{>}$
-such that:
-\begin{itemize}
-\item $G$ is a mapping from $C_1$ to $\Lambda(\Sigma_2)$ assigning to
-  every constant of the abstract signature a term in the object
-  signature, which can be understood as its
-  interpretation/implementation/realization.
-\item $F$ is a mapping from $A_1$ to $\mathcal{T}(A_2)$ which links the
-  abstract-level types with the object-level types that they realized
-  in.
-\item $F$ and $G$ are compatible, meaning that for any $c \in C_1$, we
-  have $\vdash_{\Sigma_2} G(c) : \hat{F}(\tau_1(c))$ (we will be using
-  $\hat{F}$ and $\hat{G}$ to refer to the homomorphic extensions of $F$
-  and $G$ to $\mathcal{T}(A_1)$ and $\Lambda(\Sigma_1)$ respectively).
-\end{itemize}
-
-An abstract categorial grammar for us will then be just a collection of
-signatures with their distinguished types and lexicons connecting these
-signatures. A common pattern will have us using two signatures for the
-surface forms of utterances (strings) and their logical forms (logical
-propositions) and an abstract signature which is connected to both of
-the object signatures via lexicons. Parsing is then just a matter of
-inverting the surface lexicon to get the abstract term and then applying
-to it the logical lexicon. Generation is symmetric, we simply invert the
-logical lexicon and apply the surface lexicon.
-
-\subsubsection{Example ACG}
-
-We will look at an example of an ACG from
-
-$\Sigma_{Synt}$ will be our abstract signature with $A_{Synt} = \{np, n,
-s\}$, $C_{Synt} = \{C_{every}, C_{some}, C_{love}, C_{man}, C_{woman}\}$
-and $\tau_{Synt}$ as follows:
-\begin{itemize}
-\item
-\end{itemize}
-
-...
-
-\subsection{Interaction Grammars}
+\input{acg}
 

### 2013-05-06 thesis.tex
Moved motivation into motivation.tex

@@
 
-\subsection{Motivation}
-
-To develop applications capable of understanding natural language, we
-have to analyze the discourse as a complex structure. The structure of
-the discourse and the rhetorical relations between its constituent
-propositions have shown to have an important effect on anaphora and on
-the semantic content of the discourse.
-
-Let us consider these examples from \cite{asher2003logics}.
-
-\begin{exe}
-  \ex \label{narxexp-nar} \begin{xlist}
-    \ex \label{narxexp-nar-a} Max fell.
-    \ex \label{narxexp-nar-b} John helped him up.
-  \end{xlist}
-  \ex \label{narxexp-exp} \begin{xlist}
-    \ex \label{narxexp-exp-a} Max fell.
-    \ex \label{narxexp-exp-b} John pushed him.
-  \end{xlist}
-\end{exe}
-
-In (\ref{narxexp-nar}), we intuitively recognize that the second
-proposition (\ref{narxexp-nar-b}) serves as a narrative continuation of
-(\ref{narxexp-nar-a}), whereas in (\ref{narxexp-exp}), the proposition
-(\ref{narxexp-exp-b}) serves as an explanation to the event described in
-(\ref{narxexp-exp-a}). This distinction has interesting consequences as
-to what we can infer from these two discourse excerpts. In the case of
-(\ref{narxexp-nar}), we can infer that the event described in
-(\ref{narxexp-nar-b}) occurred after the event described in
-(\ref{narxexp-nar-a}). In the latter case (\ref{narxexp-exp}), we can
-conversely infer that the events happened in the opposite order. We feel
-that a complete understanding of the two examples above entails
-inferring the correct temporal order and we would welcome a principled
-way to make this kind of decisions.
-
-\begin{exe}
-  \ex \label{salmon} \begin{xlist}
-    \ex Max had a lovely evening last night.
-    \ex He had a great meal.
-    \ex \label{salmon-here} He ate salmon.
-    \ex He devoured lots of cheese.
-    \ex He then won a dancing competition.
-  \end{xlist}
-\end{exe}
-
-%% \sdrtree{
-%%  &                      & \LAB{Max had a lovely evening} \\
-%%           & \LAB{He had a great meal} & & \LAB{He won a dancing competition} \\
-%% \LAB{He ate salmon} &     & \LAB{He devoured cheese}
-%% }
-
-The discourse in (\ref{salmon}) and its structure in (TODO) demonstrate
-another feature of discourse structures. First off, seeing the
-hiearchical structure of the discourse gives us important information
-about the granularity of the descriptions given in the individual
-propositions, information that could be useful for performing tasks such
-as text summarization.
-
-Furthermore, the discourse structure has grammatical consequences. It is
-thanks to our knowledge of the discourse structure that we can predict
-that a proposition like ``It was a beautiful pink.''  could not
-coherently follow our excerpt. SDRT, \cite{asher2003logics}, the theory
-of discourse structure that we adhere to, would not license a discourse
-structure in which the new proposition connects to (\ref{salmon-here})
-as its \emph{Elaboration}. As a consequence, it states that the pronoun
-``it'' cannot have the salmon as its antecedent, which is a constraint
-that is not enforced by prior theories of discourse such as DRT.
-
-\begin{exe}
-  \ex \label{diaimp} \begin{xlist}
-    \ex A: Smith doesn't seem to have a girlfriend.
-    \ex \label{diaimp-b} B: He's been paying lots of visits to New York lately.
-  \end{xlist}
-\end{exe}
-
-In the example dialogue (\ref{diaimp}), based on the prosody of
-proposition (\ref{diaimp-b}) the discourse structure would link the two
-propositions with an \emph{Evidence} or \emph{Counter-evidence}
-relation. Knowing this structure would allow us to either infer that B
-believes that Smith has a girlfriend in New York or that Smith doesn't
-have a girlfriend because he is too busy in New York.
-
-Based on the findings in \cite{asher2003logics}, we surmise that a clear
-picture of the discourse structure is essential to capturing the
-intended meaning of a discourse.
-
-To support our approach, we will stand on the shoulders of many giants,
-the first of them being Richard Montague. As in his approach
-\cite{montague1973proper}, we will assign functions and values as
-denotations of wordforms and use function application to compose them
-together to yield the denotations of phrases, propositions and
-discourses. To account for the discourse-level phenomena, we will defer
-to the theories of DRT \cite{kamp1993discourse} and SDRT
-\cite{asher2003logics}.
-
-The grammatical formalism of our choice for this task will be the
-Abstract Categorial Grammars (ACG) \cite{de2001towards}. This framework
-lets use lambda calculus to express the syntax-semantics interface in a
-fashion similar to Montague's. Furthermore, elegant techniques for using
-ACGs to implemenent DRT and SDRT using continuations have been
-discovered \cite{de2006towards} \cite{asher2011sdrt}
-\cite{asher2011montagovian}.
-
-Finally, our work will be supported by the existence of a corpus
-annotated with the rhetorical relations of SDRT, the Annodis corpus
-\cite{afantenos2012empirical}.
+\input{motivation}
 

### 2013-05-06 motivation.tex
Adding stray motivation.tex

\subsection{Motivation}

To develop applications capable of understanding natural language, we
have to analyze the discourse as a complex structure. The structure of
the discourse and the rhetorical relations between its constituent
propositions have shown to have an important effect on anaphora and on
the semantic content of the discourse.

Let us consider these examples from \cite{asher2003logics}.

\begin{exe}
  \ex \label{narxexp-nar} \begin{xlist}
    \ex \label{narxexp-nar-a} Max fell.
    \ex \label{narxexp-nar-b} John helped him up.
  \end{xlist}
  \ex \label{narxexp-exp} \begin{xlist}
    \ex \label{narxexp-exp-a} Max fell.
    \ex \label{narxexp-exp-b} John pushed him.
  \end{xlist}
\end{exe}

In (\ref{narxexp-nar}), we intuitively recognize that the second
proposition (\ref{narxexp-nar-b}) serves as a narrative continuation of
(\ref{narxexp-nar-a}), whereas in (\ref{narxexp-exp}), the proposition
(\ref{narxexp-exp-b}) serves as an explanation to the event described in
(\ref{narxexp-exp-a}). This distinction has interesting consequences as
to what we can infer from these two discourse excerpts. In the case of
(\ref{narxexp-nar}), we can infer that the event described in
(\ref{narxexp-nar-b}) occurred after the event described in
(\ref{narxexp-nar-a}). In the latter case (\ref{narxexp-exp}), we can
conversely infer that the events happened in the opposite order. We feel
that a complete understanding of the two examples above entails
inferring the correct temporal order and we would welcome a principled
way to make this kind of decisions.

\begin{exe}
  \ex \label{salmon} \begin{xlist}
    \ex Max had a lovely evening last night.
    \ex He had a great meal.
    \ex \label{salmon-here} He ate salmon.
    \ex He devoured lots of cheese.
    \ex He then won a dancing competition.
  \end{xlist}
\end{exe}

%% \sdrtree{
%%  &                      & \LAB{Max had a lovely evening} \\
%%           & \LAB{He had a great meal} & & \LAB{He won a dancing competition} \\
%% \LAB{He ate salmon} &     & \LAB{He devoured cheese}
%% }

The discourse in (\ref{salmon}) and its structure in (TODO) demonstrate
another feature of discourse structures. First off, seeing the
hiearchical structure of the discourse gives us important information
about the granularity of the descriptions given in the individual
propositions, information that could be useful for performing tasks such
as text summarization.

Furthermore, the discourse structure has grammatical consequences. It is
thanks to our knowledge of the discourse structure that we can predict
that a proposition like ``It was a beautiful pink.''  could not
coherently follow our excerpt. SDRT, \cite{asher2003logics}, the theory
of discourse structure that we adhere to, would not license a discourse
structure in which the new proposition connects to (\ref{salmon-here})
as its \emph{Elaboration}. As a consequence, it states that the pronoun
``it'' cannot have the salmon as its antecedent, which is a constraint
that is not enforced by prior theories of discourse such as DRT.

\begin{exe}
  \ex \label{diaimp} \begin{xlist}
    \ex A: Smith doesn't seem to have a girlfriend.
    \ex \label{diaimp-b} B: He's been paying lots of visits to New York lately.
  \end{xlist}
\end{exe}

In the example dialogue (\ref{diaimp}), based on the prosody of
proposition (\ref{diaimp-b}) the discourse structure would link the two
propositions with an \emph{Evidence} or \emph{Counter-evidence}
relation. Knowing this structure would allow us to either infer that B
believes that Smith has a girlfriend in New York or that Smith doesn't
have a girlfriend because he is too busy in New York.

Based on the findings in \cite{asher2003logics}, we surmise that a clear
picture of the discourse structure is essential to capturing the
intended meaning of a discourse.

To support our approach, we will stand on the shoulders of many giants,
the first of them being Richard Montague. As in his approach
\cite{montague1973proper}, we will assign functions and values as
denotations of wordforms and use function application to compose them
together to yield the denotations of phrases, propositions and
discourses. To account for the discourse-level phenomena, we will defer
to the theories of DRT \cite{kamp1993discourse} and SDRT
\cite{asher2003logics}.

The grammatical formalism of our choice for this task will be the
Abstract Categorial Grammars (ACG) \cite{de2001towards}. This framework
lets use lambda calculus to express the syntax-semantics interface in a
fashion similar to Montague's. Furthermore, elegant techniques for using
ACGs to implemenent DRT and SDRT using continuations have been
discovered \cite{de2006towards} \cite{asher2011sdrt}
\cite{asher2011montagovian}.

However, abstract categorial grammars are quite young and no significant
grammar has been developed under this framework. To facilitate the
creation of such a grammar, we will borrow heavily from the Frigram
Interaction Grammar \cite{perrier2007french} and its linguistic
resources.

Finally, our work will be supported by the existence of a corpus
annotated with the rhetorical relations of SDRT, the Annodis corpus
\cite{afantenos2012empirical}.

### 2013-05-06 motivation.tex
Motivating IGs

@@
 Interaction Grammar \cite{perrier2007french} and its linguistic
-resources.
+resources. Namely, we will rely on the fact that Frigram is defined
+separately from its lexicon, Frilex, which we will use in its
+entirety. Frigram, the grammar itself, can also serve us as a guideline
+when designing our own grammar thanks to the close ties between the
+formalisms of interaction grammars and abstract categorial grammars
+\cite{perrier1999intuitionistic}.
 

### 2013-05-10 ig.tex
Wrote up the exposition on IGs

\subsection{Interaction Grammars}

We will briefly discuss one more grammar formalism and that is the
formalism of Interaction Grammars \cite{guillaume2009interaction}.
Interaction grammars are a grammatical formalism centered around the
concept of \emph{polarity}. An interaction grammar is a set of
(underspecified) tree descriptions which define a set of trees that can
be constructed by superposing some of these tree descriptions while
respecting the polarities described in the nodes of the individual tree
descriptions.

We will illustrate the objects and mechanisms of interaction grammars on
a simple example sentence.

\begin{exe}
  \ex \label{ex:ig} Jean la voit.
\end{exe}

\begin{figure}
  \centering
  \includegraphics[scale=0.25]{ig-parse.pdf}
  \caption{\label{fig:ig-parse} Syntactic tree of sentence
    (\ref{ex:ig}). The numbers at the top of the nodes are labels of
    nodes from the elementary polarized tree descriptions that generated
    the syntactic tree (see Figure \ref{fig:ig-eptds}).}
\end{figure}

Figure \ref{fig:ig-parse} shows the syntactic tree assigned by Frigram
to the sentence in (\ref{ex:ig}). It is a rooted ordered tree with the
topmost node being the root. Solid lines indicate immediate dominance
with the higher node being the parent and the lower the child. The
ordering of children is given by the arrows, which signify the immediate
precedence relation between sister nodes.

The nodes in the tree are of three different kinds with respect to the
phonological form of their subtree. We recognize the \emph{anchor
  nodes}, which are displayed in vivid yellow and which are leaf nodes
containing some non-empty string as their phonological content (written
in a gray rectangle at the top of the node). Then we have the
\emph{empty} nodes, which are drawn in white and whose phonological form
is empty. Finally, there are the pale yellow \emph{non-empty} nodes,
internal nodes whose phonological content (the yield of the subtree) is
not empty.

The nodes of the syntactic tree are also decorated with
features. Interaction grammars are a formalism enabling us to define
sets of the structures described above, the \emph{syntactic
  trees}. Similar to Tree Adjoining Grammars, interaction grammars are
formulated in terms of a set of elementary structures which can combine
to produce the final output structures. However, unlike in tree
adjoining grammars, the output structure is not constructed from the
elementary structures via some set of algebraic operations (substitution
and adjunction in case of TAG). Instead, \emph{polarized tree
  descriptions} (PTDs), the elementary structures of interaction
grammars, impose constraints on the final structure and a structure is
said to be generated by some PTDs if it is a minimal structure
satisfying those constraints (it is a \emph{minimal model}).

\begin{figure}
  \centering
  \includegraphics[scale=0.25]{ig-eptds.pdf}
  \caption{\label{fig:ig-eptds} The elementary polarized tree
    descriptions used to generate the sentence (\ref{ex:ig}).}
\end{figure}

In Figure~\ref{fig:ig-eptds}, we can see the elementary polarized tree
descriptions (EPTDs) that generated the parse tree in
Figure~\ref{fig:ig-parse}.

As we said before, a PTD is a set of constraints on some syntactic
tree. Let us expound on what constraints the structures of
Figure~\ref{fig:ig-eptds} impose.

A node in a PTD can be read as a statement that there must exist a node
in the final syntactic tree that has compatible values for all the
features and that carries the same phonological string, if any. Such a
node of the syntactic tree is then called the \emph{interpretation} of
the PTD node (in Figure~\ref{fig:ig-parse}, every node of the syntactic
tree bears a list of the PTD nodes that it interprets, so you can see
exactly how the interpretation function works in our example).

A solid line between two nodes in a PTD means that the interpretations
of these two nodes must be in a parent-child relationship
(\emph{immediate dominance}). A dashed green arrow tells us that the
interpretations have to be sister nodes with the former preceding the
latter in the ordered tree (\emph{precedence}). Note that not all sister
nodes in a PTD have to be linked with this precedence relation. See for
example the nodes (1,1) and (1,5) in the EPTD of the clitic pronoun
$la$.

The formalism also allows us to specify \emph{immediate precedence} and
general \emph{dominance}. The latter is useful for modelling unbounded
dependencies such as those between a relative pronoun and its trace in
some embedded clause of the relative clause, but it is not used in our
present example. Finally, one kind of constraint that is used in our
example is the orange rectangle in node (3,0), which requires that the
interpretation of (3,0) is the right-most daughter amongst its siblings.

Now that we have covered the structural constraints imposed by the tree
descriptions, we will turn our attention to the defining characteristic
of interaction grammars, the polarities. As you have noticed on
Figure~\ref{fig:ig-eptds}, some of the features in our PTDs are
annotated with special symbols and colors. These denote the different
polarities and are used by the formalism for two distinct purposes: the
positive ($\textcolor{red}{\rightarrow}$) and negative
($\textcolor{blue}{\leftarrow}$) polarities are used to model the
resource sensitivity of language, while the virtual polarities
($\textcolor{purple}{\sim}$) are used to match against the
surrounding context.

The way the polarities are handled is that every model (output syntactic
tree) is required to have only saturated polarities on its features
(i.e. no positive, negative or virtual polarities). Whenever more than
one node of the PTD is mapped onto a single node of the syntactic tree,
the polarities of each feature are combined. The combination mechanism
allows to combine a positively polarized feature with a negatively
polarized one to yield a saturated one. Virtual features can be factored
out only by combining with a saturated polarity.

If we look at the PTDs of Figure~\ref{fig:ig-eptds}, we can see this
polarity mechanism in action. The EPTD of $Jean$ has a positive $cat$
feature in its root node saying that it provides one $np$, and a
negative $funct$ feature saying that it expects some function. In the
EPTD of $voit$, we have two nodes for the subject and object, both of
them expecting an $np$ and providing them the $subj$ and $obj$
functions, respectively. The root of the EPTD than provides a complete
sentence of category $s$ and expects some function that this sentence
will play. The full stop EPTD finalizes the sentence by accepting a node
with category $s$ and giving it a $void$ function.

The clitic pronoun $la$ participates in the positive/negative resource
management system as well, since using the clitic fills up the object
slot in the valency of a verb. The EPTD of $la$ also relies heavily on
virtual features and uses them to specify in more detail the context in
which this rule is applicable, so as to avoid over-generation.

Now that we understand the constraints imposed by PTDs, we can start to
see that the syntactic tree in Figure~\ref{fig:ig-parse} is truly a
model of the PTDS given in Figure~\ref{fig:ig-eptds} and more than that,
it is the only minimal model. To illustrate more clearly how the
polarities and constraints of the individual PTDs end up generating the
syntactic tree of Figure~\ref{fig:ig-parse}, we can look at the result
of superposing the EPTDs of $la$ and $voit$ by merging the nodes (1,2)
and (1,5) with (2,2) and (2,4) respectively on
Figure~\ref{fig:ig-partial}.

\begin{figure}
  \centering
  \includegraphics[scale=0.25]{ig-partial.pdf}
  \caption{\label{fig:ig-partial} The result of merging the EPTDs of
    $la$ and $voit$ by merging the nodes (1,5) and (1,5) with (2,2) and
    (2,4), respectively.}
\end{figure}

### 2013-05-13 ig.tex
Talked about Frigram a little

@@
 polarized one to yield a saturated one. Virtual features can be factored
-out only by combining with a saturated polarity.
+out only by combining with a saturated polarity (they are somewhat
+analogous to the $=_{c}$ constraints of LFG \cite{kaplan1982lexical}).
 
@@
 them expecting an $np$ and providing them the $subj$ and $obj$
-functions, respectively. The root of the EPTD than provides a complete
+functions, respectively. The root of the EPTD then provides a complete
 sentence of category $s$ and expects some function that this sentence
@@
 see that the syntactic tree in Figure~\ref{fig:ig-parse} is truly a
-model of the PTDS given in Figure~\ref{fig:ig-eptds} and more than that,
+model of the PTDs given in Figure~\ref{fig:ig-eptds} and more than that,
 it is the only minimal model. To illustrate more clearly how the
@@
 \end{figure}
+
+\subsubsection{Frigram, an interaction grammar}
+
+The reason we are interested in interaction grammars is because of the
+existence of Frigram, a large scale grammar of French which is
+lexicalized by Frilex, an independent lexical resource that we will be
+building our grammar on as well. Furthermore, as we will see
+in~\ref{sssec:link-ig-acg}, interaction grammars are closely linked with
+abstract categorial grammars, which makes Frigram a suitable grammar to
+use as a guide when developing our grammar.
+
+In this section, we will introduce the metagrammatical structure of
+Frigram and talk about how it uses the formalism of interaction grammars
+to solve several tricky linguistic phenomena.
+
+Frigram is a wide-coverage interaction grammar of French. As we
+mentioned before, an interaction grammar is a set of elementary
+polarized tree descriptions. In a lexicalized interaction grammar, each
+of these EPTDs is connected to a specific wordform whose particular use
+it describes. Given the amount of wordforms that a useable grammar of
+French would need to cover, the number of EPTDs in our grammar could
+easily reach hundreds of thousands, if not millions.
+
+The first step in fighting this explosion is to factor out irrelevant
+differences between similar wordforms. For this purpose, Frilex was
+created. Frilex is a morpho-syntactic lexicon of French compiled from
+various other pre-existing lexicons of French. It is in effect a large
+relation linking the wordforms of French to \emph{hypertags}, feature
+structures describing the morphological properties of the wordforms and
+their syntactic valencies.
+
+With Frilex in place, Frigram can be defined as a set of unanchored
+EPTDs which are paired with feature structures which delimit the subset
+of the Frilex items to which the EPTDs apply. This kind of simple
+``metagrammar'' already saves us a lot of effort, the number of
+unanchored EPTDs defined by Frigram is somewhere around 4000 (TODO: Cite
+the Frig documentation somehow).
+
+However, defining some 4000 EPTDs manually still seems like a very
+tedious, error prone and hard to maintain approach. Many of the EPTDs
+have to repeatedly describe common phenomena such as subject-object
+agreement or predicate-argument saturation. Similarly as in other
+software endeavors, it would be preferable to define these common
+patterns in some reusable module and compose EPTDs from these building
+blocks. This is where XMG steps in.
+
+XMG is a metagrammar compiler adding yet another level of indirection
+between what the grammar author writes and what ends up in the
+bottom-level interaction grammar. XMG provides a language that lets the
+grammar author define the lexical items of his grammar and to combine
+these definitions to yield new and more elaborate lexical items.
+
+In the case of Frigram, the lexical items (termed \emph{classes} in XMG)
+take the shape of PTDs coupled with feature structures which
+circumscribe the interface to Frilex. Both tree descriptions and feature
+structures have very natural ways of composing together, by conjunction
+and unification, respectively. This, alongside with disjunction, are the
+chief tools that the grammar authors can use (and in the case of
+Frigram, have used) to succintly define their grammar.
+
+With the added capability of composing classes, the definition of
+Frigram reduces to about 400 class definitions. Of these classes, 160
+are terminal, meaning that they actually describe lexical items that are
+to be included in the final grammar. Thanks to the disjunction
+composition operator of XMG, these 160 classes end up generating the
+staggering 4000 EPTDs.
+
+\subsubsection{The link with abstract categorial grammars}
+\label{sssec:link-ig-acg}

### 2013-05-13 ig.tex
Wrote about negation in Frigram

@@
 The nodes of the syntactic tree are also decorated with
-features. Interaction grammars are a formalism enabling us to define
-sets of the structures described above, the \emph{syntactic
-  trees}. Similar to Tree Adjoining Grammars, interaction grammars are
-formulated in terms of a set of elementary structures which can combine
-to produce the final output structures. However, unlike in tree
-adjoining grammars, the output structure is not constructed from the
-elementary structures via some set of algebraic operations (substitution
-and adjunction in case of TAG). Instead, \emph{polarized tree
-  descriptions} (PTDs), the elementary structures of interaction
-grammars, impose constraints on the final structure and a structure is
-said to be generated by some PTDs if it is a minimal structure
-satisfying those constraints (it is a \emph{minimal model}).
+features.
+
+Interaction grammars are a formalism enabling us to define sets of the
+structures described above, the \emph{syntactic trees}. Similar to Tree
+Adjoining Grammars, interaction grammars are formulated in terms of a
+set of elementary structures which can combine to produce the final
+output structures. However, unlike in tree adjoining grammars, the
+output structure is not constructed from the elementary structures via
+some set of algebraic operations (substitution and adjunction in case of
+TAG). Instead, \emph{polarized tree descriptions} (PTDs), the elementary
+structures of interaction grammars, impose constraints on the final
+structure and a structure is said to be generated by some PTDs if it is
+a minimal structure satisfying those constraints (it is a \emph{minimal
+  model}).
 
@@
 The formalism also allows us to specify \emph{immediate precedence} and
-general \emph{dominance}. The latter is useful for modelling unbounded
+(large) \emph{dominance}. The latter is useful for modelling unbounded
 dependencies such as those between a relative pronoun and its trace in
 some embedded clause of the relative clause, but it is not used in our
-present example. Finally, one kind of constraint that is used in our
-example is the orange rectangle in node (3,0), which requires that the
-interpretation of (3,0) is the right-most daughter amongst its siblings.
+present example.
+
+Finally, one kind of constraint that is used in our example is the
+orange rectangle in node (3,0), which requires that the interpretation
+of (3,0) is the rightmost daughter amongst its siblings.
 
@@
 ($\textcolor{blue}{\leftarrow}$) polarities are used to model the
-resource sensitivity of language, while the virtual polarities
-($\textcolor{purple}{\sim}$) are used to match against the
-surrounding context.
+resource sensitivity of languages, while the virtual polarities
+($\textcolor{purple}{\sim}$) are used to for pattern matching against
+the context.
 
@@
 the polarities of each feature are combined. The combination mechanism
-allows to combine a positively polarized feature with a negatively
-polarized one to yield a saturated one. Virtual features can be factored
-out only by combining with a saturated polarity (they are somewhat
+allows us to combine a positively polarized feature with a negatively
+polarized one to yield a saturated one. Virtual polarities can only be
+eliminated by combining them with a saturated polarity (they are
 analogous to the $=_{c}$ constraints of LFG \cite{kaplan1982lexical}).
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.25]{ig-neg.pdf}
+  \caption{\label{fig:ig-neg} EPTDs of $aucun$ and $ne$ demonstrating
+    the way polarities are used to model French negation.}
+\end{figure}
+
+We will briefly discuss how interaction grammars, and Frigram in
+particular, handle a tricky linguistic phenomenon in an interesting way.
+In French, negation is signalled by the particle $ne$ which must be
+accompanied by one of several designated determiners, pronouns or
+adverbs. The particle $ne$ assumes a position right before the inflected
+verb, but its partner, e.g. the determiner $aucun$ can appear in a rich
+variety of positions:
+
+\begin{exe}
+  \ex \label{ex:aucun-suj} Aucun tatou ne court.
+  \ex \label{ex:aucun-obj} Jean n'aime l'odeur d'aucun tatou.
+  \ex \label{ex:aucun-bad} * Le tatou qu'aucun loup chasse ne court.
+  \ex \label{ex:aucun-good} Le tatou qu'aucun loup ne chasse court.
+\end{exe}
+
+Sentences (\ref{ex:aucun-suj}) and (\ref{ex:aucun-obj}) demonstrate that
+the $aucun$ determiner can be used both in the subject and object
+positions, as a determiner of either one of the arguments directly or of
+a noun phrase which complements one of them. However, it is not
+admissible to place the determiner in an embedded clause as it is in
+(\ref{ex:aucun-bad}). If the determiner $aucun$ is used in an embedded
+clause, it is the embedded clause itself that must be negated, as in
+sentence (\ref{ex:aucun-good}).
+
+This example demonstrates two points of interest to us.
+
+First, it uses a new polarized feature, $neg$, to model the fact that
+the two wordforms which enable negation must always co-occur. The
+particle $ne$ provides a positive $neg$ feature to the clause in which
+it occurs and this positive polarity must be saturated by a negative
+polarity contributed by one of the possible partner words, such as
+$aucun$. This shows that the usefulness of positive and negative
+polarities extends beyond the $cat$/$funct$ pair that we have already
+seen and which is used to model predicate/argument structure in a
+similar way as was already done in tree adjoining grammars or categorial
+grammars.
+
+Second, the determiner $aucun$ can occur deep, e.g. inside a complement
+of one of the arguments, but it cannot occur everywhere, as is
+demonstrated by (\ref{ex:aucun-bad}), where we try to put the determiner
+inside another embedded clause. There is a way to express this kind of
+constraint in interaction grammars that we have not revealed
+yet. Whenever we use a (large) dominance (a top-down green dashed line),
+we usually want the range of nodes that this dominance relation can
+cross to be somehow restricted. In interaction grammars, we can
+associate a feature structure with a dominance constraint which will
+force all the nodes that will span the distance between the two nodes
+linked by the dominance constraint to unify with that feature
+structure.
+
+In the notation used in our illustrations (which were generated by the
+Leopar parser\footnote{\url{http://leopar.loria.fr/}}), this is conveyed
+by putting all the restricting features as virtual features in a new
+node and splicing the new node in between the two nodes linked by the
+dominance constraint before (see the node (0,4) in
+Figure~\ref{fig:ig-neg}, which merely serves to ensure that all the
+nodes on the path from (0,3) to (0,2) have either $np$ or $pp$ as the
+value of their $cat$ feature). When using this notation, the (large)
+dominance constraint should then be read as not only stating that the
+top node must be an ancestor to the bottom node, but also that any
+intermediate nodes between the ancestor and successor must match the
+ancestor's features.
+
 \subsubsection{The link with abstract categorial grammars}

### 2013-05-15 ig.tex
Wrote about the IG-ACG link

@@
 \label{sssec:link-ig-acg}
+
+Both abstract categorial grammars and interaction grammars, having come
+out of the same research team, are both closely related to linear
+logic. Not only are these two formalisms based on the same logical
+framework, there is a surprisingly direct connection between the
+principles guiding the syntactic composition of constants having
+intuitionistic implicative linear types in abstract categorial grammars
+and of polarized tree descriptions in interaction grammars.
+
+Perrier \cite{perrier1999intuitionistic} proved an interesting result
+which bridges these two worlds. His theorem states that every IILL
+(intuitionistic implicative linear logic) sequent $F_1, ..., F_n \vdash
+G$ is provable if and only if the syntactic description $D((F_1 \limp
+... \limp F_n \limp G)^+)$ is valid. The second premise deserves some
+elaboration.
+
+What we mean by a \emph{syntactic description} is something similar to a
+polarized tree description, albeit more minimalistic. Syntactic
+descriptions only talk about immediate dominance and (large) dominance,
+there are no precedence constraints on sister nodes and (large)
+dominance cannot be restricted by a feature structure. Nodes themselves
+are no longer even represented by feature structures, but by simple
+atomic categories paired with polarities. Every node in a syntactic
+description is thus associated with a single atomic category and a
+polarity which is either positive or negative.
+
+A syntactic description that is \emph{valid} is one for which there
+exists a model. A model of a syntactic description is defined in the
+same manner as for polarized tree descriptions, but instead of unifying
+feature structures, we simply demand that all nodes of the description
+which map to a node in the model must have the same category and the
+number of positively polarized nodes must equal the number of negatively
+polarized nodes.
+
+Finally, we explain the term $D((F_1 \limp ... \limp F_n \limp
+G)^+)$. The $+$ operation recursively assigns positive and negative
+polarities to all formula occurrences within $F_1 \limp ... \limp F_n
+\limp G$. These polarities then drive the recursive definition of $D$,
+which maps this polarized formula into a syntactic description. Since we
+will not cover the theorem deeply in our treatise, we will settle for
+knowing that the composition of the positive polarization operation $+$
+and the syntactic description producing operation $D$ maps an IILL
+formula into a syntactic description such that above-stated theorem
+holds. Furthermore, this mapping can be easily constructed since Perrier
+gives straightforward ways of computing the results of both operations.
+
+This theorem establishes a striking symmetry between the two
+formalisms. Let us consider the case of parsing a sentence using a
+lexicalized grammar in both formalism. First, the sentence is tokenized
+into wordforms and for each wordform, a lexical item is selected, be it
+a typed constant in some abstract signature for ACGs or an EPTD for IGs.
+
+In ACGs, we will try to take the typed constants and to use them to
+produce a term having some distinguished type $s$. Thanks to the
+Curry-Howard correspondence, constructing a term having a given type is
+nothing more than proving that term's type as a formula in the logic of
+our type system. What this means is that our parsing problem boils down
+to proving the IILL sequent $\tau_1, ..., \tau_n \vdash s$ (we omit the
+inclusion of intuitionistic non-linear implications since they are not
+relevant for parsing sentences). The sentence thus becomes parseable by
+our grammar under the given lexical selection if and only if this
+sequent is provable and furthermore, the proof of the sequent is our
+desired syntactic structure in a disguise.
+
+Now in IGs, we will conjoin all the EPTDs into a single complex
+polarized tree description. Parsing is then just a matter of finding a
+model for this description. The sentence is therefore parseable by our
+grammar under the given lexical selection if and only if this
+description is valid and furthermore, the model which proves the
+description valid is our desired syntactic structure.
+
+This symmetry turns Perrier's theorem into a strong claim that lets us
+transform the types of a signature into elementary syntactic
+descriptions such that some selection of descriptions will generate some
+tree if and only if the selection of the corresponging typed constants
+generated some properly-typed term. Inverting this transformation would
+then let us take the elementary syntactic descriptions of an IG like
+Frigram and turn them into types for an abstract signature of our ACG.
+
+However, this approach is not sufficient. The $D \circ +$ transformation
+is far from a bijection.
+
+Firstly, it is not injective and several types end up being represented
+by the same syntactic description. What this means is that the
+transformation ignores the insignificant differences between formulas
+such as $a \limp (b \limp c)$ and $b \limp (a \limp c)$. This would not
+pose a serious issue though, as when we would try to invert $D \circ +$,
+we could simply choose one of the possible formulas which fit the
+syntactic description as our canonical representation.
+
+A bigger problem is posed by the fact that $D \circ +$ is not
+surjective. Perrier \cite{perrier2001intuitionistic} characterized the
+set of syntactic descriptions that can be obtained as images of IILL
+sequents using our transformation $D \circ +$, the set of \emph{IILL
+  tree descriptions}, which is a proper subset of the set of all
+syntactic descriptions. Specifically, in an IILL tree descriptions, all
+immediate dominance constraints go from negative nodes to positive
+nodes, all (large) dominance constraints lead from positive nodes to
+negative nodes and the root has a positive polarity.
+
+While this property may not hold for the PTDs we find in Frigram, it
+still gives us a general plan for recovering IILL types. The strategy
+suggested by the description-formula connection given by the theorem
+yields a mapping $F$ from the under-specified trees to IILL formulas.
+
+$$
+F(N) = F(N_1) \limp ... \limp F(N_k) \limp cat(N)
+$$
+
+where $N_1, ..., N_k$ are the children of $N$ and $cat(N)$ is the
+category of $N$. If $N$ is a leaf, then $F(N)$ is simply $cat(N)$. In
+this definition, we treat the tree descriptions as trees with the union
+of the immediate dominance and (large) dominance constraints serving as
+the edges of the tree.
+
+This strategy corresponds to our intuitions about tree descriptions
+should be expressed using IILL types. We will consider two very simple
+cases to demonstrate it.
+
+First, picture a simple tree description with a positive root $A$ and
+two (necessarily) negative children $B$ and $C$. Our strategy would give
+this tree the type $B \limp C \limp A$ (or $C \limp B \limp A$), which
+correctly expresses the fact that this tree needs to consume some node
+of category $B$ and another one of category $C$ and is able to provide a
+node of category $A$. This tree description can also be easily
+translated to TAG by turning the negative leaf nodes into substitution
+nodes. The conventional translation of TAG into categorial grammar then
+confirms the type proposed by our strategy.
+
+The second example will demonstrate the translation of a Frigram
+EPTD. Before we do that, however, we will first discuss some of the ways
+the actual PTDs of interaction grammars and Frigram differ from the IILL
+tree descriptions we just talked about.
+
+The PTDs of IG include precedence constraints. We can see them as
+serving two purposes, to both provide an ordering of the constituents
+and to further constrain syntactic composition. In our translation to an
+ACG, we could imagine handling the first role, that of providing word
+order, by defining a lexicon from our abstract signature. Each constant
+would combine its arguments in the order in which their corresponding
+subtrees appeared in the ordered PTD. However, the second role of
+precedence constraints, to establish further restrictions on syntactic
+composition, would be lost, as the only restricting factor on syntactic
+composition in ACGs is the type system.
+
+Another large difference is the use of multiple independent polarized
+features per node. A possible solution here is to salvage the most
+salient polarized feature, $cat$ (possibly with its complementary
+feature, $funct$), and ignore the others. The results and intuitions we
+have about intepreting tree descriptions as types do not offer an easy
+way of modeling polarities across different dimensions, features, at the
+same time. This would mean that phenomena, such as the paired
+grammatical words for negation in French, which were handled using
+special polarized features will need new solutions in the framework of
+ACGs.
+
+Next up are the virtual features. They serve multiple purposes as
+well. Most often they are used to select nodes in the context and to
+link them to the nodes contributed by the EPTD.
+
+A common pattern using virtual features is Frigram's handling of
+modifiers. The way we would model a modifier in categorial grammars
+would be to have a function of type $X \limp X$ where $X$ is the type of
+the consituent being modified. In TAG, we would have an adjunction tree
+with a root and a foot node of category $X$ and which pins the modifier
+on the $X$ phrase. This approach would work in interaction grammars as
+well but the formalism offers other solutions too. The designers of
+Frigram have opted for an alternate treatment in which modifying a
+constituent does not increase the depth of the parse tree. Modifiers are
+adjoined as sister nodes of the constituents they modify and it is
+through virtual features that IGs can ensure that a modifier is inserted
+only in the proper context.
+
+See Figure~\ref{fig:ig-adj} for an example of an adjective modifier in
+Frigram. Nodes (1,3) and (1,2) select the noun to be modified and its
+projection, respectively. Singling out this pair of nodes lets us hang
+the adjective as a child of the projection and state that it must
+precede the noun in the word order. We can see that this does not
+increase the depth of the noun in the parse tree, since the noun is only
+referred to using virtual features in this EPTD, meaning that some other
+EPTD must provide the nodes that will saturate these virtual features.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.25]{ig-adj.pdf}
+  \caption{\label{fig:ig-adj} An EPTD for the adjective $grand$ in
+    Frigram.}
+\end{figure}
+
+Understanding this pattern can give us a useful tool for reading
+EPTDs. Consider the EPTD of $la$ we saw in
+Figure~\ref{fig:ig-eptds}. The EPTD can now be seen as applying this
+pattern twice at the same time to both cliticize the pronoun to the
+verbal kernel and to fill the object valency slot of the
+verb.
+
+Another use of virtual features in IGs is to restrict the range of
+(large) dominance relations. We have seen this before in the EPTD of the
+determiner $aucun$ (Figure~\ref{fig:ig-neg}), which had to reach up to
+the nearest containing clause and hang a negative $neg$ feature there.
+It is also present in the EPTD of the relative pronoun $que$
+(Figure~\ref{fig:ig-que}) which will bring us back to the translation
+strategy proposed above.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.25]{ig-que.pdf}
+  \caption{\label{fig:ig-que} An EPTD of the relative pronoun $que$.}
+\end{figure}
+
+Once again, at the root of the EPTD we see the modifier pattern in
+action. If we take the subtree formed by the nodes (2,8) and (2,9) to
+represent a modification of some constituent, say a noun, then we get a
+type like $n \limp n$. However, the EPTD contains also positive and
+negative $cat$ polarities. The node (2,0) has a negative $cat$ polarity
+and we can consider it as an argument, meaning that the type we would
+like to derive would look like $F((2,0)) \limp (n \limp n)$. Our
+strategy $F$ has nothing to say about virtual features or (large)
+dominance constraints. If we look only at the positive and negative
+$cat$ features, $F((2,0))$ becomes $np \limp s$, which would give us
+$(np \limp s) \limp (n \limp n)$ as the final type.
+
+However, this translation is very lossy in terms of its discriminatory
+power. Many of the constraints which were forced by the original EPTD
+are now ignored. We can imagine boosting our strategy by using the
+feature structures as our types instead of atomic categories, which
+would not only ensure that e.g. the $sent\_type$ of the embedded clause
+must be $decl$, but would also capture the information in the polarized
+$funct$ feature so that the relative pronoun $que$ is not permitted to
+extract subjects. There are still constraints that we would miss out
+though, such as the fact that the subject (2,3) of the clause (2,0) must
+precede any embedded clauses (2,6)...(2,5), the fact that any such
+embedded clause must serve either the $modal$ or the $obj$ function and
+the fact that the embedded clause (2,5) must have a subject (2,4).
+
+Other constraints that we did not mention also do not have a direct
+equivalent in ACGs. One example would be a feature of IGs that we have
+not even mentioned which allows us the grammar author to provide an
+\emph{exhaustive} list of children of a node in a PTD. Another kind of
+tricky constraint are the orange rectangles that we can see in
+Figure~\ref{fig:ig-que} which enforce that a constituent must be the
+left-most or right-most daughter. The problem with these is similar as
+with precedence relations. To model their impact on word order would be
+easy using an ACG lexicon, but to model their effect on restricting
+syntactic composition would be more tricky.
+
+In the end, even though both formalisms (IGs and ACGs) are built on the
+framework of linear logic and their fundamental principles of syntactic
+composition have been linked together by Perrier's theorem, we conclude
+that embedding Frigram into an ACG or constructing some automated
+technique for translating the Frigram EPTDs into ACG types and terms
+would not be feasible. Perhaps using some of the other connectives of
+linear logic would allows us to model the Frigram EPTDs, and the ways
+they manipulate multiple resources at the same time, in a more efficient
+manner. However, this would lead us well out of the framework of ACGs in
+which we wish to conduct our research.
+
+In our grammar, we will still gladly use the lexicon Frilex, which
+proves the usefulness of clearly defining the interface between the
+lexicon from the grammar in the architecture of Frigram. In our grammar,
+Frigram can still serve as a guide on how to interface with Frilex, on
+which phenomena need to be covered and in what detail, and what is a
+good way to structure the reusable components of a grammar of French.

### 2013-05-15 acg.tex
First reading/revision of section I

@@
   \begin{prooftree}
-    \AxiomC{$\emptyset, \Gamma_i \vdash_\Sigma c : \tau(c)$ (cons)}
+    \AxiomC{$\emptyset; \Gamma_i \vdash_\Sigma c : \tau(c)$ (cons)}
   \end{prooftree}
   \begin{prooftree}
-    \AxiomC{$(x : \alpha), \Gamma_i \vdash_\Sigma x : \alpha$ (l-var)}
+    \AxiomC{$(x : \alpha); \Gamma_i \vdash_\Sigma x : \alpha$ (l-var)}
   \end{prooftree}
   \begin{prooftree}
-    \AxiomC{$\emptyset, (\Gamma_i, x : \alpha) \vdash_\Sigma x : \alpha$ (i-var)}
+    \AxiomC{$\emptyset; (\Gamma_i, x : \alpha) \vdash_\Sigma x : \alpha$ (i-var)}
   \end{prooftree}
   \begin{prooftree}
-    \AxiomC{$(\Gamma_l, x : \alpha), \Gamma_i \vdash_\Sigma t : \beta$}
+    \AxiomC{$(\Gamma_l, x : \alpha); \Gamma_i \vdash_\Sigma t : \beta$}
     \RightLabel{(l-abs)}
-    \UnaryInfC{$\Gamma_l, \Gamma_i \vdash_\Sigma \lambda^{\circ} x. t : \alpha \limp \beta$}
+    \UnaryInfC{$\Gamma_l; \Gamma_i \vdash_\Sigma \lambda^{\circ} x. t : \alpha \limp \beta$}
   \end{prooftree}
   \begin{prooftree}
-    \AxiomC{$\Gamma_l, (\Gamma_i, x : \alpha) \vdash_\Sigma t : \beta$}
+    \AxiomC{$\Gamma_l; (\Gamma_i, x : \alpha) \vdash_\Sigma t : \beta$}
     \RightLabel{(i-abs)}
-    \UnaryInfC{$\Gamma_l, \Gamma_i \vdash_\Sigma \lambda x. t : \alpha \to \beta$}
+    \UnaryInfC{$\Gamma_l; \Gamma_i \vdash_\Sigma \lambda x. t : \alpha \to \beta$}
   \end{prooftree}
   \begin{prooftree}
-    \AxiomC{$\Gamma_l, \Gamma_i \vdash_\Sigma t : \alpha \limp \beta$}
-    \AxiomC{$\Delta_l, \Delta_i \vdash_\Sigma u : \alpha$}
+    \AxiomC{$\Gamma_l; \Gamma_i \vdash_\Sigma t : \alpha \limp \beta$}
+    \AxiomC{$\Delta_l; \Delta_i \vdash_\Sigma u : \alpha$}
     \RightLabel{(l-app)}
-    \BinaryInfC{$(\Gamma_l, \Delta_l), (\Gamma_i, \Delta_i) \vdash_\Sigma (t\ u) : \beta$}
+    \BinaryInfC{$(\Gamma_l, \Delta_l); (\Gamma_i, \Delta_i) \vdash_\Sigma (t\ u) : \beta$}
   \end{prooftree}
   \begin{prooftree}
-    \AxiomC{$\Gamma_l, \Gamma_i \vdash_\Sigma t : \alpha \to \beta$}
-    \AxiomC{$\emptyset, \Delta_i \vdash_\Sigma u : \alpha$}
+    \AxiomC{$\Gamma_l; \Gamma_i \vdash_\Sigma t : \alpha \to \beta$}
+    \AxiomC{$\emptyset; \Delta_i \vdash_\Sigma u : \alpha$}
     \RightLabel{(i-app)}
-    \BinaryInfC{$\Gamma_l, (\Gamma_i, \Delta_i) \vdash_\Sigma (t\ u) : \beta$}
+    \BinaryInfC{$\Gamma_l; (\Gamma_i, \Delta_i) \vdash_\Sigma (t\ u) : \beta$}
   \end{prooftree}
@@
 signatures with their distinguished types and lexicons connecting these
-signatures. A common pattern will have us using two signatures for the
-surface forms of utterances (strings) and their logical forms (logical
-propositions) and an abstract signature which is connected to both of
-the object signatures via lexicons. Parsing is then just a matter of
-inverting the surface lexicon to get the abstract term and then applying
-to it the logical lexicon. Generation is symmetric, we simply invert the
-logical lexicon and apply the surface lexicon.
+signatures. A common pattern will have us using two object signatures
+for the surface forms of utterances (strings) and their logical forms
+(logical propositions) and an abstract signature which is connected to
+both of the object signatures via lexicons. Parsing is then just a
+matter of inverting the surface lexicon to get the abstract term and
+then applying to it the logical lexicon. Generation is symmetric, we
+simply invert the logical lexicon and apply the surface lexicon.
 
@@
 can write as $\mathcal{L}_{syntax}(n) = \mathcal{L}_{syntax}(np) =
-\mathcal{L}_{syntax}(s)$\footnote{Whenever we talk about some lexicon,
-  we will use the lexicon itself, which is formally a pair of mappings
-  $\mathopen{<}F, G\mathclose{>}$, to mean either $F$, $G$ or their
-  homomorphic extensions $\hat{F}$ and $\hat{G}$ depending on whether
-  the argument is an atomic type, a constant, a complex type or a term
-  respectively.}. From that, we can use the homomorphic extension of our
-type mapping to arrive at the object-level interpretations of complex
-types, such as the type of predicates, $np \limp s$, which becomes a
-unary string function, $\textsc{String} \limp \textsc{String}$, or the
-type of quantified noun phrases, $(np \limp s) \limp s$, which becomes a
-higher-order string function, $(\textsc{String} \limp \textsc{String})
-\limp \textsc{String}$. With the types out of the way, we can give the
-interpretation of the individual constants of $\Sigma_{Synt}$:
+\mathcal{L}_{syntax}(s) = \textsc{String}$\footnote{Whenever we talk
+  about some lexicon, we will use the lexicon itself, which is formally
+  a pair of mappings $\mathopen{<}F, G\mathclose{>}$, to mean either
+  $F$, $G$ or their homomorphic extensions $\hat{F}$ and $\hat{G}$
+  depending on whether the argument is an atomic type, a constant, a
+  complex type or a term respectively.}. From that, we can use the
+homomorphic extension of our type mapping to arrive at the object-level
+interpretations of complex types, such as the type of predicates, $np
+\limp s$, which becomes a unary string function, $\textsc{String} \limp
+\textsc{String}$, or the type of quantified noun phrases, $(np \limp s)
+\limp s$, which becomes a higher-order string function,
+$(\textsc{String} \limp \textsc{String}) \limp \textsc{String}$. With
+the types out of the way, we can give the interpretation of the
+individual constants of $\Sigma_{Synt}$:
 \begin{itemize}
@@
   argument, $\mathcal{L}_{syntax}(t_4) =_{\beta} \lambda^{\circ}
-  x.\ x\ +\ loves\ +\ some\ +\ woman$
+  x.\ x\ +\ loves\ +\ some\ +\ woman$.
 \item $t_5 = (\synt{every}\ \synt{man})\ (\lambda^{\circ}
@@
 object terms, for which there exists a term in the abstract language
-such that its image given by the lexicon yields the same object term.
+such that its image given by the lexicon yields the same object term,
+i.e. it contains only terms which denote strings that spell out
+sentences of our fragment of English.
 

### 2013-05-15 ig.tex
First reading/revision of section I

@@
 concept of \emph{polarity}. An interaction grammar is a set of
-(underspecified) tree descriptions which define a set of trees that can
+(under-specified) tree descriptions which define a set of trees that can
 be constructed by superposing some of these tree descriptions while
@@
   \includegraphics[scale=0.25]{ig-parse.pdf}
-  \caption{\label{fig:ig-parse} Syntactic tree of sentence
+  \caption{\label{fig:ig-parse} The syntactic tree of sentence
     (\ref{ex:ig}). The numbers at the top of the nodes are labels of
@@
 Figure \ref{fig:ig-parse} shows the syntactic tree assigned by Frigram
-to the sentence in (\ref{ex:ig}). It is a rooted ordered tree with the
-topmost node being the root. Solid lines indicate immediate dominance
-with the higher node being the parent and the lower the child. The
-ordering of children is given by the arrows, which signify the immediate
-precedence relation between sister nodes.
+to sentence (\ref{ex:ig}). It is a rooted ordered tree with the topmost
+node being the root. Solid lines indicate immediate dominance with the
+higher node being the parent and the lower the child. The ordering of
+children is given by the arrows, which signify the immediate precedence
+relation between sister nodes.
 
@@
 
-The nodes of the syntactic tree are also decorated with
-features.
+The nodes of the syntactic tree are also decorated with features.
 
 Interaction grammars are a formalism enabling us to define sets of the
-structures described above, the \emph{syntactic trees}. Similar to Tree
-Adjoining Grammars, interaction grammars are formulated in terms of a
+structures described above, the \emph{syntactic trees}. Similar to tree
+adjoining grammars, interaction grammars are formulated in terms of a
 set of elementary structures which can combine to produce the final
@@
 a minimal structure satisfying those constraints (it is a \emph{minimal
-  model}).
+  model}). This distinction separates TAG as a formalism in the
+generative-enumerative framework of syntax with IG as a formalism in the
+model-theoretic framework.
 
@@
 management system as well, since using the clitic fills up the object
-slot in the valency of a verb. The EPTD of $la$ also relies heavily on
-virtual features and uses them to specify in more detail the context in
-which this rule is applicable, so as to avoid over-generation.
+slot in the valency of a verb. The EPTD of $la$ also uses virtual
+features heavily to select the right place where to hang the pronoun.
 
@@
 see that the syntactic tree in Figure~\ref{fig:ig-parse} is truly a
-model of the PTDs given in Figure~\ref{fig:ig-eptds} and more than that,
-it is the only minimal model. To illustrate more clearly how the
-polarities and constraints of the individual PTDs end up generating the
-syntactic tree of Figure~\ref{fig:ig-parse}, we can look at the result
-of superposing the EPTDs of $la$ and $voit$ by merging the nodes (1,2)
-and (1,5) with (2,2) and (2,4) respectively on
-Figure~\ref{fig:ig-partial}.
+model of the PTDs given in Figure~\ref{fig:ig-eptds} (furthermore, it is
+the only minimal model). To illustrate more clearly how the polarities
+and constraints of the individual PTDs end up generating the syntactic
+tree of Figure~\ref{fig:ig-parse}, we can look at the result of
+superposing the EPTDs of $la$ and $voit$ by merging the nodes (1,2) and
+(1,5) with (2,2) and (2,4) respectively on Figure~\ref{fig:ig-partial}.
 
@@
 However, defining some 4000 EPTDs manually still seems like a very
-tedious, error prone and hard to maintain approach. Many of the EPTDs
-have to repeatedly describe common phenomena such as subject-object
+tedious, error prone and hard to maintain process. Many of the EPTDs
+have to repeatedly describe common phenomena such as subject-predicate
 agreement or predicate-argument saturation. Similarly as in other
@@
 composition operator of XMG, these 160 classes end up generating the
-staggering 4000 EPTDs.
+4000 EPTDs.
 
@@
 descriptions such that some selection of descriptions will generate some
-tree if and only if the selection of the corresponging typed constants
-generated some properly-typed term. Inverting this transformation would
+tree if and only if the selection of the corresponding typed constants
+generated some properly typed term. Inverting this transformation would
 then let us take the elementary syntactic descriptions of an IG like
@@
 still gives us a general plan for recovering IILL types. The strategy
-suggested by the description-formula connection given by the theorem
+suggested by the description/formula connection given by the theorem
 yields a mapping $F$ from the under-specified trees to IILL formulas.
@@
 
-This strategy corresponds to our intuitions about tree descriptions
-should be expressed using IILL types. We will consider two very simple
-cases to demonstrate it.
+This strategy corresponds to our intuitions about how should tree
+descriptions be expressed using IILL types. We will consider two very
+simple cases to demonstrate it.
 
@@
 and we can consider it as an argument, meaning that the type we would
-like to derive would look like $F((2,0)) \limp (n \limp n)$. Our
-strategy $F$ has nothing to say about virtual features or (large)
-dominance constraints. If we look only at the positive and negative
-$cat$ features, $F((2,0))$ becomes $np \limp s$, which would give us
-$(np \limp s) \limp (n \limp n)$ as the final type.
+like to derive would look like $F((2,0)) \limp (n \limp n)$. However,
+our strategy $F$ has nothing to say about virtual features or (large)
+dominance constraints, so we will have to ignore those. If we look only
+at the positive and negative $cat$ features, $F((2,0))$ becomes $np
+\limp s$, which would give us $(np \limp s) \limp (n \limp n)$ as the
+final type.
 
@@
 power. Many of the constraints which were forced by the original EPTD
-are now ignored. We can imagine boosting our strategy by using the
+are now ignored. We can imagine improving our strategy by using the
 feature structures as our types instead of atomic categories, which
@@
 $funct$ feature so that the relative pronoun $que$ is not permitted to
-extract subjects. There are still constraints that we would miss out
+extract subjects. There are still constraints that we would miss out on
 though, such as the fact that the subject (2,3) of the clause (2,0) must
@@
 Figure~\ref{fig:ig-que} which enforce that a constituent must be the
-left-most or right-most daughter. The problem with these is similar as
-with precedence relations. To model their impact on word order would be
-easy using an ACG lexicon, but to model their effect on restricting
-syntactic composition would be more tricky.
+leftmost or rightmost daughter of its parent. The problem with these is
+similar as with precedence relations. To model their impact on word
+order would be easy using an ACG lexicon, but to model their effect on
+restricting syntactic composition would be more tricky.
 
@@
 technique for translating the Frigram EPTDs into ACG types and terms
-would not be feasible. Perhaps using some of the other connectives of
-linear logic would allows us to model the Frigram EPTDs, and the ways
-they manipulate multiple resources at the same time, in a more efficient
-manner. However, this would lead us well out of the framework of ACGs in
-which we wish to conduct our research.
+would not be feasible. We suspect that the incompatibility of these two
+formalisms is due to ACG being a formalism based in the
+generative-enumerative framework of syntax while IG is based in the
+model-theoretic approach. Perhaps we could model the operational view of
+IG, the combination of under-specified trees via superposition, by
+employing some of the other connectives of linear logic to model the
+fact that superposing PTDs usually involves many resources. However,
+this would lead us well out of the framework of ACGs in which we wish to
+conduct our research because of its elegant approach to discourse
+denotations.
 
-In our grammar, we will still gladly use the lexicon Frilex, which
-proves the usefulness of clearly defining the interface between the
-lexicon from the grammar in the architecture of Frigram. In our grammar,
-Frigram can still serve as a guide on how to interface with Frilex, on
-which phenomena need to be covered and in what detail, and what is a
-good way to structure the reusable components of a grammar of French.
+In our grammar, we will still gladly use the lexicon Frilex though,
+which proves the usefulness of clearly defining the interface between
+the lexicon and the grammar in the architecture of Frigram. In our
+grammar, Frigram can still serve as a guide on how to interface with
+Frilex, on which phenomena need to be covered and in what detail, and
+what is a good way to structure the reusable components of a grammar of
+French.

### 2013-05-15 motivation.tex
First reading/revision of section I

@@
 another feature of discourse structures. First off, seeing the
-hiearchical structure of the discourse gives us important information
+hierarchical structure of the discourse gives us important information
 about the granularity of the descriptions given in the individual

### 2013-05-17 my-work.tex
Imported report on my work

\section{Wide-coverage abstract categorial grammars}

As part of my internship, I will have to construct a wide-coverage
abstract categorial grammar of French. Abstract categorial grammars
are defined in terms of signatures, universes of typed constants, and
lexicons, which map elements of signatures to lambda terms in other
signatures. Our grammar will thus end up as an ensemble of signatures,
each containing a number of elements proportional to the size of our
dictionary, and some lexicons that will link the signatures together.

From the above exposition, it should be obvious that giving the
grammar in its literal form, i.e. listing the contents of all the
signatures and the mappings of the lexicons, would be unfeasible. The
approach that we have chosen is to employ a pre-existing database of
lexical items and describe how entries in this database generate the
elements of our signatures and the mappings in our lexicons.

The lexical database that we will be using is Frilex. Frilex was
itself constructed by collating several lexicographic sources and its
purpose was to serve as the lexical database for Frigram, a
wide-coverage interaction grammar of French, much like the way we will
use it to populate our grammar. In terms of structure, Frilex is a
relation between wordforms and hypertags, which are feature structures
giving the lemma, syntactic category, agreement, valency and other
pertinent morpho-syntactic features of the wordform.

\subsection{Defining signatures}

The task of the grammar author is then to describe for each signature
how the entries of the lexical database correspond to typed citizens
of the signature and analogously so for the lexicons. If we look at
how this correspondence should be modelled, we quickly arrive at the
conclusion that it cannot be a total function (e.g. in the case of a
signature of semantic predicates, we will most likely not want to have
predicates that correspond to determiners) nor in fact a partial
function (e.g. in the syntactic signature, we might very well desire
to have two different elements for transitive verbs depending on
whether they are to be interpreted with subject-wide or object-wide
scope). In the end, the correspondence seems to be most aptly
described as a relation. Therefore in our computational representation
of the grammar, the signature is defined via a relational (logic)
program. The relational programming system we have chosen is
miniKanren, particularly its Clojure implementation dubbed core.logic.

The next question we will have to answer is how exactly will this
signature-defining relation look like. To answer that, we first have
to be more precise about how we want the elements of our signature
yielded by this relation to be structured. As was said before, we want
all the elements in our signatures to have a type. Furthermore, we
also need a way how to distinguish elements of the same type, so we
will need to provide each element with some kind of identifier. We
could state that these identifiers are completely arbitrary or up to
the grammar author to generate, but this would make specifying the
mappings of lexicons burdensome, so we will try to reach something
better.

For starters, we might consider the hypertag as a candidate for the
identifier. However, it is possible that our lexical database will
hold two entries with the same hypertag but different wordforms (e.g.
alternative spellings) and we would like to distinguish the two to be
able to produce either of the two spellings using a lexicon. So next,
we can think of the entire lexical entry, the wordform and its
hypertag, as an identifier. This breaks down in the case of transitive
verbs having two projections which differ in subject/object scope. To
resolve cases like these, we add a synthetic part to our identifier
which we call a specifier and which is the grammar author's duty to
assign.

In the end, the identifiers for the elements of our signatures are
composed of their ``naturally-occurring'' part (the wordform and the
hypertag of their lexical entry) and their synthetic part (the
author-defined specifier). To answer our question, the relation which
defines the elements of a signature is a 4-ary relation between
wordforms, hypertags, spec(ifier)s and types. This relation posits
some constraints on the wordform and hypertag and then licenses the
possible combinations of values for the spec and the type. Since the
definition of this relation takes place within a general purpose
programming language, the grammar author is able to define any
abstractions which facilitate his statement of the relation.

It might seem we have solved the problem of defining signatures, but
there is still one thing that we have to address. In many of our
signatures, we will desire to have some elements of the signature that
don't correspond to any lexical item but instead provide some kind of
furniture for working in the domain. In the signature of semantics,
these might be constants corresponding to the boolean operators and
quantifiers, whereas in the domain of strings, we might think of the
string concatenation operator. Therefore, in our formalism, we must
acknowledge that elements of signatures are not only generated by the
above-described lexical relation, but also by contain some set of
constant elements. These elements also have a type and an identifier,
however, in their case, the identifier is not composed of some lexical
entry, but is simply the symbol that was used by the grammar author to
name the constant.

\subsection{Defining lexicons}

Now that we have a way of giving a concise computational
representation of a wide-coverage signature, we turn our attention to
the lexicons, the mappings relating the signatures. A lexicon is
defined as a mapping from elements of one signature to terms over the
elements of another signature. In our formalism, we simply follow this
definition, though for ease of implementation, we do not implement the
function as a functional program, but as a relational program. So far,
it is up to the grammar author to make sure that his lexicon assigns
only one term to a single element of the signature.

### 2013-05-21 my-work-rewritten.tex
Started rewriting my-work

\section{Wide-coverage abstract categorial grammars}

Our goal is to develop large abstract categorial grammars which cover a
variety of linguistic phenomena. All of these phenomena have been
studied in detail but we would like to have a grammar which gracefully
combines all these solutions. At this scale, working out the grammar on
paper stops being sufficient to detect all the subtle interactions and
it is desirable to have a formalized representation that can be reasoned
about or tested by a computer.

To some extent, this need is already met by the \textbf{ACG development
  toolkit}\footnote{\url{http://www.loria.fr/equipes/calligramme/acg/}}. The
toolkit reads in signature and lexicon definitions and offers two modes
of operation. In the first, it checks the validity of the defined
ACG. That is to say, it checks whether the terms assigned by the
lexicons are well-typed and whether their types are consistent with the
object types dictated by the lexicon. The second mode of operation
offers an interactive experience in which the user is free to run type
inference on terms built on one of their signatures to find out whether
they are well-typed and if so, what is their principal type. The
interactive mode then lets the user apply a lexicon, or a composition of
lexicons, to a term and get the $\beta$-normal form of the object term.

We consider the above capabilities vital for doing any involved grammar
engineering in the framework of ACGs. However, there are some
limitations which make it not sufficient for our needs. The foremost of
those is that the signatures and lexicons are all explicitly realized in
(primary) memory and the toolkit expects them to be given directly in a
file. The grammars that we would like to develop will cover an
exhaustive list of wordforms and the size of our grammar definitions
would therefore grow proportionately with the size of our dictionary.

Given this setup, there are several approaches we might try, none of
them too appealing. We might write our metagrammar in a different format
and use a tool which would combine the metagrammar and the lexical
database\footnote{We will be using the term \emph{lexical database} to
  refer to what is usually called a lexicon, in order not to cause
  confusion with the lexicons of ACGs.} to produce the lexicalized ACG
in a direct representation, ready to be loaded by the toolkit. This
workflow would make development on the grammar tedious as every time we
would change the metagrammar, we would have to regenerate the direct
representation and then load it in the ACG toolkit. With extra effort,
the former cost could be alleviated by devising a system for
regenerating only parts of the grammars which will have changed since
last time. However, we would still have to pay the price of loading the
entire grammar into the toolkit before being able to interact with it
again.

Furthermore, we could pick a small representative fragment of the
dictionary and test only on that fragment during the development of our
grammar, to reduce the time it takes to both generate the direct
representation of the grammar and the time it takes to load it in the
toolkit. This would have made developing the grammar already
tenable.

However, in our approach, we have opted to spend more time on tool
development to ease subsequent work and we have developed a system which
makes defining and experimenting with lexicalized ACGs easier. The
metagrammars are defined as relational procedures which link the items
of the lexical database to elements of signatures and lexicons. The
grammars can be redefined without having to reload the lexical database
or any file of similar scale. Type inference and lexicon application
with $\beta$-normalization are provided as relations which can be used
interactively or as part of a larger program. En example of such a
larger program might be a routine to check the validity of an ACG or to
test some of its properties.

In this chapter, we will spend time explaining the decisions which went
into designing the system and the form it has now.

\subsection{}

### 2013-05-21 my-work-rewritten.tex
Wrote tools and techniques

@@
 
-\subsection{}
+\subsection{The tools and techniques}
+
+The dominant decision which we made during the development of the system
+was to use \emph{relational programming}. Relational programming is a
+discipline of logic programming where relations are written in a
+``pure'' style \cite{byrd2010relational}. This means that relations
+cannot be picky about which arguments are passed in as inputs and which
+as outputs, all modes must be permissible. Accepting such a discipline
+prohibits us from using some of the special operators of Prolog such as
+cut (\texttt{!}) and \texttt{copy\_term/2}. Instead, we must resort to
+other techniques which do not break purity, like \emph{constraint logic
+  programming}, which allows us to express properties outside of the
+scope of unification by attaching constraints to terms and verifying
+them at the proper time.
+
+By insisting on purity, I/O becomes more difficult as we would not allow
+the side-effectful pseudorelations of Prolog. Therefore, we will not
+force ourselves to implement the entire program front-to-back using only
+relational programming. Instead, we will use a relational programming
+system, miniKanren, which is embedded inside a functional programming
+language. This lets us reap the benefits of relational programming
+inside the core logic of our program while deferring to the functional
+programming language to provide I/O and other bookkeeping operations.
+
+\subsubsection{The case for relational programming}
+
+There are several reasons why we might want to choose a relational
+programming system for our implementation. First off, the problems which
+consiture the algorithmic core of our program are often mere textbook
+examples of how to use a relational programming system. This goes for
+type inference/checking/inhabitation, which are all implemented by a
+single relation which is often used as a kind of ``Hello World'' program
+in the world of relational programming. Similarly, implementing
+$\beta$-reduction cleanly (without having to handle variable clashes by
+generating new variables and $\alpha$-converting) is a motivational
+example for introducing \emph{nominal logic programming}. To reach
+feature parity with the ACG toolkit, one would only need to implement
+the mapping of a lexicon over a term, which is a simple functor.
+
+The use of logic programming techniques is also very popular in our
+domain. If we look at the system demonstrations at the \emph{Logical
+  Aspects of Computational Linguistics 2012} conference, we see that 3
+out of the 5 proposals have been implemented in a declarative/logic
+programming system (Oz, Prolog). 2 of the 3 systems are parsers for
+categorial grammars (automated theorem provers), which might turn out to
+be an interesting direction for our system to go into as well, as
+practical applications appear on the horizon.\footnote{Theorem provers
+  are another problem area where miniKanren fits well, see
+  $\alpha$leanTAP \cite{near2008alpha}.}
+
+The third system was XMG, a language for writing metagrammars, which is
+another goal of our system, the one which distinguishes it from the
+original ACG toolkit. We believe that using a declarative programming
+system such as miniKanren with the ability to extend it using a
+practical functional programming language presents an empowering way of
+defining metagrammars. This allows the grammar designer to compose his
+grammar in any way he sees fit and to use the metaprogramming facilities
+of the underlying programming language to transcribe the grammar in a
+way that makes the most sense in her case.
+
+Finally, some of the structures that we will encounter in our domain are
+more directly modelled in mathematics as a relation instead of a
+function. This has the benefit that to implement a relation in
+relational programming, one just writes the relation, while if we were
+forced to encode everything in functions, we would have to manage this
+multiplicity of values manually. Examples of structures that have this
+kind of structures if Frilex, our lexical database, where a single
+wordform can be mapped to multiple hypertags. We will see another
+example of a relation within our domain in the next subsection and we
+will also discuss the advantages and drawbacks of modelling functions as
+relations.
+
+\subsubsection{Choice of implementation language}
+
+There are two languages which play host to a substantial implementation
+of miniKanren (meaning one that has extensible unification and/or
+constraints and that includes support for nominal logic),
+Racket\footnote{\url{https://github.com/calvis/cKanren}} and
+Clojure\footnote{\url{https://github.com/clojure/core.logic}}. The two
+languages are very similar to each other and therefore the choice is of
+no great significance. An appealing feature of Clojure is that it uses
+abstract collection types pervasively instead of tying the standard
+operations to specific structures such as lists and cons cells and that
+it follows other similarly enlightened design decisions. Racket on the
+other hand has had more time to mature, has better metaprogramming
+facilities and is more recognized in academia. In developing our system,
+we have settled for Clojure. One of the advantages of doing so is that
+the Clojure implementation of miniKanren, dubbed core.logic, has
+inspired the software community and the project has attracted
+contributors that submit issue reports and
+patches\footnote{\url{http://dev.clojure.org/jira/browse/LOGIC}}.
+
+\subsection{Defining signatures and lexicons}

### 2013-05-22 my-work-rewritten.tex
Almost finished writing about defining signatures

@@
 \subsection{Defining signatures and lexicons}
+
+\subsubsection{Defining signatures}
+
+In~\ref{sssec:sig}, we formally presented signatures as triples
+$\mathopen{<}A, C, \tau\mathclose{>}$. The set $C$ can be easily
+reconstructed as the domain of $\tau$ and the set $A$ is just the set of
+all atomic types occurring in the range of $\tau$. Therefore, to define
+the signature $\mathopen{<}A, C, \tau\mathclose{>}$, it is enough to
+provide the function $\tau$, which in its set-theoretical realization is
+just a set of pairs.  In relational programming, a set can be easily
+encoded as a unary relation that answers the membership test. Therefore,
+our goal will be to construct to answer this relation and this relation
+will be our final representation of a signature.
+
+We have our lexical database, a set of pairs of wordforms and hypertags,
+and we want to arrive at a set of pairs of constants and their
+types. This mapping decomposed into smaller parts by considering the
+contribution of every part of the lexical database individually. What
+will be the right model for the link between some part of the lexical
+database and some part of our signature?
+
+\begin{description}
+  \item[Function of wordform] We could imagine that every different
+    wordform of the lexical database will end up being mapped to some
+    constant in the signature. However, this is unsuitable since the
+    lexical database is a relation that can assign more than one
+    distinct hypertag to a single wordform (e.g. with different part of
+    speech) and we would like to have constants of different types for
+    each of these hypertags.
+  \item[Function of hypertag] Conversely, we might imagine that the
+    hypertags of the lexical database are mapped via a function to the
+    constants of the signature. Here we run into trouble when we
+    consider variations in spelling or phonology. We might have two
+    items in our lexical database that share the same hypertag but whose
+    wordforms still differ. We would like to have two distinct constants
+    in our signature then, so that when these two constants are mapped
+    by the surface lexicon, they will yield two different strings.
+  \item[Function of lexical entry] Since neither the wordform nor the
+    hypertag is enough, we could consider the entire lexical entry (an
+    element of the lexical database, a pair of a wordform and one of its
+    hypertags). This solves both of the above problems but it is still
+    not perfect. Consider the case of a determiner and some semantic
+    signature. Since our link between elements of the lexical database
+    and of our signature is a function, there must be a constant in the
+    signature for every item of our lexical database. If you recall
+    in~\ref{sssec:acg-sem}, the constants of the semantic signature
+    $\Sigma_{Sem}$ contained predicates for the common nouns $man$ and
+    $woman$ and for the verb $loves$ and some common logical
+    furniture. However, it did not contain any constants which would
+    correspond to the determiners $every$ and $some$.
+  \item[Partial function of lexical entry] We could fix the issue above
+    by using a partial function. This accounts for cases when a lexical
+    entry has no constants to represent it in a signature. Nevertheless,
+    this only fixes half of the problem. We will consider the single
+    lexical entry which represents e.g. the 3rd person singular present
+    tense form $loves$ of the transitive verb $love$. A legitimate
+    account of subject/object scope ambiguity would be to consider two
+    constants in our abstract syntactic signature that have distinct
+    images in some semantic lexicon, one giving the outer scope to the
+    subject and the other to the object. This means that in order to
+    define such a signature, we would need to link a single lexical
+    entry in our database to more than one constant in the signature.
+  \item[Relation between lexical entry and constant] Finally,
+    considering all the above common scenarios, the only mathematical
+    model general enough to handle all of the above cases is that of a
+    relation. In our system, this will correspond exactly to a relation
+    that will be provided by the grammar author.
+\end{description}
+
+We have established that lexicons will be defined by a relation, a
+relation between a lexical entry (a wordform and a hypertag) and a
+constant (an identity\footnote{The identity of a constant is simply what
+  distinguishes it from other constants of the same signature having the
+  same type.} and a type). The identity of a lexical constant also has a
+complex structure. We know that the lexical constants we define using
+the relation described above are always linked to a lexical entry. This
+lexical entry at least partially identifies the constant. For cases
+where we have more than one constant per lexical entry (such as the
+example of the two different scope readings of $loves$), we need some
+extra piece of information to specify which constant we are referring to
+(we call this extra piece of information a \emph{specifier}). This
+specifier can have any shape that the grammar author desires. We can see
+an example of a lexical constant in Figure~\ref{fig:lex-const}.
+
+\begin{figure}
+  % TODO: Center this.
+  \centering
+  \begin{verbatim}
+{:type [-> NP [-> NP S]]
+ :id {:lex-entry {:wordform "loves"
+                  :hypertag ...}
+      :spec {:scope :subject-wide}}}
+  \end{verbatim}
+  \caption{\label{fig:lex-const} An example of a lexical constant from a
+    hypothetical syntactic signature (the hypertag is elided for brevity).}
+\end{figure}
+
+The system takes care of wiring up the relations and assembling the
+signature. The grammar author's duty is to provide a 4-ary relation that
+links the four parts of a lexical constant: the wordform, the hypertag,
+the specifier and the type. This relation can be thought of as a
+multi-valued ``function'' that receives as input a wordform and a
+hypertag (that is, a single lexical entry) and outputs (possibly
+multiple) pairs of specifiers and types. Since we are working in a
+functional programming language that supports manipulation of
+higher-order functions, we can provide helper functions which construct
+these 4-ary relations for the grammar author in some simple cases. Our
+library thus provides functions like \texttt{unitypedr}, which expects a
+type and returns the 4-ary relation that assigns to every lexical item a
+single constant whose specifier is $nil$ and whose type is the supplied
+type, or \texttt{ht->typer}, which expects a mapping from hypertag
+patterns to types and returns the 4-ary relation which tries to map the
+hypertag against the patterns and yields constants with the
+corresponding types and $nil$ specifiers. When other patterns emerge,
+the grammar author is completely free to implement her new abstractions
+using the full power of the functional programming language. See
+Figure~\ref{fig:lex-sig-impl} for an example of implementing a
+lexicalized version of the $\Sigma_{Synt}$ signature
+from~\ref{sssec:example-sig} using the \texttt{ht->typer} abstraction.
+
+\begin{figure}
+  \centering
+  \begin{verbatim}
+(ht->typer {{:head {:cat "n"}}      'N
+            {:head {:cat "v"
+                    :trans "true"}} (-> 'NP 'NP 'S)
+            {:head {:cat "det"}}    (-> 'N (-> 'NP 'S) 'S)})
+  \end{verbatim}
+  \caption{\label{fig:lex-sig-impl} A lexicalized version of the
+    $\Sigma$%_{Synt} TODO TODO TODO
+ signature from~\ref{sssec:example-sig}.}
+\end{figure}
+
+In the above paragraphs, we have been referring to the constants that we
+were defining as \emph{lexical constants}. That was to distinguish them
+from the \emph{non-lexical constants} that we will introduce now. The
+reason for having non-lexical constants is that not every constant of
+every signature is linked to (generated by) some item of our lexical
+database. Consider, for example, the empty string $\epsilon$ and the
+string concatenation operator $+$ of the $\Sigma_{String}$ signature we
+gave in~\ref{sssec:example-sig} or the quantifiers and logical
+connectives of the $\Sigma_{Sem}$ signature in~\ref{sssec:acg-sem}.
+
+\subsection{Testing signatures and lexicons}

### 2013-05-23 my-work-rewritten.tex
Finished writing 'Defining signatures'

@@
 connectives of the $\Sigma_{Sem}$ signature in~\ref{sssec:acg-sem}.
+These constants are not associated to any lexical entry or specifier and
+therefore their representation differs. See
+Figure~\ref{fig:nonlex-const} for an example of an non-lexicalized
+constant from a semantic signature.
+
+\begin{figure}
+  \centering
+  \begin{verbatim}
+{:type [-> T [-> T T]]
+ :id {:constant-name and?}}
+  \end{verbatim}
+  \caption{\label{fig:nonlex-const} An example of the representation of
+    a non-lexicalized constant (the conjunction operator in a semantic
+    signature).}
+\end{figure}
+
+Signatures of non-lexical constants can be simply defined by a mapping
+from the constant names to their types as
+in~\ref{fig:nonlex-sig-impl}. Such signatures will generally not be
+sufficient by themselves and will need to be complemented by other
+signatures which contain the lexical constants. Since we represent
+signatures as unary relations, combining them to produce their unions or
+intersections is just a matter of taking their disjunctions or
+conjunctions, respectively. With this, we have enough to fully define
+the other signatures of subsection~\ref{ssec:acg} in a lexicalized
+manner, see Figure~\ref{fig:example-sig-impl}.
+
+\begin{figure}
+  \centering
+  \begin{verbatim}
+(nonlex-sigr {'and?    (-> 'T 'T 'T)
+              'imp?    (-> 'T 'T 'T)
+              'forall? (-> (=> 'E 'T) 'T)
+              'exists? (-> (=> 'E 'T) 'T)})
+  \end{verbatim}
+  \caption{\label{fig:nonlex-sig-impl} A signature of non-lexical
+    constants belonging to $\Sigma$%_{Synt}
+    from~\ref{sssec:acg-sem}. The single arrow \texttt{->} corresponds
+    to linear implication while the double arrow \texttt{=>} corresponds
+    to intuitionistic implication.}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{verbatim}
+(def synt-sig
+  (ht->typer {{:head {:cat "n"}}      'N
+              {:head {:cat "v"
+                      :trans "true"}} (-> 'NP 'NP 'S)
+              {:head {:cat "det"}}    (-> 'N (-> 'NP 'S) 'S)}))
+
+(def string-sig
+  (ors (nonlex-sigr {'++ (-> 'Str 'Str 'Str)})
+       (unitypedr 'Str)))
+
+(def sem-sig
+  (ors (nonlex-sigr {'and?    (-> 'T 'T 'T)
+                     'imp?    (-> 'T 'T 'T)
+                     'forall? (-> (=> 'E 'T) 'T)
+                     'exists? (-> (=> 'E 'T) 'T)})
+       (ht->typer {{:head {:cat "n"}}      (-> 'E 'T)
+                   {:head {:cat "v"
+                           :trans "true"}} (-> 'E 'E 'T)})))
+  \end{verbatim}
+  \caption{\label{fig:example-sig-impl} The definition of the
+    lexicalized versions of the example signatures of
+    subsection~\ref{ssec:acg}.}
+\end{figure}
+
+\subsubsection{Defining lexicons}
 

### 2013-05-23 my-work-rewritten.tex
Wrote 'Defining lexicons'

@@
   \caption{\label{fig:lex-sig-impl} A lexicalized version of the
-    $\Sigma$%_{Synt} TODO TODO TODO
-    signature from~\ref{sssec:example-sig}.}
+    $\Sigma$(\_{Synt}) signature from~\ref{sssec:example-sig}.}
 \end{figure}
@@
   \caption{\label{fig:nonlex-sig-impl} A signature of non-lexical
-    constants belonging to $\Sigma$%_{Synt}
+    constants belonging to $\Sigma$(\_{Sem})
     from~\ref{sssec:acg-sem}. The single arrow \texttt{->} corresponds
@@
 
+We have explained how we can use elements of our lexical database to
+define signatures. Now, we will turn to lexicons. Similar to how we
+reduced signatures from the formal triples $\mathopen{<}A, C,
+\tau\mathclose{>}$ to just the type assignment function $\tau$, we can
+similarly reduce lexicons from the pairs $\mathopen{<}F, G\mathclose{>}$
+to only $G$, the part which operates on terms. $F$, or at least its
+relevant subset, can be easily retrieved by putting the types of the
+abstract constants side-by-side with the types of their object terms to
+get a sufficient subset of $\hat{F}$ and finding $F$, the mapping from
+atomic abstract types whose homomorphic extension includes $\hat{F}$.
+
+In our relational framework, a mapping is most directly represented as a
+binary relation, which is the representation that we have chosen. In the
+previous subsubsection, we talked a lot about how individual entries in
+the lexical database generate the items of a signature. Solving the
+problem for signatures also solved it for lexicons, since their
+structure is very similar: a signature is a set of pairs of constants
+and types, a lexicon is a set of pairs of constants and
+terms. Therefore, a lexical entry contributes the object terms for the
+constants which it generated in the abstract signature.
+
+Our representation of lexical constants was purposefully engineered so
+as to make the lexical entry directly accessible to the lexicon
+implementation. This way, the lexicon author can inspect the hypertag
+and the wordform of the lexical entry, the specifier of the concrete
+constant that it is mapping right now and even the type she assigned to
+the constant in the abstract signature to allow her to choose the
+correct object term. Furthermore, since lexicons are still
+relations/sets, we can compose them together by disjunction. See
+Figure~\ref{lex-impl} for the implementation of the lexicalized versions
+of the $\mathcal{L}_{Syntax}$ and $\mathcal{L}_{Sem}$ lexicons.
+
+\begin{figure}
+  \centering
+\begin{verbatim}
+(defn syntax-lexo
+  [synt-constant string-term]
+  (with-sig-consts string-sig
+    (l/fresh [hypertag string-constant]
+      (has-hypertago synt-constant hypertag)
+      (share-lex-entryo synt-constant string-constant)
+      (string-sig string-constant)
+      (fs-assigne hypertag string-term
+                  {:head {:cat "n"}}
+                  ,(rt string-constant)
+                  {:head {:cat "v"
+                          :trans "true"}}
+                  ,(rt (ll [x y] (++ (++ x string-constant) y)))
+                  {:head {:cat "det"}}
+                  ,(ll [x R] (R (++ ua-stx-constant x)))))))
+
+(defn sem-lexo
+  [synt-constant sem-term]
+  (with-sig-consts sem-sig
+    (l/fresh [hypertag sem-constant]
+      (has-hypertago synt-constant hypertag)
+      (l/conde [(share-lex-entryo synt-constant sem-constant)
+                (sem-sig sem-constant)
+                (fs-assigne hypertag sem-term
+                            {:head {:cat "n"}}
+                            ,(rt sem-constant)
+                            {:head {:cat "v"
+                                    :trans "true"}}
+                            ,(rt sem-constant))]
+               [(fs-assigne hypertag sem-term
+                            {:head {:cat "det"
+                                    :det_type "indef"}}
+                            ,(rt (ll [p q] (exists? (il [x] (and? (p x)
+                                                                  (q x))))))
+                            {:head {:cat "det"
+                                    :det_type "def"}}
+                            ,(rt (ll [p q] (forall? (il [x] (imp? (p x)
+                                                                  (q x)))))))]))))
+
+\end{verbatim}
+  \caption{\label{fig:lex-impl} The implementation of the lexicons from
+    subsection~\ref{ssec:acg}. NOTE: This is too verbose both to put on
+    the paper and also to code. These examples demonstrates two common
+    and related shapes of lexicons. The $\mathcal{L}$(\_{Syntax}) and
+    the first half of the $\mathcal{L}$(\_{Sem}) are lexicalized, in the
+    sense that the object term contains a lexical constant with the same
+    lexical entry as the abstract constant.\footnote{This notion of
+      ``lexicalized'' is quite different to the one I use in the rest of
+      the thesis so far (which roughly means ``generated by a lexical
+      database'') and also different from the one used by Kanazawa
+      (FIX!).} This means that before constructing the object term, we
+    have to acquire the object constant that is contained in the term
+    and then we just fit it in of the recipes, which we choose according
+    to the hypertag. The second half of $\mathcal{L}$(\_{Sem}) simply
+    assigns object terms based on the hypertag. A more concise and
+    desirable implementation would allow us simply write the pairs of
+    hypertags and object terms (recipes encoded as lambdas). My plan is
+    to have exactly this, but I have not thought yet about how to split
+    the various aspects into different reusable abstractions.}
+\end{figure}
+
 \subsection{Testing signatures and lexicons}

### 2013-06-04 my-work-rewritten.tex
Prettier implementation of lexicons

@@
 (def string-sig
-  (ors (nonlex-sigr {'++ (-> 'Str 'Str 'Str)})
+  (ors (nonlex-sigr {'++ (-> 'Str 'Str 'Str)
+                     'empty-str 'Str})
        (unitypedr 'Str)))
@@
 relations/sets, we can compose them together by disjunction. See
-Figure~\ref{lex-impl} for the implementation of the lexicalized versions
-of the $\mathcal{L}_{Syntax}$ and $\mathcal{L}_{Sem}$ lexicons.
+Figure~\ref{fig:lex-impl} for the implementation of the lexicalized
+versions of the $\mathcal{L}_{Syntax}$ and $\mathcal{L}_{Sem}$ lexicons.
 
@@
 \begin{verbatim}
-(defn syntax-lexo
-  [synt-constant string-term]
+(def syntax-lexo
   (with-sig-consts string-sig
-    (l/fresh [hypertag string-constant]
-      (has-hypertago synt-constant hypertag)
-      (share-lex-entryo synt-constant string-constant)
-      (string-sig string-constant)
-      (fs-assigne hypertag string-term
-                  {:head {:cat "n"}}
-                  ,(rt string-constant)
+    (lexicalizer string-sig
+                 (ht-lexiconr {{:head {:cat "n"}}
+                               ,(rt (ll [_] _))
                                {:head {:cat "v"
                                        :trans "true"}}
-                  ,(rt (ll [x y] (++ (++ x string-constant) y)))
+                               ,(rt (ll [_ x y] (++ (++ x _) y)))
                                {:head {:cat "det"}}
-                  ,(ll [x R] (R (++ ua-stx-constant x)))))))
+                               ,(rt (ll [_ x R] (R (++ _ x))))}))))
 
-(defn sem-lexo
-  [synt-constant sem-term]
+(def sem-lexo
   (with-sig-consts sem-sig
-    (l/fresh [hypertag sem-constant]
-      (has-hypertago synt-constant hypertag)
-      (l/conde [(share-lex-entryo synt-constant sem-constant)
-                (sem-sig sem-constant)
-                (fs-assigne hypertag sem-term
-                            {:head {:cat "n"}}
-                            ,(rt sem-constant)
+    (orr (lexicalizer sem-sig
+                      (ht-lexiconr {{:head {:cat "n"}}
+                                    ,(rt (ll [_ x] (_ x)))
                                     {:head {:cat "v"
                                             :trans "true"}}
-                            ,(rt sem-constant))]
-               [(fs-assigne hypertag sem-term
+                                    ,(rt (ll [_ s o] (_ s o)))}))
+         (ht-lexiconr {{:head {:cat "det"
+                               :lemma "un"}}
+                       ,(rt (ll [P Q] (exists? (il [x] (and? (P x)
+                                                             (Q x))))))
                        {:head {:cat "det"
-                                    :det_type "indef"}}
-                            ,(rt (ll [p q] (exists? (il [x] (and? (p x)
-                                                                  (q x))))))
-                            {:head {:cat "det"
-                                    :det_type "def"}}
-                            ,(rt (ll [p q] (forall? (il [x] (imp? (p x)
-                                                                  (q x)))))))]))))
-
+                               :lemma "chaque"}}
+                       ,(rt (ll [P Q] (forall? (il [x] (imp? (P x)
+                                                             (Q x))))))}))))
 \end{verbatim}
-  \caption{\label{fig:lex-impl} The implementation of the lexicons from
-    subsection~\ref{ssec:acg}. NOTE: This is too verbose both to put on
-    the paper and also to code. These examples demonstrates two common
-    and related shapes of lexicons. The $\mathcal{L}$(\_{Syntax}) and
-    the first half of the $\mathcal{L}$(\_{Sem}) are lexicalized, in the
-    sense that the object term contains a lexical constant with the same
-    lexical entry as the abstract constant.\footnote{This notion of
-      ``lexicalized'' is quite different to the one I use in the rest of
-      the thesis so far (which roughly means ``generated by a lexical
-      database'') and also different from the one used by Kanazawa
-      (FIX!).} This means that before constructing the object term, we
-    have to acquire the object constant that is contained in the term
-    and then we just fit it in of the recipes, which we choose according
-    to the hypertag. The second half of $\mathcal{L}$(\_{Sem}) simply
-    assigns object terms based on the hypertag. A more concise and
-    desirable implementation would allow us simply write the pairs of
-    hypertags and object terms (recipes encoded as lambdas). My plan is
-    to have exactly this, but I have not thought yet about how to split
-    the various aspects into different reusable abstractions.}
+  \caption{\label{fig:lex-impl} The implementation of the
+    $\mathcal{L}$(\_{Syntax}) and $\mathcal{L}$(\_{Sem}) lexicons from
+    subsection~\ref{ssec:acg}.}
 \end{figure}
 
+ NOTE: This is too verbose both to put on the paper and also to
+ code. These examples demonstrate two common and related shapes of
+ lexicons. The $\mathcal{L}$(\_{Syntax}) and the first half of the
+ $\mathcal{L}$(\_{Sem}) are lexicalized, in the sense that the object
+ term contains a lexical constant with the same lexical entry as the
+ abstract constant.\footnote{This notion of ``lexicalized'' is quite
+   different to the one I use in the rest of the thesis so far (which
+   roughly means ``generated by a lexical database'') and also different
+   from the one used by Kanazawa (FIX!).} This means that before
+ constructing the object term, we have to acquire the object constant
+ that is contained in the term and then we just fit it in of the
+ recipes, which we choose according to the hypertag. The second half of
+ $\mathcal{L}$(\_{Sem}) simply assigns object terms based on the
+ hypertag. A more concise and desirable implementation would allow us
+ simply write the pairs of hypertags and object terms (recipes encoded
+ as lambdas). My plan is to have exactly this, but I have not thought
+ yet about how to split the various aspects into different reusable
+ abstractions.
+
 \subsection{Testing signatures and lexicons}

### 2013-06-10 the-last-part.tex
Talked about the 3 mechs and their naive combo

\section{The Last Part}

In this section, we will attempt to incorporate accounts of several
linguistic phenomena in a single framework, highlight the modularity
issues this brings up and propose some solutions to them.

\subsection{Negation}

We will start with an account of the paired grammatical words mechanism
for negation in French. Recall the EPTDs we have seen
in~\ref{sssec:frigram}, repeated here on Figure~\ref{fig:ig-neg-rep}.

\begin{figure}
  \centering
  \includegraphics[scale=0.25]{ig-neg.pdf}
  \caption{\label{fig:ig-neg-rep} Frigram EPTDs for a negative
    determiner and the paired grammatical word $ne$.}
\end{figure}

At a basic level, $aucun$ functions like any other determiner and $ne$
as any verb modifier. However, we want to encode the constraint that
whenever a verb is modified by $ne$, the modified verb will demand that
one of its arguments contains a negative determiner. Since we need to
make a distinction between phrases that contain negative determiners and
those that do not, we will need to provide different types for both (two
terms with the same type are indistinguishable w.r.t. syntactic
composition in a type-logical grammar such as ACGs). Our type for
$aucun$ will thus end up being $N \limp ((NP\_NEG \limp S) \limp S)$
instead of the usual $N \limp ((NP \limp S) \limp S)$.\footnote{Strictly
  following the EPTD for $aucun$ would lead us to give it a type
  $DET\_NEG$ and then give the two types $DET \limp ((NP \limp S) \limp
  S)$ and $DET\_NEG \limp ((NP\_NEG \limp S) \limp S)$ to nouns. In this
  demonstration, we prefer to align our examples with the treatment
  which is customary to ACG research.}  The type of $ne$ will be more
verbose, as it will consume a verb and transform its valency so that it
demands that exactly one of its $NP$ arguments contains a negative
determiner.

$$
N_{ne_{tv_1}} : (NP \limp NP \limp S) \limp (NP\_NEG \limp NP \limp S)
$$
$$
N_{ne_{tv_2}} : (NP \limp NP \limp S) \limp (NP \limp NP\_NEG \limp S)
$$

Furthermore, we could add the case when both the subject and the object
contain negative determiners, such as in~(\ref{ex:aucun2}). This case is
currently not covered by Frigram (though it could be), but is considered
grammatical by French speakers. Finally, in a complete grammar, $ne$
would have to provide types capable of transforming all the other
syntactic valencies.

\begin{exe}
  \ex \label{ex:aucun2} Aucune fourmi n'aime aucun tatou.
\end{exe}

$$
N_{ne_{tv_3}} : (NP \limp NP \limp S) \limp (NP\_NEG \limp NP\_NEG \limp S)
$$

\begin{figure}
  \centering
  \includegraphics[scale=0.25]{parse-aucun.pdf}
  \caption{\label{fig:parse-aucun} One of the parse trees assigned to
    sentence (\ref{ex:aucun-obj}) by Frigram/Leopar.}
\end{figure}

However, this is not the only change we have to effect on our grammar in
order to properly handle the paired grammatical words for
negation. Consider the example sentence (\ref{ex:aucun-obj}) and the
parse tree assigned to it by Frigram on Figure~\ref{fig:parse-aucun}. It
is not enough that the embedded noun phrase $aucun\ tatou$ has a type
that tells us it contains a negative determiner. We need this bit of
internal information also at the level of the enclosing noun phrase
$l'odeur\ d'aucun\ tatou$ in which the former noun phrase is embedded
since this is a property of the enclosing noun phrase that is crucial to
its syntactic combinatorics. This presupposes the existence of a
mechanism for propagating this kind of information up the parse tree. In
the case of our ACG, this means that all the functions which modify
phrases must be aware whether or not their argument contains a negative
determiner and to convey this information in its result type as
well. Moreover, since the negative determiner can also come from a
complement of a prepositional phrase, the types of prepositions will
need to take into account the presence of negative determiners in both
their complements and the phrases they modify.

$$
N_{que_1} : (NP \limp S) \limp N \limp N
$$
$$
N_{que_2} : (NP \limp S) \limp N\_NEG \limp N\_NEG
$$
$$
N_{de_1} : NP \limp N \limp N
$$
$$
N_{de_2} : NP \limp N\_NEG \limp N\_NEG
$$
$$
N_{de_3} : NP\_NEG \limp N \limp N\_NEG
$$

If we suppose the presence of the appropriate lexical items typed as in
the examples of~\ref{ssec:acg}, we have a type system in which we can
type the syntactic terms corresponding to the sentences
(\ref{ex:good-neg-double}), (\ref{ex:good-neg-embed}) and
(\ref{ex:good-neg-rel}) while making it impossible to type the
structures corresponding to sentences (\ref{ex:bad-neg-noneg}) and
({\ref{ex:bad-neg-rel}}).

In sentence (\ref{ex:bad-neg-noneg}), we have no constant
$N_{ne_{tv_?}}$ for $ne$ which would yield a verb capable of consuming
two noun phrases that contain no negative determiners.

In sentence (\ref{ex:bad-neg-rel}), there are multiple problems with
typing. First, $N_{chasse}$ requires its first argument to be of type
$NP$, a noun phrase with no negative determiner. This means that
$\lambda^{\circ} x.\ N_{chasse}\ x\ y$ has type $NP \limp S$ which makes
it an invalid argument for the function $N_{aucun}\ N_{loup}$, which has
type $(NP\_NEG \limp S) \limp S$. Even supposing that we would be able
to type the relative clause with $NP \limp S$, the complementizer $que$
would map $N_{le}\ N_{tatou}$ from type $NP$ to type $NP$, preserving
the property that the noun phrase does not contain a negative
determiner. This would then make it incompatible with
$N_{ne_{iv}}\ N_{court}$ which expects its argument to be an
$NP\_NEG$. By misplacing the paired grammatical words, both the main
clause and the relative clause become untypable.

\begin{exe}
  \ex \label{ex:good-neg-double} Aucune fourmi n'aime aucun tatou. \\
      $(N_{aucune}\ N_{fourmi})\ (\lambda^{\circ} x.\ (N_{aucun}\ N_{tatou})\ (\lambda^{\circ} y.\ N_{ne_{tv_3}}\ N_{aime}\ x\ y))$
  \ex \label{ex:good-neg-embed} Jean n'aime l'odeur d'aucun tatou. \\
      $(N_{le}\ (N_{de_3}\ (N_{aucun}\ N_{tatou})\ N_{odeur}))\ (\lambda^{\circ} y.\ N_{ne_{tv_2}}\ N_{aime}\ N_{Jean}\ y)$
  \ex \label{ex:good-neg-rel} Le tatou qu'aucun loup ne chasse court. \\
      $(N_{le}\ (N_{que}\ (\lambda^{\circ} y.\ (N_{aucun}\ N_{loup})\ (\lambda^{\circ} x.\ N_{ne_{tv_1}}\ N_{chasse}\ x\ y))\ N_{tatou}))\ (\lambda^{\circ} x.\ N_{court}\ x)$
  \ex * \label{ex:bad-neg-noneg} Jean n'aime le tatou. \\
      ? $(N_{le}\ N_{tatou})\ (\lambda^{\circ} y.\ N_{ne_{tv_?}}\ N_{aime}\ N_{Jean}\ y)$
  \ex * \label{ex:bad-neg-rel} Le tatou qu'aucun loup chasse ne court. \\
      ? $(N_{le}\ (N_{que}\ (\lambda^{\circ} y.\ (N_{aucun}\ N_{loup})\ (\lambda^{\circ} x.\ N_{chasse_?}\ x\ y))\ N_{tatou}))\ (\lambda^{\circ} x.\ N_{ne_{iv_?}}\ N_{court}\ x)$
\end{exe}

There are two things that we would like to highlight about this
grammatical treatment. First off, we had to refine our types to make
them convey more than one piece of information in a single type
(e.g. something is a noun phrase and at the same time it is something
that contains a negative determiner). Second, a property that we wish to
express in the type of a term may be due to one of its subconstituents
and it is therefore necessary to encode the propagation of this property
up the parse tree in the types of all the intervening operators. In our
case, it meant that handling negative determiners and the paired
grammatical word for negation required us to not only define and type
constants for these two categories, but to also change the types of noun
modifiers ($que\ ...$) and prepositions ($de$).

\subsection{Extraction}
\label{ssec:extraction}

We will now turn to another phenomenon, extraction in relative clauses,
focus on one of its properties, the fact that subjects can be extracted
using $qui$ only from the relative clause itself whereas objects can be
extracted using $que$ from embedded clauses contained therein, and
present an implementation of this constraint by way of types in an
ACG. The solution that we will show was published in
\cite{pogodalla2012controlling}. Our presentation is slightly adapted to
conform to the patterns seen in previous examples and it does not use
dependent types as they are not essential to this specific constraint.

The key idea in the treatment presented below is to give a special type
to extracted constituents. We will thus have two types for noun phrases:
$NP\_VAR$, which stands for ``empty'' noun phrases introduced by
extraction, and $NP$, which stands for the other noun phrases, i.e.
proper names and nouns with determiners. The grammar gives no constant
capable of constructing a value of type $NP\_VAR$. The only way to
obtain a term with this type is to abstract over a variable having the
type. Then, the types of relative pronouns end up being $(NP\_VAR \limp
S) \limp N \limp N$, ensuring that the relative clause given as the
first argument is a $\lambda$-abstraction whose argument has type
$NP\_VAR$.

Once we have this kind of information in the type system, we can start
distinguishing clauses containing rooted extraction, where a subject or
object is extracted directly from the relative clause, embedded
extraction, where an object is extracted from a clause embedded inside
the relative clause, or no extraction. Types of functions which produce
values of type $S$ will be sensitive to the presence of extracted
variables in their noun phrase and clausal arguments.

$$
E_{dort_1} : NP \limp S
$$
$$
E_{dort_2} : NP\_VAR \limp S\_ROOT
$$
$$
E_{aime_1} : NP \limp NP \limp S
$$
$$
E_{aime_2} : NP\_VAR \limp NP \limp S\_ROOT
$$
$$
E_{aime_3} : NP \limp NP\_VAR \limp S\_ROOT
$$
$$
E_{dit\ que_1} : NP \limp S \limp S
$$
$$
E_{dit\ que_2} : NP\_VAR \limp S \limp S\_ROOT
$$
$$
E_{dit\ que_3} : NP \limp S\_ROOT \limp S\_EMB
$$
$$
E_{dit\ que_4} : NP \limp S\_EMB \limp S\_EMB
$$

Now all of the infrastructure is in place for us to express the type of
the relative pronouns $qui$ and $que$.

$$
E_{qui} : (NP\_VAR \limp S\_ROOT) \limp N \limp N
$$
$$
E_{que_1} : (NP\_VAR \limp S\_ROOT) \limp N \limp N
$$
$$
E_{que_2} : (NP\_VAR \limp S\_EMB) \limp N \limp N
$$

The above types ensure that the abstracted variables corresponding to
traces of extracted constituents will be given the type $NP\_VAR$, that
$qui$ can only be used to extract constituents from root positions while
$que$ allows for extraction from both root and embedded
positions.\footnote{NB: The constraint that $qui$ can only extract
  subjects and that $que$ can only extract objects is not enforced
  here. Likewise, this fragment does not handle multiple extraction.}

Given this type system, we can assign the type $S$ to the term
expressing the syntactic structure of (\ref{ex:good-ext}) but we cannot
a find a typable term for sentence (\ref{ex:bad-ext}), where $qui$ is
used to extract a subject from a clause embedded in the relative clause.
The problem with sentence (\ref{ex:bad-ext}) is that the type of the
relative clause is $NP\_VAR \limp S\_EMB$ which is not a valid argument
for any constant representing the relative pronoun $qui$. Typing the
variable $t$ with $NP$ instead of $NP\_VAR$ and using $E_{aime_1}$
instead of $E_{aime_2}$ would not help, since the resulting relative
clause would then have type $NP \limp S$, which would not admitted by
any relative pronoun.

\begin{exe}
  \ex * \label{ex:bad-ext} Le tatou qui Jean dit que \_ aime Marie dort. \\
      ? $(E_{le}\ (E_{qui_?}\ (\lambda^{\circ} t.\ E_{dit\ que_3}\ E_{Jean}\ (E_{aime_2}\ t\ E_{Marie}))\ E_{tatou}))\ (\lambda^{\circ} x.\ E_{dort_1}\ x)$
  \ex \label{ex:good-ext} Le tatou que Jean dit que Marie aime \_ dort. \\
      $(E_{le}\ (E_{que_2}\ (\lambda^{\circ} t.\ E_{dit\ que_3}\ E_{Jean}\ (E_{aime_3}\ E_{Marie}\ t))\ E_{tatou}))\ (\lambda^{\circ} x.\ E_{dort_1}\ x)$
\end{exe}

As was the case with our treatment of negation, we refined our familiar
types by tacking on new kinds of information and defined rules for how
these pieces of information percolate up the parse tree all the way to
the topmost level still pertinent for ensuring grammaticality.

\subsection{Agreement}

Finally, we will cover the phenomenon of number agreement. This
motivates a straightforward refinement of our types, where all noun and
noun phrase types carry number information.

$$
A_{Marie} : NP\_SG
$$
$$
A_{tatou} : N\_SG
$$
$$
A_{tatous} : N\_PL
$$
$$
A_{le} : N\_SG \limp ((NP\_SG \limp S) \limp S)
$$
$$
A_{les} : N\_PL \limp ((NP\_PL \limp S) \limp S)
$$
$$
A_{dort} : NP\_SG \limp S
$$
$$
A_{dorment} : NP\_PL \limp S
$$
$$
A_{aime_1} : NP\_SG \limp NP\_SG \limp S
$$
$$
A_{aime_2} : NP\_SG \limp NP\_PL \limp S
$$
$$
A_{aiment_1} : NP\_PL \limp NP\_SG \limp S
$$
$$
A_{aiment_2} : NP\_PL \limp NP\_PL \limp S
$$
$$
A_{qui_1} : (NP\_SG \limp S) \limp N\_SG \limp N\_SG
$$
$$
A_{qui_2} : (NP\_PL \limp S) \limp N\_PL \limp N\_PL
$$
$$
A_{de_1} : NP\_SG \limp N\_SG \limp N\_SG
$$
$$
A_{de_2} : NP\_PL \limp N\_SG \limp N\_SG
$$
$$
A_{de_3} : NP\_SG \limp N\_PL \limp N\_PL
$$
$$
A_{de_4} : NP\_PL \limp N\_Pl \limp N\_PL
$$

Given the above signature, we can assign the type $S$ to the syntactic
term of sentence (\ref{ex:good-agr}) but not to the one of sentence
(\ref{ex:bad-agr}). In the latter sentence, the type of the variable $t$
has to be $NP\_SG$ since it is used as an argument to $A_{dort}$. This
means that the relative clause $dort$ has type $NP\_SG \limp
S$. However, there is no type for $qui$ which would allow us to modify
the plural noun $A_{tatous}$ of type $N\_PL$ by a relative clause with a
missing singular noun phrase, which has the type $NP\_SG \limp S$.

\begin{exe}
  \ex \label{ex:good-agr} Les tatous de Marie dorment. \\
      $(A_{les}\ (A_{de}\ A_{Marie}\ A_{tatous}))\ (\lambda^{\circ} x.\ A_{dorment}\ x)$
  \ex * \label{ex:bad-agr} Marie aime les tatous qui dort. \\
      ? $(A_{les}\ (A_{qui_?}\ (\lambda^{\circ} t.\ A_{dort}\ t)\ A_{tatous}))\ (\lambda^{\circ} y.\ A_{aime_2}\ A_{Marie}\ y)$
\end{exe}

Once again, we have expressed a linguistic constraint by refining our
types, augmenting them with a new bit of information, we constrained the
arguments of functions using these new types and we have described how
this new type information propagates from the given arguments to the
produced values.

In the next subsection, we will see what it takes to enforce all of the
constraints above (negation, extraction and agreement) at the same time,
which will give us some idea on how to go about building a grammar which
handles a wide variety of such phenomena.

\subsection{Putting It All Together}

In a grammar that handles negation, extraction and agreement, our types
will need to carry all of the refinements we introduced before. To
recapitulate, $NP$ will be parameterized by whether or not it contains a
negative determiner ($\_\textcolor{red}{NEG{=}T}$ and
$\_\textcolor{red}{NEG{=}F}$ respectively), whether or not it is a trace
from an extraction ($\_\textcolor{green}{VAR{=}T}$ and
$\_\textcolor{green}{VAR{=}F}$) and its number
($\_\textcolor{blue}{NUM{=}SG}$ or $\_\textcolor{blue}{NUM{=}PL}$). For
$N$, we have the same refinements except for $\textcolor{green}{VAR}$
since extraction only moves $NP$s. Clauses, of type $S$, will be
discriminated based on the level of extraction within them
($\_\textcolor{green}{EXT{=}NO}$, $\_\textcolor{green}{EXT{=}ROOT}$ and
$\_\textcolor{green}{EXT{=}EMB}$).

The individual types will have to respect all of the constraints at the
same time, each one being a complete specification of a possible
situation (complete w.r.t. the features listed above). To see and
appreciate what this means in practice, we give the types for the
preposition $de$:

$$
C_{de_1} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG})
$$
$$
C_{de_2} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL})
$$
$$
C_{de_3} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
$$
$$
C_{de_4} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
$$
$$
C_{de_5} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
$$
$$
C_{de_6} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
$$
$$
C_{de_7} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG})
$$
$$
C_{de_8} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL})
$$
$$
C_{de_9} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
$$
$$
C_{de_{10}} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
$$
$$
C_{de_{11}} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
$$
$$
C_{de_{12}} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
$$

This highlights two important problems with the current approach. One is
the untenable complexity. Even though the three mechanisms are
completely independent of each other, here we are forced to refer to all
of them in every type. It is no longer easy to glance at the types
assigned to $de$ and see how it behaves with respect to, for example,
the mechanism of negation (compare with the type assignments we gave to
$de$ in previous subsections). Another cause of unreadability of this
signature is due to the sheer number of types on display, which leads us
to our second issue.

The signature demonstrates an exponential growth in the number of typed
constants included. In cases where the behaviors of a wordform across
the different mechanisms are independent, the complete type signature
will need to provide a type for every combination of situations in all
of the mechanisms. Assuming that every mechanism will assign on average
more than one type to every wordform, the number of the wordform's types
in the final signature will be exponential w.r.t. the number of
different mechanisms handled.

We can try making the enumeration of these types easier for the grammar
author by introducing metavariables for the values of the features. This
means that we could write the following 3 templates instead of the 12
types above:

$$
\forall \textcolor{blue}{x}, \textcolor{blue}{y} \in \{\textcolor{blue}{SG}, \textcolor{blue}{PL}\}
$$
$$
C_{de_{1(\textcolor{blue}{x}, \textcolor{blue}{y})}} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y})
$$
$$
C_{de_{2(\textcolor{blue}{x}, \textcolor{blue}{y})}} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y})
$$
$$
C_{de_{3(\textcolor{blue}{x}, \textcolor{blue}{y})}} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y})
$$

This is getting more concise and easier to comprehend. We can take this
approach further by not only quantifying over variables but also
computing some of the values using functions:

$$
\forall \textcolor{blue}{x}, \textcolor{blue}{y} \in\{\textcolor{blue}{SG}, \textcolor{blue}{PL}\}, \forall \textcolor{red}{p}, \textcolor{red}{q} \in \{\textcolor{red}{F}, \textcolor{red}{T}\}, \textcolor{red}{p \land q} \neq \textcolor{red}{T}
$$
$$
C_{de(\textcolor{blue}{x}, \textcolor{blue}{y}, \textcolor{red}{p}, \textcolor{red}{q})} : (NP\_\textcolor{red}{NEG{=}p}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}q}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}(p \lor q)}\_\textcolor{blue}{NUM{=}y})
$$

The above is vastly more concise and manages to convey the intent more
clearly (the intent being that it cannot be the case that there are
negative determiners in both the complement noun phrase and the modified
noun and that the value of the result's \textcolor{red}{NEG} feature is
always the disjunction of the \textcolor{red}{NEG} features of the two
arguments).

As of right now, we have introduced this kind of rule only as
meta-notation. This means that the real underlying grammar that will end
up being used by some algorithm still suffers from the same exponential
cardinality of the signature. Fortunately, there exists a well-studied
construction in type theory that lets us assign to terms generic types
such as the one defined in the template above. This construction is the
\emph{dependent product} and its utility in ACGs has already been
noted.\footnote{See \cite{de2007two} for its introduction into the
  domain of ACGs, \cite{de2007type} for a small example grammar, which
  handles the interaction between number and gender agreement and
  coordination, and finally see \cite{pogodalla2012controlling} for
  using dependent types to handle several constraints regarding
  extraction (including the one we covered in~\ref{ssec:extraction}).}

While extending the type system by including dependent types is
definitely useful and it helps to solve the problem of having our
signatures grow exponentially in size, it does not solve our problem
with the complexity of the new grammar. Granted, the new meta-type
written in the dependent style is more readable and easier to reason
about than the enumeration of 12 basic types we presented at the
beginning of this subsection, however, the result is still complex in
that it forces us to describe the behavior of three independent
mechanisms in one place. This one place happens to be a wordform which
has seemingly little connection to these mechanisms, the preposition
$de$. We can therefore imagine that it will participate in many other
phenomena and that the type we assign to it will grow beyond our
abilities to reason about it effectively.

Furthermore, the current notation does not even make it obvious that
these are independent mechanisms. One has to examine the type to make
sure that there is no interaction. In the next two subsections, we will
devote ourselves to investigating ways of disentangling these.

\subsection{Intersections of Types}

### 2013-06-11 the-last-part.tex
Fleshing out the type systems for the 3 mechs

@@
 composition in a type-logical grammar such as ACGs). Our type for
-$aucun$ will thus end up being $N \limp ((NP\_NEG \limp S) \limp S)$
-instead of the usual $N \limp ((NP \limp S) \limp S)$.\footnote{Strictly
-  following the EPTD for $aucun$ would lead us to give it a type
-  $DET\_NEG$ and then give the two types $DET \limp ((NP \limp S) \limp
-  S)$ and $DET\_NEG \limp ((NP\_NEG \limp S) \limp S)$ to nouns. In this
+$aucun$ will thus end up being $N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S)
+\limp S)$ instead of the usual $N \limp ((NP \limp S) \limp
+S)$.\footnote{Strictly following the EPTD for $aucun$ would lead us to
+  give it a type $DET\_NEG{=}F$ and then give the two types
+  $DET\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S)$ and $DET\_NEG{=}T
+  \limp ((NP\_NEG{=}T \limp S) \limp S)$ to nouns. In this
   demonstration, we prefer to align our examples with the treatment
-  which is customary to ACG research.}  The type of $ne$ will be more
-verbose, as it will consume a verb and transform its valency so that it
-demands that exactly one of its $NP$ arguments contains a negative
-determiner.
+  which is customary to ACG research.} The type for $le$ will need to be
+able handle cases where its argument noun either already contains a
+negative determiner or not.
 
 $$
-N_{ne_{tv_1}} : (NP \limp NP \limp S) \limp (NP\_NEG \limp NP \limp S)
+N_{aucun} : N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S) \limp S)
 $$
 $$
-N_{ne_{tv_2}} : (NP \limp NP \limp S) \limp (NP \limp NP\_NEG \limp S)
+N_{le_1} : N\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S)
+$$
+$$
+N_{le_2} : N\_NEG{=}T \limp ((NP\_NEG{=}T \limp S) \limp S)
+$$
+
+The type of $ne$ will be more verbose, as it will consume a verb and
+transform its valency so that it demands that exactly one of its $NP$
+arguments contains a negative determiner.
+
+$$
+N_{ne_{tv_1}} : (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}F \limp S)
+$$
+$$
+N_{ne_{tv_2}} : (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}F \limp NP\_NEG{=}T \limp S)
 $$
@@
 $$
-N_{ne_{tv_3}} : (NP \limp NP \limp S) \limp (NP\_NEG \limp NP\_NEG \limp S)
+N_{ne_{tv_3}} : (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}T \limp S)
 $$
@@
 $$
-N_{que_1} : (NP \limp S) \limp N \limp N
+N_{que_1} : (NP\_NEG{=}F \limp S) \limp N\_NEG{=}F \limp N\_NEG{=}F
 $$
 $$
-N_{que_2} : (NP \limp S) \limp N\_NEG \limp N\_NEG
+N_{que_2} : (NP\_NEG{=}F \limp S) \limp N\_NEG{=}T \limp N\_NEG{=}T
 $$
 $$
-N_{de_1} : NP \limp N \limp N
+N_{de_1} : NP\_NEG{=}F \limp N\_NEG{=}F \limp N\_NEG{=}F
 $$
 $$
-N_{de_2} : NP \limp N\_NEG \limp N\_NEG
+N_{de_2} : NP\_NEG{=}F \limp N\_NEG{=}T \limp N\_NEG{=}T
 $$
 $$
-N_{de_3} : NP\_NEG \limp N \limp N\_NEG
+N_{de_3} : NP\_NEG{=}T \limp N\_NEG{=}F \limp N\_NEG{=}T
 $$
@@
 typing. First, $N_{chasse}$ requires its first argument to be of type
-$NP$, a noun phrase with no negative determiner. This means that
-$\lambda^{\circ} x.\ N_{chasse}\ x\ y$ has type $NP \limp S$ which makes
-it an invalid argument for the function $N_{aucun}\ N_{loup}$, which has
-type $(NP\_NEG \limp S) \limp S$. Even supposing that we would be able
-to type the relative clause with $NP \limp S$, the complementizer $que$
-would map $N_{le}\ N_{tatou}$ from type $NP$ to type $NP$, preserving
-the property that the noun phrase does not contain a negative
-determiner. This would then make it incompatible with
-$N_{ne_{iv}}\ N_{court}$ which expects its argument to be an
-$NP\_NEG$. By misplacing the paired grammatical words, both the main
-clause and the relative clause become untypable.
+$NP\_NEG{=}F$, a noun phrase with no negative determiner. This means
+that $\lambda^{\circ} x.\ N_{chasse}\ x\ y$ has type $NP\_NEG{=}F \limp
+S$ which makes it an invalid argument for the function
+$N_{aucun}\ N_{loup}$, which has type $(NP\_NEG{=}T \limp S) \limp
+S$. Even if we were able to assign it a type compatible with
+$N_{que_1}$, the final relative clause would map $N_{tatou}$ to another
+term of type $N\_NEG{=}F$, which has no ``free'' negative
+determiners. $N_{le_1}$ will map this to a term typed $(NP\_NEG{=}F
+\limp S) \limp S$. Finally, this term is not capable of taking
+$(\lambda^{\circ} x.\ N_{ne_{iv}}\ N_{court}\ x)$ as its argument, since
+its type is $NP\_NEG{=}T \limp S$.
 
@@
   \ex \label{ex:good-neg-embed} Jean n'aime l'odeur d'aucun tatou. \\
-      $(N_{le}\ (N_{de_3}\ (N_{aucun}\ N_{tatou})\ N_{odeur}))\ (\lambda^{\circ} y.\ N_{ne_{tv_2}}\ N_{aime}\ N_{Jean}\ y)$
+      $(N_{le_2}\ (N_{de_3}\ (N_{aucun}\ N_{tatou})\ N_{odeur}))\ (\lambda^{\circ} y.\ N_{ne_{tv_2}}\ N_{aime}\ N_{Jean}\ y)$
   \ex \label{ex:good-neg-rel} Le tatou qu'aucun loup ne chasse court. \\
-      $(N_{le}\ (N_{que}\ (\lambda^{\circ} y.\ (N_{aucun}\ N_{loup})\ (\lambda^{\circ} x.\ N_{ne_{tv_1}}\ N_{chasse}\ x\ y))\ N_{tatou}))\ (\lambda^{\circ} x.\ N_{court}\ x)$
+      $(N_{le_1}\ (N_{que_1}\ (\lambda^{\circ} y.\ (N_{aucun}\ N_{loup})\ (\lambda^{\circ} x.\ N_{ne_{tv_1}}\ N_{chasse}\ x\ y))\ N_{tatou}))\ (\lambda^{\circ} x.\ N_{court}\ x)$
   \ex * \label{ex:bad-neg-noneg} Jean n'aime le tatou. \\
-      ? $(N_{le}\ N_{tatou})\ (\lambda^{\circ} y.\ N_{ne_{tv_?}}\ N_{aime}\ N_{Jean}\ y)$
+      * $(N_{le_1}\ N_{tatou})\ (\lambda^{\circ} y.\ N_{ne_{tv_?}}\ N_{aime}\ N_{Jean}\ y)$
   \ex * \label{ex:bad-neg-rel} Le tatou qu'aucun loup chasse ne court. \\
-      ? $(N_{le}\ (N_{que}\ (\lambda^{\circ} y.\ (N_{aucun}\ N_{loup})\ (\lambda^{\circ} x.\ N_{chasse_?}\ x\ y))\ N_{tatou}))\ (\lambda^{\circ} x.\ N_{ne_{iv_?}}\ N_{court}\ x)$
+      * $(N_{le_1}\ (N_{que}\ (\lambda^{\circ} y.\ (N_{aucun}\ N_{loup})\ (\lambda^{\circ} x.\ N_{chasse}\ x\ y))\ N_{tatou}))\ (\lambda^{\circ} x.\ N_{ne_{iv}}\ N_{court}\ x)$
 \end{exe}
@@
 to extracted constituents. We will thus have two types for noun phrases:
-$NP\_VAR$, which stands for ``empty'' noun phrases introduced by
-extraction, and $NP$, which stands for the other noun phrases, i.e.
-proper names and nouns with determiners. The grammar gives no constant
-capable of constructing a value of type $NP\_VAR$. The only way to
-obtain a term with this type is to abstract over a variable having the
-type. Then, the types of relative pronouns end up being $(NP\_VAR \limp
-S) \limp N \limp N$, ensuring that the relative clause given as the
-first argument is a $\lambda$-abstraction whose argument has type
-$NP\_VAR$.
+$NP\_VAR{=}T$, which stands for ``empty'' noun phrases introduced by
+extraction, and $NP\_VAR{=}F$, which stands for the other noun phrases,
+i.e.  proper names and nouns with determiners. The grammar gives no
+constant capable of constructing a value of type $NP\_VAR{=}T$. The only
+way to obtain a term with this type is to abstract over a variable
+having the type. Then, the types of relative pronouns end up being
+$(NP\_VAR{=}T \limp S) \limp N \limp N$, ensuring that the incomplete
+clause given as the first argument is a $\lambda$-abstraction whose
+argument has type $NP\_VAR{=}T$.
 
@@
 $$
-E_{dort_1} : NP \limp S
+E_{dort_1} : NP\_VAR{=}F \limp S\_EXT{=}NO
 $$
 $$
-E_{dort_2} : NP\_VAR \limp S\_ROOT
+E_{dort_2} : NP\_VAR{=}T \limp S\_EXT{=}ROOT
 $$
 $$
-E_{aime_1} : NP \limp NP \limp S
+E_{aime_1} : NP\_VAR{=}F \limp NP\_VAR{=}F \limp S\_EXT{=}NO
 $$
 $$
-E_{aime_2} : NP\_VAR \limp NP \limp S\_ROOT
+E_{aime_2} : NP\_VAR{=}T \limp NP\_VAR{=}F \limp S\_EXT{=}ROOT
 $$
 $$
-E_{aime_3} : NP \limp NP\_VAR \limp S\_ROOT
+E_{aime_3} : NP\_VAR{=}F \limp NP\_VAR{=}T \limp S\_EXT{=}ROOT
 $$
 $$
-E_{dit\ que_1} : NP \limp S \limp S
+E_{dit\ que_1} : NP\_VAR{=}F \limp S\_EXT{=}NO \limp S\_EXT{=}NO
 $$
 $$
-E_{dit\ que_2} : NP\_VAR \limp S \limp S\_ROOT
+E_{dit\ que_2} : NP\_VAR{=}T \limp S\_EXT{=}NO \limp S\_EXT{=}ROOT
 $$
 $$
-E_{dit\ que_3} : NP \limp S\_ROOT \limp S\_EMB
+E_{dit\ que_3} : NP\_VAR{=}F \limp S\_EXT{=}ROOT \limp S\_EXT{=}EMB
 $$
 $$
-E_{dit\ que_4} : NP \limp S\_EMB \limp S\_EMB
+E_{dit\ que_4} : NP\_VAR{=}F \limp S\_EXT{=}EMB \limp S\_EXT{=}EMB
 $$
@@
 $$
-E_{qui} : (NP\_VAR \limp S\_ROOT) \limp N \limp N
+E_{qui} : (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N
 $$
 $$
-E_{que_1} : (NP\_VAR \limp S\_ROOT) \limp N \limp N
+E_{que_1} : (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N
 $$
 $$
-E_{que_2} : (NP\_VAR \limp S\_EMB) \limp N \limp N
+E_{que_2} : (NP\_VAR{=}T \limp S\_EXT{=}EMB) \limp N \limp N
 $$
@@
 The above types ensure that the abstracted variables corresponding to
-traces of extracted constituents will be given the type $NP\_VAR$, that
-$qui$ can only be used to extract constituents from root positions while
-$que$ allows for extraction from both root and embedded
+traces of extracted constituents will be given the type $NP\_VAR{=}T$,
+that $qui$ can only be used to extract constituents from root positions
+while $que$ allows for extraction from both root and embedded
 positions.\footnote{NB: The constraint that $qui$ can only extract
@@
 
-Given this type system, we can assign the type $S$ to the term
+We also note that the splitting of the clause type $S$ into three finer
+types $S\_EXT={NO}$, $S\_EXT{=}ROOT$ and $S\_EXT{=}EMB$ means that types
+that work with $S$ regardless of its extraction status will have to
+provide alternatives for all the possible variants.
+
+$$
+E_{le_1} : N \limp ((NP\_VAR{=}F \limp S\_EXT{=}NO) \limp S\_EXT{=}NO)
+$$
+$$
+E_{le_2} : N \limp ((NP\_VAR{=}F \limp S\_EXT{=}ROOT) \limp S\_EXT{=}ROOT)
+$$
+$$
+E_{le_3} : N \limp ((NP\_VAR{=}F \limp S\_EXT{=}EMB) \limp S\_EXT{=}EMB)
+$$
+
+While we have introduced the notion of an $NP\_VAR{=}T$ for describing
+traces, we can cover another element of our fragment, the preposition
+$de$, and enforce the constraint that $qui$/$que$ cannot extract
+prepositional complements.
+
+$$
+E_{de} : NP\_VAR{=}F \limp N \limp N
+$$
+
+Given this type system, we can assign the type $S\_EXT{=}NO$ to the term
 expressing the syntactic structure of (\ref{ex:good-ext}) but we cannot
 a find a typable term for sentence (\ref{ex:bad-ext}), where $qui$ is
-used to extract a subject from a clause embedded in the relative clause.
+used to extract a subject from a clause embedded in a relative clause.
 The problem with sentence (\ref{ex:bad-ext}) is that the type of the
-relative clause is $NP\_VAR \limp S\_EMB$ which is not a valid argument
-for any constant representing the relative pronoun $qui$. Typing the
-variable $t$ with $NP$ instead of $NP\_VAR$ and using $E_{aime_1}$
-instead of $E_{aime_2}$ would not help, since the resulting relative
-clause would then have type $NP \limp S$, which would not admitted by
-any relative pronoun.
+relative clause is $NP\_VAR{=}T \limp S\_EXT{=}EMB$ which is not a valid
+argument for any constant representing the relative pronoun
+$qui$. Typing the variable $t$ with $NP\_VAR{=}F$ instead of
+$NP\_VAR{=}T$ and using $E_{aime_1}$ instead of $E_{aime_2}$ would not
+help, since the resulting relative clause would then have type
+$NP\_VAR{=}F \limp S\_EXT{=}NO$, which would not be admitted by any
+relative pronoun.
 
@@
   \ex * \label{ex:bad-ext} Le tatou qui Jean dit que \_ aime Marie dort. \\
-      ? $(E_{le}\ (E_{qui_?}\ (\lambda^{\circ} t.\ E_{dit\ que_3}\ E_{Jean}\ (E_{aime_2}\ t\ E_{Marie}))\ E_{tatou}))\ (\lambda^{\circ} x.\ E_{dort_1}\ x)$
+      * $(E_{le_1}\ (E_{qui_?}\ (\lambda^{\circ} t.\ E_{dit\ que_3}\ E_{Jean}\ (E_{aime_2}\ t\ E_{Marie}))\ E_{tatou}))\ (\lambda^{\circ} x.\ E_{dort_1}\ x)$
   \ex \label{ex:good-ext} Le tatou que Jean dit que Marie aime \_ dort. \\
-      $(E_{le}\ (E_{que_2}\ (\lambda^{\circ} t.\ E_{dit\ que_3}\ E_{Jean}\ (E_{aime_3}\ E_{Marie}\ t))\ E_{tatou}))\ (\lambda^{\circ} x.\ E_{dort_1}\ x)$
+      $(E_{le_1}\ (E_{que_2}\ (\lambda^{\circ} t.\ E_{dit\ que_3}\ E_{Jean}\ (E_{aime_3}\ E_{Marie}\ t))\ E_{tatou}))\ (\lambda^{\circ} x.\ E_{dort_1}\ x)$
 \end{exe}
@@
 these pieces of information percolate up the parse tree all the way to
-the topmost level still pertinent for ensuring grammaticality.
+the topmost level still pertinent for ensuring grammaticality. We will
+now look at another similar example.
 
@@
 $$
-A_{Marie} : NP\_SG
+A_{Marie} : NP\_NUM{=}SG
 $$
 $$
-A_{tatou} : N\_SG
+A_{tatou} : N\_NUM{=}SG
 $$
 $$
-A_{tatous} : N\_PL
+A_{tatous} : N\_NUM{=}PL
 $$
 $$
-A_{le} : N\_SG \limp ((NP\_SG \limp S) \limp S)
+A_{le} : N\_NUM{=}SG \limp ((NP\_NUM{=}SG \limp S) \limp S)
 $$
 $$
-A_{les} : N\_PL \limp ((NP\_PL \limp S) \limp S)
+A_{les} : N\_NUM{=}PL \limp ((NP\_NUM{=}PL \limp S) \limp S)
 $$
 $$
-A_{dort} : NP\_SG \limp S
+A_{dort} : NP\_NUM{=}SG \limp S
 $$
 $$
-A_{dorment} : NP\_PL \limp S
+A_{dorment} : NP\_NUM{=}PL \limp S
 $$
 $$
-A_{aime_1} : NP\_SG \limp NP\_SG \limp S
+A_{aime_1} : NP\_NUM{=}SG \limp NP\_NUM{=}SG \limp S
 $$
 $$
-A_{aime_2} : NP\_SG \limp NP\_PL \limp S
+A_{aime_2} : NP\_NUM{=}SG \limp NP\_NUM{=}PL \limp S
 $$
 $$
-A_{aiment_1} : NP\_PL \limp NP\_SG \limp S
+A_{aiment_1} : NP\_NUM{=}PL \limp NP\_NUM{=}SG \limp S
 $$
 $$
-A_{aiment_2} : NP\_PL \limp NP\_PL \limp S
+A_{aiment_2} : NP\_NUM{=}PL \limp NP\_NUM{=}PL \limp S
 $$
 $$
-A_{qui_1} : (NP\_SG \limp S) \limp N\_SG \limp N\_SG
+A_{qui_1} : (NP\_NUM{=}SG \limp S) \limp N\_NUM{=}SG \limp N\_NUM{=}SG
 $$
 $$
-A_{qui_2} : (NP\_PL \limp S) \limp N\_PL \limp N\_PL
+A_{qui_2} : (NP\_NUM{=}PL \limp S) \limp N\_NUM{=}PL \limp N\_NUM{=}PL
 $$
 $$
-A_{de_1} : NP\_SG \limp N\_SG \limp N\_SG
+A_{de_1} : NP\_NUM{=}SG \limp N\_NUM{=}SG \limp N\_NUM{=}SG
 $$
 $$
-A_{de_2} : NP\_PL \limp N\_SG \limp N\_SG
+A_{de_2} : NP\_NUM{=}PL \limp N\_NUM{=}SG \limp N\_NUM{=}SG
 $$
 $$
-A_{de_3} : NP\_SG \limp N\_PL \limp N\_PL
+A_{de_3} : NP\_NUM{=}SG \limp N\_NUM{=}PL \limp N\_NUM{=}PL
 $$
 $$
-A_{de_4} : NP\_PL \limp N\_Pl \limp N\_PL
+A_{de_4} : NP\_NUM{=}PL \limp N\_Pl \limp N\_NUM{=}PL
 $$
@@
 (\ref{ex:bad-agr}). In the latter sentence, the type of the variable $t$
-has to be $NP\_SG$ since it is used as an argument to $A_{dort}$. This
-means that the relative clause $dort$ has type $NP\_SG \limp
+has to be $NP\_NUM{=}SG$ since it is used as an argument to $A_{dort}$. This
+means that the relative clause $dort$ has type $NP\_NUM{=}SG \limp
 S$. However, there is no type for $qui$ which would allow us to modify
-the plural noun $A_{tatous}$ of type $N\_PL$ by a relative clause with a
-missing singular noun phrase, which has the type $NP\_SG \limp S$.
+the plural noun $A_{tatous}$ of type $N\_NUM{=}PL$ by a relative clause with a
+missing singular noun phrase, which has the type $NP\_NUM{=}SG \limp S$.
 
@@
   \ex * \label{ex:bad-agr} Marie aime les tatous qui dort. \\
-      ? $(A_{les}\ (A_{qui_?}\ (\lambda^{\circ} t.\ A_{dort}\ t)\ A_{tatous}))\ (\lambda^{\circ} y.\ A_{aime_2}\ A_{Marie}\ y)$
+      * $(A_{les}\ (A_{qui_?}\ (\lambda^{\circ} t.\ A_{dort}\ t)\ A_{tatous}))\ (\lambda^{\circ} y.\ A_{aime_2}\ A_{Marie}\ y)$
 \end{exe}

### 2013-06-11 the-last-part.tex
Working on introducing the first subtyping

@@
 \subsection{Intersections of Types}
+
+Our goal is to split the types into smaller parts that only talk about
+features belonging to one of the three mechanisms covered above. This
+means that we would like to take some phrase about which we know a lot
+of disparate pieces of information (presence of negative determiner,
+trace status, number\ldots), having type
+e.g. $NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}$,
+and use it as an argument for some function which only cares about one
+channel of information and would have a type like
+$NP\_\textcolor{red}{NEG{=}T} \limp N\_\textcolor{red}{NEG{=}F} \limp
+N\_\textcolor{red}{NEG{=}T}$. To make a type such as
+$NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}$
+be a legal argument to a function which expects an argument of type
+$NP\_\textcolor{red}{NEG{=}T}$, we will need to introduce some notion of
+\emph{subtyping}.
+
+The types we have been showing up to now propose a very straightforward
+notion of subtyping. All of the types can be seen as conjunctions of
+properties where $\_$ (underscore) is the conjunction operator. A type
+$\tau$ can then be seen as a subtype of type $\sigma$ whenever the
+conjuncts of $\tau$ form a superset of the conjuncts of $\sigma$. This
+is not unlike the structural typing of records in OCaml (CITE SOME OCAML
+SPEC HERE), where a record type is a subtype of another record type if
+the former provides, among others, all the fields provided by the latter
+and gives them the same type. In our case, one type is a subtype of
+another if the former gives, among others, all the properties given by
+the latter. This notion of subtyping can be expressed by adding the rule
+of inference in Figure~\ref{fig:type-superset} to our type
+system.\footnote{In the rule }
+
+\begin{figure}
+  \begin{prooftree}
+    \AxiomC{$\Gamma_l; \Gamma_i \vdash_\Sigma M : A_1 \land \ldots A_{i-1} \land A_i \land A_{i+1} \ldots \land A_n$}
+    \RightLabel{(subtype)}
+    \UnaryInfC{$\Gamma_l; \Gamma_i \vdash_\Sigma M : A_1 \land \ldots A_{i-1} \land A_{i+1} \ldots \land A_n$}
+  \end{prooftree}
+  \caption{\label{fig:type-superset} A rule of inference giving us
+    subtyping for conjunctions of atomic types. For readability, we use
+    $\land$ instead of $\_$ (underscore) to designate the conjunction of
+    properties.}
+\end{figure}
+
+With the new subtyping rule, we can
+
+$$
+S_{tatou} : N \land NEG{=}F \land NUM{=}SG
+$$
+$$
+S_{le_{NEG_1}} : (N \land NEG{=}F) \limp (((NP \land NEG{=}F) \limp S) \limp S)
+$$
+$$
+S_{le_{NEG_2}} : (N \land NEG{=}T) \limp (((NP \land NEG{=}T) \limp S) \limp S)
+$$
+$$
+S_{le_{NUM}} : (N \land NUM{=}SG) \limp (((NP \land NEG{=}SG) \limp S) \limp S)
+$$
+$$
+S_{le_{VAR}} : N \limp (((NP \land VAR{=}F) \limp S) \limp S)
+$$

### 2013-06-12 the-last-part.tex
Added remark about "de" and NP_VAR
Brought back the "de" constraint to the E fragment"

@@
 While we have introduced the notion of an $NP\_VAR{=}T$ for describing
-traces, we can cover another element of our fragment, the preposition
-$de$, and enforce the constraint that $qui$/$que$ cannot extract
-prepositional complements.
+traces, we can cover another element of our running fragment, the
+preposition $de$, and enforce the constraint that $qui$/$que$ cannot
+extract prepositional complements by providing only a single type for
+$de$ that accepts only unmoved noun phrases as complements.
 

### 2013-06-13 the-last-part.tex
Buildup to the definition of GCGs

@@
 of inference in Figure~\ref{fig:type-superset} to our type
-system.\footnote{In the rule }
+system.\footnote{In the rule \ldots (talk about $\land$ instead of $\_$
+  (underscore) here...)}
 
@@
 
+(Talk about whether or not this is a conservative extension...)
+
 With the new subtyping rule, we can
@@
 $$
-S_{le_{NUM}} : (N \land NUM{=}SG) \limp (((NP \land NEG{=}SG) \limp S) \limp S)
+S_{le_{NUM}} : (N \land NUM{=}SG) \limp (((NP \land NUM{=}SG) \limp S) \limp S)
 $$
@@
 $$
+
+\ldots
+
+\subsection{Intersections of Languages}
+
+We will now turn to another solution to the complexity of grammars
+implementing multiple independent mechanisms.
+
+When ACGs are discussed, one of the oft-praised properties is
+compositionality. Let us consider a situation where we have three
+signatures $\Sigma_1$, $\Sigma_2$ and $\Sigma_3$, their respective
+distinguished types $S_1$, $S_2$ and $S_3$ and two lexicons
+$\mathcal{L}_{12} : \Sigma_1 \to \Sigma_2$ and $\mathcal{L}_{23} :
+\Sigma_2 \to \Sigma_3$ such that $\mathcal{L}_{12}(S_1) = S_2$ and
+$\mathcal{L}_{23}(S_2) = S_3$. (TODO: INCLUDE A DIAGRAM OF THIS
+SITUATION)
+
+We can now define languages $A_1$, $A_2$ and $A_3$ where $A_i$ is the
+set of terms from $\Lambda(\Sigma_i)$ that have the type $S_i$. Such
+languages, which are defined only in terms of a signature and some
+distinguished type, are called \emph{abstract languages} in the
+terminology of ACGs. They consist of terms whose structure is
+constrained only by the types assigned by the signature.
+
+Further than that, we can look at the languages $\mathcal{L}_{12}(A_1)$
+and $\mathcal{L}_{23}(A_2)$, images of the abstract languages given by
+the lexicons. Since the lexicons are homomorphic mappings w.r.t. the
+types, all elements of $\mathcal{L}_{12}(A_1)$ will have the type
+$\mathcal{L}_{12}(S_1) = S_2$ and will therefore also belong to $A_2$
+(similarly for $\mathcal{L}_{23}(A_2)$ and $A_3$). These new languages,
+which are formed by mapping an ACG abstract language using a lexicon,
+are called \emph{object languages}. Their terms are constrained not only
+by the types of the signature they are built upon but also by the type
+system of the abstract signature.
+
+Finally, we can also take the composition of lexicons $\mathcal{L}_{13}
+= \mathcal{L}_{23} \circ \mathcal{L}_{12}$, which itself is also a
+lexicon, and map it over $A_1$. This gives us a language
+$\mathcal{L}_{13}(A_1) = \mathcal{L}_{23}(\mathcal{L}_{12}(A_1))$ which
+is at the same time the image of the object language
+$\mathcal{L}_{12}(A_1)$ given by $\mathcal{L}_{23}$ and also the image
+of the abstract language $A_1$ given by $\mathcal{L}_{13}$. This points
+to an interesting property of ACG languages: they are closed under
+transformation by a lexicon.
+
+In the ACG literature, ACGs are defined as quadruples $\mathcal{G} =
+\mathopen{<}\Sigma_A, \Sigma_O, \mathcal{L}, S\mathclose{>}$ where
+$\Sigma_A$ is the abstract signature, $\Sigma_O$ is the object
+signature, $\mathcal{L} : \Sigma_A \to \Sigma_O$ is a lexicon linking
+the two and $S$ is a distinguished type of the abstract signature
+$\Sigma_A$. From this grammar, we can define two languages. The abstract
+language $\mathcal{A}(\mathcal{G}) = \{t \in \Lambda(\Sigma_A) \mid
+\ \vdash_{\Sigma_A} t : S\}$ and the object language
+$\mathcal{O}(\mathcal{G}) = \mathcal{L}(\mathcal{A}(\mathcal{G})) = \{t
+\in \Lambda(\Sigma_O) \mid \exists u \in \mathcal{A}(\mathcal{G}),
+\mathcal{L}(u) = t\}$. The composition of two ACGs $\mathcal{G}_1 =
+\mathopen{<}\Sigma_1, \Sigma_2, \mathcal{L}_{12}, S_1\mathclose{>}$ and
+$\mathcal{G}_2 = \mathopen{<}\Sigma_2, \Sigma_3, \mathcal{L}_{23},
+S_2\mathclose{>}$ is defined as $\mathcal{G}_2 \circ \mathcal{G}_1 =
+\mathopen{<}\Sigma_1, \Sigma_3, \mathcal{L}_{23} \circ \mathcal{L}_{12},
+S_1\mathclose{>}$.
+
+The definition of ACGs given above is quite practical in that it is
+complete, it gives a precise account of what set of languages are
+generated by ACGs (abstract and object languages for which there exist
+the quadruples described above). This contrasts with the presentation of
+ACGs that we gave in~\ref{ssec:acg}, where we defined ACGs as
+collections of signatures with associated distinguished types and
+connected with lexicons. Our definition only introduced the requisite
+machinery, but delayed the delimitation of the way languages can be
+defined. Leaving this question open will let us answer it now by
+providing an extension/generalization of ACGs which will help us
+overcome the modularity issues that have been the theme of this
+section.\footnote{Incidentally, this was not the reason for the way we
+  presented ACGs in~\ref{ssec:acg}. We chose this style as it grouped
+  together the information in a more natural way. We approach signatures
+  as descriptions of domains (strings, syntax, semantics), each one
+  being associated with a distinguished type telling us which objects
+  the domain set out to describe (strings, sentences, truth
+  values). Having distinguished types directly attached to signatures
+  allows us to omit them from the ACG quadruples. However, this only
+  leaves us with triples containing the two signatures and the lexicon
+  connecting them, and since we can infer the signatures themselves from
+  the domain and the range of the lexicon, we do not have much
+  motivation to introduce these tuples. Instead of composing ACGs, we
+  compose just the lexicons.
+
+On top of that, we will want to define signatures and lexicons that
+yield different object languages (at least some string representation of
+the form and some semantic representation of the meaning), which would
+mean having at least two different ACGs which are linked (e.g. by
+sharing the same abstract signature). However, since these two grammars
+are very closely related, we would prefer to reason about our grammar as
+\emph{one} system/graph of signatures and lexicons which defines
+multiple languages instead of talking about \emph{multiple} grammars
+which share some similarities.}
+
+Before we turn to our proposal, we will quickly recall some of the other
+styles of composition used in ACGs (a more detailed exposition can be
+found in \cite{pogodalla2012controlling}). Besides making the object
+signature of one grammar be the abstract signature of another as we saw
+above (TODO: INSERT REFERENCE TO FUTURE DIAGRAM), we can have two
+grammars share the same abstract signature allowing us to transduce
+between terms in the two object signatures. See Figure~(TODO: INSERT A
+DIAGRAM FOR THIS SITUATION) for a diagrammatic example.
+
+These two modes of composition can be fruitfully mixed. Consider (TODO:
+INSERT A DIAGRAM), where we introduce a syntactic signature TODO where
+different terms are only allowed to yield the same string whenever they
+represent different syntactic structures of a syntactically ambiguous
+phrase. This means that purely semantic distinctions such as scope
+ambiguity have to be moved to a more abstract signature from whose terms
+the semantic representation can be derived by a function. This approach
+is studied more deeply in \cite{pogodalla2007generalizing}.
+
+Another interesting composition pattern is to take an existing structure
+of grammars and add a new abstract signature on top. This signature can
+be used to further constrain the items in (TODO: what was before the
+abstract-most signature/language). In \cite{pogodalla2012controlling},
+this is what the authors use to develop their constraints on
+extraction. Each of their constraints is presented as a new abstract
+signature constraining the original syntactic signature.
+
+However, a question that the authors do not pose is how to combine all
+these constraints. If we were to apply the same pattern repeatedly to
+get a structure like the one on (TODO: INSERT A DIAGRAM HERE), we would
+have to deal with successively more complex type systems on each
+level. This is due to the condition that a lexicon has to be a
+homomorphism and thus it must, among other things, always map well-typed
+terms to well-typed terms. To illustrate the problem, imagine a
+structure such as the one on (TODO) that handles some different
+constraints in each of the vertically stacked signatures including
+e.g. agreement. If we tried to add a new abstract signature at the top
+to implement another constraint, we would have to ensure that all the
+terms we licence in our new abstract signature have a well-typed
+counterpart in the existing abstract language. However, this means that
+our new signature must make sure that the terms it approves do not break
+agreement or any of the previously established constraints. Effectively,
+our new signature would be forced to reimplement all of the existing
+constraints, rendering vain all our effort to split up the constraints
+into multiple signatures.
+
+What we would like instead is to put the signatures side-by-side such
+that they do not depend on each other (as on diagram TODO). However, how
+does this actually define a language? We can arrive at the answer by
+examining how are the graphical representations of ACGs connected to the
+languages they define. We will associate a language to every node of the
+graph. Before we begin, we notice that every ACG diagram formed by the
+above rules of composition is a directed tree and therefore every node
+has at most one parent/predecessor. To the root of the tree, the
+abstract-most signature in the graph, we will assign the abstract
+language generated by that signature and its distinguished type. To all
+the other nodes, we will assign the result of mapping their parent's
+language by the lexicon linking the two. By homomorphism of the
+lexicons, we can also deduce the distinguished types for all the
+non-root nodes. They are the images of their parents' distinguished
+types by the connecting lexicons. For all of the nodes, it is true that
+every term in the language assigned to the node has the node's
+distinguished type.
+
+We can confirm that the above graphical model of defining ACGs is
+coherent. First, we can show that our interpretation is sound and so
+that any directed tree adorned with signatures on the nodes, compatible
+lexicons on the edges and a distinguished type on the root can only
+define ACG languages. This is trivial, since the language corresponding
+to the root is an ACG abstract language and the class of ACG languages
+is closed under transformation by a lexicon and therefore all of the
+languages corresponding to the descendants of the root are ACG languages
+as well. Second, we can try and show some kind of completeness. However,
+there is not much formalization done to describe a ``system of ACGs that
+share common signatures'' (something that our generalization will
+fix). We can nevertheless show that it is possible to express any ACG
+using our graphical model. Given a grammar $\mathcal{G} =
+\mathopen{<}\Sigma_A, \Sigma_O, \mathcal{L}, S\mathclose{>}$, we just
+produce a tree whose root has the signature $\Sigma_A$ and distinguished
+type $S$ and which is connected to a single child node having the
+signature $\Sigma_O$ by the lexicon $\mathcal{L}$. Then the two nodes of
+our tree define both the abstract and the object language defined by
+$\mathcal{G}$.
+
+We have now given some formal account about how we can read language
+definitions in diagrams. Still, we have kept our analysis to the tree
+diagrams used in current ACG literature, where every node has at most
+one parent. To make the generalization more visible, we will switch from
+an algebraic manner of presentation, in which we stated that the
+language of a node is equal to an image of its parent's language, to a
+calculus-based manner of presentation, in which we will state that the
+language of a node is the set of all terms that have an antecedent in
+the parent's language. Besides this property, there is also the fact
+that every term in a node's language has the node's distinguished
+type. This property is trivially true for nodes with parents because of
+the homomorphic property of lexicons. However, it is crucial for the
+root node for which this is the single defining property of its
+language. To sum up, we have that for every node of a tree diagram, its
+language is the set of all terms that have the node's distinguished type
+and that have an antecedent in the parent's language, if there is
+any. This rule is already general enough to handle the case of zero or
+one parents, we can easily generalize it to handle an arbitrary count:
+\emph{the language of a node is the set of all terms that have the
+  node's distinguished type and that have antecedents in the languages
+  of all of the node's parents}. If we transpose this back to the
+algebraic style of presentation, we say that the language of a node is
+the intersection of the set of terms having the node's distinguished
+type and the intersection of all its parents' languages.
+
+TODO: In the previous paragraph, add a notice saying that the terms must
+be also built on the signature of the node.
+
+We will now turn to a formal definition of this generalized graph-based
+formalism and discuss some its values and properties.

### 2013-06-13 the-last-part.tex
Arborescence!
Writing about GACGs...

@@
   being associated with a distinguished type telling us which objects
-  the domain set out to describe (strings, sentences, truth
+  the domain has set out to describe (strings, sentences, truth
   values). Having distinguished types directly attached to signatures
@@
 multiple languages instead of talking about \emph{multiple} grammars
-which share some similarities.}
+which share some similarities.
+
+TODO: Somehow express that its nice to have the links between the
+signatures and lexicons that are shown in diagrams \emph{explicit}
+rather than having them \emph{implicit}.}
 
@@
 structure such as the one on (TODO) that handles some different
-constraints in each of the vertically stacked signatures including
+constraint in each of the vertically stacked signatures, including
 e.g. agreement. If we tried to add a new abstract signature at the top
@@
 our new signature would be forced to reimplement all of the existing
-constraints, rendering vain all our effort to split up the constraints
-into multiple signatures.
+constraints, rendering in vain all our effort to split up the
+constraints into multiple signatures.
 
@@
 type $S$ and which is connected to a single child node having the
-signature $\Sigma_O$ by the lexicon $\mathcal{L}$. Then the two nodes of
-our tree define both the abstract and the object language defined by
-$\mathcal{G}$.
+signature $\Sigma_O$ by an edge labelled by the lexicon
+$\mathcal{L}$. Then the two nodes of our tree define both the abstract
+and the object language defined by $\mathcal{G}$.
 
-We have now given some formal account about how we can read language
+We have now given some formal account of how we can read language
 definitions in diagrams. Still, we have kept our analysis to the tree
@@
 
+TODO: Use the correct term arborescence in the above paragraphs!
+% http://en.wikipedia.org/wiki/Arborescence_(graph_theory)
+
 We will now turn to a formal definition of this generalized graph-based
 formalism and discuss some its values and properties.
+
+\subsubsection{Graphical Abstract Categorial Grammars}
+
+We define a \emph{graphical abstract categorial grammar} as a quadruple
+$\mathcal{G} = \mathopen{<} G, \Sigma, S, {\mathcal{L}} \mathclose{>}$
+where $G$ is a directed graph with vertices $V(G)$ and edges $E(G)$,
+$\Sigma$ and $S$ are labelings assigning signatures and distinguished
+types, respectively, to the vertices in $V$ and ${\mathcal{L}}$ is a
+labeling assigning lexicons to the edges in $E$. Furthermore, a
+well-formed graphical ACG satisfies the following conditions:
+
+\begin{itemize}
+\item $G$ is a directed acyclic graph.
+\item For all $(u,v) \in E(G)$, $\mathcal{L}_{(u,v)} : \Sigma_u \to
+  \Sigma_v$.
+\item For all $(u,v) \in E(G)$, $\mathcal{L}_{(u,v)}(S_u) = S_v$.
+\end{itemize}
+
+We will say that a node $u$ is \emph{more abstract} than a node $v$
+whenever $(u,v)$ belongs to the strict partial order induced by the
+edges of $G$ (i.e. there is a path from $u$ to $v$).
+
+We let the nodes in a graphical ACG $\mathcal{G} = \mathopen{<} G,
+\Sigma, S, \mathcal{L} \mathclose{>}$ define two kinds of
+languages. The \emph{intrinsic languages} $\mathcal{I}_{\mathcal{G}}$,
+which are constrained only by the type signature described in the node
+itself, and the \emph{extrinsic languages} $\mathcal{E}_{\mathcal{G}}$,
+which are also constrained by type systems in more abstract nodes of the
+graph. These two notions correspond to those of abstract languages and
+object languages in ACGs.
+
+$$
+\mathcal{I}_{\mathcal{G}}(v) = \{t \in \Lambda(\Sigma_v)
+\mid\ \vdash_{\Sigma_u} t : S_v\}
+$$
+$$
+\mathcal{E}_{\mathcal{G}}(v) = \mathcal{I}_{\mathcal{G}}(v) \cap
+\bigcap_{(u,v) \in E} \mathcal{L}_{(u,v)}(\mathcal{E}_{\mathcal{G}}(u)).
+$$
+
+From the above definition, we see that the ACG diagrams we introduced
+before are a special case of graphical ACGs where the graph is an
+arborescence.
+
+\subsubsubsection{Formal Properties of Graphical ACGs}
+
+Observation: The set of intrinsic languages definable by graphical ACGs
+is the same as the set of abstract languages definable by ACGs.
+
+Theorem: The set of extrinsic languages definable by graphical ACGs is
+the set of object languages definable by ACGs closed under intersection
+and transformation by a lexicon.
+
+Proof:
+
+First, we prove that an extrinsic language can be always constructed
+from object languages by intersections and transformations by
+lexicons. Let us consider any graphical ACG $\mathcal{G} = \mathopen{<}
+G, \Sigma, S, \mathcal{L} \mathclose{>}$ and any topological ordering
+$v_1, \ldots, v_n$ of the nodes of $G$ (an ordering such that more
+abstract nodes always precede less abstract ones). Before we start, we
+will just note that the all the intrinsic languages defined by $v_i$ are
+abstract languages which are a special case of object languages.
+
+\begin{itemize}
+\item For $v_1$, the case is trivial since its extrinsic language is
+  just its intrinsic language as the node $v_1$ has no predecessors in
+  $G$.
+\item For any other node $v_n$, we look at the definition of an
+  extrinsic language and remark that the only operators are intersection
+  and application of a lexicon with the operands being the intrinsic
+  language of $v_n$ and the extrinsic languages of more abstract
+  nodes. By induction hypothesis, we have that all of the operands have
+  been constructed from object languages using intersections and
+  transformations by lexicons and therefore even the extrinsic language
+  defined by $v_n$ has this property.
+\end{itemize}
+
+Now we have to prove the converse. We do so by showing that extrinsic
+languages contain object languages and are closed under intersection and
+transformation by a lexicon.
+
+\begin{itemize}
+\item Let $L$ be an object language defined by the ACG $\mathopen{<}
+  \Sigma_A, \Sigma_O, \mathcal{L}, S\mathclose{>}$. We take the graph
+  $G$ with $V(G) = \{A, O\}$ and $E(G) = \{(A,O)\}$, we label the nodes
+  $A$ and $O$ with $\Sigma_A$ and $\Sigma_O$ as signatures and $S$ and
+  $\mathcal{L}(S)$ as distinguished types respectively. Finally, we
+  label the edge $(A,O)$ with $\mathcal{L}$ and we have a graphical ACG
+  in which the node $O$ defines the extrinsic language $L$.
+
+\item Let $L_1$ and $L_2$ be two extrinsic languages defined by the
+  nodes $u_1$ and $u_2$ respectively belonging to two graphical ACGs. We
+  can construct a new graphical ACG by taking the union of the two ACGs
+  (union of graphs (union of vertices and edges) and unions of the
+  labelings).\footnote{Assuming, without loss of generality, that the
+    sets of vertices of the two graphs are disjoint.} We then add a new
+  node $v$ which will define the language $L_1 \cap L_2$.
+
+  The signature labeling $v$ will be the union of the two signatures
+  labeling $u_1$ and $u_2$ such that the intrinsic language of $v$ will
+  be a superset of the intrinsic languages, and by extension the
+  extrinsic languages, of both $u_1$ and $u_2$. Either of the two
+  distinguished types that label $u_1$ or $u_2$ can be used to label $v$
+  (if they are actually different, then the intersection is trivially
+  empty). Finally, we will add two edges leading from $u_1$ and $u_2$ to
+  $v$, both labeled with the identity lexicon. We will now show that
+  the extrinsic language of $v$ in the newly formed graphical ACG
+  $\mathcal{G}'$ is $L_1 \cap L_2$.
+
+  \begin{align*}
+    \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v) \cap
+    \mathcal{L}_{(u_1,v)}(\mathcal{E}_{\mathcal{G}'}(u_1)) \cap
+    \mathcal{L}_{(u_2,v)}(\mathcal{E}_{\mathcal{G}'}(u_2)) \\ &=
+    \mathcal{I}_{\mathcal{G}'}(v) \cap \mathcal{E}_{\mathcal{G}'}(u_1) \cap
+    \mathcal{E}_{\mathcal{G}'}(u_2) \\ &= \mathcal{E}_{\mathcal{G}'}(u_1)
+    \cap \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
+    \mathcal{E}_{\mathcal{G}}(u_1) \cap \mathcal{E}_{\mathcal{G}}(u_2) \\ &=
+    L_1 \cap L_2
+  \end{align*}
+
+  The simplifications use the facts that: the two new lexicons are
+  identities, that the intrinsic language of $v$ is a superset of the
+  extrinsic languages of $u_1$ and $u_2$, that the extrinsic languages
+  of $u_1$ and $u_2$ are the same in the new graphical ACG as in the two
+  original graphical ACGs and finally that $L_1$ and $L_2$ are the
+  extrinsic languages defined by $u_1$ and $u_2$.
+
+\item Let $L \subset \Lambda(\Sigma_A)$ be an extrinsic language and
+  $\mathcal{L} : \Sigma_A \to \Sigma_O$ a lexicon from its signature
+  $\Sigma_A$ to the signature $\Sigma_O$. We will show that the language
+  $\mathcal{L}(L) \subset \Lambda(\Sigma_O)$ is also an extrinsic
+  language. Let $u$ be the node that defines the extrinsic language $L$
+  in some graphical ACG $\mathcal{G}$. We construct a new graphical ACG
+  by adding a new vertex $v$ to $\mathcal{G}$ labeled with the signature
+  $\Sigma_O$ and the distinguished type $\mathcal{L}(S)$ where $S$ is
+  the distinguished type labeling $u$. To this graph, we conjoin the
+  edge $(u,v)$ and label it with $\mathcal{L}$. Now we verify that the
+  extrinsic language of $v$ is the language $\mathcal{L}(L).$
+
+  \begin{align*}
+    \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v) \cap
+    \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
+    &= \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
+    &= \mathcal{L}(\mathcal{E}_{\mathcal{G}}(u)) \\
+    &= \mathcal{L}(L)
+  \end{align*}
+
+\end{itemize}
+
+Corollary: The set of extrinsic languages definable by graphical ACGs is
+the same as the set of object languages definable by ACGs if and only if
+the set of object languages definable by ACGs is closed under
+intersection.

### 2013-06-13 acg.tex
Almost finished writing about defining signatures
Finished writing 'Defining signatures'
Prettier implementation of lexicons
Switched to the report class, start using theorem envs

@@
-\subsection{Abstract Categorial Grammars}
+\section{Abstract Categorial Grammars}
+\label{sec:acg}
 
@@
 
-\subsubsection{Higher-order signatures}
+\subsection{Higher-order signatures}
+\label{ssec:sig}
 
@@
 
-\subsubsection{Example signature}
-\label{sssec:example-sig}
+\subsection{Example signature}
+\label{ssec:example-sig}
 
@@
 
-\subsubsection{Lexicons}
+\subsection{Lexicons}
 
@@
 
-\subsubsection{Example lexicon}
-\label{sssec:example-lex}
+\subsection{Example lexicon}
+\label{ssec:example-lex}
 
@@
 To illustrate the ideas of a lexicon and an abstract categorial grammar,
-we will expand our example from \ref{sssec:example-sig}. We will
+we will expand our example from \ref{ssec:example-sig}. We will
 consider another signature, $\Sigma_{Synt}$, which will describe the
@@
 
-\subsubsection{Example semantic lexicon}
+\subsection{Example semantic lexicon}
+\label{ssec:acg-sem}
 
@@
 abstract term can be interpreted multiple ways, using different
-lexicons. In the previous section, we considered the abstract terms of
+lexicons. In the previous chapter, we considered the abstract terms of
 the $\Sigma_{Synt}$ signature and interpreted them as computations on
@@
   x.\ ((P\ x) \land (Q\ x))$
-\item $\mathcal{L}_{sem}(\synt{love}) = \textbf{love}$
-\item $\mathcal{L}_{sem}(\synt{man}) = \textbf{man}$
-\item $\mathcal{L}_{sem}(\synt{woman}) = \textbf{woman}$
+\item $\mathcal{L}_{sem}(\synt{love}) = \lambda^{\circ} s o. (\textbf{love}\ s\ o)$
+\item $\mathcal{L}_{sem}(\synt{man}) = \lambda^{\circ} x. (\textbf{man}\ x)$
+\item $\mathcal{L}_{sem}(\synt{woman}) = \lambda^{\circ} x. (\textbf{woman}\ x)$
 \end{itemize}
@@
 Let us now consider the example terms of the $\Sigma_{Synt}$ signature
-from \ref{sssec:example-lex}. $t_3 = \synt{some}\ \synt{woman}$ will be
+from \ref{ssec:example-lex}. $t_3 = \synt{some}\ \synt{woman}$ will be
 interpreted as $\mathcal{L}_{sem}(t_3) =_{\beta} \lambda^{\circ}

### 2013-06-13 ig.tex
Talked about the 3 mechs and their naive combo
Switched to the report class, start using theorem envs

@@
-\subsection{Interaction Grammars}
+\section{Interaction Grammars}
 
@@
 
-\subsubsection{Frigram, an interaction grammar}
+\subsection{Frigram, an interaction grammar}
+\label{ssec:frigram}
 
@@
 building our grammar on as well. Furthermore, as we will see
-in~\ref{sssec:link-ig-acg}, interaction grammars are closely linked with
+in~\ref{ssec:link-ig-acg}, interaction grammars are closely linked with
 abstract categorial grammars, which makes Frigram a suitable grammar to
@@
 
-In this section, we will introduce the metagrammatical structure of
+In this chapter, we will introduce the metagrammatical structure of
 Frigram and talk about how it uses the formalism of interaction grammars
@@
 
-\subsubsection{The link with abstract categorial grammars}
-\label{sssec:link-ig-acg}
+\subsection{The link with abstract categorial grammars}
+\label{ssec:link-ig-acg}
 

### 2013-06-13 my-work-rewritten.tex
Switched to the report class, start using theorem envs

@@
-\section{Wide-coverage abstract categorial grammars}
+\chapter{Wide-coverage abstract categorial grammars}
 
@@
 
-\subsection{The tools and techniques}
+\section{The tools and techniques}
 
@@
 
-\subsubsection{The case for relational programming}
+\subsection{The case for relational programming}
 
@@
 wordform can be mapped to multiple hypertags. We will see another
-example of a relation within our domain in the next subsection and we
+example of a relation within our domain in the next section and we
 will also discuss the advantages and drawbacks of modelling functions as
@@
 
-\subsubsection{Choice of implementation language}
+\subsection{Choice of implementation language}
 
@@
 
-\subsection{Defining signatures and lexicons}
+\section{Defining signatures and lexicons}
 
-\subsubsection{Defining signatures}
+\subsection{Defining signatures}
 
-In~\ref{sssec:sig}, we formally presented signatures as triples
+In~\ref{ssec:sig}, we formally presented signatures as triples
 $\mathopen{<}A, C, \tau\mathclose{>}$. The set $C$ can be easily
@@
     signature for every item of our lexical database. If you recall
-    in~\ref{sssec:acg-sem}, the constants of the semantic signature
+    in~\ref{ssec:acg-sem}, the constants of the semantic signature
     $\Sigma_{Sem}$ contained predicates for the common nouns $man$ and
@@
 lexicalized version of the $\Sigma_{Synt}$ signature
-from~\ref{sssec:example-sig} using the \texttt{ht->typer} abstraction.
+from~\ref{ssec:example-sig} using the \texttt{ht->typer} abstraction.
 
@@
   \caption{\label{fig:lex-sig-impl} A lexicalized version of the
-    $\Sigma$(\_{Synt}) signature from~\ref{sssec:example-sig}.}
+    $\Sigma$(\_{Synt}) signature from~\ref{ssec:example-sig}.}
 \end{figure}
@@
 string concatenation operator $+$ of the $\Sigma_{String}$ signature we
-gave in~\ref{sssec:example-sig} or the quantifiers and logical
-connectives of the $\Sigma_{Sem}$ signature in~\ref{sssec:acg-sem}.
+gave in~\ref{ssec:example-sig} or the quantifiers and logical
+connectives of the $\Sigma_{Sem}$ signature in~\ref{ssec:acg-sem}.
 These constants are not associated to any lexical entry or specifier and
@@
 conjunctions, respectively. With this, we have enough to fully define
-the other signatures of subsection~\ref{ssec:acg} in a lexicalized
+the other signatures of section~\ref{sec:acg} in a lexicalized
 manner, see Figure~\ref{fig:example-sig-impl}.
@@
     constants belonging to $\Sigma$(\_{Sem})
-    from~\ref{sssec:acg-sem}. The single arrow \texttt{->} corresponds
+    from~\ref{ssec:acg-sem}. The single arrow \texttt{->} corresponds
     to linear implication while the double arrow \texttt{=>} corresponds
@@
     lexicalized versions of the example signatures of
-    subsection~\ref{ssec:acg}.}
+    section~\ref{sec:acg}.}
 \end{figure}
 
-\subsubsection{Defining lexicons}
+\subsection{Defining lexicons}
 
@@
 binary relation, which is the representation that we have chosen. In the
-previous subsubsection, we talked a lot about how individual entries in
+previous subsection, we talked a lot about how individual entries in
 the lexical database generate the items of a signature. Solving the
@@
     $\mathcal{L}$(\_{Syntax}) and $\mathcal{L}$(\_{Sem}) lexicons from
-    subsection~\ref{ssec:acg}.}
+    section~\ref{sec:acg}.}
 \end{figure}
@@
 
-\subsection{Testing signatures and lexicons}
+\section{Testing signatures and lexicons}

### 2013-06-13 the-last-part.tex
Switched to the report class, start using theorem envs

@@
-\section{The Last Part}
+\chapter{The Last Part}
 
-In this section, we will attempt to incorporate accounts of several
+In this chapter, we will attempt to incorporate accounts of several
 linguistic phenomena in a single framework, highlight the modularity
@@
 
-\subsection{Negation}
+\section{Negation}
 
@@
 for negation in French. Recall the EPTDs we have seen
-in~\ref{sssec:frigram}, repeated here on Figure~\ref{fig:ig-neg-rep}.
+in~\ref{ssec:frigram}, repeated here on Figure~\ref{fig:ig-neg-rep}.
 
@@
 If we suppose the presence of the appropriate lexical items typed as in
-the examples of~\ref{ssec:acg}, we have a type system in which we can
+the examples of~\ref{sec:acg}, we have a type system in which we can
 type the syntactic terms corresponding to the sentences
@@
 
-\subsection{Extraction}
-\label{ssec:extraction}
+\section{Extraction}
+\label{sec:extraction}
 
@@
 
-\subsection{Agreement}
+\section{Agreement}
 
@@
 
-In the next subsection, we will see what it takes to enforce all of the
+In the next section, we will see what it takes to enforce all of the
 constraints above (negation, extraction and agreement) at the same time,
@@
 
-\subsection{Putting It All Together}
+\section{Putting It All Together}
 
@@
   using dependent types to handle several constraints regarding
-  extraction (including the one we covered in~\ref{ssec:extraction}).}
+  extraction (including the one we covered in~\ref{sec:extraction}).}
 
@@
 about than the enumeration of 12 basic types we presented at the
-beginning of this subsection, however, the result is still complex in
+beginning of this section, however, the result is still complex in
 that it forces us to describe the behavior of three independent
@@
 
-\subsection{Intersections of Types}
+\section{Intersections of Types}
 
@@
 
-\subsection{Intersections of Languages}
+\section{Intersections of Languages}
 
@@
 the quadruples described above). This contrasts with the presentation of
-ACGs that we gave in~\ref{ssec:acg}, where we defined ACGs as
+ACGs that we gave in~\ref{sec:acg}, where we defined ACGs as
 collections of signatures with associated distinguished types and
@@
 overcome the modularity issues that have been the theme of this
-section.\footnote{Incidentally, this was not the reason for the way we
-  presented ACGs in~\ref{ssec:acg}. We chose this style as it grouped
+chapter.\footnote{Incidentally, this was not the reason for the way we
+  presented ACGs in~\ref{sec:acg}. We chose this style as it grouped
   together the information in a more natural way. We approach signatures
@@
 
-\subsubsection{Graphical Abstract Categorial Grammars}
+\subsection{Graphical Abstract Categorial Grammars}
 
@@
 
-\subsubsubsection{Formal Properties of Graphical ACGs}
+\subsubsection{On the Expressivity of Graphical ACGs}
 
-Observation: The set of intrinsic languages definable by graphical ACGs
-is the same as the set of abstract languages definable by ACGs.
+\begin{observation}
+  The set of intrinsic languages definable by graphical ACGs is the same
+  as the set of abstract languages definable by ACGs.
+\end{observation}
 
-Theorem: The set of extrinsic languages definable by graphical ACGs is
-the set of object languages definable by ACGs closed under intersection
-and transformation by a lexicon.
+\begin{theorem}
+  The set of extrinsic languages definable by graphical ACGs is the set
+  of object languages definable by ACGs closed under intersection and
+  transformation by a lexicon.
+\end{theorem}
 
-Proof:
-
+\begin{proof}
   First, we prove that an extrinsic language can be always constructed
   from object languages by intersections and transformations by
-lexicons. Let us consider any graphical ACG $\mathcal{G} = \mathopen{<}
-G, \Sigma, S, \mathcal{L} \mathclose{>}$ and any topological ordering
-$v_1, \ldots, v_n$ of the nodes of $G$ (an ordering such that more
-abstract nodes always precede less abstract ones). Before we start, we
-will just note that the all the intrinsic languages defined by $v_i$ are
-abstract languages which are a special case of object languages.
+  lexicons. Let us consider any graphical ACG $\mathcal{G} =
+  \mathopen{<} G, \Sigma, S, \mathcal{L} \mathclose{>}$ and any
+  topological ordering $v_1, \ldots, v_n$ of the nodes of $G$ (an
+  ordering such that more abstract nodes always precede less abstract
+  ones). Before we start, we will just note that the all the intrinsic
+  languages defined by $v_i$ are abstract languages which are a special
+  case of object languages.
 
@@
     \item For $v_1$, the case is trivial since its extrinsic language is
-  just its intrinsic language as the node $v_1$ has no predecessors in
-  $G$.
+      just its intrinsic language as the node $v_1$ has no predecessors
+      in $G$.
     \item For any other node $v_n$, we look at the definition of an
-  extrinsic language and remark that the only operators are intersection
-  and application of a lexicon with the operands being the intrinsic
-  language of $v_n$ and the extrinsic languages of more abstract
-  nodes. By induction hypothesis, we have that all of the operands have
-  been constructed from object languages using intersections and
-  transformations by lexicons and therefore even the extrinsic language
-  defined by $v_n$ has this property.
+      extrinsic language and remark that the only operators are
+      intersection and application of a lexicon with the operands being
+      the intrinsic language of $v_n$ and the extrinsic languages of
+      more abstract nodes. By induction hypothesis, we have that all of
+      the operands have been constructed from object languages using
+      intersections and transformations by lexicons and therefore even
+      the extrinsic language defined by $v_n$ has this property.
   \end{itemize}
@@
   Now we have to prove the converse. We do so by showing that extrinsic
-languages contain object languages and are closed under intersection and
-transformation by a lexicon.
+  languages contain object languages and are closed under intersection
+  and transformation by a lexicon.
 
@@
     \item Let $L$ be an object language defined by the ACG $\mathopen{<}
-  \Sigma_A, \Sigma_O, \mathcal{L}, S\mathclose{>}$. We take the graph
-  $G$ with $V(G) = \{A, O\}$ and $E(G) = \{(A,O)\}$, we label the nodes
-  $A$ and $O$ with $\Sigma_A$ and $\Sigma_O$ as signatures and $S$ and
-  $\mathcal{L}(S)$ as distinguished types respectively. Finally, we
-  label the edge $(A,O)$ with $\mathcal{L}$ and we have a graphical ACG
-  in which the node $O$ defines the extrinsic language $L$.
+      \Sigma_A, \Sigma_O, \mathcal{L}, S\mathclose{>}$. We take the
+      graph $G$ with $V(G) = \{A, O\}$ and $E(G) = \{(A,O)\}$, we label
+      the nodes $A$ and $O$ with $\Sigma_A$ and $\Sigma_O$ as signatures
+      and $S$ and $\mathcal{L}(S)$ as distinguished types
+      respectively. Finally, we label the edge $(A,O)$ with
+      $\mathcal{L}$ and we have a graphical ACG in which the node $O$
+      defines the extrinsic language $L$.
 
     \item Let $L_1$ and $L_2$ be two extrinsic languages defined by the
-  nodes $u_1$ and $u_2$ respectively belonging to two graphical ACGs. We
-  can construct a new graphical ACG by taking the union of the two ACGs
-  (union of graphs (union of vertices and edges) and unions of the
-  labelings).\footnote{Assuming, without loss of generality, that the
-    sets of vertices of the two graphs are disjoint.} We then add a new
-  node $v$ which will define the language $L_1 \cap L_2$.
+      nodes $u_1$ and $u_2$ respectively belonging to two graphical
+      ACGs. We can construct a new graphical ACG by taking the union of
+      the two ACGs (union of graphs (union of vertices and edges) and
+      unions of the labelings).\footnote{Assuming, without loss of
+        generality, that the sets of vertices of the two graphs are
+        disjoint.} We then add a new node $v$ which will define the
+      language $L_1 \cap L_2$.
 
       The signature labeling $v$ will be the union of the two signatures
-  labeling $u_1$ and $u_2$ such that the intrinsic language of $v$ will
-  be a superset of the intrinsic languages, and by extension the
-  extrinsic languages, of both $u_1$ and $u_2$. Either of the two
-  distinguished types that label $u_1$ or $u_2$ can be used to label $v$
-  (if they are actually different, then the intersection is trivially
-  empty). Finally, we will add two edges leading from $u_1$ and $u_2$ to
-  $v$, both labeled with the identity lexicon. We will now show that
-  the extrinsic language of $v$ in the newly formed graphical ACG
-  $\mathcal{G}'$ is $L_1 \cap L_2$.
+      labeling $u_1$ and $u_2$ such that the intrinsic language of $v$
+      will be a superset of the intrinsic languages, and by extension
+      the extrinsic languages, of both $u_1$ and $u_2$. Either of the
+      two distinguished types that label $u_1$ or $u_2$ can be used to
+      label $v$ (if they are actually different, then the intersection
+      is trivially empty). Finally, we will add two edges leading from
+      $u_1$ and $u_2$ to $v$, both labeled with the identity lexicon. We
+      will now show that the extrinsic language of $v$ in the newly
+      formed graphical ACG $\mathcal{G}'$ is $L_1 \cap L_2$.
 
       \begin{align*}
-    \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v) \cap
-    \mathcal{L}_{(u_1,v)}(\mathcal{E}_{\mathcal{G}'}(u_1)) \cap
+        \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v)
+        \cap \mathcal{L}_{(u_1,v)}(\mathcal{E}_{\mathcal{G}'}(u_1)) \cap
         \mathcal{L}_{(u_2,v)}(\mathcal{E}_{\mathcal{G}'}(u_2)) \\ &=
-    \mathcal{I}_{\mathcal{G}'}(v) \cap \mathcal{E}_{\mathcal{G}'}(u_1) \cap
-    \mathcal{E}_{\mathcal{G}'}(u_2) \\ &= \mathcal{E}_{\mathcal{G}'}(u_1)
-    \cap \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
-    \mathcal{E}_{\mathcal{G}}(u_1) \cap \mathcal{E}_{\mathcal{G}}(u_2) \\ &=
-    L_1 \cap L_2
+        \mathcal{I}_{\mathcal{G}'}(v) \cap
+        \mathcal{E}_{\mathcal{G}'}(u_1) \cap
+        \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
+        \mathcal{E}_{\mathcal{G}'}(u_1) \cap
+        \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
+        \mathcal{E}_{\mathcal{G}}(u_1) \cap
+        \mathcal{E}_{\mathcal{G}}(u_2) \\ &= L_1 \cap L_2
       \end{align*}
@@
       The simplifications use the facts that: the two new lexicons are
-  identities, that the intrinsic language of $v$ is a superset of the
-  extrinsic languages of $u_1$ and $u_2$, that the extrinsic languages
-  of $u_1$ and $u_2$ are the same in the new graphical ACG as in the two
-  original graphical ACGs and finally that $L_1$ and $L_2$ are the
-  extrinsic languages defined by $u_1$ and $u_2$.
+      identities, that the intrinsic language of $v$ is a superset of
+      the extrinsic languages of $u_1$ and $u_2$, that the extrinsic
+      languages of $u_1$ and $u_2$ are the same in the new graphical ACG
+      as in the two original graphical ACGs and finally that $L_1$ and
+      $L_2$ are the extrinsic languages defined by $u_1$ and $u_2$.
 
@@
     $\mathcal{L} : \Sigma_A \to \Sigma_O$ a lexicon from its signature
-  $\Sigma_A$ to the signature $\Sigma_O$. We will show that the language
-  $\mathcal{L}(L) \subset \Lambda(\Sigma_O)$ is also an extrinsic
-  language. Let $u$ be the node that defines the extrinsic language $L$
-  in some graphical ACG $\mathcal{G}$. We construct a new graphical ACG
-  by adding a new vertex $v$ to $\mathcal{G}$ labeled with the signature
-  $\Sigma_O$ and the distinguished type $\mathcal{L}(S)$ where $S$ is
-  the distinguished type labeling $u$. To this graph, we conjoin the
-  edge $(u,v)$ and label it with $\mathcal{L}$. Now we verify that the
-  extrinsic language of $v$ is the language $\mathcal{L}(L).$
+    $\Sigma_A$ to the signature $\Sigma_O$. We will show that the
+    language $\mathcal{L}(L) \subset \Lambda(\Sigma_O)$ is also an
+    extrinsic language. Let $u$ be the node that defines the extrinsic
+    language $L$ in some graphical ACG $\mathcal{G}$. We construct a new
+    graphical ACG by adding a new vertex $v$ to $\mathcal{G}$ labeled
+    with the signature $\Sigma_O$ and the distinguished type
+    $\mathcal{L}(S)$ where $S$ is the distinguished type labeling
+    $u$. To this graph, we conjoin the edge $(u,v)$ and label it with
+    $\mathcal{L}$. Now we verify that the extrinsic language of $v$ is
+    the language $\mathcal{L}(L).$
 
@@
     \end{align*}
+  \end{itemize}
+\end{proof}
 
-\end{itemize}
-
-Corollary: The set of extrinsic languages definable by graphical ACGs is
-the same as the set of object languages definable by ACGs if and only if
-the set of object languages definable by ACGs is closed under
+\begin{corollary}
+  The set of extrinsic languages definable by graphical ACGs is the same
+  as the set of object languages definable by ACGs if and only if the
+  set of object languages definable by ACGs is closed under
   intersection.
+\end{corollary}

### 2013-06-13 the-last-part.tex
More formal talking about GACGs

@@
 \section{Putting It All Together}
+\label{sec:together}
 
@@
     \item Let $L_1$ and $L_2$ be two extrinsic languages defined by the
-      nodes $u_1$ and $u_2$ respectively belonging to two graphical
-      ACGs. We can construct a new graphical ACG by taking the union of
-      the two ACGs (union of graphs (union of vertices and edges) and
-      unions of the labelings).\footnote{Assuming, without loss of
-        generality, that the sets of vertices of the two graphs are
-        disjoint.} We then add a new node $v$ which will define the
+      nodes $u_1$ and $u_2$, respectively, and belonging to some two
+      graphical ACGs. We can construct a new graphical ACG by taking the
+      union of the two ACGs (union of graphs (union of vertices and
+      edges) and unions of the labelings).\footnote{Assuming, without
+        loss of generality, that the sets of vertices of the two graphs
+        are disjoint.} We then add a new node $v$ which will define the
       language $L_1 \cap L_2$.
@@
     \end{align*}
+
+    The facts used here are in turn: the inclusion of
+    $\mathcal{I}_{\mathcal{G}'}(v)$ in
+    $\mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u))$, the observation that
+    modifying our graphical grammar did not change the extrinsic
+    language of $u$ and the assumption that $L$ is the extrinsic
+    language of $u$.
   \end{itemize}
@@
 \end{corollary}
+
+We conclude this subsubsection with a few words to motivate the need we
+feel for this (possibly) more expressive formalism. We believe that in
+order to manage the development of wide-coverage grammars, the grammars
+need to be kept as simple as possible, which means, among other things,
+that independent constraints should be implemented separately and not
+complected together (we saw the difficulties caused by this already on a
+microscopic fragment in section~\ref{sec:together}). This means that if
+we were to define, e.g., a partial grammar in which we would only
+enforce agreement and another one in which we would only enforce the
+proper use of relative pronouns, we would like the language in which
+both agreement and proper use of relative pronouns are respected to be
+expressible in our formalism and on top of that, we would like this
+language to be easily expressed in terms of the two grammars for the
+individual constraints. We now have a requirement on the formalism in
+which we would write such grammars, we want a formalism where the set of
+all the ``useful'' definable languages is closed on intersection. If the
+object languages of ACGs are indeed closed on intersection, then our
+extension is just a convenience for expressing intersection more
+easily. However, if they are not, we believe our extension adds a
+critical level of expressivity needed for simple and modular
+wide-coverage grammars.
+
+\subsubsection{On the Decidability of Graphical ACGs}
+
+We turn to the problem of parsing with graphical ACGs. Parsing with ACGs
+is a problem on the limits of decidability. When we consider ACGs whose
+abstract signatures only contain types of order 2 or less, parsing is
+possible in polynomial type for both linear \cite{salvati2005problemes}
+and non-linear \cite{kanazawa2007parsing} type systems (such as the one
+we introduced in~\ref{sec:acg}). If the order of the abstract signature
+exceeds 2, the membership problem for linear ACGs becomes decidable if
+and only if the satisfiability problem in the multiplicative exponential
+fragment linear logic is decidable \cite{de2004vector}, which is an open
+problem. On the other hand, by extending the type system to include
+dependent types, parsing can become undecidable too since type checking
+in such a system is itself undecidable.
+
+Our intent here is to show that by extending the formalism, we have not
+made the problem of parsing with the grammar more difficult, at least in
+terms of decidability if not in terms of complexity. Let us imagine a
+graphical ACG where every edge (two signatures, distinguished types and
+a lexicon) is an ACG for which there is an effective parsing algorithm
+and all type checking is decidable in all the signatures. Then we can
+build a simple effective parsing algorithm for any of the extrinsic
+languages defined in the graphical ACG.
+
+We will be testing whether the term $\gamma$ belongs to the extrinsic
+language of some node $v$. Our algorithm first checks whether $\gamma$
+belongs to the intrinsic language of $v$ by checking its type. If it
+does, it proceeds to verify that it is also in the extrinsic language by
+finding antecedents in all of its abstract predecessors $u_i$, if there
+are any. We can use the presupposed parsing algorithm for the ACG
+$\mathopen{<} \Sigma_{u_i}, \Sigma_v, \mathcal{L}_{(u_i,v)},
+S_{u_i}\mathclose{>}$ to find all the abstract terms $\alpha_{i,j}$
+having type $S_{u_i}$ and being antecedents to $\gamma$ w.r.t. the
+lexicon $\mathcal{L}_{(u_i,v)}$. Such terms belong to the intrinsic
+language of $u_i$ but we are interested only in antecedents coming from
+the extrinsic language. Therefore, what we have to do is to recursively
+apply our parsing algorithm to all of the antecedent-candidates
+$\{\alpha_{i,j} \mid \forall j\}$ until we find one which belongs to the
+extrinsic language of $u_i$. Since the graph is acyclic, we are
+guaranteed that this recursive traversal will terminate. We repeat this
+for all the predecessors $u_i$ and if we succeed to find antecedents in
+all of them, we have confirmed the membership of $\gamma$ in
+$\mathcal{E}_{\mathcal{G}}(v)$ and also built its abstract
+antecedent(s).
+
+\subsubsection{On Grammar Engineering with Graphical ACGs}
+
+%% \subsubsection{On Alternative Interpretations of Graphical ACGs}
+%
+% Mention the example with the levels of linguistic representations
+% linked by many-to-many transductions and explain how a different
+% theory of graphs defining languages might be suitable to that
+% architecture.
+%
+% Also point out the weirdness of the diamond shape.

### 2013-06-15 the-last-part.tex
Finished drafting Graphical ACGs

@@
 \section{Negation}
+\label{sec:negation}
 
@@
 \section{Agreement}
+\label{sec:agreement}
 
@@
 \section{Intersections of Languages}
+\label{sec:sects-of-langs}
 
@@
 and the object language defined by $\mathcal{G}$.
+TODO: Prove here that the modes and patterns of composition in ACG have
+the predictable semantics as well.
 
@@
 
-%% \subsubsection{On Alternative Interpretations of Graphical ACGs}
-%
-% Mention the example with the levels of linguistic representations
-% linked by many-to-many transductions and explain how a different
-% theory of graphs defining languages might be suitable to that
-% architecture.
-%
-% Also point out the weirdness of the diamond shape.
+Having covered some of the formal properties of graphical ACGs, we will
+get back to building grammars. Armed with this formalism, we can solve
+the puzzle of incorporating the three constraints of sections
+\ref{sec:negation}, \ref{sec:extraction} and \ref{sec:agreement}. Our
+graph $G$ will consist of 6 nodes, $V(G) = \{Neg, Ext, Agr, Synt, Str,
+Sem\}$ with edges going from $Neg$, $Ext$ and $Agr$ to $Synt$ and from
+$Synt$ to $Str$ and $Sem$ (see Figure~TODO). The signatures for the
+first three named vertices are exactly those that have been presented in
+their dedicated sections and it is the case also for the distinguished
+types. The lexicons for the edges connecting these constraint nodes to
+$Synt$ all behave according to the scheme below:
+
+\begin{align*}
+  \mathcal{L}_{(constraint,Synt)}(X_{wordform_{subscript}}) &= C_{wordform} \\
+  \mathcal{L}_{(constraint,Synt)}(TYPE\_FEATURES)     &= TYPE
+\end{align*}
+
+The $Synt$-$Str$-$Sem$ subgraph is identical to the pair of ACGs
+displayed in \ref{sec:acg} except for some additions that need to be
+made so that the fragment covers the wordforms that we introduced when
+discussing the constraints (prepositions, relative pronouns, the
+negative particle $ne$\ldots). However, their types will not be tied to
+any of the constraints.
+
+\begin{align*}
+  C_{de} &: NP \limp N \limp N \\
+  C_{qui} &: (NP \limp S) \limp N \limp N \\
+  C_{dit\ que} &: NP \limp S \limp S \\
+  C_{dorment} &: NP \limp S \\
+  C_{ne_{tv}} &: (NP \limp NP \limp S) \limp (NP \limp NP \limp S) \\
+  C_{ne_{iv}} &: (NP \limp S) \limp (NP \limp S) \\
+  C_{ne_{cv}} &: (NP \limp S \limp S) \limp (NP \limp S \limp S)
+\end{align*}
+
+This lets us describe the high-level combinatorics of language that are
+of interest to the categorial grammarian in a manner which is idiomatic
+in academic literature (simple focused types $N$, $NP$ and $S$). At the
+same time, we can actually use this textbook grammar as a basis for a
+more realistic grammar which handles constraints such as agreement and
+extraction islands. Furthermore, these additional constraints are also
+defined in a manner that has its precedent in existing research. When
+the authors of \cite{pogodalla2012controlling} decided to put the
+constraints of extraction under a microscope and try to encode them in
+ACGs, they defined their constraints in the same way as our architecture
+does. This graphical architecture lets us write the syntactic kernel and
+the additional constraints independently and using two different styles
+that have both been shown useful by their own merits, which we see as
+indicative of enabling modularity.
+
+In the above approach, we have defined a language by providing a grammar
+which provides us with useful structures but which licences
+incorrect/ungrammatical structures. We have then refined this grammar by
+tacking on constraints ``on the side''. This approach is generally
+useful and can be seen e.g. in \cite{van2004concepts} to define the
+formal programming language kernel on which a formal semantics is
+established. In our case, we are interested in having intelligible
+syntactic structures too as we would like to compute their semantic
+representations. The syntax-semantics interface already presents a
+significant challenge and having less-than-ideal syntactic structures to
+start with does not make formalizing it any easier.
+
+We can imagine a scenario in which graphical ACGs become indispensable
+to have a clean syntax-semantics interface. In
+\ref{sec:sects-of-langs}(TODO: put here a more specific label after
+splitting up the section into subsections), we have seen a pattern of
+composing ACGs wherein ambiguities which are of a semantic nature are
+moved to a more abstract signature (i.e. the syntactic terms are
+transduced to the semantic terms). But there is also another pattern,
+that of adding a new abstract signature to constrain the syntactic
+terms. Since in classical ACGs, one does not define languages by
+providing more than one abstract signature, the two abstract signatures
+would have to be stacked on top of each other.
+
+Consider the case when we first add the abstract signature to constrain
+our syntactic terms. If we then try to transduce the terms of this new
+abstract signature, we will have to deal with the whole medley of types
+that were introduced in the constraining signature. This problem occurs
+since in our type-logical grammar the type system of the syntactic
+signature does not provide just the interface of the syntactic layer,
+but also its implementation. By putting the constraining signature in a
+different subgraph, aside from the syntax-semantics transducer, we can
+shield the syntax-semantics interface from the gnarly implementation
+details of the rules of syntax. This in turn gives us an answer as to
+what level of information should be expressed in the syntactic signature
+itself and which other constraints should be tucked away in hidden
+abstract signatures: the syntactic signature to which the
+syntax-semantics transducer is connected should only contain information
+relevant to the syntax-semantics interface. (TODO: Think about this some
+more, about whether there is any important difference to the
+syntax-semantics interface when the syntactic signature does not enforce
+anything, e.g..)
+
+As for the other case where we first build the syntax-semantics
+transducer and than constrain its abstract signature with a more
+abstract one is not much better. Now we have a converse problem where
+the implementation of the syntactic constraints is complected with bits
+of semantics since the constraining signature will have to map onto a
+signature built around types combining both syntax and semantics. (TODO:
+Go to bed.)
+
+% We could have talked here about making the graph of constraints deeper
+% to ``reuse'' common concepts and types. However due to the need for
+% more abstract signatures to enforce and basically reimplement the
+% constraints in less abstract signatures this would not be so
+% useful.
+
+% We could also talk about how constraints could be added to semantic or
+% morphological signatures, but we have nothing interesting to say about
+% that.
+
+\subsubsection{On Alternative Interpretations of Graphical ACGs}
+
+Our definitions of intrinsic and extrinsic have consequences which might
+be counter-intuitive and not correspond to the grammarian's intent. We
+will describe what these consequences are and propose an alternative
+notion of a language defined by a graphical ACG which aims to reconcile
+them.
+
+The first scenario which can be tricky is when two paths from nodes
+$u_1$ and $u_2$ converge on a single node $v$. If we then take a term
+$\alpha$ from $\mathcal{E}_{\mathcal{G}}(u_1)$, the term
+$\mathcal{L}_{(u_1,v)}(\alpha)$ belongs to
+$\mathcal{I}_{\mathcal{G}}(v)$ but not necessarily to
+$\mathcal{E}_{\mathcal{G}}(v)$ since it might be the case that there is
+no antecedent for $\mathcal{L}_{(u_1,v)}(\alpha)$ in
+$\mathcal{E}_{\mathcal{G}}(u_2)$. To sum up, mapping a member of some
+$\mathcal{E}_{\mathcal{G}}(u)$ using some $\mathcal{L}_{(u,v)}$ does not
+guarantee that the result belongs to $\mathcal{E}_{\mathcal{G}}(v)$,
+which is contrary to the case of object languages in ACGs.
+
+This highlights some limitations of our notion of an extrinsic
+language. Consider the graphical ACG of Figure~TODO. We have more than
+one path converging to $Synt$. We will consider a term $\alpha$ of
+$\mathcal{E}_{\mathcal{G}}(Sem)$. $\alpha$ has an antecedent in
+$\mathcal{E}_{\mathcal{G}}(SyntSem)$ which has a descendant $\beta$ in
+$\mathcal{I}_{\mathcal{G}}(Synt)$. However, because of the constraints
+on $Synt$ ($Neg$, $Ext$ and $Agr$), $\beta$ can be syntactic term
+violating one of these constraints which yields an ungrammatical
+sentence. Therefore, the notion of an extrinsic language is insufficient
+to express the set of meaning terms from $\Lambda(\Sigma_{Sem})$ which
+are expressible by grammatical sentences. Such definitions would be
+useful in an example like the one in Figure~TODO where we have a chain
+of signatures representing the different levels of linguistic
+description whose terms are in many-to-many relationships expressed
+through transducers. In such a scenario, it would be desirable to be
+able to express the set of descriptions belonging to one level,
+e.g. morphology, that are related to valid descriptions on all the other
+levels.
+
+Compared to ACG object languages, we lose another guarantee. In ACGs,
+the central role of the root abstract-most signature guarantees us that
+whenever a term belongs to an object language, we can find terms in all
+the other object languages that all share the same abstract-most
+term. On the other hand in the intrinsic and extrinsic languages we
+defined on graphical ACGs, having a term in an extrinsic language of
+some node $v$ only guarantees the presence of corresponding terms in the
+extrinsic languages of its ancestors and the intrinsic languages of
+their descendants.
+
+The second puzzling scenario is the diamond shape on Figure~TODO. If we
+consider a term $\delta$ of $\mathcal{E}_{\mathcal{G}}(Bottom)$, we know
+that there exist antecedents $\beta$ and $\gamma$ in
+$\mathcal{E}_{\mathcal{G}}(Left)$ and
+$\mathcal{E}_{\mathcal{G}}(Right)$. And while there definitely exist
+antecedents $\alpha$ and $\alpha'$ to both $\beta$ and $\gamma$ in
+$\mathcal{E}_{\mathcal{G}}(Top)$, these antecedents can be different
+disjoint, meaning there does not have to be a single $\alpha \in
+\mathcal{E}_{\mathcal{G}}(Top)$ which yields $\beta \in
+\mathcal{E}_{\mathcal{G}}(Left)$ and $\gamma \in
+\mathcal{E}_{\mathcal{G}}(Right)$ such that they both yield
+$\delta$. This means that the result of parsing a term in some
+$\mathcal{E}_{\mathcal{G}}(v)$ is a collection of abstract terms with
+one term not just for every node $u$ that is more abstract than $v$ but
+for every path leading to $v$. Parsing in our diamond grammar would thus
+produce four terms $\beta$, $\gamma$, $\alpha$ and $\alpha'$
+corresponding to the four paths $(Left,Bottom)$, $(Right,Bottom)$,
+$(Top,Left,Bottom)$ and $(Top,Right,Bottom)$, respectively. We contrast
+this again with the case of ACGs where the result of parsing can be
+represented by associating a term to every signature in the
+arborescence.\footnote{One could also say that the result of parsing in
+  an ACG is just the antecedent in the abstract-most signature and that
+  the other terms can then be easily computed by application of
+  lexicons. Similarly, we could say that the result of parsing in a
+  general graphical ACG is a collection of abstract terms for every
+  \emph{maximal} path leading to the node defining the language in
+  question. What might be more intuitive, though, is to have some theory
+  of graphical ACGs where the result of parsing is a collection of
+  antecedents in all the abstract-most nodes.}
+
+We will propose a new way of assigning languages to nodes in a graphical
+ACG that will conserve some of the intuitive behavior we are accustomed
+to from ACGs. We need not modify the definition of graphical ACGs to do
+so since it is only considered with the structure of the graph and its
+decoration with ACG paraphernalia, we only define a new kind of
+languages.
+
+Let $\mathcal{G} = \mathopen{<} G, \Sigma, S, \mathcal{L} \mathclose{>}$
+be a graphical ACG. We define the \emph{pangraphical language} of node
+$u$, $\mathcal{P}_{\mathcal{G}}(u)$, to be the set of terms $t$ such
+that there exists a labeling of nodes with terms $T$ satisfying the
+following conditions:
+
+\begin{itemize}
+  \item $T_u = t$.
+  \item For all $v \in V(G)$, $\vdash_{\Sigma_v} T_v : S_v$.
+  \item For all $(v,w) \in E(G)$, $\mathcal{L}_{(v,w)}(T_v) = T_w$.
+\end{itemize}
+
+TODO!: Show that the pangraphical languages solve all the problems
+outlined above.
+
+TODO: $\mathcal{P}_{\mathcal{G}}(u) \subset \mathcal{E}_{\mathcal{G}}(u)
+\subset \mathcal{I}_{\mathcal{G}}(u)$
+
+TODO: On arborescences, $\mathcal{P}_{\mathcal{G}}(u) =
+\mathcal{E}_{\mathcal{G}}(u)$.
+
+TODO?: Show that extrinsic languages are included in pangraphical
+languages.
+
+TODO?: Are pangraphical languages also included in extrinsic languages?
+
+TODO?: Write a parsing algorithm for pangraphical languages similar to
+the one for extrinsic languages.
+
+\subsubsection{On the Practical Consequences of Graphical ACGs}
+
+TODO: The trivial lexicons we have for our constraints can be
+effectively inverted, giving us effective constraint verification. Using
+filtering and/or supertagging could also make it efficient.
+
+TODO: Constraint signatures can use specific type extensions, their
+added complexity is then contained within the constraint. Furthermore,
+we can treat the constraint signature as a black box and in the context
+of an implementation, we replace it with a special purpose procedure
+which verifies the constraint efficiently.
+
+TODO?: Practical applications: robustness by being able to parse
+sentences which violate some constraints, fine-grained syntax checking
+(pinpoint the constraint that is being violated).
+
+TODO?: We will need to define a lot of signatures which look exactly
+like some other signature but with some small refinements. The ways we
+will refine the signature will hopefully not be so diverse and we would
+hope to develop some tools to make formulating these refinements
+feasible compared to having to restate the entire grammar every time we
+add a new constraint.

### 2013-06-16 the-last-part.tex
Finished the graphical ACG chapter

--- 2013-06-13 the-last-part.tex
@@
 
-$$
-N_{aucun} : N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S) \limp S)
-$$
-$$
-N_{le_1} : N\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S)
-$$
-$$
-N_{le_2} : N\_NEG{=}T \limp ((NP\_NEG{=}T \limp S) \limp S)
-$$
+\begin{align*}
+N_{aucun} &: N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S) \limp S) \\
+N_{le_1} &: N\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S) \\
+N_{le_2} &: N\_NEG{=}T \limp ((NP\_NEG{=}T \limp S) \limp S) \\
+\end{align*}
 
@@
 
-$$
-N_{ne_{tv_1}} : (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}F \limp S)
-$$
-$$
-N_{ne_{tv_2}} : (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}F \limp NP\_NEG{=}T \limp S)
-$$
+\begin{align*}
+N_{ne_{tv_1}} &: (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}F \limp S) \\
+N_{ne_{tv_2}} &: (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}F \limp NP\_NEG{=}T \limp S)
+\end{align*}
 
@@
 
-$$
-N_{ne_{tv_3}} : (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}T \limp S)
-$$
+\begin{align*}
+N_{ne_{tv_3}} &: (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}T \limp S)
+\end{align*}
 
@@
 
-$$
-N_{que_1} : (NP\_NEG{=}F \limp S) \limp N\_NEG{=}F \limp N\_NEG{=}F
-$$
-$$
-N_{que_2} : (NP\_NEG{=}F \limp S) \limp N\_NEG{=}T \limp N\_NEG{=}T
-$$
-$$
-N_{de_1} : NP\_NEG{=}F \limp N\_NEG{=}F \limp N\_NEG{=}F
-$$
-$$
-N_{de_2} : NP\_NEG{=}F \limp N\_NEG{=}T \limp N\_NEG{=}T
-$$
-$$
-N_{de_3} : NP\_NEG{=}T \limp N\_NEG{=}F \limp N\_NEG{=}T
-$$
+\begin{align*}
+N_{que_1} &: (NP\_NEG{=}F \limp S) \limp N\_NEG{=}F \limp N\_NEG{=}F \\
+N_{que_2} &: (NP\_NEG{=}F \limp S) \limp N\_NEG{=}T \limp N\_NEG{=}T \\
+N_{de_1} &: NP\_NEG{=}F \limp N\_NEG{=}F \limp N\_NEG{=}F \\
+N_{de_2} &: NP\_NEG{=}F \limp N\_NEG{=}T \limp N\_NEG{=}T \\
+N_{de_3} &: NP\_NEG{=}T \limp N\_NEG{=}F \limp N\_NEG{=}T
+\end{align*}
 
@@
 
-$$
-E_{dort_1} : NP\_VAR{=}F \limp S\_EXT{=}NO
-$$
-$$
-E_{dort_2} : NP\_VAR{=}T \limp S\_EXT{=}ROOT
-$$
-$$
-E_{aime_1} : NP\_VAR{=}F \limp NP\_VAR{=}F \limp S\_EXT{=}NO
-$$
-$$
-E_{aime_2} : NP\_VAR{=}T \limp NP\_VAR{=}F \limp S\_EXT{=}ROOT
-$$
-$$
-E_{aime_3} : NP\_VAR{=}F \limp NP\_VAR{=}T \limp S\_EXT{=}ROOT
-$$
-$$
-E_{dit\ que_1} : NP\_VAR{=}F \limp S\_EXT{=}NO \limp S\_EXT{=}NO
-$$
-$$
-E_{dit\ que_2} : NP\_VAR{=}T \limp S\_EXT{=}NO \limp S\_EXT{=}ROOT
-$$
-$$
-E_{dit\ que_3} : NP\_VAR{=}F \limp S\_EXT{=}ROOT \limp S\_EXT{=}EMB
-$$
-$$
-E_{dit\ que_4} : NP\_VAR{=}F \limp S\_EXT{=}EMB \limp S\_EXT{=}EMB
-$$
+\begin{align*}
+E_{dort_1} &: NP\_VAR{=}F \limp S\_EXT{=}NO \\
+E_{dort_2} &: NP\_VAR{=}T \limp S\_EXT{=}ROOT \\
+E_{aime_1} &: NP\_VAR{=}F \limp NP\_VAR{=}F \limp S\_EXT{=}NO \\
+E_{aime_2} &: NP\_VAR{=}T \limp NP\_VAR{=}F \limp S\_EXT{=}ROOT \\
+E_{aime_3} &: NP\_VAR{=}F \limp NP\_VAR{=}T \limp S\_EXT{=}ROOT \\
+E_{dit\ que_1} &: NP\_VAR{=}F \limp S\_EXT{=}NO \limp S\_EXT{=}NO \\
+E_{dit\ que_2} &: NP\_VAR{=}T \limp S\_EXT{=}NO \limp S\_EXT{=}ROOT \\
+E_{dit\ que_3} &: NP\_VAR{=}F \limp S\_EXT{=}ROOT \limp S\_EXT{=}EMB \\
+E_{dit\ que_4} &: NP\_VAR{=}F \limp S\_EXT{=}EMB \limp S\_EXT{=}EMB
+\end{align*}
 
@@
 
-$$
-E_{qui} : (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N
-$$
-$$
-E_{que_1} : (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N
-$$
-$$
-E_{que_2} : (NP\_VAR{=}T \limp S\_EXT{=}EMB) \limp N \limp N
-$$
+\begin{align*}
+E_{qui} &: (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N \\
+E_{que_1} &: (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N \\
+E_{que_2} &: (NP\_VAR{=}T \limp S\_EXT{=}EMB) \limp N \limp N
+\end{align*}
 
@@
 
-$$
-E_{le_1} : N \limp ((NP\_VAR{=}F \limp S\_EXT{=}NO) \limp S\_EXT{=}NO)
-$$
-$$
-E_{le_2} : N \limp ((NP\_VAR{=}F \limp S\_EXT{=}ROOT) \limp S\_EXT{=}ROOT)
-$$
-$$
-E_{le_3} : N \limp ((NP\_VAR{=}F \limp S\_EXT{=}EMB) \limp S\_EXT{=}EMB)
-$$
+\begin{align*}
+E_{le_1} &: N \limp ((NP\_VAR{=}F \limp S\_EXT{=}NO) \limp S\_EXT{=}NO) \\
+E_{le_2} &: N \limp ((NP\_VAR{=}F \limp S\_EXT{=}ROOT) \limp S\_EXT{=}ROOT) \\
+E_{le_3} &: N \limp ((NP\_VAR{=}F \limp S\_EXT{=}EMB) \limp S\_EXT{=}EMB)
+\end{align*}
 
@@
 
-$$
-E_{de} : NP\_VAR{=}F \limp N \limp N
-$$
+\begin{align*}
+E_{de} &: NP\_VAR{=}F \limp N \limp N
+\end{align*}
 
@@
 these pieces of information percolate up the parse tree all the way to
-the topmost level still pertinent for ensuring grammaticality. We will
-now look at another similar example.
+the topmost level still pertinent for ensuring grammaticality.
 
 \section{Agreement}
-
-Finally, we will cover the phenomenon of number agreement. This
-motivates a straightforward refinement of our types, where all noun and
-noun phrase types carry number information.
-
-$$
-A_{Marie} : NP\_NUM{=}SG
-$$
-$$
-A_{tatou} : N\_NUM{=}SG
-$$
-$$
-A_{tatous} : N\_NUM{=}PL
-$$
-$$
-A_{le} : N\_NUM{=}SG \limp ((NP\_NUM{=}SG \limp S) \limp S)
-$$
-$$
-A_{les} : N\_NUM{=}PL \limp ((NP\_NUM{=}PL \limp S) \limp S)
-$$
-$$
-A_{dort} : NP\_NUM{=}SG \limp S
-$$
-$$
-A_{dorment} : NP\_NUM{=}PL \limp S
-$$
-$$
-A_{aime_1} : NP\_NUM{=}SG \limp NP\_NUM{=}SG \limp S
-$$
-$$
-A_{aime_2} : NP\_NUM{=}SG \limp NP\_NUM{=}PL \limp S
-$$
-$$
-A_{aiment_1} : NP\_NUM{=}PL \limp NP\_NUM{=}SG \limp S
-$$
-$$
-A_{aiment_2} : NP\_NUM{=}PL \limp NP\_NUM{=}PL \limp S
-$$
-$$
-A_{qui_1} : (NP\_NUM{=}SG \limp S) \limp N\_NUM{=}SG \limp N\_NUM{=}SG
-$$
-$$
-A_{qui_2} : (NP\_NUM{=}PL \limp S) \limp N\_NUM{=}PL \limp N\_NUM{=}PL
-$$
-$$
-A_{de_1} : NP\_NUM{=}SG \limp N\_NUM{=}SG \limp N\_NUM{=}SG
-$$
-$$
-A_{de_2} : NP\_NUM{=}PL \limp N\_NUM{=}SG \limp N\_NUM{=}SG
-$$
-$$
-A_{de_3} : NP\_NUM{=}SG \limp N\_NUM{=}PL \limp N\_NUM{=}PL
-$$
-$$
-A_{de_4} : NP\_NUM{=}PL \limp N\_Pl \limp N\_NUM{=}PL
-$$
+\label{sec:agreement}
+
+Finally, we will look at another example of implementing a linguistic
+constraint in ACGs by covering the phenomenon of number
+agreement.
+
+Number agreement motivates a straightforward refinement of our types,
+where all noun and noun phrase types carry number information.
+
+\begin{align*}
+A_{Marie} &: NP\_NUM{=}SG \\
+A_{tatou} &: N\_NUM{=}SG \\
+A_{tatous} &: N\_NUM{=}PL \\
+A_{le} &: N\_NUM{=}SG \limp ((NP\_NUM{=}SG \limp S) \limp S) \\
+A_{les} &: N\_NUM{=}PL \limp ((NP\_NUM{=}PL \limp S) \limp S) \\
+A_{dort} &: NP\_NUM{=}SG \limp S \\
+A_{dorment} &: NP\_NUM{=}PL \limp S \\
+A_{aime_1} &: NP\_NUM{=}SG \limp NP\_NUM{=}SG \limp S \\
+A_{aime_2} &: NP\_NUM{=}SG \limp NP\_NUM{=}PL \limp S \\
+A_{aiment_1} &: NP\_NUM{=}PL \limp NP\_NUM{=}SG \limp S \\
+A_{aiment_2} &: NP\_NUM{=}PL \limp NP\_NUM{=}PL \limp S \\
+A_{qui_1} &: (NP\_NUM{=}SG \limp S) \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
+A_{qui_2} &: (NP\_NUM{=}PL \limp S) \limp N\_NUM{=}PL \limp N\_NUM{=}PL \\
+A_{de_1} &: NP\_NUM{=}SG \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
+A_{de_2} &: NP\_NUM{=}PL \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
+A_{de_3} &: NP\_NUM{=}SG \limp N\_NUM{=}PL \limp N\_NUM{=}PL \\
+A_{de_4} &: NP\_NUM{=}PL \limp N\_Pl \limp N\_NUM{=}PL
+\end{align*}
 
@@
 
-In the next section, we will see what it takes to enforce all of the
-constraints above (negation, extraction and agreement) at the same time,
-which will give us some idea on how to go about building a grammar which
-handles a wide variety of such phenomena.
-
 \section{Putting It All Together}
+\label{sec:together}
+
+In this section, we will see what it takes to enforce all of the
+constraints introduced in the preceding sections (negation, extraction
+and agreement) at the same time, which will give us some idea on how to
+go about building a grammar which handles a wide variety of such
+phenomena.
 
@@
 
-$$
-C_{de_1} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_2} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_3} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_4} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_5} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_6} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_7} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_8} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_9} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_{10}} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_{11}} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_{12}} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
-$$
+\begin{align*}
+C_{de_1} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
+(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \\
+C_{de_2} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
+(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \\
+C_{de_3} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
+C_{de_4} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
+C_{de_5} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
+C_{de_6} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
+C_{de_7} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
+(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \\
+C_{de_8} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
+(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \\
+C_{de_9} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
+C_{de_{10}} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
+C_{de_{11}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
+C_{de_{12}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
+\end{align*}
 
@@
 $$
-$$
-C_{de_{1(\textcolor{blue}{x}, \textcolor{blue}{y})}} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y})
-$$
-$$
-C_{de_{2(\textcolor{blue}{x}, \textcolor{blue}{y})}} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y})
-$$
-$$
-C_{de_{3(\textcolor{blue}{x}, \textcolor{blue}{y})}} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y})
-$$
+\begin{align*}
+C_{de_{1(\textcolor{blue}{x}, \textcolor{blue}{y})}} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \\
+C_{de_{2(\textcolor{blue}{x}, \textcolor{blue}{y})}} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y}) \\
+C_{de_{3(\textcolor{blue}{x}, \textcolor{blue}{y})}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y})
+\end{align*}
 
@@
 $$
-$$
-C_{de(\textcolor{blue}{x}, \textcolor{blue}{y}, \textcolor{red}{p}, \textcolor{red}{q})} : (NP\_\textcolor{red}{NEG{=}p}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}q}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}(p \lor q)}\_\textcolor{blue}{NUM{=}y})
-$$
+\begin{align*}
+C_{de(\textcolor{blue}{x}, \textcolor{blue}{y}, \textcolor{red}{p}, \textcolor{red}{q})} &: (NP\_\textcolor{red}{NEG{=}p}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}q}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}(p \lor q)}\_\textcolor{blue}{NUM{=}y})
+\end{align*}
 
@@
 these are independent mechanisms. One has to examine the type to make
-sure that there is no interaction. In the next two subsections, we will
+sure that there is no interaction. In the rest of this chapter, we will
 devote ourselves to investigating ways of disentangling these.
@@
 \section{Intersections of Languages}
-
-We will now turn to another solution to the complexity of grammars
-implementing multiple independent mechanisms.
-
-When ACGs are discussed, one of the oft-praised properties is
-compositionality. Let us consider a situation where we have three
-signatures $\Sigma_1$, $\Sigma_2$ and $\Sigma_3$, their respective
-distinguished types $S_1$, $S_2$ and $S_3$ and two lexicons
-$\mathcal{L}_{12} : \Sigma_1 \to \Sigma_2$ and $\mathcal{L}_{23} :
-\Sigma_2 \to \Sigma_3$ such that $\mathcal{L}_{12}(S_1) = S_2$ and
-$\mathcal{L}_{23}(S_2) = S_3$. (TODO: INCLUDE A DIAGRAM OF THIS
-SITUATION)
+\label{sec:sects-of-langs}
+
+We will turn to another solution to the problem of complex grammars that
+try to implement multiple independent constraints. This solution will be
+based on a generalization of ACGs, but in order to introduce this
+generalization, we will first need to present the manner in which ACGs
+and their compositions are formulated in ACG literature.
+
+\subsection{The Usual Formalization of ACGs}
+\label{ssec:usual-acgs}
+
+Let us consider a system with three signatures $\Sigma_1$, $\Sigma_2$
+and $\Sigma_3$, their respective distinguished types $S_1$, $S_2$ and
+$S_3$ and two lexicons $\mathcal{L}_{12} : \Sigma_1 \to \Sigma_2$ and
+$\mathcal{L}_{23} : \Sigma_2 \to \Sigma_3$ such that
+$\mathcal{L}_{12}(S_1) = S_2$ and $\mathcal{L}_{23}(S_2) = S_3$. See
+Figure~\ref{fig:acg-serial-comp} for a diagrammatic visualization of
+this structure.
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{diagrams/acg-serial-comp.pdf}
+    \caption{\label{fig:acg-serial-comp} Serial composition of ACGs.}
+  \end{subfigure}
+  \qquad
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{diagrams/acg-parallel-comp.pdf}
+    \caption{\label{fig:acg-parallel-comp} Parallel composition of ACGs.}
+  \end{subfigure}
+  \caption{\label{fig:acg-comp-modes} Two modes of composition for
+    ACGs.}
+\end{figure}
 
@@
 $\mathcal{L}_{12}(S_1) = S_2$ and will therefore also belong to $A_2$
-(similarly for $\mathcal{L}_{23}(A_2)$ and $A_3$). These new languages,
-which are formed by mapping an ACG abstract language using a lexicon,
-are called \emph{object languages}. Their terms are constrained not only
-by the types of the signature they are built upon but also by the type
-system of the abstract signature.
+(similarly, we have that $\mathcal{L}_{23}(A_2) \subseteq A_3$). These
+new languages, which are formed by mapping an ACG abstract language
+using a lexicon, are called \emph{object languages}. Their terms are
+constrained not only by the types of the signature they are built upon
+but also by the type system of the abstract signature.
 
@@
 $\mathcal{L}_{12}(A_1)$ given by $\mathcal{L}_{23}$ and also the image
-of the abstract language $A_1$ given by $\mathcal{L}_{13}$. This points
-to an interesting property of ACG languages: they are closed under
-transformation by a lexicon.
+of the abstract language $A_1$ given by $\mathcal{L}_{13}$ (thus it is
+itself also an object language). This points to an interesting property
+of ACG languages: they are closed under transformation by a lexicon.
 
@@
 The definition of ACGs given above is quite practical in that it is
-complete, it gives a precise account of what set of languages are
+complete, i.e. it gives a precise account of what set of languages are
 generated by ACGs (abstract and object languages for which there exist
 the quadruples described above). This contrasts with the presentation of
-ACGs that we gave in~\ref{sec:acg}, where we defined ACGs as
-collections of signatures with associated distinguished types and
-connected with lexicons. Our definition only introduced the requisite
-machinery, but delayed the delimitation of the way languages can be
-defined. Leaving this question open will let us answer it now by
-providing an extension/generalization of ACGs which will help us
-overcome the modularity issues that have been the theme of this
-chapter.\footnote{Incidentally, this was not the reason for the way we
-  presented ACGs in~\ref{sec:acg}. We chose this style as it grouped
-  together the information in a more natural way. We approach signatures
-  as descriptions of domains (strings, syntax, semantics), each one
-  being associated with a distinguished type telling us which objects
-  the domain has set out to describe (strings, sentences, truth
-  values). Having distinguished types directly attached to signatures
-  allows us to omit them from the ACG quadruples. However, this only
-  leaves us with triples containing the two signatures and the lexicon
-  connecting them, and since we can infer the signatures themselves from
-  the domain and the range of the lexicon, we do not have much
-  motivation to introduce these tuples. Instead of composing ACGs, we
-  compose just the lexicons.
-
-On top of that, we will want to define signatures and lexicons that
-yield different object languages (at least some string representation of
-the form and some semantic representation of the meaning), which would
-mean having at least two different ACGs which are linked (e.g. by
-sharing the same abstract signature). However, since these two grammars
-are very closely related, we would prefer to reason about our grammar as
-\emph{one} system/graph of signatures and lexicons which defines
-multiple languages instead of talking about \emph{multiple} grammars
-which share some similarities.
-
-TODO: Somehow express that its nice to have the links between the
-signatures and lexicons that are shown in diagrams \emph{explicit}
-rather than having them \emph{implicit}.}
+ACGs that we gave in~\ref{sec:acg}, where we defined ACGs as collections
+of signatures associated with distinguished types and connected with
+lexicons. Our definition only introduced the requisite machinery, but
+delayed the delimitation of the way languages can be defined. Leaving
+this question open will let us answer it now by providing an extension
+of ACGs which will help us overcome the modularity issues that have been
+the theme of this chapter.\footnote{Incidentally, this was not the
+  reason for the way we presented ACGs in~\ref{sec:acg}. We chose this
+  style as it grouped together the information in a more natural way. We
+  approach signatures as descriptions of domains (strings, syntax,
+  semantics), each one being associated with a distinguished type
+  telling us which objects the domain has set out to describe (strings,
+  sentences, truth values). Having distinguished types directly attached
+  to signatures allows us to omit them from the ACG quadruples,
+  $\mathopen{<}\Sigma_A, \Sigma_O, \mathcal{L},
+  \cancel{S}\mathclose{>}$. However, this only leaves us with triples
+  containing the two signatures and the lexicon connecting them, and
+  since we can infer the signatures themselves from the domain and the
+  range of the lexicon $\mathcal{L} : \Sigma_A \to \Sigma_O$, we do not
+  have much motivation to introduce these tuples,
+  $\mathopen{<}\cancel{\Sigma_A}, \cancel{\Sigma_O}, \mathcal{L},
+  \cancel{S}\mathclose{>}$. Instead of composing ACGs, we compose just
+  the lexicons.
+
+On top of that, we will want to define signatures and lexicons to yield
+multiple different object languages (at least one for the string
+representations of forms and one for the semantic representations of
+meaning), which would mean having at least two different ACGs which are
+somehow linked together (e.g. by sharing the same abstract
+signature). However, since these two grammars are so interrelated, we
+would like to reason about them in a way that makes these relationships
+explicit and thus easier to study as opposed to considering a collection
+of grammars which are somehow implicitly similar.}
+
+\subsection{Patterns of Composition in ACGs}
+\label{ssec:acg-patterns}
 
@@
 signature of one grammar be the abstract signature of another as we saw
-above (TODO: INSERT REFERENCE TO FUTURE DIAGRAM), we can have two
-grammars share the same abstract signature allowing us to transduce
-between terms in the two object signatures. See Figure~(TODO: INSERT A
-DIAGRAM FOR THIS SITUATION) for a diagrammatic example.
-
-These two modes of composition can be fruitfully mixed. Consider (TODO:
-INSERT A DIAGRAM), where we introduce a syntactic signature TODO where
-different terms are only allowed to yield the same string whenever they
-represent different syntactic structures of a syntactically ambiguous
-phrase. This means that purely semantic distinctions such as scope
-ambiguity have to be moved to a more abstract signature from whose terms
-the semantic representation can be derived by a function. This approach
-is studied more deeply in \cite{pogodalla2007generalizing}.
+before (Figure~\ref{fig:acg-serial-comp}), we can have two grammars
+share the same abstract signature allowing us to transduce between terms
+in the two object signatures (Figure~\ref{fig:acg-parallel-comp}).
+
+NB: \emph{Transduction} is a process in which mapping an object $o_1$ to
+(potentially many) other objects $o_2$ is realized by first inverting
+one function $f_1$ to find the object's antecedents $a$ and then
+applying to them another function $f_2$, i.e. an object $o_1$ is
+transduced to an object $o_2$ whenever there exists an $a$ such that
+$f_1(a) = o_1$ and $f_2(a) = o_2$. In our example, terms from
+$\mathcal{O}( \mathopen{<} \Sigma_1, \Sigma_2, \mathcal{L}_{12}, S
+\mathclose{>} )$ are transduced to terms from $\mathcal{O}( \mathopen{<}
+\Sigma_1, \Sigma_3, \mathcal{L}_{13}, S \mathclose{>} )$ by inverting
+the lexicon $\mathcal{L}_{12}$ to find an antecedent in $\mathcal{A}(
+\mathopen{<} \Sigma_1, \Sigma_2, \mathcal{L}_{12}, S \mathclose{>} )$
+and then applying to it the lexicon $\mathcal{L}_{13}$.
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{diagrams/parallel-over-serial.pdf}
+    \caption{\label{fig:parallel-over-serial} Transduction from an
+      abstract signature.}
+  \end{subfigure}
+  \qquad
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{diagrams/serial-over-parallel.pdf}
+    \caption{\label{fig:serial-over-parallel} Constraint on an abstract
+      signature.}
+  \end{subfigure}
+  \caption{\label{fig:acg-comp-patterns} Mixing the ACG modes of
+    composition.}
+\end{figure}
+
+These two modes of composition can be fruitfully mixed. Consider the
+graph on Figure~\ref{fig:parallel-over-serial} in which we introduce a
+syntactic signature $Syntax$ where different terms are only allowed to
+yield the same string whenever they represent different parses of a
+syntactically ambiguous phrase. This means that purely semantic
+distinctions such as scope ambiguity have to be moved to a more abstract
+signature ($SyntSem$) from whose terms the semantic representation can
+be derived by a function ($\mathcal{L}_{InterSem}$). This approach is
+studied more deeply in \cite{pogodalla2007generalizing}.
 
 Another interesting composition pattern is to take an existing structure
-of grammars and add a new abstract signature on top. This signature can
-be used to further constrain the items in (TODO: what was before the
-abstract-most signature/language). In \cite{pogodalla2012controlling},
+of grammars and add a new abstract signature on top
+(Figure~\ref{fig:serial-over-parallel}). This new signature ($Constr$)
+can be used to further constrain the items in the previously
+abstract-most signature ($Synt$). In \cite{pogodalla2012controlling},
 this is what the authors use to develop their constraints on
@@
 
-However, a question that the authors do not pose is how to combine all
-these constraints. If we were to apply the same pattern repeatedly to
-get a structure like the one on (TODO: INSERT A DIAGRAM HERE), we would
-have to deal with successively more complex type systems on each
-level. This is due to the condition that a lexicon has to be a
-homomorphism and thus it must, among other things, always map well-typed
-terms to well-typed terms. To illustrate the problem, imagine a
-structure such as the one on (TODO) that handles some different
-constraint in each of the vertically stacked signatures, including
-e.g. agreement. If we tried to add a new abstract signature at the top
-to implement another constraint, we would have to ensure that all the
-terms we licence in our new abstract signature have a well-typed
-counterpart in the existing abstract language. However, this means that
-our new signature must make sure that the terms it approves do not break
-agreement or any of the previously established constraints. Effectively,
-our new signature would be forced to reimplement all of the existing
-constraints, rendering in vain all our effort to split up the
-constraints into multiple signatures.
-
-What we would like instead is to put the signatures side-by-side such
-that they do not depend on each other (as on diagram TODO). However, how
-does this actually define a language? We can arrive at the answer by
-examining how are the graphical representations of ACGs connected to the
-languages they define. We will associate a language to every node of the
-graph. Before we begin, we notice that every ACG diagram formed by the
-above rules of composition is a directed tree and therefore every node
-has at most one parent/predecessor. To the root of the tree, the
-abstract-most signature in the graph, we will assign the abstract
-language generated by that signature and its distinguished type. To all
-the other nodes, we will assign the result of mapping their parent's
-language by the lexicon linking the two. By homomorphism of the
-lexicons, we can also deduce the distinguished types for all the
-non-root nodes. They are the images of their parents' distinguished
-types by the connecting lexicons. For all of the nodes, it is true that
-every term in the language assigned to the node has the node's
-distinguished type.
+However, a question that the authors do not pose, but which is of great
+interest to us, is how to combine all these constraints. If we were to
+apply the same pattern repeatedly to get a structure like the one on
+Figure~\ref{fig:stacked-constraints}, we would have to deal with
+successively more complex type systems on each level. This is due to the
+condition that a lexicon has to be a homomorphism and thus it must,
+among other things, always map well-typed terms to well-typed
+terms. Because of this, the terms in the abstract language of the
+signature we introduce to handle the second constraint must also satisfy
+the first constraint, since mapping them by $\mathcal{L}_{Constr_2}$
+must yield a well-typed term in the signature corresponding to the first
+constraint. In the end, whenever we wish to add a new signature to
+handle another constraint, we have to acknowledge and reimplement in it
+all of the existing constraints. This effectively renders the pattern
+useless since we might just as well consider the final abstract-most
+signature handling all the constraints ($Constr_{1 \land 2}$) and attach
+it directly to the syntactic signature ($Synt$) by taking the
+composition of the lexicons in between ($\mathcal{L}_{Constr_1} \circ
+\mathcal{L}_{Constr_2}$).
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{diagrams/stacked-constraints.pdf}
+    \caption{\label{fig:stacked-constraints} Stacking the constraints
+      vertically.}
+  \end{subfigure}
+  \qquad
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{diagrams/constraints-side-by-side.pdf}
+    \caption{\label{fig:constraints-side-by-side} Laying out the
+      constraints side by side.}
+  \end{subfigure}
+  \caption{Constraining a syntactic signature with multiple
+    constraints.}
+\end{figure}
+
+What we would like instead is to put the signatures side by side such
+that they do not depend on each other
+(Figure~\ref{fig:constraints-side-by-side}). However, our language of
+syntactic structures is no longer defined by mapping a single abstract
+language through a lexicon. It is outside of the scope of the common
+formalization of ACGs that we presented in~\ref{ssec:usual-acgs}. How do
+we define this language in a manner that is consistent with our current
+definitions of ACGs? We can arrive at the answer by examining how the
+graphical representations of ACGs are connected to the languages they
+define.
+
+\subsection{Interpreting ACG Diagrams}
+
+We will associate a language to every node of an ACG diagram. Before we
+begin, we notice that every ACG diagram formed by the above modes of
+composition is an arborescence\footnote{An arborescence is a directed
+  rooted tree, where every edge points away from the root.} and
+therefore every node has at most one predecessor (the parent). To the
+root of the tree, the abstract-most signature in the graph, we will
+assign the abstract language generated by that signature and its
+distinguished type. To all the other nodes, we will assign the result of
+mapping their parent's language by the lexicon linking the two. By
+homomorphism of the lexicons, we can find distinguished types for all
+the non-root nodes. They are simply the images of their parents'
+distinguished types by the connecting lexicons. For all of the nodes, it
+is true that every term in the language assigned to the node has the
+node's distinguished type.
 
 We can confirm that the above graphical model of defining ACGs is
-coherent. First, we can show that our interpretation is sound and so
-that any directed tree adorned with signatures on the nodes, compatible
-lexicons on the edges and a distinguished type on the root can only
-define ACG languages. This is trivial, since the language corresponding
-to the root is an ACG abstract language and the class of ACG languages
-is closed under transformation by a lexicon and therefore all of the
-languages corresponding to the descendants of the root are ACG languages
-as well. Second, we can try and show some kind of completeness. However,
-there is not much formalization done to describe a ``system of ACGs that
-share common signatures'' (something that our generalization will
+coherent.
+
+First, we can show that our interpretation is sound meaning that any
+arborescence adorned with signatures on the nodes, compatible lexicons
+on the edges and a distinguished type on the root can only define ACG
+languages (abstract or object languages). This is trivial, since the
+language corresponding to the root is an ACG abstract language and the
+class of ACG languages is closed under transformation by a lexicon and
+therefore all of the languages corresponding to the descendants of the
+root are ACG languages as well.
+
+Second, we can try and show some kind of completeness. However, there is
+not much formalization done to describe a ``system of ACGs that share
+some common signatures'' (something that our generalization will
 fix). We can nevertheless show that it is possible to express any ACG
@@
 $\mathcal{L}$. Then the two nodes of our tree define both the abstract
-and the object language defined by $\mathcal{G}$.
-
-We have now given some formal account of how we can read language
-definitions in diagrams. Still, we have kept our analysis to the tree
-diagrams used in current ACG literature, where every node has at most
-one parent. To make the generalization more visible, we will switch from
-an algebraic manner of presentation, in which we stated that the
-language of a node is equal to an image of its parent's language, to a
-calculus-based manner of presentation, in which we will state that the
-language of a node is the set of all terms that have an antecedent in
-the parent's language. Besides this property, there is also the fact
-that every term in a node's language has the node's distinguished
-type. This property is trivially true for nodes with parents because of
-the homomorphic property of lexicons. However, it is crucial for the
-root node for which this is the single defining property of its
-language. To sum up, we have that for every node of a tree diagram, its
-language is the set of all terms that have the node's distinguished type
-and that have an antecedent in the parent's language, if there is
-any. This rule is already general enough to handle the case of zero or
-one parents, we can easily generalize it to handle an arbitrary count:
-\emph{the language of a node is the set of all terms that have the
-  node's distinguished type and that have antecedents in the languages
-  of all of the node's parents}. If we transpose this back to the
-algebraic style of presentation, we say that the language of a node is
-the intersection of the set of terms having the node's distinguished
-type and the intersection of all its parents' languages.
-
-TODO: In the previous paragraph, add a notice saying that the terms must
-be also built on the signature of the node.
-
-TODO: Use the correct term arborescence in the above paragraphs!
-% http://en.wikipedia.org/wiki/Arborescence_(graph_theory)
-
-We will now turn to a formal definition of this generalized graph-based
-formalism and discuss some its values and properties.
-
-\subsection{Graphical Abstract Categorial Grammars}
+and the object language defined by $\mathcal{G}$. Similarly, we could
+argue that the languages we aim to define using ACG diagrams (such as
+the ones on Figures \ref{fig:acg-comp-modes} and
+\ref{fig:acg-comp-patterns}) are covered by our assignment of languages
+to nodes of the diagrams.
+
+We have now given a formal account of how we can read language
+definitions from ACG diagrams. Still, we have kept our analysis to the
+arborescent diagrams used in current ACG literature, where every node
+has at most one parent. We will now proceed to generalize it to cases
+where nodes have more than just one parent.
+
+To make the generalization more visible, we will switch from an
+algebraic manner of presentation to a relational one. Before, we stated
+that the language of a node is the image of its parent's language.
+
+$$
+Language(u) = Lexicon(Language(Parent(u)))
+$$
+
+Now, we will restate this as saying that the language of a node is the
+set of terms that have an antecedent in the parent's language.
+
+$$
+Language(u) = \{x \in \Lambda(Signature(u)) \mid \exists a \in
+Language(Parent(u)), \; Lexicon(a) = x\}
+$$
+
+This definition works for non-root nodes of an ACG diagram. In the root
+case, we have to constrain the language by stating that that the terms
+of the language must have the root's distinguished type. This is a
+property that is also true in the non-root cases due to the homomorphic
+properties of lexicons. If we wanted to unify the definitions of the
+root language and the non-root languages, we would have to check that a
+term of the language has the distinguished type and that it has an
+antecedent in the parent, if it has any parent. A relational statement
+of this unification could look like this.
+
+\begin{align*}
+Language(u) = \{&x \in \Lambda(Signature(u)) \mid Type(x) = DistinguishedType(u) \\
+&\land \forall p \in Parents(u), \; \exists a \in Language(p), \; Lexicon_{(p,u)}(a) = x\}
+\end{align*}
+
+If the node has no parents, no constraint is enforced. If it does have
+some, we enforce the antecedent constraint. This shows us a natural
+generalization to the case where nodes can have an arbitrary number of
+parents. If we transpose this back to the algebraic style of
+presentation, we get the following expression.
+
+$$
+Language(u) = DistinguishedlyTyped(u) \quad \cap \bigcap_{p \in Parents(u)} Lexicon_{(p,u)}(Language(p))
+$$
+
+\section{Graphical Abstract Categorial Grammars}
+
+In this section, we distill the reasoning about generalizing ACG
+diagrams to graphs in a definition of an extension for ACGs. We then
+examine its key properties and implications.
 
@@
 $\Sigma$ and $S$ are labelings assigning signatures and distinguished
-types, respectively, to the vertices in $V$ and ${\mathcal{L}}$ is a
-labeling assigning lexicons to the edges in $E$. Furthermore, a
+types, respectively, to the vertices $V(G)$ and ${\mathcal{L}}$ is a
+labeling assigning lexicons to the edges $E(G)$. Furthermore, a
 well-formed graphical ACG satisfies the following conditions:
@@
 We let the nodes in a graphical ACG $\mathcal{G} = \mathopen{<} G,
-\Sigma, S, \mathcal{L} \mathclose{>}$ define two kinds of
-languages. The \emph{intrinsic languages} $\mathcal{I}_{\mathcal{G}}$,
-which are constrained only by the type signature described in the node
-itself, and the \emph{extrinsic languages} $\mathcal{E}_{\mathcal{G}}$,
-which are also constrained by type systems in more abstract nodes of the
+\Sigma, S, \mathcal{L} \mathclose{>}$ define two kinds of languages. The
+\emph{intrinsic languages} $\mathcal{I}_{\mathcal{G}}$, which are
+constrained only by the type signature described in the node itself, and
+the \emph{extrinsic languages} $\mathcal{E}_{\mathcal{G}}$, which are
+also constrained by type signatures in more abstract nodes of the
 graph. These two notions correspond to those of abstract languages and
@@
 \mathcal{I}_{\mathcal{G}}(v) = \{t \in \Lambda(\Sigma_v)
-\mid\ \vdash_{\Sigma_u} t : S_v\}
+\mid\ \vdash_{\Sigma_v} t : S_v\}
 $$
@@
 
-\subsubsection{On the Expressivity of Graphical ACGs}
+\subsection{On the Expressivity of Graphical ACGs}
 
@@
   lexicons. Let us consider any graphical ACG $\mathcal{G} =
-  \mathopen{<} G, \Sigma, S, \mathcal{L} \mathclose{>}$ and any
-  topological ordering $v_1, \ldots, v_n$ of the nodes of $G$ (an
+  \mathopen{<} G, \Sigma, S, \mathcal{L} \mathclose{>}$ and some
+  topological ordering $v_1, \ldots, v_n$ of its nodes $V(G)$ (an
   ordering such that more abstract nodes always precede less abstract
-  ones). Before we start, we will just note that the all the intrinsic
-  languages defined by $v_i$ are abstract languages which are a special
-  case of object languages.
+  ones). Before we start, we will remark that all the intrinsic
+  languages $\mathcal{I}_{\mathcal{G}}(v_i)$ are object languages, since
+  intrinsic languages are abstract languages which are in turn just a
+  special case of object languages.
+
+  We proceed by induction on the topological ordering:
 
   \begin{itemize}
-    \item For $v_1$, the case is trivial since its extrinsic language is
-      just its intrinsic language as the node $v_1$ has no predecessors
-      in $G$.
-    \item For any other node $v_n$, we look at the definition of an
-      extrinsic language and remark that the only operators are
-      intersection and application of a lexicon with the operands being
-      the intrinsic language of $v_n$ and the extrinsic languages of
-      more abstract nodes. By induction hypothesis, we have that all of
-      the operands have been constructed from object languages using
-      intersections and transformations by lexicons and therefore even
-      the extrinsic language defined by $v_n$ has this property.
+    \item For $v_1$, the case is trivial. $v_1$ has no predecessors and
+      so $\mathcal{E}_{\mathcal{G}}(v_1) =
+      \mathcal{I}_{\mathcal{G}}(v_1)$, which is an object language.
+    \item For any other node $v_n$, we look at the definition of
+      $\mathcal{E}_{\mathcal{G}}(v_n)$ and remark that the only
+      operators are intersection and application of a lexicon with the
+      operands being $\mathcal{I}_{\mathcal{G}}(v_n)$ (which is an
+      object language) and the extrinsic languages of its predecessors
+      (which can be constructed from object languages using
+      intersections and lexicons thanks to the induction
+      hypothesis). Therefore even $\mathcal{E}_{\mathcal{G}}(v_n)$ can
+      be constructed from object languages using only intersections and
+      lexicons.
   \end{itemize}
@@
   \begin{itemize}
-    \item Let $L$ be an object language defined by the ACG $\mathopen{<}
-      \Sigma_A, \Sigma_O, \mathcal{L}, S\mathclose{>}$. We take the
-      graph $G$ with $V(G) = \{A, O\}$ and $E(G) = \{(A,O)\}$, we label
-      the nodes $A$ and $O$ with $\Sigma_A$ and $\Sigma_O$ as signatures
-      and $S$ and $\mathcal{L}(S)$ as distinguished types
-      respectively. Finally, we label the edge $(A,O)$ with
-      $\mathcal{L}$ and we have a graphical ACG in which the node $O$
-      defines the extrinsic language $L$.
+    \item Let $L$ be an object language defined by the ACG $\mathcal{G}
+      = \mathopen{<} \Sigma_A, \Sigma_O, \mathcal{L},
+      S\mathclose{>}$. We take the graph $G$ with $V(G) = \{A, O\}$ and
+      $E(G) = \{(A,O)\}$, we label the nodes $A$ and $O$ with $\Sigma_A$
+      and $\Sigma_O$ as signatures and $S$ and $\mathcal{L}(S)$ as
+      distinguished types respectively. Finally, we label the edge
+      $(A,O)$ with $\mathcal{L}$ and we get a graphical ACG $G'$ in
+      which $\mathcal{E}_{\mathcal{G'}}(O) = \mathcal{O}(G) = L$.
 
     \item Let $L_1$ and $L_2$ be two extrinsic languages defined by the
-      nodes $u_1$ and $u_2$ respectively belonging to two graphical
-      ACGs. We can construct a new graphical ACG by taking the union of
-      the two ACGs (union of graphs (union of vertices and edges) and
-      unions of the labelings).\footnote{Assuming, without loss of
-        generality, that the sets of vertices of the two graphs are
-        disjoint.} We then add a new node $v$ which will define the
-      language $L_1 \cap L_2$.
+      nodes $u_1$ and $u_2$ in some graphical ACGs $\mathcal{G}_1$ and
+      $\mathcal{G}_2$, respectively. We can construct a new graphical
+      ACG $\mathcal{G}'$ by taking the union of the two ACGs (union of
+      graphs (union of vertices and edges) and unions of the
+      labelings).\footnote{Assuming, without loss of generality, that
+        the sets of vertices of the two graphs are disjoint.} We will
+      then add a new node $v$ such that $\mathcal{E}_{\mathcal{G}'}(v) =
+      L_1 \cap L_2$.
 
       The signature labeling $v$ will be the union of the two signatures
-      labeling $u_1$ and $u_2$ such that the intrinsic language of $v$
-      will be a superset of the intrinsic languages, and by extension
-      the extrinsic languages, of both $u_1$ and $u_2$. Either of the
-      two distinguished types that label $u_1$ or $u_2$ can be used to
-      label $v$ (if they are actually different, then the intersection
-      is trivially empty). Finally, we will add two edges leading from
+      labeling $u_1$ and $u_2$, ensuring that $\forall u \in \{u_1,
+      u_2\}$, $\mathcal{I}_{\mathcal{G}'}(v) \supseteq
+      \mathcal{I}_{\mathcal{G}'}(u) \supseteq
+      \mathcal{E}_{\mathcal{G}'}(u)$. We can use either of the two
+      distinguished types that label $u_1$ or $u_2$ to label $v$ (if
+      they are different, then the intersection $L_1 \cap L_2$ is
+      trivially empty). Finally, we will add two edges leading from
       $u_1$ and $u_2$ to $v$, both labeled with the identity lexicon. We
-      will now show that the extrinsic language of $v$ in the newly
-      formed graphical ACG $\mathcal{G}'$ is $L_1 \cap L_2$.
+      will now show that $\mathcal{E}_{\mathcal{G}'}(v) = L_1 \cap L_2$.
 
@@
       The simplifications use the facts that: the two new lexicons are
-      identities, that the intrinsic language of $v$ is a superset of
-      the extrinsic languages of $u_1$ and $u_2$, that the extrinsic
-      languages of $u_1$ and $u_2$ are the same in the new graphical ACG
-      as in the two original graphical ACGs and finally that $L_1$ and
-      $L_2$ are the extrinsic languages defined by $u_1$ and $u_2$.
-
-  \item Let $L \subset \Lambda(\Sigma_A)$ be an extrinsic language and
-    $\mathcal{L} : \Sigma_A \to \Sigma_O$ a lexicon from its signature
-    $\Sigma_A$ to the signature $\Sigma_O$. We will show that the
-    language $\mathcal{L}(L) \subset \Lambda(\Sigma_O)$ is also an
-    extrinsic language. Let $u$ be the node that defines the extrinsic
-    language $L$ in some graphical ACG $\mathcal{G}$. We construct a new
-    graphical ACG by adding a new vertex $v$ to $\mathcal{G}$ labeled
+      identities, that $\mathcal{I}_{\mathcal{G}'}(v) \supseteq
+      \mathcal{E}_{\mathcal{G}'}(u_1) \cap
+      \mathcal{E}_{\mathcal{G}'}(u_2)$, that
+      $\mathcal{E}_{\mathcal{G}'}(u_1) = \mathcal{E}_{\mathcal{G}}(u_1)$
+      and $\mathcal{E}_{\mathcal{G}'}(u_2) =
+      \mathcal{E}_{\mathcal{G}}(u_2)$ and finally that
+      $\mathcal{E}_{\mathcal{G}}(u_1) = L_1$ and
+      $\mathcal{E}_{\mathcal{G}}(u_2) = L_2$.
+
+  \item Let $L \subseteq \Lambda(\Sigma_O)$ be a language which is the
+    result of mapping an extrinsic language $L' \subseteq
+    \Lambda(\Sigma_A)$ through the lexicon $\mathcal{L} : \Sigma_A \to
+    \Sigma_O$. We will show that $L$ is also an extrinsic language. Let
+    $u$ be the node that defines the extrinsic language $L$ in some
+    graphical ACG $\mathcal{G}$. We construct a new graphical ACG
+    $\mathcal{G}'$ by adding a new vertex $v$ to $\mathcal{G}$ labeled
     with the signature $\Sigma_O$ and the distinguished type
@@
     $u$. To this graph, we conjoin the edge $(u,v)$ and label it with
-    $\mathcal{L}$. Now we verify that the extrinsic language of $v$ is
-    the language $\mathcal{L}(L).$
+    $\mathcal{L}$. Now we verify that $\mathcal{E}_{\mathcal{G}'}(v) =
+    L$.
 
@@
       &= \mathcal{L}(\mathcal{E}_{\mathcal{G}}(u)) \\
-      &= \mathcal{L}(L)
-    \end{align*}
+      &= \mathcal{L}(L') \\
+      &= L
+    \end{align*}
+
+    The assumptions used here are in turn: that
+    $\mathcal{I}_{\mathcal{G}'}(v) \supseteq
+    \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u))$, that
+    $\mathcal{E}_{\mathcal{G}'}(u) = \mathcal{E}_{\mathcal{G}}(u)$, that
+    $\mathcal{E}_{\mathcal{G}}(u) = L'$ and that $\mathcal{L}(L') = L$.
   \end{itemize}
@@
 \end{corollary}
+
+We conclude this subsection with a few words to motivate the need we
+feel for this (possibly) more expressive formalism. We believe that in
+order to manage the development of wide-coverage grammars, the grammars
+need to be kept as simple as possible, which means, among other things,
+that independent constraints should be implemented separately and not
+complected together (we saw the difficulties caused by this already on a
+microscopic fragment in section~\ref{sec:together}). This means that if
+we were to define, e.g., a partial grammar in which we would only
+enforce agreement and another grammar in which we would only enforce the
+proper use of relative pronouns, we would like the language in which
+both agreement and proper use of relative pronouns are respected to be
+expressible in our formalism, and on top of that, we would like this
+language to be easily expressed in terms of the two grammars for the
+individual constraints. We now have a requirement on the formalism in
+which we would like to use: we want a formalism where the set of all the
+``useful'' definable languages is closed on intersection. If the object
+languages of ACGs are indeed closed on intersection, then our extension
+is just a convenience for expressing intersection more easily. However,
+if they are not, we believe our extension adds a critical level of
+expressivity needed for simple and modular wide-coverage grammars.
+
+\subsection{On the Decidability of Graphical ACGs}
+\label{ssec:graphical-decidability}
+
+We turn to the problem of parsing with graphical ACGs. Parsing with ACGs
+is a problem on the limits of decidability. When we consider ACGs whose
+abstract signatures only contain types of order 2 or less, parsing is
+possible in polynomial type for both linear \cite{salvati2005problemes}
+and non-linear \cite{kanazawa2007parsing} type systems (such as the one
+we introduced in~\ref{sec:acg}). If the order of the abstract signature
+exceeds 2, the membership problem for linear ACGs becomes decidable if
+and only if the satisfiability problem in the multiplicative exponential
+fragment of linear logic is decidable \cite{de2004vector}, which is an
+open problem. On the other hand, by extending the type system to include
+dependent types, parsing can become undecidable too since in such a
+system type checking itself is undecidable.
+
+Our intent here is to show that by extending the formalism we have not
+made the problem of parsing more difficult, at least in terms of
+decidability if not in terms of complexity. Let us therefore imagine a
+graphical ACG where every edge (two signatures, distinguished types and
+a lexicon) is an ACG for which there is an effective parsing algorithm
+and type checking is decidable in all the signatures of the graphical
+ACG. Then we can prove the existence of a simple effective parsing
+algorithm for any of the extrinsic languages defined in the graphical
+ACG.
+
+We will be testing whether the term $o$ belongs to
+$\mathcal{E}_{\mathcal{G}}(v)$. Our algorithm first checks whether $o$
+belongs to $\mathcal{I}_{\mathcal{G}}(v)$ by checking its type. If it
+does, it proceeds to verify that it is also in
+$\mathcal{E}_{\mathcal{G}}(v)$ by finding antecedents in all of its
+abstract predecessors $u_i$, if there are any. We can use the
+presupposed parsing algorithm for the ACG $\mathopen{<} \Sigma_{u_i},
+\Sigma_v, \mathcal{L}_{(u_i,v)}, S_{u_i}\mathclose{>}$ to find all the
+abstract terms $\alpha_{i,j}$ having type $S_{u_i}$ and being
+antecedents to $o$ w.r.t. the lexicon $\mathcal{L}_{(u_i,v)}$. Such
+terms belong to $\mathcal{I}_{\mathcal{G}}(u_i)$, but in order to prove
+that $o$ is in $\mathcal{E}_{\mathcal{G}}(v)$, we need to find an
+antecedent that belongs to $\mathcal{E}_{\mathcal{G}}(u_i)$. What we do
+is we recursively apply our parsing algorithm to all of the
+antecedent-candidates $\{\alpha_{i,j} \mid \forall j\}$ until we find
+one which belongs $\mathcal{E}_{\mathcal{G}}(u_i)$. Since the graph is
+acyclic, we are guaranteed that this recursive traversal will
+terminate. We repeat this for all the predecessors $u_i$ and if we
+succeed to find antecedents in all of their extrinsic languages, we have
+confirmed the membership of $o$ in $\mathcal{E}_{\mathcal{G}}(v)$ and
+also built its abstract antecedent(s).
+
+\subsection{On Grammar Engineering with Graphical ACGs}
+\label{ssec:graphical-engineering}
+
+Having covered some of the formal properties of graphical ACGs, we will
+get back to building grammars. Armed with this formalism, we can solve
+the puzzle of incorporating the three constraints of sections
+\ref{sec:negation}, \ref{sec:extraction} and \ref{sec:agreement}. Our
+graph $G$ will consist of 6 nodes, $V(G) = \{Neg, Ext, Agr, Synt, Str,
+Sem\}$ with edges going from $Neg$, $Ext$ and $Agr$ to $Synt$ and from
+$Synt$ to $Str$ and $Sem$ (see Figure~\ref{fig:puzzle-solution}). The
+signatures and distinguished types for the first three vertices are
+exactly those that have been presented in the first three sections of
+this chapter. The lexicons for the edges connecting these constraint
+nodes to $Synt$ all behave according to the scheme below:
+
+\begin{align*}
+  \mathcal{L}_{(constraint,Synt)}(X_{constant_{index}}) &= C_{constant} \\
+  \mathcal{L}_{(constraint,Synt)}(TYPE\_FEATURES)  &= TYPE
+\end{align*}
+
+The $\{Synt, Str, Sem\}$ subgraph is identical to the pair of ACGs
+described in \ref{sec:acg} except for some additions that need to be
+made so that the fragment covers the wordforms that we have introduced
+in the course of discussing the three constraints (prepositions,
+relative pronouns, the negative particle $ne$\ldots). However, the types
+of these new constants will not be tied to any of those constraints.
+
+\begin{align*}
+  C_{de} &: NP \limp N \limp N \\
+  C_{qui} &: (NP \limp S) \limp N \limp N \\
+  C_{dit\ que} &: NP \limp S \limp S \\
+  C_{dorment} &: NP \limp S \\
+  C_{ne_{tv}} &: (NP \limp NP \limp S) \limp (NP \limp NP \limp S) \\
+  C_{ne_{iv}} &: (NP \limp S) \limp (NP \limp S) \\
+  C_{ne_{cv}} &: (NP \limp S \limp S) \limp (NP \limp S \limp S)
+\end{align*}
+
+This approach lets us describe the high-level combinatorics of language
+that are of interest to the categorial grammarian in a manner which is
+idiomatic in academic literature (using the simple and focused types
+$N$, $NP$ and $S$). At the same time, we can actually use this textbook
+grammar as a basis for a more realistic grammar which handles
+constraints such as agreement and extraction islands. Furthermore, these
+additional constraints are also defined in a manner that has its
+precedent in existing research. When the authors of
+\cite{pogodalla2012controlling} decided to put the constraints of
+extraction under a microscope and try to encode them in ACGs, they
+defined their constraints in exactly the same way as our architecture
+does. This graphical architecture lets us write the syntactic kernel and
+the additional constraints independently and using two different styles
+that have both been shown useful by their own merits, which we see as
+indicative of enabling good modularity.
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{diagrams/puzzle-solution.pdf}
+    \caption{{\label{fig:puzzle-solution} The graph of a grammar
+        enforcing all of the constraints discussed in this chapter.}}
+  \end{subfigure}
+  \qquad
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{diagrams/both-patterns.pdf}
+    \caption{{\label{fig:both-patterns} Combining the two ACG patterns
+        from~\ref{ssec:acg-patterns} in a single graphical ACG.}}
+  \end{subfigure}
+  \caption{Practical architectures for graphical ACGs.}
+\end{figure}
+
+In the above approach, we have defined a language by writing a grammar
+which provides us with useful structures but which does not enforce
+complete correctness. We have then refined this grammar by tacking on
+constraints ``on the side''. This approach is generally useful and can
+be seen, e.g., in \cite{van2004concepts} to define the formal
+programming language kernel on which a formal semantics is
+established. In our case, we are interested in having intelligible
+syntactic structures too as we would like to compute their semantic
+representations. The syntax-semantics interface already presents a
+significant challenge and having less-than-ideal syntactic structures to
+start with does not make formalizing it any easier.
+
+We will show one more persuasive scenario whose goal is to show the
+importance of graphical ACGs for having a clean syntax-semantics
+interface. In \ref{ssec:acg-patterns}, we have seen an ACG pattern that
+put semantic ambiguities into an abstract signature that is higher than
+syntax. But there was also another pattern, that of adding a new
+abstract signature to constrain syntactic terms. We can combine both of
+them in graphical ACGs by superposing their two graphs to get the
+structure we can see on Figure~\ref{fig:both-patterns}.
+
+If we try to reap the benefits of both patterns and try to replicate the
+same in classical ACGs (arborescent graphs), we hit a snag: we cannot
+have both the $Constr$ and $SyntSem$ nodes as the direct predecessors of
+the $Syntax$ node. We can make $Constr$ be the predecessor of $Syntax$,
+but then $SyntSem$ will be connecting $Constr$ and $Sem$ and the
+syntax-semantics interface will have to deal with the types which were
+needed to implement syntactic constraints and have no place in a
+syntax-semantics interface. If we try it the other way, it is not much
+better. $Constr$ will end up having to work with the more complicated
+type system of $SyntSem$ instead of being able to talk only about the
+$Syntax$ it is interested in.
+
+By putting the constraining signature in a part of the graph unreachable
+from the syntax-semantics transducer, we can shield the syntax-semantics
+interface from the gnarly implementation details of the rules of
+syntax. This in turn gives us a guideline as to what information should
+be expressed in the syntactic signature itself and which should be
+hidden in constraints: the syntactic signature should only contain
+information relevant to the syntax-semantics interface.
+
+This might then raise a question whether the chosen syntactic signature
+will end up being suitable for expressing all of the syntactic
+constraints. However, this can be resolved by deepening the constraint
+subgraph and adding a ``compatibility layer'' in the form of a new
+signature which provides different syntactic structures and maps them to
+the canonical ones. Constraints can then opt in to controlling the
+syntactic structures offered by this new signature.
+
+\subsection{On Alternative Interpretations of Graphical ACGs}
+
+Our definitions of intrinsic and extrinsic languages have consequences
+which might be counter-intuitive and not correspond to the grammarian's
+intent. We will describe what these consequences are and propose an
+alternative notion of a graphical ACG language which aims to reconcile
+them.
+
+The first phenomenon that deserves pointing out occurs whenever two
+paths from nodes $u_1$ and $u_2$ converge on a single node $v$. If we
+then take a term $\alpha$ from $\mathcal{E}_{\mathcal{G}}(u_1)$, the
+term $\beta = \mathcal{L}_{(u_1,v)}(\alpha)$ belongs to
+$\mathcal{I}_{\mathcal{G}}(v)$ but not necessarily to
+$\mathcal{E}_{\mathcal{G}}(v)$ since it might be the case that there is
+no antecedent for $\beta$ in $\mathcal{E}_{\mathcal{G}}(u_2)$. In other
+words, mapping a member of $\mathcal{E}_{\mathcal{G}}(u)$ using the
+lexicon $\mathcal{L}_{(u,v)}$ does not guarantee that the result belongs
+to $\mathcal{E}_{\mathcal{G}}(v)$, which is a different behavior from
+that of object languages in ACGs (or that of extrinsic languages in
+arborescent graphical ACGs).
+
+This highlights some limitations of our notion of an extrinsic
+language. Consider once more the graphical ACG of
+Figure~\ref{fig:both-patterns}. We have more than one path converging to
+$Syntax$. We will consider a term $\alpha$ of
+$\mathcal{E}_{\mathcal{G}}(Sem)$. $\alpha$ has an antecedent in
+$\mathcal{E}_{\mathcal{G}}(SyntSem)$ which has a descendant $\beta$ in
+$\mathcal{I}_{\mathcal{G}}(Syntax)$. However, because of the constraint
+$Constr$ on $Syntax$, $\beta$ can be a syntactic term outside of
+$\mathcal{E}_{\mathcal{G}}(Syntax)$ and yielding an ungrammatical
+sentence. The notion of an extrinsic language is insufficient to express
+the set of meanings from $\Lambda(\Sigma_{Sem})$ which are expressible
+only by grammatical sentences. Defining languages like this would be of
+particular interest in cases like the one in
+Figure~\ref{fig:transducer-grammar} where we have a chain of signatures
+representing the different levels of linguistic description whose terms
+are in many-to-many relationships expressed through transducers. In such
+a scenario, it would be desirable to be able to express the set of
+descriptions belonging to one level, e.g. morphology, that correspond to
+valid descriptions on all the other levels.
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{diagrams/transducer-grammar.pdf}
+    \caption{{\label{fig:transducer-grammar} A graphical ACG connecting
+        (constrained) levels of linguistic description through
+        transducers.}}
+  \end{subfigure}
+  \qquad
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{diagrams/diamond-grammar.pdf}
+    \caption{{\label{fig:diamond-grammar} A graphical ACG demonstrating
+        two different paths between two nodes.}}
+  \end{subfigure}
+  \caption{Graphical ACGs posing challenges to extrinsic languages.}
+\end{figure}
+
+Compared to ACG object languages (or to extrinsic languages of
+arborescent graphical ACGs), we lose another guarantee. In
+arborescences, the central role of the root abstract-most signature
+guarantees us that whenever a term belongs to an extrinsic language, we
+can find terms in all the other extrinsic languages that share the same
+abstract-most term. On the other hand, in non-arborescent graphical
+ACGs, having a term in an extrinsic language of some node $v$ only
+guarantees the presence of corresponding terms in the extrinsic
+languages of its ancestors and the intrinsic languages of their
+descendants.
+
+The second puzzling scenario is the diamond shape on
+Figure~\ref{fig:diamond-grammar}. If we consider a term $\delta$ of
+$\mathcal{E}_{\mathcal{G}}(Bottom)$, we know that there exist
+antecedents $\beta$ and $\gamma$ in $\mathcal{E}_{\mathcal{G}}(Left)$
+and $\mathcal{E}_{\mathcal{G}}(Right)$. And while there definitely exist
+antecedents $\alpha$ and $\alpha'$ to both $\beta$ and $\gamma$ in
+$\mathcal{E}_{\mathcal{G}}(Top)$, these antecedents can be different,
+meaning there does not have to be a single $\alpha \in
+\mathcal{E}_{\mathcal{G}}(Top)$ which yields $\beta \in
+\mathcal{E}_{\mathcal{G}}(Left)$ and $\gamma \in
+\mathcal{E}_{\mathcal{G}}(Right)$ such that they both yield
+$\delta$. This means that the result of parsing a term in some
+$\mathcal{E}_{\mathcal{G}}(v)$ is a collection of abstract terms with
+one term not just per every node $u$ that is more abstract than $v$ but
+per every different path leading to $v$. Parsing a term $\delta \in
+\mathcal{E}_{\mathcal{G}}(Bottom)$ in our diamond grammar would thus
+produce four antecedents $\beta$, $\gamma$, $\alpha$ and $\alpha'$
+corresponding to the four paths $[Left,Bottom]$, $[Right,Bottom]$,
+$[Top,Left,Bottom]$ and $[Top,Right,Bottom]$, respectively. We contrast
+this again with the case of classical ACGs where the result of parsing
+can always be represented by associating a term to every node of the
+arborescence.\footnote{One could also say that the result of parsing in
+  an ACG is just the antecedent in the abstract-most signature and that
+  the other terms can then be easily computed by application of
+  lexicons. Similarly, we could say that the result of parsing in a
+  general graphical ACG is a collection of abstract terms for every
+  \emph{maximal} path leading to the node defining the language in
+  question. What might be more intuitive, though, is to have some theory
+  of graphical ACGs where the result of parsing is a collection of
+  antecedents in all the abstract-most nodes.}
+
+We will propose a new way of assigning languages to nodes in a graphical
+ACG that will conserve the intuitive properties of ACGs we talked about
+above. We do not need to modify the definition of a graphical ACG in
+order to do so since the definition is only concerned with the structure
+of the graph and its decoration with ACG paraphernalia, not the
+languages defined. We will define a third kind of languages based on
+graphical ACGs.
+
+Let $\mathcal{G} = \mathopen{<} G, \Sigma, S, \mathcal{L} \mathclose{>}$
+be a graphical ACG. We define the \emph{pangraphical language} of node
+$u$, $\mathcal{P}_{\mathcal{G}}(u)$, to be the set of terms $t$ such
+that there exists a labeling of nodes with terms $T$ satisfying the
+following conditions:
+
+\begin{itemize}
+  \item $T_u = t$.
+  \item For all $v \in V(G)$, $\vdash_{\Sigma_v} T_v : S_v$.
+  \item For all $(v,w) \in E(G)$, $\mathcal{L}_{(v,w)}(T_v) = T_w$.
+\end{itemize}
+
+It is easy to see that the newly introduced pangraphical languages solve
+all of the counter-intuitive quirks of object languages that we
+discussed in this subsection.
+
+\begin{itemize}
+\item Applying the lexicon $\mathcal{L}_{(u,v)}$ to an element of
+  $\mathcal{P}_{\mathcal{G}}(u)$ always yields an element of
+  $\mathcal{P}_{\mathcal{G}}(v)$, since the labeling of nodes with terms
+  $T$ that proved the former term's membership in the former language
+  also proves the latter term's membership in the latter language.
+
+  This definition allows us to directly establish the language of
+  meanings which are expressible only by grammatical sentences in
+  Figure~\ref{fig:both-patterns} and the language of morphological terms
+  that have valid corresponding representations on all the other levels
+  of linguistic description in Figure~\ref{fig:transducer-grammar}.
+\item The diamond grammar of Figure~\ref{fig:diamond-grammar} is also no
+  longer problematic since the labeling of nodes with terms guarantees
+  that there is always a single term per node that can generate the term
+  being parsed.
+\end{itemize}
+
+In the previous section, we took a stab at formalizing ACG diagrams and
+finding the simplest extension capable of handling non-arborescent
+graphs. The notion of extrinsic languages that we arrived at allowed us
+to define a nicely delimited set of languages, the object languages of
+ACGs with intersection. Discrepancies in the behavior of extrinsic
+languages in arborescent and non-arborescent graphs led us to
+pangraphical languages, which behave more in line with our
+intuitions. However, the expressivity of pangraphical languages,
+especially w.r.t. to extrinsic languages, still remains to be
+explored. Due to limitations of both space and time, we will cover the
+interesting properties of pangraphical languages only briefly and
+partially here.
+
+\begin{description}
+  \item[Extrinsic languages can be defined similarly to pangraphical
+    languages] \hfill \\
+
+    In the definition of $\mathcal{P}_{\mathcal{G}}(u)$, we replace the
+    labeling of nodes with terms by a labeling of all the paths leading
+    to $u$.
+
+  \item[$\mathcal{P}_{\mathcal{G}}(u) \subseteq
+    \mathcal{E}_{\mathcal{G}}(u) \subseteq
+    \mathcal{I}_{\mathcal{G}}(u)$] \hfill \\
+
+    Given a labeling of nodes with terms such that it satisfies the
+    conditions in the definition of a pangraphical language, we can show
+    by induction on the topological ordering of the graph that every
+    term assigned by the labeling belongs to the extrinsic language of
+    the same node. Therefore, $\alpha \in \mathcal{P}_{\mathcal{G}}(u)
+    \implies \alpha \in \mathcal{E}_{\mathcal{G}}(u)$.
+
+    $\mathcal{E}_{\mathcal{G}}(u) \subseteq
+    \mathcal{I}_{\mathcal{G}}(u)$ comes straight from the definition of
+    $\mathcal{E}_{\mathcal{G}}(u)$.
+
+    We have thus came up with a series of successively more constrained
+    languages: intrinsic languages are constrained by one node,
+    extrinsic languages are also constrained by its ancestors and
+    pangraphical languages are constrained by the entire graph.
+
+  \item[On arborescences, $\mathcal{P}_{\mathcal{G}}(u) =
+    \mathcal{E}_{\mathcal{G}}(u)$.] \hfill \\
+
+    On an arborescence, being a member of an extrinsic language means
+    there exists an antecedent in the abstract-most language of the
+    arborescence. By virtue of the one-parent property of arborescences,
+    we can simply apply all the lexicons in the arborescence to this
+    abstract-most antecedent to find a labeling of nodes with terms
+    which proves that the member of the extrinsic language is also a
+    member of the pangraphical language.
+
+    This confirms that both extrinsic languages and pangraphical
+    languages are reasonable generalizations of ACG diagrams since their
+    behaviors on arborescences are consistent with each other and with
+    that of ACG object languages.
+
+  \item[The set of extrinsic languages is included in the set of
+    pangraphical languages.] \hfill \\
+
+    Let us have some node $u$ in a graphical ACG $\mathcal{G}$. We build
+    a graphical ACG $\mathcal{G}'$ where every node corresponds to a
+    path in $\mathcal{G}$ which ends at $u$ and two nodes are connected
+    with an edge whenever one path is an immediate extension of the
+    other. We then have $\mathcal{P}_{\mathcal{G'}}([u]) =
+    \mathcal{E}_{\mathcal{G}}(u)$.
+
+  \item[Is the set of pangraphical languages included in the set of
+    extrinsic languages?] \hfill \\
+
+    That is an interesting question indeed, whose resolution is left as
+    an exercise for the reader.
+
+  \item[Pangraphical languages are not ``less decidable'' than object
+    languages.] \hfill \\
+
+    Presuming that type checking is decidable for all the signatures of
+    the graphical ACGs and that we can parse effectively for all the ACG
+    edges in the graph, we can reuse the parsing algorithm we introduced
+    in \ref{ssec:graphical-decidability}. We apply this algorithm to
+    enumerate all the possible extrinsic parses in a language
+    $\mathcal{E}_{\mathcal{G}}(v)$ and keep only those where for each
+    ancestor $u$ we have the same abstract term for all paths from $u$
+    to $v$.
+\end{description}
+
+\subsection{On the Practical Consequences of Graphical ACGs}
+
+In this final subsection, we will enumerate some of the practical
+benefits and challenges of using graphical ACGs for implementing
+grammar-based systems.
+
+\begin{itemize}
+\item The lexicons that go from our constraint signatures ($Neg$, $Ext$
+  and $Agr$ in subsection \ref{ssec:graphical-engineering}) are all
+  trivial in the sense they map constants to constants. Finding an
+  antecedent to a term is then just a question of finding an antecedent
+  for every constant in the term (i.e. ``tagging'' the constants of the
+  term with antecedent constants). This means that there is an effective
+  parsing algorithm for verifying the constraints. Furthermore, this
+  algorithm could also be made efficient by introducing filtering
+  techniques \cite{guillaume2008toolchain} and statistical supertagging
+  \cite{moot2012categorial}.
+
+\item Constraints can introduce type extensions in their abstract
+  signatures and the added complexity of these extensions is then
+  contained within the constraint. Furthermore, the constraint signature
+  can be treated as a ``black box'' that verifies a constraint and in
+  the context of an implementation, it can be replaced with a
+  special-purpose hand-written procedure that does the same more
+  efficiently.
+
+\item The distributed nature of graphical ACGs has its own practical
+  applications. Parsing can be made more robust by being able to parse
+  sentences which violate some of the constraints. Parsing an
+  ungrammatical sentence and identifying the constraints it violates can
+  also serve as a basis for a syntax checker.
+
+\item If we decide to define all of our syntactic constraints in
+  separate signatures, we will end up with a lot of signatures which
+  look exactly like the syntactic signature but with some small
+  refinements (see the signatures in sections \ref{sec:negation},
+  \ref{sec:extraction} and \ref{sec:agreement}). Thus we will desire
+  some concise way of formulating these refinements so that we will not
+  need to replicate the entire syntactic signature for every
+  constraint.
+
+  We hope that the class of useful refinements is small and that these
+  constraint signatures could be systematically constructed from the
+  syntactic signature by applying a series of refining
+  transformations. This is an interesting engineering challenge awaiting
+  anyone willing to design a truly wide-coverage graphical ACG using
+  constraints.
+\end{itemize}

### 2013-06-17 abstract.tex
Last night shift

\begin{abstract}
We present work whose ultimate goal is the creation of wide-coverage
abstract categorial grammars (ACGs) that could be used to automatically
build discourse-level representations. In our work, we advance towards
that goal by laying down the foundations necessary for building
wide-coverage ACGs.

We first examine existing language resources, in particular the Frigram
interaction grammar and its lexicon Frilex, and assess their utility to
building a wide-coverage ACG. We then present our implementation of the
ACG machinery which allows us to experiment with grammars lexicalized by
Frilex. Finally, we consider the challenge of integrating the treatment
of disparate linguistic constraints in a single ACG and propose a
generalization of the formalism: graphical abstract categorial
grammars. The report concludes with an exploration of some of the formal
properties of graphical ACGs.
\end{abstract}

### 2013-06-17 acg.tex
Last night shift

@@
 
-\subsection{Higher-order signatures}
+\subsection{Higher-Order Signatures}
 \label{ssec:sig}
@@
   \caption{\label{fig:type-judgments} Type judgment schemas of the
-    well-typed lambda terms $\Lambda(\Sigma)$ for a signature $\Sigma =
-    \mathopen{<}A, C, \tau\mathclose{>}$}
+    well-typed lambda terms $\Lambda(\Sigma)$ built on a signature
+    $\Sigma = \mathopen{<}A, C, \tau\mathclose{>}$.}
 \end{figure}
@@
 
-\subsection{Example signature}
+\subsection{Example Signature}
 \label{ssec:example-sig}
@@
 type, $A_{String} = \{\textsc{String}\}$. Our constants will be the
-wordforms of our example fragment ($every$, $some$, $man$, $woman$,
-$loves$), the empty string $\epsilon$ and the string concatention
-operator, $+$. $\tau_{String}$ assigns the type $\textsc{String}$ to all
-the wordforms and to $\epsilon$ and the type $\textsc{String} \limp
-\textsc{String} \limp \textsc{String}$\footnote{As a convention,
-  whenever we omit parentheses in a type, we presume the $\limp$ and
-  $\to$ type constructors always bind from the right, meaning that $a
-  \to b \to c$ should be read as $a \to (b \to c)$.}
-to the concatenation operator $+$.
+wordforms of our example fragment (\emph{every}, \emph{some},
+\emph{man}, \emph{woman}, \emph{loves}), the empty string $\epsilon$ and
+the string concatenation operator, $+$. $\tau_{String}$ assigns the type
+$\textsc{String}$ to all the wordforms and to $\epsilon$ and the type
+$\textsc{String} \limp \textsc{String} \limp
+\textsc{String}$\footnote{As a convention, whenever we omit parentheses
+  in a type, we presume the $\limp$ and $\to$ type constructors always
+  bind from the right, meaning that $a \to b \to c$ should be read as $a
+  \to (b \to c)$.}  to the concatenation operator $+$.
 
@@
   convention. The unadorned way of writing this term would be $t_1 =
-  (+\ some)\ woman$. Firstly, we will allow ourselves to drop the
+  (+\ some)\ woman$. First, we will allow ourselves to drop the
   parentheses, meaning that any string of expressions $f\ x\ y$ should
-  be read as $(f\ x)\ y$. Secondly, we will admit a specific infix
+  be read as $(f\ x)\ y$. Second, we will admit a specific infix
   notation for some constants which denote functions of two
   arguments. This will then let us write $x\ +\ y$ instead of
-  $+\ x\ y$.} denotes the concatenation of $some$ and $woman$, which is
-the phrase $some\ woman$; its type is $\textsc{String}$. A term such as
-$t_2 = \lambda^{\circ} x. x\ +\ man$ denotes a function that appends the
-string $man$ to its argument; its type is $\textsc{String} \limp
-\textsc{String}$. If we restrict the set $\Lambda(\Sigma_{String})$ only
-to terms having the type $\textsc{String}$, we get the set of
-expressions which denote strings (a set that contains terms like $t_1$
-but not $t_2$).
+  $+\ x\ y$.} denotes the concatenation of \emph{some} and \emph{woman},
+which is the phrase \emph{some\ woman}; its type is $\textsc{String}$. A
+term such as $t_2 = \lambda^{\circ} x. x\ +\ man$ denotes a function
+that appends the string \emph{man} to its argument; its type is
+$\textsc{String} \limp \textsc{String}$. If we restrict the set
+$\Lambda(\Sigma_{String})$ only to terms having the distinguished type
+$\textsc{String}$, we get the set of expressions which denote strings (a
+set that contains terms like $t_1$ but not $t_2$).
 
@@
 
-The idea of a signature is coupled with that of a \textbf{lexicon},
+The idea of a signature is coupled with the one of \textbf{lexicon},
 which is a mapping between two different signatures (mapping the
 constants of one into well-typed terms of the other). Formally speaking,
-a lexicon $\mathcal{L}$ from a signature $\Sigma_1 = \mathopen{<}A_1, C_1,
-\tau_1\mathclose{>}$ (which we call the abstract signature) to a
+a lexicon $\mathcal{L}$ from a signature $\Sigma_1 = \mathopen{<}A_1,
+C_1, \tau_1\mathclose{>}$ (which we call the abstract signature) to a
 signature $\Sigma_2 = \mathopen{<}A_2, C_2, \tau_2\mathclose{>}$ (which
@@
 such that:
+
 \begin{itemize}
@@
   abstract-level types with the object-level types that they realized
-  in.
+  as.
 \item $F$ and $G$ are compatible, meaning that for any $c \in C_1$, we
@@
 
+See Figure~\ref{fig:single-acg} for the kind of diagram we will be using
+to represent signatures and lexicons that will form our grammars.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{diagrams/single-acg.pdf}
+    \caption{{\label{fig:single-acg} A lexicon with its abstract
+        and object signatures..}}
+  \end{subfigure}
+  \qquad
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{diagrams/double-acg.pdf}
+    \caption{{\label{fig:double-acg} Two lexicons sharing the same
+        abstract signature.}}
+  \end{subfigure}
+  \caption{Diagrams of abstract categorial grammars.}
+\end{figure}
+
 An abstract categorial grammar for us will then be just a collection of
@@
 (logical propositions) and an abstract signature which is connected to
-both of the object signatures via lexicons. Parsing is then just a
-matter of inverting the surface lexicon to get the abstract term and
-then applying to it the logical lexicon. Generation is symmetric, we
-simply invert the logical lexicon and apply the surface lexicon.
+both of the object signatures via lexicons (as we can see on
+Figure~\ref{fig:double-acg}). Parsing is then just a matter of inverting
+the surface lexicon to get the abstract term and then applying to it the
+logical lexicon. Generation is symmetric, we simply invert the logical
+lexicon and apply the surface lexicon.
 
-\subsection{Example lexicon}
+\subsection{Example Lexicon}
 \label{ssec:example-lex}
@@
 \cite{montague1973proper}. The atomic types $A_{Synt}$ will consist of
-the types $np$, $n$ and $s$. Our constants, $C_{Synt} = \{\synt{every},
+the types $NP$, $N$ and $S$. Our constants, $C_{Synt} = \{\synt{every},
 \synt{some}, \synt{love}, \synt{man}, \synt{woman}\}$, will have the
 following types, as predicted by $\tau_{Synt}$:
-\begin{itemize}
-\item $\tau_{Synt}(\synt{every}) = \tau_{Synt}(\synt{some}) = n \limp
-  ((np \limp s) \limp s)$
-\item $\tau_{Synt}(\synt{love}) = np \limp np \limp s$
-\item $\tau_{Synt}(\synt{man}) = \tau_{Synt}(\synt{woman}) = n$
-\end{itemize}
 
-Let us explore some terms from $\Lambda(\Sigma_{Synt})$. $t_3 =
-\synt{some}\ \synt{woman}$ has type $(np \limp s) \limp s$, which is the
-type that serves to describe quantified noun phrases in our
+\begin{align*}
+\tau_{Synt}(\synt{every}) &= N \limp ((NP \limp S) \limp S) \\
+\tau_{Synt}(\synt{some}) &= N \limp ((NP \limp S) \limp S) \\
+\tau_{Synt}(\synt{love}) &= NP \limp NP \limp S \\
+\tau_{Synt}(\synt{man}) &= N \\
+\tau_{Synt}(\synt{woman}) &= N
+\end{align*}
+
+Let us explore some terms from $\Lambda(\Sigma_{Synt})$. Let $t_3 =
+\synt{some}\ \synt{woman}$, $t_3$ has type $(NP \limp S) \limp S$, which
+is the type that serves to describe quantified noun phrases in our
 example. Terms of this type expect a verb phrase (or some other
-predicate) of type $np \limp s$, and can yield a sentence of type $s$.
+predicate) of type $NP \limp S$, and can yield a sentence of type $S$.
 
-Now we will see how this approach can handle transitive verbs. In $t_4 =
-\lambda^{\circ} x.\ (\synt{some}\ \synt{woman})\ (\lambda^{\circ}
-y.\ \synt{love}\ x\ y)$, which is a term of type $np \limp s$
-representing the verb phrase \emph{loves some woman}, we first take the
-expression $\synt{love}\ x\ y$ (\emph{$x$ loves $y$}, type $s$) and we
+Now we will see how this approach can handle transitive verbs. Let us
+consider the term $t_4 = \lambda^{\circ}
+x.\ (\synt{some}\ \synt{woman})\ (\lambda^{\circ}
+y.\ \synt{love}\ x\ y)$, which has type $NP \limp S$ and represents the
+verb phrase \emph{loves some woman}. To build it, we first take the
+expression $\synt{love}\ x\ y$ (\emph{$x$ loves $y$}, type $S$) and we
 abstract over $y$ to get the predicate $\lambda^{\circ}
-y.\ \synt{love}\ x\ y$ (\emph{is loved by $x$}, type $np \limp s$). By
+y.\ \synt{love}\ x\ y$ (\emph{is loved by $x$}, type $NP \limp S$). By
 applying the quantified noun phrase $t_3$ to this term, we get
 $(\synt{some}\ \synt{woman})\ (\lambda^{\circ} y.\ \synt{love}\ x\ y)$
-(\emph{$x$ loves some woman}, type $s$). Finally, by abstracting over
+(\emph{$x$ loves some woman}, type $S$). Finally, by abstracting over
 $x$, we get the verb phrase $t_4$.
@@
 y.\ \synt{love}\ x\ y))$ is the result of applying the quantified noun
-phrase $\synt{every}\ \synt{man}$ (of type $(np \limp s) \limp s$) to
-$t_4$. What we get is a term of type $s$ which describes the sentence
+phrase $\synt{every}\ \synt{man}$ (of type $(NP \limp S) \limp S$) to
+$t_4$. What we get is a term of type $S$ which describes the sentence
 \emph{every man loves some woman}. Note that as before, we can restrict
 the set $\Lambda(\Sigma_{Synt})$ to terms having the distinguished type
-$s$ and we will arrive at the set of terms which describe sentences
+$S$ and we will arrive at the set of terms which describe sentences
 (this set will include $t_5$, but not $t_3$ or $t_4$).
@@
 In the example above, I have been presuming some kind of implied
-connection between the terms of $\Lambda(\Sigma_{Synt})$ and phrases of
-some natural language, strings. Indeed, while we have defined a system
-of structures which describe the quantification and predicate structures
-of a microscopic fragment of English and a system for talking about
-strings in concatenation, we have not given any explicit link between
-the two. It is at this moment that we will introduce a lexicon to link
-these two levels of description.
+connection between the terms of $\Lambda(\Sigma_{Synt})$ and English
+phrases. Indeed, while we have defined a system of structures which
+describe the quantification and predicate structures of a microscopic
+fragment of English and a system for talking about strings in
+concatenation, we have not given any explicit link between the two. It
+is at this moment that we will introduce a lexicon to link these two
+levels of description.
 
@@
 So, how does our lexicon $\mathcal{L}_{syntax}$ look like? It will map
-all the atomic types of $\Sigma_{Synt}$ into $\textsc{String}$, which we
-can write as $\mathcal{L}_{syntax}(n) = \mathcal{L}_{syntax}(np) =
-\mathcal{L}_{syntax}(s) = \textsc{String}$\footnote{Whenever we talk
-  about some lexicon, we will use the lexicon itself, which is formally
-  a pair of mappings $\mathopen{<}F, G\mathclose{>}$, to mean either
-  $F$, $G$ or their homomorphic extensions $\hat{F}$ and $\hat{G}$
-  depending on whether the argument is an atomic type, a constant, a
-  complex type or a term respectively.}. From that, we can use the
-homomorphic extension of our type mapping to arrive at the object-level
-interpretations of complex types, such as the type of predicates, $np
-\limp s$, which becomes a unary string function, $\textsc{String} \limp
-\textsc{String}$, or the type of quantified noun phrases, $(np \limp s)
-\limp s$, which becomes a higher-order string function,
-$(\textsc{String} \limp \textsc{String}) \limp \textsc{String}$. With
-the types out of the way, we can give the interpretation of the
-individual constants of $\Sigma_{Synt}$:
-\begin{itemize}
-\item $\mathcal{L}_{syntax}(\synt{every}) = \lambda^{\circ} x
-  R.\ R\ (every\ +\ x)$
-\item $\mathcal{L}_{syntax}(\synt{some}) = \lambda^{\circ} x
-  R.\ R\ (some\ +\ x)$
-\item $\mathcal{L}_{syntax}(\synt{love}) = \lambda^{\circ} x
-  y.\ x\ +\ loves\ +\ y$
-\item $\mathcal{L}_{syntax}(\synt{man}) = man$
-\item $\mathcal{L}_{syntax}(\synt{woman}) = woman$
-\end{itemize}
+all the atomic types of $\Sigma_{Synt}$ into
+$\textsc{String}$.\footnote{Whenever we talk about some lexicon, we will
+  use the lexicon itself, which is formally a pair of mappings
+  $\mathopen{<}F, G\mathclose{>}$, to mean either $F$, $G$ or their
+  homomorphic extensions $\hat{F}$ and $\hat{G}$ depending on whether
+  the argument is an atomic type, a constant, a complex type or a term
+  respectively.}
+
+\begin{align*}
+\mathcal{L}_{syntax}(N) &= \textsc{String} \\
+\mathcal{L}_{syntax}(NP) &= \textsc{String} \\
+\mathcal{L}_{syntax}(S) &= \textsc{String}
+\end{align*}
+
+Given the above mappings, we can use the homomorphic properties of our
+type mapping to arrive at the object-level interpretations of complex
+abstract types, such as the type of predicates, $NP \limp S$, which
+becomes a unary string function, $\textsc{String} \limp
+\textsc{String}$, or the type of quantified noun phrases, $(NP \limp S)
+\limp S$, which becomes a higher-order string function,
+$(\textsc{String} \limp \textsc{String}) \limp \textsc{String}$.
+
+With the types out of the way, we can give the interpretation of the
+individual constants of $\Sigma_{Synt}$.
+
+\begin{align*}
+\mathcal{L}_{syntax}(\synt{every}) &= \lambda^{\circ} x R.\ R\ (every\ +\ x) \\
+\mathcal{L}_{syntax}(\synt{some}) &= \lambda^{\circ} x R.\ R\ (some\ +\ x) \\
+\mathcal{L}_{syntax}(\synt{love}) &= \lambda^{\circ} x y.\ x\ +\ loves\ +\ y \\
+\mathcal{L}_{syntax}(\synt{man}) &= man \\
+\mathcal{L}_{syntax}(\synt{woman}) &= woman \\
+\end{align*}
 
@@
 look at what they look like after applying the lexicon to them.
+
 \begin{itemize}
@@
   language} generated by our signature $\Sigma_{Synt}$ and the
-distinguished type $s$, that is, elements of $\Lambda(\Sigma_{Synt})$
-having type $s$. Then we can define the \emph{object language}, which is
+distinguished type $S$, that is, elements of $\Lambda(\Sigma_{Synt})$
+having type $S$. Then we can define the \emph{object language}, which is
 the image of the abstract language when transformed using the
-lexicon. The object language contains strings, terms from
+lexicon. The object language contains strings, i.e. terms from
 $\Lambda(\Sigma_{String})$ having the type $\textsc{String}$. However,
@@
 
-\subsection{Example semantic lexicon}
+\subsection{Example Semantic Lexicon}
 \label{ssec:acg-sem}
@@
 One perk of working with abstract categorial grammars is that one
-abstract term can be interpreted multiple ways, using different
+abstract term can be interpreted in multiple ways, using different
 lexicons. In the previous chapter, we considered the abstract terms of
 the $\Sigma_{Synt}$ signature and interpreted them as computations on
-strings. Now we will take the same abstract terms, but try to intepret
-them as computations on entities and truth values.
+strings. Now we will take the same abstract terms, but try to interpret
+them as computations on entities and truth values. A diagram of this
+setup can be seen on the Figure~\ref{fig:double-acg}.
 
@@
 entities, $e$, and a type for truth values, $t$. The constants $C_{Sem}$
-will contain atomic predicates from our fragment, $\textbf{man}$,
+will contain the atomic predicates from our fragment, $\textbf{man}$,
 $\textbf{woman}$ and $\textbf{love}$, logical connectives, $\Rightarrow$
 and $\land$, and quantifiers, $\forall$ and $\exists$. Their types as
-given by $\tau_{Sem}$ are:
-\begin{itemize}
-\item $\tau_{Sem}(\textbf{man}) = \tau_{Sem}(\textbf{woman}) = e \limp
-  t$
-\item $\tau_{Sem}(\textbf{love}) = e \limp e \limp t$
-\item $\tau_{Sem}(\Rightarrow) = \tau_{Sem}(\land) = t \limp t \limp t$
-\item $\tau_{Sem}(\forall) = \tau_{Sem}(\exists) = (e \to t) \limp t$
-\end{itemize}
+given by $\tau_{Sem}$ are as follows:\footnote{The non-linear
+  implication ($\to$) in the types of $\forall$ and $\exists$, $(e \to
+  t) \limp t$, means that the variable ``introduced'' by the quantifier
+  can be used multiple times in the predicate, which has type $e \to
+  t$. We will rely on this property later when defining the meanings of
+  determiners.}
+
+\begin{align*}
+\tau_{Sem}(\textbf{man}) &= e \limp t \\
+\tau_{Sem}(\textbf{woman}) &= e \limp t \\
+\tau_{Sem}(\textbf{love}) &= e \limp e \limp t \\
+\tau_{Sem}(\Rightarrow) &= t \limp t \limp t \\
+\tau_{Sem}(\land) &= t \limp t \limp t \\
+\tau_{Sem}(\forall) &= (e \to t) \limp t \\
+\tau_{Sem}(\exists) &= (e \to t) \limp t
+\end{align*}
 
@@
 $\Sigma_{Synt}$, we will give a lexicon from the latter to the
-former. First, we give the type mappings. $\mathcal{L}_{sem}(n) = e
+former. First, we give the type mappings. $\mathcal{L}_{sem}(N) = e
 \limp t$ meaning that nouns will yield boolean functions of one argument
 (unary \emph{predicates} in functional programming
-terminology). $\mathcal{L}_{sem}(np) = e$, simple noun phrases will be
+terminology). $\mathcal{L}_{sem}(NP) = e$, simple noun phrases will be
 mapped into terms which yield some entity. Finally,
-$\mathcal{L}_{sem}(s) = t$, sentences will be assigned terms which
+$\mathcal{L}_{sem}(S) = t$, sentences will be assigned terms which
 compute their truthiness. Let us see how this plan is achieved in the
-$\mathcal{L}_{sem}$ lexicon.
-\begin{itemize}
-\item $\mathcal{L}_{sem}(\synt{every}) = \lambda^{\circ} P Q.\ \forall
-  x.\ ((P\ x) \Rightarrow (Q\ x))$\footnote{We will be using $\forall
-    x.\ M$ as a shortcut for $\forall\ (\lambda x.\ M)$ and $\exists
-    x.\ M$ for $\exists\ (\lambda x.\ M)$.}
-\item $\mathcal{L}_{sem}(\synt{some}) = \lambda^{\circ} P Q.\ \exists
-  x.\ ((P\ x) \land (Q\ x))$
-\item $\mathcal{L}_{sem}(\synt{love}) = \lambda^{\circ} s o. (\textbf{love}\ s\ o)$
-\item $\mathcal{L}_{sem}(\synt{man}) = \lambda^{\circ} x. (\textbf{man}\ x)$
-\item $\mathcal{L}_{sem}(\synt{woman}) = \lambda^{\circ} x. (\textbf{woman}\ x)$
-\end{itemize}
+$\mathcal{L}_{sem}$ lexicon.\footnote{We will be using $\forall x.\ M$
+  as a shortcut for $\forall\ (\lambda x.\ M)$ and $\exists x.\ M$ for
+  $\exists\ (\lambda x.\ M)$.}
+
+\begin{align*}
+\mathcal{L}_{sem}(\synt{every}) &= \lambda^{\circ} P Q.\ \forall x.\ ((P\ x) \Rightarrow (Q\ x)) \\
+\mathcal{L}_{sem}(\synt{some}) &= \lambda^{\circ} P Q.\ \exists x.\ ((P\ x) \land (Q\ x)) \\
+\mathcal{L}_{sem}(\synt{love}) &= \lambda^{\circ} so.\ (\textbf{love}\ s\ o) \\
+\mathcal{L}_{sem}(\synt{man}) &= \lambda^{\circ} x.\ (\textbf{man}\ x) \\
+\mathcal{L}_{sem}(\synt{woman}) &= \lambda^{\circ} x.\ (\textbf{woman}\ x)
+\end{align*}
 
@@
 interpreted as $\mathcal{L}_{sem}(t_3) =_{\beta} \lambda^{\circ}
-Q.\ \exists x.\ ((\textbf{woman}\ x) \land (Q\ x))$ having type $(e \limp
-t) \limp t$. It is a function that expects some predicate function and
-tests whether that predicate holds at least for some woman (some entity
+Q.\ \exists x.\ ((\textbf{woman}\ x) \land (Q\ x))$ having type $(e
+\limp t) \limp t$. It is a function that expects some predicate and
+tests whether that predicate holds for at least some woman (some entity
 for which the $\textbf{woman}$ predicate holds).
 
-In $t_4 = \lambda^{\circ}
+When we consider $t_4 = \lambda^{\circ}
 x.\ (\synt{some}\ \synt{woman})\ (\lambda^{\circ}
@@
 $\Lambda(\Sigma_{Synt})$ for which $\mathcal{L}_{syntax}(t_5) =_{\beta}
-every\ +\ man\ +\ loves\ +\ some\ +\ woman$. We could just as well take $t_6
-= (\synt{some}\ \synt{woman})\ (\lambda^{\circ}
+every\ +\ man\ +\ loves\ +\ some\ +\ woman$. We could just as well take
+$t_6 = (\synt{some}\ \synt{woman})\ (\lambda^{\circ}
 y.\ (\synt{every}\ \synt{man})\ (\lambda^{\circ}
@@
 
-$$
-\mathcal{L}_{sem}(t_5) =_{\beta} \forall x.\ ((\textbf{man}\ x)
+\begin{align*}
+\mathcal{L}_{sem}(t_5) &=_{\beta} \forall x.\ ((\textbf{man}\ x)
 \Rightarrow (\exists y.\ ((\textbf{woman}\ y) \land
-(\textbf{love}\ x\ y))))
-$$
-$$
-\mathcal{L}_{sem}(t_6) =_{\beta} \exists y.\ ((\textbf{woman}\ y)
+(\textbf{love}\ x\ y)))) \\
+\mathcal{L}_{sem}(t_6) &=_{\beta} \exists y.\ ((\textbf{woman}\ y)
 \land (\forall x.\ ((\textbf{man}\ x) \Rightarrow
 (\textbf{love}\ x\ y))))
-$$
+\end{align*}
 
@@
 object term. This is basically trying to invert the lexicon function,
-modulo $\beta$-equivalence. It is therefore no surprise then that we can
-find multiple abstract terms which all map to the same string term but
-which can be mapped to distinct terms using the semantic lexicon. It is
-this mechanism that enables abstract categorial grammars to handle
+modulo $\beta$-equivalence. It is therefore not surprising then that we
+can find multiple abstract terms which all map to the same string term
+but which can be mapped to distinct terms using the semantic lexicon. It
+is this mechanism that enables abstract categorial grammars to handle
 ambiguity. When we reverse the scenario and go from semantic

### 2013-06-17 conclusion.tex
Last night shift

--- 2013-06-13 the-last-part.tex
@@
-\chapter{The Last Part}
+\chapter{In Conclusion}
+\label{chap:conclusion}
 
-In this chapter, we will attempt to incorporate accounts of several
-linguistic phenomena in a single framework, highlight the modularity
-issues this brings up and propose some solutions to them.
+In this chapter, we first discuss some the consequences of using G-ACGs
+for writing grammars. We then summarize the contributions that we have
+put forward and suggest problems we believe to be interesting.
 
-\section{Negation}
+\section{On the Practical Consequences of Graphical ACGs}
+\label{sec:practical}
 
-We will start with an account of the paired grammatical words mechanism
-for negation in French. Recall the EPTDs we have seen
-in~\ref{ssec:frigram}, repeated here on Figure~\ref{fig:ig-neg-rep}.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.25]{ig-neg.pdf}
-  \caption{\label{fig:ig-neg-rep} Frigram EPTDs for a negative
-    determiner and the paired grammatical word $ne$.}
-\end{figure}
-
-At a basic level, $aucun$ functions like any other determiner and $ne$
-as any verb modifier. However, we want to encode the constraint that
-whenever a verb is modified by $ne$, the modified verb will demand that
-one of its arguments contains a negative determiner. Since we need to
-make a distinction between phrases that contain negative determiners and
-those that do not, we will need to provide different types for both (two
-terms with the same type are indistinguishable w.r.t. syntactic
-composition in a type-logical grammar such as ACGs). Our type for
-$aucun$ will thus end up being $N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S)
-\limp S)$ instead of the usual $N \limp ((NP \limp S) \limp
-S)$.\footnote{Strictly following the EPTD for $aucun$ would lead us to
-  give it a type $DET\_NEG{=}F$ and then give the two types
-  $DET\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S)$ and $DET\_NEG{=}T
-  \limp ((NP\_NEG{=}T \limp S) \limp S)$ to nouns. In this
-  demonstration, we prefer to align our examples with the treatment
-  which is customary to ACG research.} The type for $le$ will need to be
-able handle cases where its argument noun either already contains a
-negative determiner or not.
-
-$$
-N_{aucun} : N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S) \limp S)
-$$
-$$
-N_{le_1} : N\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S)
-$$
-$$
-N_{le_2} : N\_NEG{=}T \limp ((NP\_NEG{=}T \limp S) \limp S)
-$$
-
-The type of $ne$ will be more verbose, as it will consume a verb and
-transform its valency so that it demands that exactly one of its $NP$
-arguments contains a negative determiner.
-
-$$
-N_{ne_{tv_1}} : (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}F \limp S)
-$$
-$$
-N_{ne_{tv_2}} : (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}F \limp NP\_NEG{=}T \limp S)
-$$
-
-Furthermore, we could add the case when both the subject and the object
-contain negative determiners, such as in~(\ref{ex:aucun2}). This case is
-currently not covered by Frigram (though it could be), but is considered
-grammatical by French speakers. Finally, in a complete grammar, $ne$
-would have to provide types capable of transforming all the other
-syntactic valencies.
-
-\begin{exe}
-  \ex \label{ex:aucun2} Aucune fourmi n'aime aucun tatou.
-\end{exe}
-
-$$
-N_{ne_{tv_3}} : (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}T \limp S)
-$$
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.25]{parse-aucun.pdf}
-  \caption{\label{fig:parse-aucun} One of the parse trees assigned to
-    sentence (\ref{ex:aucun-obj}) by Frigram/Leopar.}
-\end{figure}
-
-However, this is not the only change we have to effect on our grammar in
-order to properly handle the paired grammatical words for
-negation. Consider the example sentence (\ref{ex:aucun-obj}) and the
-parse tree assigned to it by Frigram on Figure~\ref{fig:parse-aucun}. It
-is not enough that the embedded noun phrase $aucun\ tatou$ has a type
-that tells us it contains a negative determiner. We need this bit of
-internal information also at the level of the enclosing noun phrase
-$l'odeur\ d'aucun\ tatou$ in which the former noun phrase is embedded
-since this is a property of the enclosing noun phrase that is crucial to
-its syntactic combinatorics. This presupposes the existence of a
-mechanism for propagating this kind of information up the parse tree. In
-the case of our ACG, this means that all the functions which modify
-phrases must be aware whether or not their argument contains a negative
-determiner and to convey this information in its result type as
-well. Moreover, since the negative determiner can also come from a
-complement of a prepositional phrase, the types of prepositions will
-need to take into account the presence of negative determiners in both
-their complements and the phrases they modify.
-
-$$
-N_{que_1} : (NP\_NEG{=}F \limp S) \limp N\_NEG{=}F \limp N\_NEG{=}F
-$$
-$$
-N_{que_2} : (NP\_NEG{=}F \limp S) \limp N\_NEG{=}T \limp N\_NEG{=}T
-$$
-$$
-N_{de_1} : NP\_NEG{=}F \limp N\_NEG{=}F \limp N\_NEG{=}F
-$$
-$$
-N_{de_2} : NP\_NEG{=}F \limp N\_NEG{=}T \limp N\_NEG{=}T
-$$
-$$
-N_{de_3} : NP\_NEG{=}T \limp N\_NEG{=}F \limp N\_NEG{=}T
-$$
-
-If we suppose the presence of the appropriate lexical items typed as in
-the examples of~\ref{sec:acg}, we have a type system in which we can
-type the syntactic terms corresponding to the sentences
-(\ref{ex:good-neg-double}), (\ref{ex:good-neg-embed}) and
-(\ref{ex:good-neg-rel}) while making it impossible to type the
-structures corresponding to sentences (\ref{ex:bad-neg-noneg}) and
-({\ref{ex:bad-neg-rel}}).
-
-In sentence (\ref{ex:bad-neg-noneg}), we have no constant
-$N_{ne_{tv_?}}$ for $ne$ which would yield a verb capable of consuming
-two noun phrases that contain no negative determiners.
-
-In sentence (\ref{ex:bad-neg-rel}), there are multiple problems with
-typing. First, $N_{chasse}$ requires its first argument to be of type
-$NP\_NEG{=}F$, a noun phrase with no negative determiner. This means
-that $\lambda^{\circ} x.\ N_{chasse}\ x\ y$ has type $NP\_NEG{=}F \limp
-S$ which makes it an invalid argument for the function
-$N_{aucun}\ N_{loup}$, which has type $(NP\_NEG{=}T \limp S) \limp
-S$. Even if we were able to assign it a type compatible with
-$N_{que_1}$, the final relative clause would map $N_{tatou}$ to another
-term of type $N\_NEG{=}F$, which has no ``free'' negative
-determiners. $N_{le_1}$ will map this to a term typed $(NP\_NEG{=}F
-\limp S) \limp S$. Finally, this term is not capable of taking
-$(\lambda^{\circ} x.\ N_{ne_{iv}}\ N_{court}\ x)$ as its argument, since
-its type is $NP\_NEG{=}T \limp S$.
-
-\begin{exe}
-  \ex \label{ex:good-neg-double} Aucune fourmi n'aime aucun tatou. \\
-      $(N_{aucune}\ N_{fourmi})\ (\lambda^{\circ} x.\ (N_{aucun}\ N_{tatou})\ (\lambda^{\circ} y.\ N_{ne_{tv_3}}\ N_{aime}\ x\ y))$
-  \ex \label{ex:good-neg-embed} Jean n'aime l'odeur d'aucun tatou. \\
-      $(N_{le_2}\ (N_{de_3}\ (N_{aucun}\ N_{tatou})\ N_{odeur}))\ (\lambda^{\circ} y.\ N_{ne_{tv_2}}\ N_{aime}\ N_{Jean}\ y)$
-  \ex \label{ex:good-neg-rel} Le tatou qu'aucun loup ne chasse court. \\
-      $(N_{le_1}\ (N_{que_1}\ (\lambda^{\circ} y.\ (N_{aucun}\ N_{loup})\ (\lambda^{\circ} x.\ N_{ne_{tv_1}}\ N_{chasse}\ x\ y))\ N_{tatou}))\ (\lambda^{\circ} x.\ N_{court}\ x)$
-  \ex * \label{ex:bad-neg-noneg} Jean n'aime le tatou. \\
-      * $(N_{le_1}\ N_{tatou})\ (\lambda^{\circ} y.\ N_{ne_{tv_?}}\ N_{aime}\ N_{Jean}\ y)$
-  \ex * \label{ex:bad-neg-rel} Le tatou qu'aucun loup chasse ne court. \\
-      * $(N_{le_1}\ (N_{que}\ (\lambda^{\circ} y.\ (N_{aucun}\ N_{loup})\ (\lambda^{\circ} x.\ N_{chasse}\ x\ y))\ N_{tatou}))\ (\lambda^{\circ} x.\ N_{ne_{iv}}\ N_{court}\ x)$
-\end{exe}
-
-There are two things that we would like to highlight about this
-grammatical treatment. First off, we had to refine our types to make
-them convey more than one piece of information in a single type
-(e.g. something is a noun phrase and at the same time it is something
-that contains a negative determiner). Second, a property that we wish to
-express in the type of a term may be due to one of its subconstituents
-and it is therefore necessary to encode the propagation of this property
-up the parse tree in the types of all the intervening operators. In our
-case, it meant that handling negative determiners and the paired
-grammatical word for negation required us to not only define and type
-constants for these two categories, but to also change the types of noun
-modifiers ($que\ ...$) and prepositions ($de$).
-
-\section{Extraction}
-\label{sec:extraction}
-
-We will now turn to another phenomenon, extraction in relative clauses,
-focus on one of its properties, the fact that subjects can be extracted
-using $qui$ only from the relative clause itself whereas objects can be
-extracted using $que$ from embedded clauses contained therein, and
-present an implementation of this constraint by way of types in an
-ACG. The solution that we will show was published in
-\cite{pogodalla2012controlling}. Our presentation is slightly adapted to
-conform to the patterns seen in previous examples and it does not use
-dependent types as they are not essential to this specific constraint.
-
-The key idea in the treatment presented below is to give a special type
-to extracted constituents. We will thus have two types for noun phrases:
-$NP\_VAR{=}T$, which stands for ``empty'' noun phrases introduced by
-extraction, and $NP\_VAR{=}F$, which stands for the other noun phrases,
-i.e.  proper names and nouns with determiners. The grammar gives no
-constant capable of constructing a value of type $NP\_VAR{=}T$. The only
-way to obtain a term with this type is to abstract over a variable
-having the type. Then, the types of relative pronouns end up being
-$(NP\_VAR{=}T \limp S) \limp N \limp N$, ensuring that the incomplete
-clause given as the first argument is a $\lambda$-abstraction whose
-argument has type $NP\_VAR{=}T$.
-
-Once we have this kind of information in the type system, we can start
-distinguishing clauses containing rooted extraction, where a subject or
-object is extracted directly from the relative clause, embedded
-extraction, where an object is extracted from a clause embedded inside
-the relative clause, or no extraction. Types of functions which produce
-values of type $S$ will be sensitive to the presence of extracted
-variables in their noun phrase and clausal arguments.
-
-$$
-E_{dort_1} : NP\_VAR{=}F \limp S\_EXT{=}NO
-$$
-$$
-E_{dort_2} : NP\_VAR{=}T \limp S\_EXT{=}ROOT
-$$
-$$
-E_{aime_1} : NP\_VAR{=}F \limp NP\_VAR{=}F \limp S\_EXT{=}NO
-$$
-$$
-E_{aime_2} : NP\_VAR{=}T \limp NP\_VAR{=}F \limp S\_EXT{=}ROOT
-$$
-$$
-E_{aime_3} : NP\_VAR{=}F \limp NP\_VAR{=}T \limp S\_EXT{=}ROOT
-$$
-$$
-E_{dit\ que_1} : NP\_VAR{=}F \limp S\_EXT{=}NO \limp S\_EXT{=}NO
-$$
-$$
-E_{dit\ que_2} : NP\_VAR{=}T \limp S\_EXT{=}NO \limp S\_EXT{=}ROOT
-$$
-$$
-E_{dit\ que_3} : NP\_VAR{=}F \limp S\_EXT{=}ROOT \limp S\_EXT{=}EMB
-$$
-$$
-E_{dit\ que_4} : NP\_VAR{=}F \limp S\_EXT{=}EMB \limp S\_EXT{=}EMB
-$$
-
-Now all of the infrastructure is in place for us to express the type of
-the relative pronouns $qui$ and $que$.
-
-$$
-E_{qui} : (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N
-$$
-$$
-E_{que_1} : (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N
-$$
-$$
-E_{que_2} : (NP\_VAR{=}T \limp S\_EXT{=}EMB) \limp N \limp N
-$$
-
-The above types ensure that the abstracted variables corresponding to
-traces of extracted constituents will be given the type $NP\_VAR{=}T$,
-that $qui$ can only be used to extract constituents from root positions
-while $que$ allows for extraction from both root and embedded
-positions.\footnote{NB: The constraint that $qui$ can only extract
-  subjects and that $que$ can only extract objects is not enforced
-  here. Likewise, this fragment does not handle multiple extraction.}
-
-We also note that the splitting of the clause type $S$ into three finer
-types $S\_EXT={NO}$, $S\_EXT{=}ROOT$ and $S\_EXT{=}EMB$ means that types
-that work with $S$ regardless of its extraction status will have to
-provide alternatives for all the possible variants.
-
-$$
-E_{le_1} : N \limp ((NP\_VAR{=}F \limp S\_EXT{=}NO) \limp S\_EXT{=}NO)
-$$
-$$
-E_{le_2} : N \limp ((NP\_VAR{=}F \limp S\_EXT{=}ROOT) \limp S\_EXT{=}ROOT)
-$$
-$$
-E_{le_3} : N \limp ((NP\_VAR{=}F \limp S\_EXT{=}EMB) \limp S\_EXT{=}EMB)
-$$
-
-While we have introduced the notion of an $NP\_VAR{=}T$ for describing
-traces, we can cover another element of our running fragment, the
-preposition $de$, and enforce the constraint that $qui$/$que$ cannot
-extract prepositional complements by providing only a single type for
-$de$ that accepts only unmoved noun phrases as complements.
-
-$$
-E_{de} : NP\_VAR{=}F \limp N \limp N
-$$
-
-Given this type system, we can assign the type $S\_EXT{=}NO$ to the term
-expressing the syntactic structure of (\ref{ex:good-ext}) but we cannot
-a find a typable term for sentence (\ref{ex:bad-ext}), where $qui$ is
-used to extract a subject from a clause embedded in a relative clause.
-The problem with sentence (\ref{ex:bad-ext}) is that the type of the
-relative clause is $NP\_VAR{=}T \limp S\_EXT{=}EMB$ which is not a valid
-argument for any constant representing the relative pronoun
-$qui$. Typing the variable $t$ with $NP\_VAR{=}F$ instead of
-$NP\_VAR{=}T$ and using $E_{aime_1}$ instead of $E_{aime_2}$ would not
-help, since the resulting relative clause would then have type
-$NP\_VAR{=}F \limp S\_EXT{=}NO$, which would not be admitted by any
-relative pronoun.
-
-\begin{exe}
-  \ex * \label{ex:bad-ext} Le tatou qui Jean dit que \_ aime Marie dort. \\
-      * $(E_{le_1}\ (E_{qui_?}\ (\lambda^{\circ} t.\ E_{dit\ que_3}\ E_{Jean}\ (E_{aime_2}\ t\ E_{Marie}))\ E_{tatou}))\ (\lambda^{\circ} x.\ E_{dort_1}\ x)$
-  \ex \label{ex:good-ext} Le tatou que Jean dit que Marie aime \_ dort. \\
-      $(E_{le_1}\ (E_{que_2}\ (\lambda^{\circ} t.\ E_{dit\ que_3}\ E_{Jean}\ (E_{aime_3}\ E_{Marie}\ t))\ E_{tatou}))\ (\lambda^{\circ} x.\ E_{dort_1}\ x)$
-\end{exe}
-
-As was the case with our treatment of negation, we refined our familiar
-types by tacking on new kinds of information and defined rules for how
-these pieces of information percolate up the parse tree all the way to
-the topmost level still pertinent for ensuring grammaticality. We will
-now look at another similar example.
-
-\section{Agreement}
-
-Finally, we will cover the phenomenon of number agreement. This
-motivates a straightforward refinement of our types, where all noun and
-noun phrase types carry number information.
-
-$$
-A_{Marie} : NP\_NUM{=}SG
-$$
-$$
-A_{tatou} : N\_NUM{=}SG
-$$
-$$
-A_{tatous} : N\_NUM{=}PL
-$$
-$$
-A_{le} : N\_NUM{=}SG \limp ((NP\_NUM{=}SG \limp S) \limp S)
-$$
-$$
-A_{les} : N\_NUM{=}PL \limp ((NP\_NUM{=}PL \limp S) \limp S)
-$$
-$$
-A_{dort} : NP\_NUM{=}SG \limp S
-$$
-$$
-A_{dorment} : NP\_NUM{=}PL \limp S
-$$
-$$
-A_{aime_1} : NP\_NUM{=}SG \limp NP\_NUM{=}SG \limp S
-$$
-$$
-A_{aime_2} : NP\_NUM{=}SG \limp NP\_NUM{=}PL \limp S
-$$
-$$
-A_{aiment_1} : NP\_NUM{=}PL \limp NP\_NUM{=}SG \limp S
-$$
-$$
-A_{aiment_2} : NP\_NUM{=}PL \limp NP\_NUM{=}PL \limp S
-$$
-$$
-A_{qui_1} : (NP\_NUM{=}SG \limp S) \limp N\_NUM{=}SG \limp N\_NUM{=}SG
-$$
-$$
-A_{qui_2} : (NP\_NUM{=}PL \limp S) \limp N\_NUM{=}PL \limp N\_NUM{=}PL
-$$
-$$
-A_{de_1} : NP\_NUM{=}SG \limp N\_NUM{=}SG \limp N\_NUM{=}SG
-$$
-$$
-A_{de_2} : NP\_NUM{=}PL \limp N\_NUM{=}SG \limp N\_NUM{=}SG
-$$
-$$
-A_{de_3} : NP\_NUM{=}SG \limp N\_NUM{=}PL \limp N\_NUM{=}PL
-$$
-$$
-A_{de_4} : NP\_NUM{=}PL \limp N\_Pl \limp N\_NUM{=}PL
-$$
-
-Given the above signature, we can assign the type $S$ to the syntactic
-term of sentence (\ref{ex:good-agr}) but not to the one of sentence
-(\ref{ex:bad-agr}). In the latter sentence, the type of the variable $t$
-has to be $NP\_NUM{=}SG$ since it is used as an argument to $A_{dort}$. This
-means that the relative clause $dort$ has type $NP\_NUM{=}SG \limp
-S$. However, there is no type for $qui$ which would allow us to modify
-the plural noun $A_{tatous}$ of type $N\_NUM{=}PL$ by a relative clause with a
-missing singular noun phrase, which has the type $NP\_NUM{=}SG \limp S$.
-
-\begin{exe}
-  \ex \label{ex:good-agr} Les tatous de Marie dorment. \\
-      $(A_{les}\ (A_{de}\ A_{Marie}\ A_{tatous}))\ (\lambda^{\circ} x.\ A_{dorment}\ x)$
-  \ex * \label{ex:bad-agr} Marie aime les tatous qui dort. \\
-      * $(A_{les}\ (A_{qui_?}\ (\lambda^{\circ} t.\ A_{dort}\ t)\ A_{tatous}))\ (\lambda^{\circ} y.\ A_{aime_2}\ A_{Marie}\ y)$
-\end{exe}
-
-Once again, we have expressed a linguistic constraint by refining our
-types, augmenting them with a new bit of information, we constrained the
-arguments of functions using these new types and we have described how
-this new type information propagates from the given arguments to the
-produced values.
-
-In the next section, we will see what it takes to enforce all of the
-constraints above (negation, extraction and agreement) at the same time,
-which will give us some idea on how to go about building a grammar which
-handles a wide variety of such phenomena.
-
-\section{Putting It All Together}
-\label{sec:together}
-
-In a grammar that handles negation, extraction and agreement, our types
-will need to carry all of the refinements we introduced before. To
-recapitulate, $NP$ will be parameterized by whether or not it contains a
-negative determiner ($\_\textcolor{red}{NEG{=}T}$ and
-$\_\textcolor{red}{NEG{=}F}$ respectively), whether or not it is a trace
-from an extraction ($\_\textcolor{green}{VAR{=}T}$ and
-$\_\textcolor{green}{VAR{=}F}$) and its number
-($\_\textcolor{blue}{NUM{=}SG}$ or $\_\textcolor{blue}{NUM{=}PL}$). For
-$N$, we have the same refinements except for $\textcolor{green}{VAR}$
-since extraction only moves $NP$s. Clauses, of type $S$, will be
-discriminated based on the level of extraction within them
-($\_\textcolor{green}{EXT{=}NO}$, $\_\textcolor{green}{EXT{=}ROOT}$ and
-$\_\textcolor{green}{EXT{=}EMB}$).
-
-The individual types will have to respect all of the constraints at the
-same time, each one being a complete specification of a possible
-situation (complete w.r.t. the features listed above). To see and
-appreciate what this means in practice, we give the types for the
-preposition $de$:
-
-$$
-C_{de_1} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_2} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_3} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_4} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_5} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_6} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_7} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_8} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_9} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_{10}} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_{11}} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_{12}} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
-$$
-
-This highlights two important problems with the current approach. One is
-the untenable complexity. Even though the three mechanisms are
-completely independent of each other, here we are forced to refer to all
-of them in every type. It is no longer easy to glance at the types
-assigned to $de$ and see how it behaves with respect to, for example,
-the mechanism of negation (compare with the type assignments we gave to
-$de$ in previous subsections). Another cause of unreadability of this
-signature is due to the sheer number of types on display, which leads us
-to our second issue.
-
-The signature demonstrates an exponential growth in the number of typed
-constants included. In cases where the behaviors of a wordform across
-the different mechanisms are independent, the complete type signature
-will need to provide a type for every combination of situations in all
-of the mechanisms. Assuming that every mechanism will assign on average
-more than one type to every wordform, the number of the wordform's types
-in the final signature will be exponential w.r.t. the number of
-different mechanisms handled.
-
-We can try making the enumeration of these types easier for the grammar
-author by introducing metavariables for the values of the features. This
-means that we could write the following 3 templates instead of the 12
-types above:
-
-$$
-\forall \textcolor{blue}{x}, \textcolor{blue}{y} \in \{\textcolor{blue}{SG}, \textcolor{blue}{PL}\}
-$$
-$$
-C_{de_{1(\textcolor{blue}{x}, \textcolor{blue}{y})}} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y})
-$$
-$$
-C_{de_{2(\textcolor{blue}{x}, \textcolor{blue}{y})}} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y})
-$$
-$$
-C_{de_{3(\textcolor{blue}{x}, \textcolor{blue}{y})}} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y})
-$$
-
-This is getting more concise and easier to comprehend. We can take this
-approach further by not only quantifying over variables but also
-computing some of the values using functions:
-
-$$
-\forall \textcolor{blue}{x}, \textcolor{blue}{y} \in\{\textcolor{blue}{SG}, \textcolor{blue}{PL}\}, \forall \textcolor{red}{p}, \textcolor{red}{q} \in \{\textcolor{red}{F}, \textcolor{red}{T}\}, \textcolor{red}{p \land q} \neq \textcolor{red}{T}
-$$
-$$
-C_{de(\textcolor{blue}{x}, \textcolor{blue}{y}, \textcolor{red}{p}, \textcolor{red}{q})} : (NP\_\textcolor{red}{NEG{=}p}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}q}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}(p \lor q)}\_\textcolor{blue}{NUM{=}y})
-$$
-
-The above is vastly more concise and manages to convey the intent more
-clearly (the intent being that it cannot be the case that there are
-negative determiners in both the complement noun phrase and the modified
-noun and that the value of the result's \textcolor{red}{NEG} feature is
-always the disjunction of the \textcolor{red}{NEG} features of the two
-arguments).
-
-As of right now, we have introduced this kind of rule only as
-meta-notation. This means that the real underlying grammar that will end
-up being used by some algorithm still suffers from the same exponential
-cardinality of the signature. Fortunately, there exists a well-studied
-construction in type theory that lets us assign to terms generic types
-such as the one defined in the template above. This construction is the
-\emph{dependent product} and its utility in ACGs has already been
-noted.\footnote{See \cite{de2007two} for its introduction into the
-  domain of ACGs, \cite{de2007type} for a small example grammar, which
-  handles the interaction between number and gender agreement and
-  coordination, and finally see \cite{pogodalla2012controlling} for
-  using dependent types to handle several constraints regarding
-  extraction (including the one we covered in~\ref{sec:extraction}).}
-
-While extending the type system by including dependent types is
-definitely useful and it helps to solve the problem of having our
-signatures grow exponentially in size, it does not solve our problem
-with the complexity of the new grammar. Granted, the new meta-type
-written in the dependent style is more readable and easier to reason
-about than the enumeration of 12 basic types we presented at the
-beginning of this section, however, the result is still complex in
-that it forces us to describe the behavior of three independent
-mechanisms in one place. This one place happens to be a wordform which
-has seemingly little connection to these mechanisms, the preposition
-$de$. We can therefore imagine that it will participate in many other
-phenomena and that the type we assign to it will grow beyond our
-abilities to reason about it effectively.
-
-Furthermore, the current notation does not even make it obvious that
-these are independent mechanisms. One has to examine the type to make
-sure that there is no interaction. In the next two subsections, we will
-devote ourselves to investigating ways of disentangling these.
-
-\section{Intersections of Types}
-
-Our goal is to split the types into smaller parts that only talk about
-features belonging to one of the three mechanisms covered above. This
-means that we would like to take some phrase about which we know a lot
-of disparate pieces of information (presence of negative determiner,
-trace status, number\ldots), having type
-e.g. $NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}$,
-and use it as an argument for some function which only cares about one
-channel of information and would have a type like
-$NP\_\textcolor{red}{NEG{=}T} \limp N\_\textcolor{red}{NEG{=}F} \limp
-N\_\textcolor{red}{NEG{=}T}$. To make a type such as
-$NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}$
-be a legal argument to a function which expects an argument of type
-$NP\_\textcolor{red}{NEG{=}T}$, we will need to introduce some notion of
-\emph{subtyping}.
-
-The types we have been showing up to now propose a very straightforward
-notion of subtyping. All of the types can be seen as conjunctions of
-properties where $\_$ (underscore) is the conjunction operator. A type
-$\tau$ can then be seen as a subtype of type $\sigma$ whenever the
-conjuncts of $\tau$ form a superset of the conjuncts of $\sigma$. This
-is not unlike the structural typing of records in OCaml (CITE SOME OCAML
-SPEC HERE), where a record type is a subtype of another record type if
-the former provides, among others, all the fields provided by the latter
-and gives them the same type. In our case, one type is a subtype of
-another if the former gives, among others, all the properties given by
-the latter. This notion of subtyping can be expressed by adding the rule
-of inference in Figure~\ref{fig:type-superset} to our type
-system.\footnote{In the rule \ldots (talk about $\land$ instead of $\_$
-  (underscore) here...)}
-
-\begin{figure}
-  \begin{prooftree}
-    \AxiomC{$\Gamma_l; \Gamma_i \vdash_\Sigma M : A_1 \land \ldots A_{i-1} \land A_i \land A_{i+1} \ldots \land A_n$}
-    \RightLabel{(subtype)}
-    \UnaryInfC{$\Gamma_l; \Gamma_i \vdash_\Sigma M : A_1 \land \ldots A_{i-1} \land A_{i+1} \ldots \land A_n$}
-  \end{prooftree}
-  \caption{\label{fig:type-superset} A rule of inference giving us
-    subtyping for conjunctions of atomic types. For readability, we use
-    $\land$ instead of $\_$ (underscore) to designate the conjunction of
-    properties.}
-\end{figure}
-
-(Talk about whether or not this is a conservative extension...)
-
-With the new subtyping rule, we can
-
-$$
-S_{tatou} : N \land NEG{=}F \land NUM{=}SG
-$$
-$$
-S_{le_{NEG_1}} : (N \land NEG{=}F) \limp (((NP \land NEG{=}F) \limp S) \limp S)
-$$
-$$
-S_{le_{NEG_2}} : (N \land NEG{=}T) \limp (((NP \land NEG{=}T) \limp S) \limp S)
-$$
-$$
-S_{le_{NUM}} : (N \land NUM{=}SG) \limp (((NP \land NUM{=}SG) \limp S) \limp S)
-$$
-$$
-S_{le_{VAR}} : N \limp (((NP \land VAR{=}F) \limp S) \limp S)
-$$
-
-\ldots
-
-\section{Intersections of Languages}
-
-We will now turn to another solution to the complexity of grammars
-implementing multiple independent mechanisms.
-
-When ACGs are discussed, one of the oft-praised properties is
-compositionality. Let us consider a situation where we have three
-signatures $\Sigma_1$, $\Sigma_2$ and $\Sigma_3$, their respective
-distinguished types $S_1$, $S_2$ and $S_3$ and two lexicons
-$\mathcal{L}_{12} : \Sigma_1 \to \Sigma_2$ and $\mathcal{L}_{23} :
-\Sigma_2 \to \Sigma_3$ such that $\mathcal{L}_{12}(S_1) = S_2$ and
-$\mathcal{L}_{23}(S_2) = S_3$. (TODO: INCLUDE A DIAGRAM OF THIS
-SITUATION)
-
-We can now define languages $A_1$, $A_2$ and $A_3$ where $A_i$ is the
-set of terms from $\Lambda(\Sigma_i)$ that have the type $S_i$. Such
-languages, which are defined only in terms of a signature and some
-distinguished type, are called \emph{abstract languages} in the
-terminology of ACGs. They consist of terms whose structure is
-constrained only by the types assigned by the signature.
-
-Further than that, we can look at the languages $\mathcal{L}_{12}(A_1)$
-and $\mathcal{L}_{23}(A_2)$, images of the abstract languages given by
-the lexicons. Since the lexicons are homomorphic mappings w.r.t. the
-types, all elements of $\mathcal{L}_{12}(A_1)$ will have the type
-$\mathcal{L}_{12}(S_1) = S_2$ and will therefore also belong to $A_2$
-(similarly for $\mathcal{L}_{23}(A_2)$ and $A_3$). These new languages,
-which are formed by mapping an ACG abstract language using a lexicon,
-are called \emph{object languages}. Their terms are constrained not only
-by the types of the signature they are built upon but also by the type
-system of the abstract signature.
-
-Finally, we can also take the composition of lexicons $\mathcal{L}_{13}
-= \mathcal{L}_{23} \circ \mathcal{L}_{12}$, which itself is also a
-lexicon, and map it over $A_1$. This gives us a language
-$\mathcal{L}_{13}(A_1) = \mathcal{L}_{23}(\mathcal{L}_{12}(A_1))$ which
-is at the same time the image of the object language
-$\mathcal{L}_{12}(A_1)$ given by $\mathcal{L}_{23}$ and also the image
-of the abstract language $A_1$ given by $\mathcal{L}_{13}$. This points
-to an interesting property of ACG languages: they are closed under
-transformation by a lexicon.
-
-In the ACG literature, ACGs are defined as quadruples $\mathcal{G} =
-\mathopen{<}\Sigma_A, \Sigma_O, \mathcal{L}, S\mathclose{>}$ where
-$\Sigma_A$ is the abstract signature, $\Sigma_O$ is the object
-signature, $\mathcal{L} : \Sigma_A \to \Sigma_O$ is a lexicon linking
-the two and $S$ is a distinguished type of the abstract signature
-$\Sigma_A$. From this grammar, we can define two languages. The abstract
-language $\mathcal{A}(\mathcal{G}) = \{t \in \Lambda(\Sigma_A) \mid
-\ \vdash_{\Sigma_A} t : S\}$ and the object language
-$\mathcal{O}(\mathcal{G}) = \mathcal{L}(\mathcal{A}(\mathcal{G})) = \{t
-\in \Lambda(\Sigma_O) \mid \exists u \in \mathcal{A}(\mathcal{G}),
-\mathcal{L}(u) = t\}$. The composition of two ACGs $\mathcal{G}_1 =
-\mathopen{<}\Sigma_1, \Sigma_2, \mathcal{L}_{12}, S_1\mathclose{>}$ and
-$\mathcal{G}_2 = \mathopen{<}\Sigma_2, \Sigma_3, \mathcal{L}_{23},
-S_2\mathclose{>}$ is defined as $\mathcal{G}_2 \circ \mathcal{G}_1 =
-\mathopen{<}\Sigma_1, \Sigma_3, \mathcal{L}_{23} \circ \mathcal{L}_{12},
-S_1\mathclose{>}$.
-
-The definition of ACGs given above is quite practical in that it is
-complete, it gives a precise account of what set of languages are
-generated by ACGs (abstract and object languages for which there exist
-the quadruples described above). This contrasts with the presentation of
-ACGs that we gave in~\ref{sec:acg}, where we defined ACGs as
-collections of signatures with associated distinguished types and
-connected with lexicons. Our definition only introduced the requisite
-machinery, but delayed the delimitation of the way languages can be
-defined. Leaving this question open will let us answer it now by
-providing an extension/generalization of ACGs which will help us
-overcome the modularity issues that have been the theme of this
-chapter.\footnote{Incidentally, this was not the reason for the way we
-  presented ACGs in~\ref{sec:acg}. We chose this style as it grouped
-  together the information in a more natural way. We approach signatures
-  as descriptions of domains (strings, syntax, semantics), each one
-  being associated with a distinguished type telling us which objects
-  the domain has set out to describe (strings, sentences, truth
-  values). Having distinguished types directly attached to signatures
-  allows us to omit them from the ACG quadruples. However, this only
-  leaves us with triples containing the two signatures and the lexicon
-  connecting them, and since we can infer the signatures themselves from
-  the domain and the range of the lexicon, we do not have much
-  motivation to introduce these tuples. Instead of composing ACGs, we
-  compose just the lexicons.
-
-On top of that, we will want to define signatures and lexicons that
-yield different object languages (at least some string representation of
-the form and some semantic representation of the meaning), which would
-mean having at least two different ACGs which are linked (e.g. by
-sharing the same abstract signature). However, since these two grammars
-are very closely related, we would prefer to reason about our grammar as
-\emph{one} system/graph of signatures and lexicons which defines
-multiple languages instead of talking about \emph{multiple} grammars
-which share some similarities.
-
-TODO: Somehow express that its nice to have the links between the
-signatures and lexicons that are shown in diagrams \emph{explicit}
-rather than having them \emph{implicit}.}
-
-Before we turn to our proposal, we will quickly recall some of the other
-styles of composition used in ACGs (a more detailed exposition can be
-found in \cite{pogodalla2012controlling}). Besides making the object
-signature of one grammar be the abstract signature of another as we saw
-above (TODO: INSERT REFERENCE TO FUTURE DIAGRAM), we can have two
-grammars share the same abstract signature allowing us to transduce
-between terms in the two object signatures. See Figure~(TODO: INSERT A
-DIAGRAM FOR THIS SITUATION) for a diagrammatic example.
-
-These two modes of composition can be fruitfully mixed. Consider (TODO:
-INSERT A DIAGRAM), where we introduce a syntactic signature TODO where
-different terms are only allowed to yield the same string whenever they
-represent different syntactic structures of a syntactically ambiguous
-phrase. This means that purely semantic distinctions such as scope
-ambiguity have to be moved to a more abstract signature from whose terms
-the semantic representation can be derived by a function. This approach
-is studied more deeply in \cite{pogodalla2007generalizing}.
-
-Another interesting composition pattern is to take an existing structure
-of grammars and add a new abstract signature on top. This signature can
-be used to further constrain the items in (TODO: what was before the
-abstract-most signature/language). In \cite{pogodalla2012controlling},
-this is what the authors use to develop their constraints on
-extraction. Each of their constraints is presented as a new abstract
-signature constraining the original syntactic signature.
-
-However, a question that the authors do not pose is how to combine all
-these constraints. If we were to apply the same pattern repeatedly to
-get a structure like the one on (TODO: INSERT A DIAGRAM HERE), we would
-have to deal with successively more complex type systems on each
-level. This is due to the condition that a lexicon has to be a
-homomorphism and thus it must, among other things, always map well-typed
-terms to well-typed terms. To illustrate the problem, imagine a
-structure such as the one on (TODO) that handles some different
-constraint in each of the vertically stacked signatures, including
-e.g. agreement. If we tried to add a new abstract signature at the top
-to implement another constraint, we would have to ensure that all the
-terms we licence in our new abstract signature have a well-typed
-counterpart in the existing abstract language. However, this means that
-our new signature must make sure that the terms it approves do not break
-agreement or any of the previously established constraints. Effectively,
-our new signature would be forced to reimplement all of the existing
-constraints, rendering in vain all our effort to split up the
-constraints into multiple signatures.
-
-What we would like instead is to put the signatures side-by-side such
-that they do not depend on each other (as on diagram TODO). However, how
-does this actually define a language? We can arrive at the answer by
-examining how are the graphical representations of ACGs connected to the
-languages they define. We will associate a language to every node of the
-graph. Before we begin, we notice that every ACG diagram formed by the
-above rules of composition is a directed tree and therefore every node
-has at most one parent/predecessor. To the root of the tree, the
-abstract-most signature in the graph, we will assign the abstract
-language generated by that signature and its distinguished type. To all
-the other nodes, we will assign the result of mapping their parent's
-language by the lexicon linking the two. By homomorphism of the
-lexicons, we can also deduce the distinguished types for all the
-non-root nodes. They are the images of their parents' distinguished
-types by the connecting lexicons. For all of the nodes, it is true that
-every term in the language assigned to the node has the node's
-distinguished type.
-
-We can confirm that the above graphical model of defining ACGs is
-coherent. First, we can show that our interpretation is sound and so
-that any directed tree adorned with signatures on the nodes, compatible
-lexicons on the edges and a distinguished type on the root can only
-define ACG languages. This is trivial, since the language corresponding
-to the root is an ACG abstract language and the class of ACG languages
-is closed under transformation by a lexicon and therefore all of the
-languages corresponding to the descendants of the root are ACG languages
-as well. Second, we can try and show some kind of completeness. However,
-there is not much formalization done to describe a ``system of ACGs that
-share common signatures'' (something that our generalization will
-fix). We can nevertheless show that it is possible to express any ACG
-using our graphical model. Given a grammar $\mathcal{G} =
-\mathopen{<}\Sigma_A, \Sigma_O, \mathcal{L}, S\mathclose{>}$, we just
-produce a tree whose root has the signature $\Sigma_A$ and distinguished
-type $S$ and which is connected to a single child node having the
-signature $\Sigma_O$ by an edge labelled by the lexicon
-$\mathcal{L}$. Then the two nodes of our tree define both the abstract
-and the object language defined by $\mathcal{G}$.
-
-We have now given some formal account of how we can read language
-definitions in diagrams. Still, we have kept our analysis to the tree
-diagrams used in current ACG literature, where every node has at most
-one parent. To make the generalization more visible, we will switch from
-an algebraic manner of presentation, in which we stated that the
-language of a node is equal to an image of its parent's language, to a
-calculus-based manner of presentation, in which we will state that the
-language of a node is the set of all terms that have an antecedent in
-the parent's language. Besides this property, there is also the fact
-that every term in a node's language has the node's distinguished
-type. This property is trivially true for nodes with parents because of
-the homomorphic property of lexicons. However, it is crucial for the
-root node for which this is the single defining property of its
-language. To sum up, we have that for every node of a tree diagram, its
-language is the set of all terms that have the node's distinguished type
-and that have an antecedent in the parent's language, if there is
-any. This rule is already general enough to handle the case of zero or
-one parents, we can easily generalize it to handle an arbitrary count:
-\emph{the language of a node is the set of all terms that have the
-  node's distinguished type and that have antecedents in the languages
-  of all of the node's parents}. If we transpose this back to the
-algebraic style of presentation, we say that the language of a node is
-the intersection of the set of terms having the node's distinguished
-type and the intersection of all its parents' languages.
-
-TODO: In the previous paragraph, add a notice saying that the terms must
-be also built on the signature of the node.
-
-TODO: Use the correct term arborescence in the above paragraphs!
-% http://en.wikipedia.org/wiki/Arborescence_(graph_theory)
-
-We will now turn to a formal definition of this generalized graph-based
-formalism and discuss some its values and properties.
-
-\subsection{Graphical Abstract Categorial Grammars}
-
-We define a \emph{graphical abstract categorial grammar} as a quadruple
-$\mathcal{G} = \mathopen{<} G, \Sigma, S, {\mathcal{L}} \mathclose{>}$
-where $G$ is a directed graph with vertices $V(G)$ and edges $E(G)$,
-$\Sigma$ and $S$ are labelings assigning signatures and distinguished
-types, respectively, to the vertices in $V$ and ${\mathcal{L}}$ is a
-labeling assigning lexicons to the edges in $E$. Furthermore, a
-well-formed graphical ACG satisfies the following conditions:
+We will enumerate some of the practical benefits and challenges of using
+graphical ACGs for implementing grammar-based systems.
 
 \begin{itemize}
-\item $G$ is a directed acyclic graph.
-\item For all $(u,v) \in E(G)$, $\mathcal{L}_{(u,v)} : \Sigma_u \to
-  \Sigma_v$.
-\item For all $(u,v) \in E(G)$, $\mathcal{L}_{(u,v)}(S_u) = S_v$.
+\item The lexicons that go from our constraint signatures ($Neg$, $Ext$
+  and $Agr$ in subsection \ref{ssec:graphical-engineering}) are all
+  trivial in the sense they map constants to constants. Finding an
+  antecedent to a term is then just a question of finding an antecedent
+  for every constant in the term (i.e. ``tagging'' the constants of the
+  term with antecedent constants). This means that there is an effective
+  parsing algorithm for verifying the constraints. Furthermore, this
+  algorithm could also be made efficient by introducing filtering
+  techniques \cite{guillaume2008toolchain} and statistical supertagging
+  \cite{moot2012categorial}.
+
+\item Constraints can introduce type extensions in their abstract
+  signatures and the added complexity of these extensions is then
+  contained within the constraint. Furthermore, the constraint signature
+  can be treated as a ``black box'' that verifies a constraint and in
+  the context of an implementation, it can be replaced with a
+  special-purpose hand-written procedure that does the same more
+  efficiently.
+
+\item The distributed nature of graphical ACGs has its own practical
+  applications. Parsing can be made more robust by being able to parse
+  sentences which violate some of the constraints. Parsing an
+  ungrammatical sentence and identifying the constraints it violates can
+  also serve as a basis for a syntax checker.
+
+\item If we decide to define all of our syntactic constraints in
+  separate signatures, we will end up with a lot of signatures which
+  look exactly like the syntactic signature but with some small
+  refinements (see the signatures in sections \ref{sec:negation},
+  \ref{sec:extraction} and \ref{sec:agreement}). Thus we will desire
+  some concise way of formulating these refinements so that we will not
+  need to replicate the entire syntactic signature for every
+  constraint.
+
+  We hope that the class of useful refinements is small and that these
+  constraint signatures could be systematically constructed from the
+  syntactic signature by applying a series of refining
+  transformations. This is an interesting engineering challenge awaiting
+  anyone willing to design a truly wide-coverage graphical ACG using
+  constraints.
 \end{itemize}
 
-We will say that a node $u$ is \emph{more abstract} than a node $v$
-whenever $(u,v)$ belongs to the strict partial order induced by the
-edges of $G$ (i.e. there is a path from $u$ to $v$).
+\section{Conclusion}
+\label{sec:conclusion}
 
-We let the nodes in a graphical ACG $\mathcal{G} = \mathopen{<} G,
-\Sigma, S, \mathcal{L} \mathclose{>}$ define two kinds of
-languages. The \emph{intrinsic languages} $\mathcal{I}_{\mathcal{G}}$,
-which are constrained only by the type signature described in the node
-itself, and the \emph{extrinsic languages} $\mathcal{E}_{\mathcal{G}}$,
-which are also constrained by type systems in more abstract nodes of the
-graph. These two notions correspond to those of abstract languages and
-object languages in ACGs.
+The contributions of our work can be summarized in the following three
+points:
 
-$$
-\mathcal{I}_{\mathcal{G}}(v) = \{t \in \Lambda(\Sigma_v)
-\mid\ \vdash_{\Sigma_u} t : S_v\}
-$$
-$$
-\mathcal{E}_{\mathcal{G}}(v) = \mathcal{I}_{\mathcal{G}}(v) \cap
-\bigcap_{(u,v) \in E} \mathcal{L}_{(u,v)}(\mathcal{E}_{\mathcal{G}}(u)).
-$$
+\begin{enumerate}
+\item We have highlighted the challenges inherent in translating an
+  existing grammar in the formalism of IGs to the formalism of
+  ACGs. Based on the breadth of different constraints built in to the IG
+  formalism, we have conjectured that a direct translation of IG lexical
+  items to ACG lexical items based on the deep underlying similarity
+  between the two formalisms is not practical.
 
-From the above definition, we see that the ACG diagrams we introduced
-before are a special case of graphical ACGs where the graph is an
-arborescence.
+\item We have implemented a system for experimenting with (lexicalized)
+  abstract categorial grammars that can serve as the basis for
+  formulating wide-coverage graphical ACGs
+  (\url{https://github.com/jirkamarsik/acg-clj}).
 
-\subsubsection{On the Expressivity of Graphical ACGs}
+\item We have presented the graphical abstract categorial grammars
+  (G-ACGs) which generalize ACGs. We have determined the set of
+  languages definable by G-ACGs in terms of ACG languages and we have
+  shown how G-ACGs allow us to write categorial grammars that handle
+  multiple disparate linguistic constraints without sacrificing
+  simplicity.
+\end{enumerate}
 
-\begin{observation}
-  The set of intrinsic languages definable by graphical ACGs is the same
-  as the set of abstract languages definable by ACGs.
-\end{observation}
-
-\begin{theorem}
-  The set of extrinsic languages definable by graphical ACGs is the set
-  of object languages definable by ACGs closed under intersection and
-  transformation by a lexicon.
-\end{theorem}
-
-\begin{proof}
-  First, we prove that an extrinsic language can be always constructed
-  from object languages by intersections and transformations by
-  lexicons. Let us consider any graphical ACG $\mathcal{G} =
-  \mathopen{<} G, \Sigma, S, \mathcal{L} \mathclose{>}$ and any
-  topological ordering $v_1, \ldots, v_n$ of the nodes of $G$ (an
-  ordering such that more abstract nodes always precede less abstract
-  ones). Before we start, we will just note that the all the intrinsic
-  languages defined by $v_i$ are abstract languages which are a special
-  case of object languages.
+Here are some of the items that we believe warrant future investigation:
 
 \begin{itemize}
-    \item For $v_1$, the case is trivial since its extrinsic language is
-      just its intrinsic language as the node $v_1$ has no predecessors
-      in $G$.
-    \item For any other node $v_n$, we look at the definition of an
-      extrinsic language and remark that the only operators are
-      intersection and application of a lexicon with the operands being
-      the intrinsic language of $v_n$ and the extrinsic languages of
-      more abstract nodes. By induction hypothesis, we have that all of
-      the operands have been constructed from object languages using
-      intersections and transformations by lexicons and therefore even
-      the extrinsic language defined by $v_n$ has this property.
+\item Try to express the multiple independent polarized features of IGs
+  using G-ACGs.
+\item Determine whether ACG object languages are closed on intersection
+  or not.
+\item Make a closer examination of the relationship between extrinsic
+  and pangraphical languages in G-ACGs.
+\item Write grammars using the G-ACG constraint architecture to verify
+  its applicability and test the claims of modularity we have made here.
+\item Study the problem of concisely defining constraint signatures from
+  syntactic signatures and rules of feature propagation.
 \end{itemize}
-
-  Now we have to prove the converse. We do so by showing that extrinsic
-  languages contain object languages and are closed under intersection
-  and transformation by a lexicon.
-
-  \begin{itemize}
-    \item Let $L$ be an object language defined by the ACG $\mathopen{<}
-      \Sigma_A, \Sigma_O, \mathcal{L}, S\mathclose{>}$. We take the
-      graph $G$ with $V(G) = \{A, O\}$ and $E(G) = \{(A,O)\}$, we label
-      the nodes $A$ and $O$ with $\Sigma_A$ and $\Sigma_O$ as signatures
-      and $S$ and $\mathcal{L}(S)$ as distinguished types
-      respectively. Finally, we label the edge $(A,O)$ with
-      $\mathcal{L}$ and we have a graphical ACG in which the node $O$
-      defines the extrinsic language $L$.
-
-    \item Let $L_1$ and $L_2$ be two extrinsic languages defined by the
-      nodes $u_1$ and $u_2$, respectively, and belonging to some two
-      graphical ACGs. We can construct a new graphical ACG by taking the
-      union of the two ACGs (union of graphs (union of vertices and
-      edges) and unions of the labelings).\footnote{Assuming, without
-        loss of generality, that the sets of vertices of the two graphs
-        are disjoint.} We then add a new node $v$ which will define the
-      language $L_1 \cap L_2$.
-
-      The signature labeling $v$ will be the union of the two signatures
-      labeling $u_1$ and $u_2$ such that the intrinsic language of $v$
-      will be a superset of the intrinsic languages, and by extension
-      the extrinsic languages, of both $u_1$ and $u_2$. Either of the
-      two distinguished types that label $u_1$ or $u_2$ can be used to
-      label $v$ (if they are actually different, then the intersection
-      is trivially empty). Finally, we will add two edges leading from
-      $u_1$ and $u_2$ to $v$, both labeled with the identity lexicon. We
-      will now show that the extrinsic language of $v$ in the newly
-      formed graphical ACG $\mathcal{G}'$ is $L_1 \cap L_2$.
-
-      \begin{align*}
-        \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v)
-        \cap \mathcal{L}_{(u_1,v)}(\mathcal{E}_{\mathcal{G}'}(u_1)) \cap
-        \mathcal{L}_{(u_2,v)}(\mathcal{E}_{\mathcal{G}'}(u_2)) \\ &=
-        \mathcal{I}_{\mathcal{G}'}(v) \cap
-        \mathcal{E}_{\mathcal{G}'}(u_1) \cap
-        \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
-        \mathcal{E}_{\mathcal{G}'}(u_1) \cap
-        \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
-        \mathcal{E}_{\mathcal{G}}(u_1) \cap
-        \mathcal{E}_{\mathcal{G}}(u_2) \\ &= L_1 \cap L_2
-      \end{align*}
-
-      The simplifications use the facts that: the two new lexicons are
-      identities, that the intrinsic language of $v$ is a superset of
-      the extrinsic languages of $u_1$ and $u_2$, that the extrinsic
-      languages of $u_1$ and $u_2$ are the same in the new graphical ACG
-      as in the two original graphical ACGs and finally that $L_1$ and
-      $L_2$ are the extrinsic languages defined by $u_1$ and $u_2$.
-
-  \item Let $L \subset \Lambda(\Sigma_A)$ be an extrinsic language and
-    $\mathcal{L} : \Sigma_A \to \Sigma_O$ a lexicon from its signature
-    $\Sigma_A$ to the signature $\Sigma_O$. We will show that the
-    language $\mathcal{L}(L) \subset \Lambda(\Sigma_O)$ is also an
-    extrinsic language. Let $u$ be the node that defines the extrinsic
-    language $L$ in some graphical ACG $\mathcal{G}$. We construct a new
-    graphical ACG by adding a new vertex $v$ to $\mathcal{G}$ labeled
-    with the signature $\Sigma_O$ and the distinguished type
-    $\mathcal{L}(S)$ where $S$ is the distinguished type labeling
-    $u$. To this graph, we conjoin the edge $(u,v)$ and label it with
-    $\mathcal{L}$. Now we verify that the extrinsic language of $v$ is
-    the language $\mathcal{L}(L).$
-
-    \begin{align*}
-      \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v) \cap
-      \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
-      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
-      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}}(u)) \\
-      &= \mathcal{L}(L)
-    \end{align*}
-
-    The facts used here are in turn: the inclusion of
-    $\mathcal{I}_{\mathcal{G}'}(v)$ in
-    $\mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u))$, the observation that
-    modifying our graphical grammar did not change the extrinsic
-    language of $u$ and the assumption that $L$ is the extrinsic
-    language of $u$.
-  \end{itemize}
-\end{proof}
-
-\begin{corollary}
-  The set of extrinsic languages definable by graphical ACGs is the same
-  as the set of object languages definable by ACGs if and only if the
-  set of object languages definable by ACGs is closed under
-  intersection.
-\end{corollary}
-
-We conclude this subsubsection with a few words to motivate the need we
-feel for this (possibly) more expressive formalism. We believe that in
-order to manage the development of wide-coverage grammars, the grammars
-need to be kept as simple as possible, which means, among other things,
-that independent constraints should be implemented separately and not
-complected together (we saw the difficulties caused by this already on a
-microscopic fragment in section~\ref{sec:together}). This means that if
-we were to define, e.g., a partial grammar in which we would only
-enforce agreement and another one in which we would only enforce the
-proper use of relative pronouns, we would like the language in which
-both agreement and proper use of relative pronouns are respected to be
-expressible in our formalism and on top of that, we would like this
-language to be easily expressed in terms of the two grammars for the
-individual constraints. We now have a requirement on the formalism in
-which we would write such grammars, we want a formalism where the set of
-all the ``useful'' definable languages is closed on intersection. If the
-object languages of ACGs are indeed closed on intersection, then our
-extension is just a convenience for expressing intersection more
-easily. However, if they are not, we believe our extension adds a
-critical level of expressivity needed for simple and modular
-wide-coverage grammars.
-
-\subsubsection{On the Decidability of Graphical ACGs}
-
-We turn to the problem of parsing with graphical ACGs. Parsing with ACGs
-is a problem on the limits of decidability. When we consider ACGs whose
-abstract signatures only contain types of order 2 or less, parsing is
-possible in polynomial type for both linear \cite{salvati2005problemes}
-and non-linear \cite{kanazawa2007parsing} type systems (such as the one
-we introduced in~\ref{sec:acg}). If the order of the abstract signature
-exceeds 2, the membership problem for linear ACGs becomes decidable if
-and only if the satisfiability problem in the multiplicative exponential
-fragment linear logic is decidable \cite{de2004vector}, which is an open
-problem. On the other hand, by extending the type system to include
-dependent types, parsing can become undecidable too since type checking
-in such a system is itself undecidable.
-
-Our intent here is to show that by extending the formalism, we have not
-made the problem of parsing with the grammar more difficult, at least in
-terms of decidability if not in terms of complexity. Let us imagine a
-graphical ACG where every edge (two signatures, distinguished types and
-a lexicon) is an ACG for which there is an effective parsing algorithm
-and all type checking is decidable in all the signatures. Then we can
-build a simple effective parsing algorithm for any of the extrinsic
-languages defined in the graphical ACG.
-
-We will be testing whether the term $\gamma$ belongs to the extrinsic
-language of some node $v$. Our algorithm first checks whether $\gamma$
-belongs to the intrinsic language of $v$ by checking its type. If it
-does, it proceeds to verify that it is also in the extrinsic language by
-finding antecedents in all of its abstract predecessors $u_i$, if there
-are any. We can use the presupposed parsing algorithm for the ACG
-$\mathopen{<} \Sigma_{u_i}, \Sigma_v, \mathcal{L}_{(u_i,v)},
-S_{u_i}\mathclose{>}$ to find all the abstract terms $\alpha_{i,j}$
-having type $S_{u_i}$ and being antecedents to $\gamma$ w.r.t. the
-lexicon $\mathcal{L}_{(u_i,v)}$. Such terms belong to the intrinsic
-language of $u_i$ but we are interested only in antecedents coming from
-the extrinsic language. Therefore, what we have to do is to recursively
-apply our parsing algorithm to all of the antecedent-candidates
-$\{\alpha_{i,j} \mid \forall j\}$ until we find one which belongs to the
-extrinsic language of $u_i$. Since the graph is acyclic, we are
-guaranteed that this recursive traversal will terminate. We repeat this
-for all the predecessors $u_i$ and if we succeed to find antecedents in
-all of them, we have confirmed the membership of $\gamma$ in
-$\mathcal{E}_{\mathcal{G}}(v)$ and also built its abstract
-antecedent(s).
-
-\subsubsection{On Grammar Engineering with Graphical ACGs}
-
-%% \subsubsection{On Alternative Interpretations of Graphical ACGs}
-%
-% Mention the example with the levels of linguistic representations
-% linked by many-to-many transductions and explain how a different
-% theory of graphs defining languages might be suitable to that
-% architecture.
-%
-% Also point out the weirdness of the diamond shape.

### 2013-06-17 constraints.tex
Last night shift

--- 2013-06-13 the-last-part.tex
@@
-\chapter{The Last Part}
+\chapter{Treatment of Multiple Linguistic Constraints}
+\label{chap:constraints}
 
@@
 linguistic phenomena in a single framework, highlight the modularity
-issues this brings up and propose some solutions to them.
+issues this brings up.
 
 \section{Negation}
+\label{sec:negation}
 
 We will start with an account of the paired grammatical words mechanism
-for negation in French. Recall the EPTDs we have seen
+for negation in French, specifically the interaction between negative
+noun phrases and the particle \emph{ne}. Recall the EPTDs we have seen
 in~\ref{ssec:frigram}, repeated here on Figure~\ref{fig:ig-neg-rep}.
@@
   \centering
-  \includegraphics[scale=0.25]{ig-neg.pdf}
+  \includegraphics[scale=0.25]{images/ig-neg.pdf}
   \caption{\label{fig:ig-neg-rep} Frigram EPTDs for a negative
-    determiner and the paired grammatical word $ne$.}
+    determiner and the paired grammatical word \emph{ne}.}
 \end{figure}
 
-At a basic level, $aucun$ functions like any other determiner and $ne$
-as any verb modifier. However, we want to encode the constraint that
-whenever a verb is modified by $ne$, the modified verb will demand that
-one of its arguments contains a negative determiner. Since we need to
-make a distinction between phrases that contain negative determiners and
-those that do not, we will need to provide different types for both (two
-terms with the same type are indistinguishable w.r.t. syntactic
-composition in a type-logical grammar such as ACGs). Our type for
-$aucun$ will thus end up being $N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S)
-\limp S)$ instead of the usual $N \limp ((NP \limp S) \limp
-S)$.\footnote{Strictly following the EPTD for $aucun$ would lead us to
-  give it a type $DET\_NEG{=}F$ and then give the two types
+At a basic level, \emph{aucun} functions like any other determiner and
+\emph{ne} as any verb modifier. However, we want to encode the
+constraint that whenever a verb is modified by \emph{ne}, the modified
+verb will demand that one of its arguments contains a negative
+determiner. Since we need to make a distinction between phrases that
+contain negative determiners and those that do not, we will need to
+provide different types for both (two terms with the same type are
+indistinguishable w.r.t. syntactic composition in a type-logical grammar
+such as ACGs). Our type for \emph{aucun} will thus end up
+being\footnote{Strictly following the EPTD for \emph{aucun} would lead
+  us to give it a type $DET\_NEG{=}F$ and then give the two types
   $DET\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S)$ and $DET\_NEG{=}T
@@
   demonstration, we prefer to align our examples with the treatment
-  which is customary to ACG research.} The type for $le$ will need to be
-able handle cases where its argument noun either already contains a
-negative determiner or not.
+  which is customary to ACG research.} $$N\_NEG{=}F \limp ((NP\_NEG{=}T
+\limp S) \limp S)$$ instead of the usual $$N \limp ((NP \limp S) \limp
+S)$$ The type for \emph{le} will need to be able handle cases where its
+argument noun either already contains a negative determiner or not.
 
-$$
-N_{aucun} : N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S) \limp S)
-$$
-$$
-N_{le_1} : N\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S)
-$$
-$$
-N_{le_2} : N\_NEG{=}T \limp ((NP\_NEG{=}T \limp S) \limp S)
-$$
+\begin{align*}
+N_{aucun} &: N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S) \limp S) \\
+N_{le_1} &: N\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S) \\
+N_{le_2} &: N\_NEG{=}T \limp ((NP\_NEG{=}T \limp S) \limp S) \\
+\end{align*}
 
-The type of $ne$ will be more verbose, as it will consume a verb and
-transform its valency so that it demands that exactly one of its $NP$
-arguments contains a negative determiner.
+The type of \emph{ne} will be more verbose, as it will consume a verb
+and transform its valency so that it demands that exactly one of its
+$NP$ arguments contains a negative determiner.
 
-$$
-N_{ne_{tv_1}} : (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}F \limp S)
-$$
-$$
-N_{ne_{tv_2}} : (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}F \limp NP\_NEG{=}T \limp S)
-$$
+\begin{align*}
+N_{ne_{tv_1}} &: (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}F \limp S) \\
+N_{ne_{tv_2}} &: (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}F \limp NP\_NEG{=}T \limp S)
+\end{align*}
 
@@
 currently not covered by Frigram (though it could be), but is considered
-grammatical by French speakers. Finally, in a complete grammar, $ne$
-would have to provide types capable of transforming all the other
-syntactic valencies.
+grammatical by French speakers. Finally, in a complete grammar,
+\emph{ne} would have to provide types capable of transforming all the
+other syntactic valencies.
 
@@
   \centering
-  \includegraphics[scale=0.25]{parse-aucun.pdf}
+  \includegraphics[scale=0.25]{images/parse-aucun.pdf}
   \caption{\label{fig:parse-aucun} One of the parse trees assigned to
@@
 parse tree assigned to it by Frigram on Figure~\ref{fig:parse-aucun}. It
-is not enough that the embedded noun phrase $aucun\ tatou$ has a type
-that tells us it contains a negative determiner. We need this bit of
-internal information also at the level of the enclosing noun phrase
-$l'odeur\ d'aucun\ tatou$ in which the former noun phrase is embedded
+is not enough that the embedded noun phrase \emph{aucun tatou} has a
+type that tells us it contains a negative determiner. We need this bit
+of internal information also at the level of the enclosing noun phrase
+\emph{l'odeur d'aucun tatou} in which the former noun phrase is embedded
 since this is a property of the enclosing noun phrase that is crucial to
@@
 
-$$
-N_{que_1} : (NP\_NEG{=}F \limp S) \limp N\_NEG{=}F \limp N\_NEG{=}F
-$$
-$$
-N_{que_2} : (NP\_NEG{=}F \limp S) \limp N\_NEG{=}T \limp N\_NEG{=}T
-$$
-$$
-N_{de_1} : NP\_NEG{=}F \limp N\_NEG{=}F \limp N\_NEG{=}F
-$$
-$$
-N_{de_2} : NP\_NEG{=}F \limp N\_NEG{=}T \limp N\_NEG{=}T
-$$
-$$
-N_{de_3} : NP\_NEG{=}T \limp N\_NEG{=}F \limp N\_NEG{=}T
-$$
+\begin{align*}
+N_{que_1} &: (NP\_NEG{=}F \limp S) \limp N\_NEG{=}F \limp N\_NEG{=}F \\
+N_{que_2} &: (NP\_NEG{=}F \limp S) \limp N\_NEG{=}T \limp N\_NEG{=}T \\
+N_{de_1} &: NP\_NEG{=}F \limp N\_NEG{=}F \limp N\_NEG{=}F \\
+N_{de_2} &: NP\_NEG{=}F \limp N\_NEG{=}T \limp N\_NEG{=}T \\
+N_{de_3} &: NP\_NEG{=}T \limp N\_NEG{=}F \limp N\_NEG{=}T
+\end{align*}
+
+In the above types, we not only propagate information about the presence
+of a negative determiner, we also introduce a constraint in the
+mechanism stating that it cannot be the case that there is a negative
+determiner both in the modified noun and in its noun phrase complement
+(i.e. phrases like \emph{aucun tatou d'aucun homme} are not acceptable).
 
@@
 In sentence (\ref{ex:bad-neg-noneg}), we have no constant
-$N_{ne_{tv_?}}$ for $ne$ which would yield a verb capable of consuming
-two noun phrases that contain no negative determiners.
+$N_{ne_{tv_?}}$ for \emph{ne} which would yield a verb capable of
+consuming two noun phrases that contain no negative determiners.
 
@@
 constants for these two categories, but to also change the types of noun
-modifiers ($que\ ...$) and prepositions ($de$).
+modifiers (\emph{que ...}) and prepositions (\emph{de}).
 
@@
 focus on one of its properties, the fact that subjects can be extracted
-using $qui$ only from the relative clause itself whereas objects can be
-extracted using $que$ from embedded clauses contained therein, and
-present an implementation of this constraint by way of types in an
-ACG. The solution that we will show was published in
+using \emph{qui} only from the relative clause itself whereas objects
+can be extracted using \emph{que} from embedded clauses contained
+therein, and present an implementation of this constraint by way of
+types in an ACG. The solution that we will show was published in
 \cite{pogodalla2012controlling}. Our presentation is slightly adapted to
@@
 
-$$
-E_{dort_1} : NP\_VAR{=}F \limp S\_EXT{=}NO
-$$
-$$
-E_{dort_2} : NP\_VAR{=}T \limp S\_EXT{=}ROOT
-$$
-$$
-E_{aime_1} : NP\_VAR{=}F \limp NP\_VAR{=}F \limp S\_EXT{=}NO
-$$
-$$
-E_{aime_2} : NP\_VAR{=}T \limp NP\_VAR{=}F \limp S\_EXT{=}ROOT
-$$
-$$
-E_{aime_3} : NP\_VAR{=}F \limp NP\_VAR{=}T \limp S\_EXT{=}ROOT
-$$
-$$
-E_{dit\ que_1} : NP\_VAR{=}F \limp S\_EXT{=}NO \limp S\_EXT{=}NO
-$$
-$$
-E_{dit\ que_2} : NP\_VAR{=}T \limp S\_EXT{=}NO \limp S\_EXT{=}ROOT
-$$
-$$
-E_{dit\ que_3} : NP\_VAR{=}F \limp S\_EXT{=}ROOT \limp S\_EXT{=}EMB
-$$
-$$
-E_{dit\ que_4} : NP\_VAR{=}F \limp S\_EXT{=}EMB \limp S\_EXT{=}EMB
-$$
+\begin{align*}
+E_{dort_1} &: NP\_VAR{=}F \limp S\_EXT{=}NO \\
+E_{dort_2} &: NP\_VAR{=}T \limp S\_EXT{=}ROOT \\
+E_{aime_1} &: NP\_VAR{=}F \limp NP\_VAR{=}F \limp S\_EXT{=}NO \\
+E_{aime_2} &: NP\_VAR{=}T \limp NP\_VAR{=}F \limp S\_EXT{=}ROOT \\
+E_{aime_3} &: NP\_VAR{=}F \limp NP\_VAR{=}T \limp S\_EXT{=}ROOT \\
+E_{dit\ que_1} &: NP\_VAR{=}F \limp S\_EXT{=}NO \limp S\_EXT{=}NO \\
+E_{dit\ que_2} &: NP\_VAR{=}T \limp S\_EXT{=}NO \limp S\_EXT{=}ROOT \\
+E_{dit\ que_3} &: NP\_VAR{=}F \limp S\_EXT{=}ROOT \limp S\_EXT{=}EMB \\
+E_{dit\ que_4} &: NP\_VAR{=}F \limp S\_EXT{=}EMB \limp S\_EXT{=}EMB
+\end{align*}
 
 Now all of the infrastructure is in place for us to express the type of
-the relative pronouns $qui$ and $que$.
+the relative pronouns \emph{qui} and \emph{que}.
 
-$$
-E_{qui} : (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N
-$$
-$$
-E_{que_1} : (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N
-$$
-$$
-E_{que_2} : (NP\_VAR{=}T \limp S\_EXT{=}EMB) \limp N \limp N
-$$
+\begin{align*}
+E_{qui} &: (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N \\
+E_{que_1} &: (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N \\
+E_{que_2} &: (NP\_VAR{=}T \limp S\_EXT{=}EMB) \limp N \limp N
+\end{align*}
 
@@
 traces of extracted constituents will be given the type $NP\_VAR{=}T$,
-that $qui$ can only be used to extract constituents from root positions
-while $que$ allows for extraction from both root and embedded
-positions.\footnote{NB: The constraint that $qui$ can only extract
-  subjects and that $que$ can only extract objects is not enforced
-  here. Likewise, this fragment does not handle multiple extraction.}
+that \emph{qui} can only be used to extract constituents from root
+positions while \emph{que} allows for extraction from both root and
+embedded positions.\footnote{NB: The constraint that \emph{qui} can only
+  extract subjects and that \emph{que} can only extract objects is not
+  enforced here. Likewise, this fragment does not handle multiple
+  extraction.}
 
@@
 traces, we can cover another element of our running fragment, the
-preposition $de$, and enforce the constraint that $qui$/$que$ cannot
-extract prepositional complements by providing only a single type for
-$de$ that accepts only unmoved noun phrases as complements.
+preposition \emph{de}, and enforce the constraint that
+\emph{qui}/\emph{que} cannot extract prepositional complements by
+providing only a single type for \emph{de} that accepts only unmoved
+noun phrases as complements.
 
-$$
-E_{de} : NP\_VAR{=}F \limp N \limp N
-$$
+\begin{align*}
+E_{de} &: NP\_VAR{=}F \limp N \limp N
+\end{align*}
 
@@
 expressing the syntactic structure of (\ref{ex:good-ext}) but we cannot
-a find a typable term for sentence (\ref{ex:bad-ext}), where $qui$ is
-used to extract a subject from a clause embedded in a relative clause.
-The problem with sentence (\ref{ex:bad-ext}) is that the type of the
-relative clause is $NP\_VAR{=}T \limp S\_EXT{=}EMB$ which is not a valid
-argument for any constant representing the relative pronoun
-$qui$. Typing the variable $t$ with $NP\_VAR{=}F$ instead of
+a find a typable term for sentence (\ref{ex:bad-ext}), where \emph{qui}
+is used to extract a subject from a clause embedded in a relative
+clause.  The problem with sentence (\ref{ex:bad-ext}) is that the type
+of the relative clause is $NP\_VAR{=}T \limp S\_EXT{=}EMB$ which is not
+a valid argument for any constant representing the relative pronoun
+\emph{qui}. Typing the variable $t$ with $NP\_VAR{=}F$ instead of
 $NP\_VAR{=}T$ and using $E_{aime_1}$ instead of $E_{aime_2}$ would not
@@
 these pieces of information percolate up the parse tree all the way to
-the topmost level still pertinent for ensuring grammaticality. We will
-now look at another similar example.
+the topmost level still pertinent for ensuring grammaticality.
 
 \section{Agreement}
+\label{sec:agreement}
 
-Finally, we will cover the phenomenon of number agreement. This
-motivates a straightforward refinement of our types, where all noun and
-noun phrase types carry number information.
+Finally, we will look at another example of implementing a linguistic
+constraint in ACGs by covering the phenomenon of number
+agreement.
 
-$$
-A_{Marie} : NP\_NUM{=}SG
-$$
-$$
-A_{tatou} : N\_NUM{=}SG
-$$
-$$
-A_{tatous} : N\_NUM{=}PL
-$$
-$$
-A_{le} : N\_NUM{=}SG \limp ((NP\_NUM{=}SG \limp S) \limp S)
-$$
-$$
-A_{les} : N\_NUM{=}PL \limp ((NP\_NUM{=}PL \limp S) \limp S)
-$$
-$$
-A_{dort} : NP\_NUM{=}SG \limp S
-$$
-$$
-A_{dorment} : NP\_NUM{=}PL \limp S
-$$
-$$
-A_{aime_1} : NP\_NUM{=}SG \limp NP\_NUM{=}SG \limp S
-$$
-$$
-A_{aime_2} : NP\_NUM{=}SG \limp NP\_NUM{=}PL \limp S
-$$
-$$
-A_{aiment_1} : NP\_NUM{=}PL \limp NP\_NUM{=}SG \limp S
-$$
-$$
-A_{aiment_2} : NP\_NUM{=}PL \limp NP\_NUM{=}PL \limp S
-$$
-$$
-A_{qui_1} : (NP\_NUM{=}SG \limp S) \limp N\_NUM{=}SG \limp N\_NUM{=}SG
-$$
-$$
-A_{qui_2} : (NP\_NUM{=}PL \limp S) \limp N\_NUM{=}PL \limp N\_NUM{=}PL
-$$
-$$
-A_{de_1} : NP\_NUM{=}SG \limp N\_NUM{=}SG \limp N\_NUM{=}SG
-$$
-$$
-A_{de_2} : NP\_NUM{=}PL \limp N\_NUM{=}SG \limp N\_NUM{=}SG
-$$
-$$
-A_{de_3} : NP\_NUM{=}SG \limp N\_NUM{=}PL \limp N\_NUM{=}PL
-$$
-$$
-A_{de_4} : NP\_NUM{=}PL \limp N\_Pl \limp N\_NUM{=}PL
-$$
+Number agreement motivates a straightforward refinement of our types,
+where all noun and noun phrase types carry number information.
+
+\begin{align*}
+A_{Marie} &: NP\_NUM{=}SG \\
+A_{tatou} &: N\_NUM{=}SG \\
+A_{tatous} &: N\_NUM{=}PL \\
+A_{le} &: N\_NUM{=}SG \limp ((NP\_NUM{=}SG \limp S) \limp S) \\
+A_{les} &: N\_NUM{=}PL \limp ((NP\_NUM{=}PL \limp S) \limp S) \\
+A_{dort} &: NP\_NUM{=}SG \limp S \\
+A_{dorment} &: NP\_NUM{=}PL \limp S \\
+A_{aime_1} &: NP\_NUM{=}SG \limp NP\_NUM{=}SG \limp S \\
+A_{aime_2} &: NP\_NUM{=}SG \limp NP\_NUM{=}PL \limp S \\
+\end{align*}
+\begin{align*}
+A_{aiment_1} &: NP\_NUM{=}PL \limp NP\_NUM{=}SG \limp S \\
+A_{aiment_2} &: NP\_NUM{=}PL \limp NP\_NUM{=}PL \limp S \\
+A_{qui_1} &: (NP\_NUM{=}SG \limp S) \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
+A_{qui_2} &: (NP\_NUM{=}PL \limp S) \limp N\_NUM{=}PL \limp N\_NUM{=}PL \\
+A_{de_1} &: NP\_NUM{=}SG \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
+A_{de_2} &: NP\_NUM{=}PL \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
+A_{de_3} &: NP\_NUM{=}SG \limp N\_NUM{=}PL \limp N\_NUM{=}PL \\
+A_{de_4} &: NP\_NUM{=}PL \limp N\_NUM{=}PL \limp N\_NUM{=}PL
+\end{align*}
 
@@
 (\ref{ex:bad-agr}). In the latter sentence, the type of the variable $t$
-has to be $NP\_NUM{=}SG$ since it is used as an argument to $A_{dort}$. This
-means that the relative clause $dort$ has type $NP\_NUM{=}SG \limp
-S$. However, there is no type for $qui$ which would allow us to modify
-the plural noun $A_{tatous}$ of type $N\_NUM{=}PL$ by a relative clause with a
-missing singular noun phrase, which has the type $NP\_NUM{=}SG \limp S$.
+has to be $NP\_NUM{=}SG$ since it is used as an argument to
+$A_{dort}$. This means that the relative clause \emph{dort} has type
+$NP\_NUM{=}SG \limp S$. However, there is no type for \emph{qui} which
+would allow us to modify the plural noun $A_{tatous}$ of type
+$N\_NUM{=}PL$ by a relative clause with a missing singular noun phrase,
+which has the type $NP\_NUM{=}SG \limp S$.
 
@@
 
-In the next section, we will see what it takes to enforce all of the
-constraints above (negation, extraction and agreement) at the same time,
-which will give us some idea on how to go about building a grammar which
-handles a wide variety of such phenomena.
-
 \section{Putting It All Together}
 \label{sec:together}
+
+In this section, we will see what it takes to enforce all of the
+constraints introduced in the preceding sections (negation, extraction
+and agreement) at the same time, which will give us some idea on how to
+go about building a grammar which handles a wide variety of such
+phenomena.
 
@@
 appreciate what this means in practice, we give the types for the
-preposition $de$:
+preposition \emph{de}:
 
-$$
-C_{de_1} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_2} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_3} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_4} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_5} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_6} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_7} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_8} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_9} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_{10}} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
-$$
-$$
-C_{de_{11}} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
-$$
-$$
-C_{de_{12}} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
-$$
+\begin{align*}
+C_{de_1} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
+(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \\
+C_{de_2} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
+(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \\
+C_{de_3} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
+C_{de_4} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
+C_{de_5} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
+C_{de_6} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
+C_{de_7} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
+(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \\
+C_{de_8} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
+(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \\
+C_{de_9} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
+C_{de_{10}} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
+C_{de_{11}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
+C_{de_{12}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
+(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
+\end{align*}
 
@@
 of them in every type. It is no longer easy to glance at the types
-assigned to $de$ and see how it behaves with respect to, for example,
-the mechanism of negation (compare with the type assignments we gave to
-$de$ in previous subsections). Another cause of unreadability of this
-signature is due to the sheer number of types on display, which leads us
-to our second issue.
+assigned to \emph{de} and see how it behaves with respect to, for
+example, the mechanism of negation (compare with the type assignments we
+gave to \emph{de} in previous subsections). Another cause of
+unreadability of this signature is due to the sheer number of types on
+display, which leads us to our second issue.
 
@@
 about than the enumeration of 12 basic types we presented at the
-beginning of this section, however, the result is still complex in
-that it forces us to describe the behavior of three independent
-mechanisms in one place. This one place happens to be a wordform which
-has seemingly little connection to these mechanisms, the preposition
-$de$. We can therefore imagine that it will participate in many other
-phenomena and that the type we assign to it will grow beyond our
-abilities to reason about it effectively.
+beginning of this section, however, the result is still complex in that
+it forces us to describe the behavior of three independent mechanisms in
+one place. This one place happens to be a wordform which has seemingly
+little connection to these mechanisms, the preposition \emph{de}. We can
+therefore imagine that it will participate in many other phenomena and
+that the type we assign to it will grow beyond our abilities to reason
+about it effectively.
 
@@
 these are independent mechanisms. One has to examine the type to make
-sure that there is no interaction. In the next two subsections, we will
-devote ourselves to investigating ways of disentangling these.
+sure that there is no interaction.
+
+TODO: Change this after splitting the chapters.
+
+In the rest of this chapter, we will devote ourselves to investigating
+ways of disentangling these.
 
@@
 \ldots
-
-\section{Intersections of Languages}
-
-We will now turn to another solution to the complexity of grammars
-implementing multiple independent mechanisms.
-
-When ACGs are discussed, one of the oft-praised properties is
-compositionality. Let us consider a situation where we have three
-signatures $\Sigma_1$, $\Sigma_2$ and $\Sigma_3$, their respective
-distinguished types $S_1$, $S_2$ and $S_3$ and two lexicons
-$\mathcal{L}_{12} : \Sigma_1 \to \Sigma_2$ and $\mathcal{L}_{23} :
-\Sigma_2 \to \Sigma_3$ such that $\mathcal{L}_{12}(S_1) = S_2$ and
-$\mathcal{L}_{23}(S_2) = S_3$. (TODO: INCLUDE A DIAGRAM OF THIS
-SITUATION)
-
-We can now define languages $A_1$, $A_2$ and $A_3$ where $A_i$ is the
-set of terms from $\Lambda(\Sigma_i)$ that have the type $S_i$. Such
-languages, which are defined only in terms of a signature and some
-distinguished type, are called \emph{abstract languages} in the
-terminology of ACGs. They consist of terms whose structure is
-constrained only by the types assigned by the signature.
-
-Further than that, we can look at the languages $\mathcal{L}_{12}(A_1)$
-and $\mathcal{L}_{23}(A_2)$, images of the abstract languages given by
-the lexicons. Since the lexicons are homomorphic mappings w.r.t. the
-types, all elements of $\mathcal{L}_{12}(A_1)$ will have the type
-$\mathcal{L}_{12}(S_1) = S_2$ and will therefore also belong to $A_2$
-(similarly for $\mathcal{L}_{23}(A_2)$ and $A_3$). These new languages,
-which are formed by mapping an ACG abstract language using a lexicon,
-are called \emph{object languages}. Their terms are constrained not only
-by the types of the signature they are built upon but also by the type
-system of the abstract signature.
-
-Finally, we can also take the composition of lexicons $\mathcal{L}_{13}
-= \mathcal{L}_{23} \circ \mathcal{L}_{12}$, which itself is also a
-lexicon, and map it over $A_1$. This gives us a language
-$\mathcal{L}_{13}(A_1) = \mathcal{L}_{23}(\mathcal{L}_{12}(A_1))$ which
-is at the same time the image of the object language
-$\mathcal{L}_{12}(A_1)$ given by $\mathcal{L}_{23}$ and also the image
-of the abstract language $A_1$ given by $\mathcal{L}_{13}$. This points
-to an interesting property of ACG languages: they are closed under
-transformation by a lexicon.
-
-In the ACG literature, ACGs are defined as quadruples $\mathcal{G} =
-\mathopen{<}\Sigma_A, \Sigma_O, \mathcal{L}, S\mathclose{>}$ where
-$\Sigma_A$ is the abstract signature, $\Sigma_O$ is the object
-signature, $\mathcal{L} : \Sigma_A \to \Sigma_O$ is a lexicon linking
-the two and $S$ is a distinguished type of the abstract signature
-$\Sigma_A$. From this grammar, we can define two languages. The abstract
-language $\mathcal{A}(\mathcal{G}) = \{t \in \Lambda(\Sigma_A) \mid
-\ \vdash_{\Sigma_A} t : S\}$ and the object language
-$\mathcal{O}(\mathcal{G}) = \mathcal{L}(\mathcal{A}(\mathcal{G})) = \{t
-\in \Lambda(\Sigma_O) \mid \exists u \in \mathcal{A}(\mathcal{G}),
-\mathcal{L}(u) = t\}$. The composition of two ACGs $\mathcal{G}_1 =
-\mathopen{<}\Sigma_1, \Sigma_2, \mathcal{L}_{12}, S_1\mathclose{>}$ and
-$\mathcal{G}_2 = \mathopen{<}\Sigma_2, \Sigma_3, \mathcal{L}_{23},
-S_2\mathclose{>}$ is defined as $\mathcal{G}_2 \circ \mathcal{G}_1 =
-\mathopen{<}\Sigma_1, \Sigma_3, \mathcal{L}_{23} \circ \mathcal{L}_{12},
-S_1\mathclose{>}$.
-
-The definition of ACGs given above is quite practical in that it is
-complete, it gives a precise account of what set of languages are
-generated by ACGs (abstract and object languages for which there exist
-the quadruples described above). This contrasts with the presentation of
-ACGs that we gave in~\ref{sec:acg}, where we defined ACGs as
-collections of signatures with associated distinguished types and
-connected with lexicons. Our definition only introduced the requisite
-machinery, but delayed the delimitation of the way languages can be
-defined. Leaving this question open will let us answer it now by
-providing an extension/generalization of ACGs which will help us
-overcome the modularity issues that have been the theme of this
-chapter.\footnote{Incidentally, this was not the reason for the way we
-  presented ACGs in~\ref{sec:acg}. We chose this style as it grouped
-  together the information in a more natural way. We approach signatures
-  as descriptions of domains (strings, syntax, semantics), each one
-  being associated with a distinguished type telling us which objects
-  the domain has set out to describe (strings, sentences, truth
-  values). Having distinguished types directly attached to signatures
-  allows us to omit them from the ACG quadruples. However, this only
-  leaves us with triples containing the two signatures and the lexicon
-  connecting them, and since we can infer the signatures themselves from
-  the domain and the range of the lexicon, we do not have much
-  motivation to introduce these tuples. Instead of composing ACGs, we
-  compose just the lexicons.
-
-On top of that, we will want to define signatures and lexicons that
-yield different object languages (at least some string representation of
-the form and some semantic representation of the meaning), which would
-mean having at least two different ACGs which are linked (e.g. by
-sharing the same abstract signature). However, since these two grammars
-are very closely related, we would prefer to reason about our grammar as
-\emph{one} system/graph of signatures and lexicons which defines
-multiple languages instead of talking about \emph{multiple} grammars
-which share some similarities.
-
-TODO: Somehow express that its nice to have the links between the
-signatures and lexicons that are shown in diagrams \emph{explicit}
-rather than having them \emph{implicit}.}
-
-Before we turn to our proposal, we will quickly recall some of the other
-styles of composition used in ACGs (a more detailed exposition can be
-found in \cite{pogodalla2012controlling}). Besides making the object
-signature of one grammar be the abstract signature of another as we saw
-above (TODO: INSERT REFERENCE TO FUTURE DIAGRAM), we can have two
-grammars share the same abstract signature allowing us to transduce
-between terms in the two object signatures. See Figure~(TODO: INSERT A
-DIAGRAM FOR THIS SITUATION) for a diagrammatic example.
-
-These two modes of composition can be fruitfully mixed. Consider (TODO:
-INSERT A DIAGRAM), where we introduce a syntactic signature TODO where
-different terms are only allowed to yield the same string whenever they
-represent different syntactic structures of a syntactically ambiguous
-phrase. This means that purely semantic distinctions such as scope
-ambiguity have to be moved to a more abstract signature from whose terms
-the semantic representation can be derived by a function. This approach
-is studied more deeply in \cite{pogodalla2007generalizing}.
-
-Another interesting composition pattern is to take an existing structure
-of grammars and add a new abstract signature on top. This signature can
-be used to further constrain the items in (TODO: what was before the
-abstract-most signature/language). In \cite{pogodalla2012controlling},
-this is what the authors use to develop their constraints on
-extraction. Each of their constraints is presented as a new abstract
-signature constraining the original syntactic signature.
-
-However, a question that the authors do not pose is how to combine all
-these constraints. If we were to apply the same pattern repeatedly to
-get a structure like the one on (TODO: INSERT A DIAGRAM HERE), we would
-have to deal with successively more complex type systems on each
-level. This is due to the condition that a lexicon has to be a
-homomorphism and thus it must, among other things, always map well-typed
-terms to well-typed terms. To illustrate the problem, imagine a
-structure such as the one on (TODO) that handles some different
-constraint in each of the vertically stacked signatures, including
-e.g. agreement. If we tried to add a new abstract signature at the top
-to implement another constraint, we would have to ensure that all the
-terms we licence in our new abstract signature have a well-typed
-counterpart in the existing abstract language. However, this means that
-our new signature must make sure that the terms it approves do not break
-agreement or any of the previously established constraints. Effectively,
-our new signature would be forced to reimplement all of the existing
-constraints, rendering in vain all our effort to split up the
-constraints into multiple signatures.
-
-What we would like instead is to put the signatures side-by-side such
-that they do not depend on each other (as on diagram TODO). However, how
-does this actually define a language? We can arrive at the answer by
-examining how are the graphical representations of ACGs connected to the
-languages they define. We will associate a language to every node of the
-graph. Before we begin, we notice that every ACG diagram formed by the
-above rules of composition is a directed tree and therefore every node
-has at most one parent/predecessor. To the root of the tree, the
-abstract-most signature in the graph, we will assign the abstract
-language generated by that signature and its distinguished type. To all
-the other nodes, we will assign the result of mapping their parent's
-language by the lexicon linking the two. By homomorphism of the
-lexicons, we can also deduce the distinguished types for all the
-non-root nodes. They are the images of their parents' distinguished
-types by the connecting lexicons. For all of the nodes, it is true that
-every term in the language assigned to the node has the node's
-distinguished type.
-
-We can confirm that the above graphical model of defining ACGs is
-coherent. First, we can show that our interpretation is sound and so
-that any directed tree adorned with signatures on the nodes, compatible
-lexicons on the edges and a distinguished type on the root can only
-define ACG languages. This is trivial, since the language corresponding
-to the root is an ACG abstract language and the class of ACG languages
-is closed under transformation by a lexicon and therefore all of the
-languages corresponding to the descendants of the root are ACG languages
-as well. Second, we can try and show some kind of completeness. However,
-there is not much formalization done to describe a ``system of ACGs that
-share common signatures'' (something that our generalization will
-fix). We can nevertheless show that it is possible to express any ACG
-using our graphical model. Given a grammar $\mathcal{G} =
-\mathopen{<}\Sigma_A, \Sigma_O, \mathcal{L}, S\mathclose{>}$, we just
-produce a tree whose root has the signature $\Sigma_A$ and distinguished
-type $S$ and which is connected to a single child node having the
-signature $\Sigma_O$ by an edge labelled by the lexicon
-$\mathcal{L}$. Then the two nodes of our tree define both the abstract
-and the object language defined by $\mathcal{G}$.
-
-We have now given some formal account of how we can read language
-definitions in diagrams. Still, we have kept our analysis to the tree
-diagrams used in current ACG literature, where every node has at most
-one parent. To make the generalization more visible, we will switch from
-an algebraic manner of presentation, in which we stated that the
-language of a node is equal to an image of its parent's language, to a
-calculus-based manner of presentation, in which we will state that the
-language of a node is the set of all terms that have an antecedent in
-the parent's language. Besides this property, there is also the fact
-that every term in a node's language has the node's distinguished
-type. This property is trivially true for nodes with parents because of
-the homomorphic property of lexicons. However, it is crucial for the
-root node for which this is the single defining property of its
-language. To sum up, we have that for every node of a tree diagram, its
-language is the set of all terms that have the node's distinguished type
-and that have an antecedent in the parent's language, if there is
-any. This rule is already general enough to handle the case of zero or
-one parents, we can easily generalize it to handle an arbitrary count:
-\emph{the language of a node is the set of all terms that have the
-  node's distinguished type and that have antecedents in the languages
-  of all of the node's parents}. If we transpose this back to the
-algebraic style of presentation, we say that the language of a node is
-the intersection of the set of terms having the node's distinguished
-type and the intersection of all its parents' languages.
-
-TODO: In the previous paragraph, add a notice saying that the terms must
-be also built on the signature of the node.
-
-TODO: Use the correct term arborescence in the above paragraphs!
-% http://en.wikipedia.org/wiki/Arborescence_(graph_theory)
-
-We will now turn to a formal definition of this generalized graph-based
-formalism and discuss some its values and properties.
-
-\subsection{Graphical Abstract Categorial Grammars}
-
-We define a \emph{graphical abstract categorial grammar} as a quadruple
-$\mathcal{G} = \mathopen{<} G, \Sigma, S, {\mathcal{L}} \mathclose{>}$
-where $G$ is a directed graph with vertices $V(G)$ and edges $E(G)$,
-$\Sigma$ and $S$ are labelings assigning signatures and distinguished
-types, respectively, to the vertices in $V$ and ${\mathcal{L}}$ is a
-labeling assigning lexicons to the edges in $E$. Furthermore, a
-well-formed graphical ACG satisfies the following conditions:
-
-\begin{itemize}
-\item $G$ is a directed acyclic graph.
-\item For all $(u,v) \in E(G)$, $\mathcal{L}_{(u,v)} : \Sigma_u \to
-  \Sigma_v$.
-\item For all $(u,v) \in E(G)$, $\mathcal{L}_{(u,v)}(S_u) = S_v$.
-\end{itemize}
-
-We will say that a node $u$ is \emph{more abstract} than a node $v$
-whenever $(u,v)$ belongs to the strict partial order induced by the
-edges of $G$ (i.e. there is a path from $u$ to $v$).
-
-We let the nodes in a graphical ACG $\mathcal{G} = \mathopen{<} G,
-\Sigma, S, \mathcal{L} \mathclose{>}$ define two kinds of
-languages. The \emph{intrinsic languages} $\mathcal{I}_{\mathcal{G}}$,
-which are constrained only by the type signature described in the node
-itself, and the \emph{extrinsic languages} $\mathcal{E}_{\mathcal{G}}$,
-which are also constrained by type systems in more abstract nodes of the
-graph. These two notions correspond to those of abstract languages and
-object languages in ACGs.
-
-$$
-\mathcal{I}_{\mathcal{G}}(v) = \{t \in \Lambda(\Sigma_v)
-\mid\ \vdash_{\Sigma_u} t : S_v\}
-$$
-$$
-\mathcal{E}_{\mathcal{G}}(v) = \mathcal{I}_{\mathcal{G}}(v) \cap
-\bigcap_{(u,v) \in E} \mathcal{L}_{(u,v)}(\mathcal{E}_{\mathcal{G}}(u)).
-$$
-
-From the above definition, we see that the ACG diagrams we introduced
-before are a special case of graphical ACGs where the graph is an
-arborescence.
-
-\subsubsection{On the Expressivity of Graphical ACGs}
-
-\begin{observation}
-  The set of intrinsic languages definable by graphical ACGs is the same
-  as the set of abstract languages definable by ACGs.
-\end{observation}
-
-\begin{theorem}
-  The set of extrinsic languages definable by graphical ACGs is the set
-  of object languages definable by ACGs closed under intersection and
-  transformation by a lexicon.
-\end{theorem}
-
-\begin{proof}
-  First, we prove that an extrinsic language can be always constructed
-  from object languages by intersections and transformations by
-  lexicons. Let us consider any graphical ACG $\mathcal{G} =
-  \mathopen{<} G, \Sigma, S, \mathcal{L} \mathclose{>}$ and any
-  topological ordering $v_1, \ldots, v_n$ of the nodes of $G$ (an
-  ordering such that more abstract nodes always precede less abstract
-  ones). Before we start, we will just note that the all the intrinsic
-  languages defined by $v_i$ are abstract languages which are a special
-  case of object languages.
-
-  \begin{itemize}
-    \item For $v_1$, the case is trivial since its extrinsic language is
-      just its intrinsic language as the node $v_1$ has no predecessors
-      in $G$.
-    \item For any other node $v_n$, we look at the definition of an
-      extrinsic language and remark that the only operators are
-      intersection and application of a lexicon with the operands being
-      the intrinsic language of $v_n$ and the extrinsic languages of
-      more abstract nodes. By induction hypothesis, we have that all of
-      the operands have been constructed from object languages using
-      intersections and transformations by lexicons and therefore even
-      the extrinsic language defined by $v_n$ has this property.
-  \end{itemize}
-
-  Now we have to prove the converse. We do so by showing that extrinsic
-  languages contain object languages and are closed under intersection
-  and transformation by a lexicon.
-
-  \begin{itemize}
-    \item Let $L$ be an object language defined by the ACG $\mathopen{<}
-      \Sigma_A, \Sigma_O, \mathcal{L}, S\mathclose{>}$. We take the
-      graph $G$ with $V(G) = \{A, O\}$ and $E(G) = \{(A,O)\}$, we label
-      the nodes $A$ and $O$ with $\Sigma_A$ and $\Sigma_O$ as signatures
-      and $S$ and $\mathcal{L}(S)$ as distinguished types
-      respectively. Finally, we label the edge $(A,O)$ with
-      $\mathcal{L}$ and we have a graphical ACG in which the node $O$
-      defines the extrinsic language $L$.
-
-    \item Let $L_1$ and $L_2$ be two extrinsic languages defined by the
-      nodes $u_1$ and $u_2$, respectively, and belonging to some two
-      graphical ACGs. We can construct a new graphical ACG by taking the
-      union of the two ACGs (union of graphs (union of vertices and
-      edges) and unions of the labelings).\footnote{Assuming, without
-        loss of generality, that the sets of vertices of the two graphs
-        are disjoint.} We then add a new node $v$ which will define the
-      language $L_1 \cap L_2$.
-
-      The signature labeling $v$ will be the union of the two signatures
-      labeling $u_1$ and $u_2$ such that the intrinsic language of $v$
-      will be a superset of the intrinsic languages, and by extension
-      the extrinsic languages, of both $u_1$ and $u_2$. Either of the
-      two distinguished types that label $u_1$ or $u_2$ can be used to
-      label $v$ (if they are actually different, then the intersection
-      is trivially empty). Finally, we will add two edges leading from
-      $u_1$ and $u_2$ to $v$, both labeled with the identity lexicon. We
-      will now show that the extrinsic language of $v$ in the newly
-      formed graphical ACG $\mathcal{G}'$ is $L_1 \cap L_2$.
-
-      \begin{align*}
-        \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v)
-        \cap \mathcal{L}_{(u_1,v)}(\mathcal{E}_{\mathcal{G}'}(u_1)) \cap
-        \mathcal{L}_{(u_2,v)}(\mathcal{E}_{\mathcal{G}'}(u_2)) \\ &=
-        \mathcal{I}_{\mathcal{G}'}(v) \cap
-        \mathcal{E}_{\mathcal{G}'}(u_1) \cap
-        \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
-        \mathcal{E}_{\mathcal{G}'}(u_1) \cap
-        \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
-        \mathcal{E}_{\mathcal{G}}(u_1) \cap
-        \mathcal{E}_{\mathcal{G}}(u_2) \\ &= L_1 \cap L_2
-      \end{align*}
-
-      The simplifications use the facts that: the two new lexicons are
-      identities, that the intrinsic language of $v$ is a superset of
-      the extrinsic languages of $u_1$ and $u_2$, that the extrinsic
-      languages of $u_1$ and $u_2$ are the same in the new graphical ACG
-      as in the two original graphical ACGs and finally that $L_1$ and
-      $L_2$ are the extrinsic languages defined by $u_1$ and $u_2$.
-
-  \item Let $L \subset \Lambda(\Sigma_A)$ be an extrinsic language and
-    $\mathcal{L} : \Sigma_A \to \Sigma_O$ a lexicon from its signature
-    $\Sigma_A$ to the signature $\Sigma_O$. We will show that the
-    language $\mathcal{L}(L) \subset \Lambda(\Sigma_O)$ is also an
-    extrinsic language. Let $u$ be the node that defines the extrinsic
-    language $L$ in some graphical ACG $\mathcal{G}$. We construct a new
-    graphical ACG by adding a new vertex $v$ to $\mathcal{G}$ labeled
-    with the signature $\Sigma_O$ and the distinguished type
-    $\mathcal{L}(S)$ where $S$ is the distinguished type labeling
-    $u$. To this graph, we conjoin the edge $(u,v)$ and label it with
-    $\mathcal{L}$. Now we verify that the extrinsic language of $v$ is
-    the language $\mathcal{L}(L).$
-
-    \begin{align*}
-      \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v) \cap
-      \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
-      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
-      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}}(u)) \\
-      &= \mathcal{L}(L)
-    \end{align*}
-
-    The facts used here are in turn: the inclusion of
-    $\mathcal{I}_{\mathcal{G}'}(v)$ in
-    $\mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u))$, the observation that
-    modifying our graphical grammar did not change the extrinsic
-    language of $u$ and the assumption that $L$ is the extrinsic
-    language of $u$.
-  \end{itemize}
-\end{proof}
-
-\begin{corollary}
-  The set of extrinsic languages definable by graphical ACGs is the same
-  as the set of object languages definable by ACGs if and only if the
-  set of object languages definable by ACGs is closed under
-  intersection.
-\end{corollary}
-
-We conclude this subsubsection with a few words to motivate the need we
-feel for this (possibly) more expressive formalism. We believe that in
-order to manage the development of wide-coverage grammars, the grammars
-need to be kept as simple as possible, which means, among other things,
-that independent constraints should be implemented separately and not
-complected together (we saw the difficulties caused by this already on a
-microscopic fragment in section~\ref{sec:together}). This means that if
-we were to define, e.g., a partial grammar in which we would only
-enforce agreement and another one in which we would only enforce the
-proper use of relative pronouns, we would like the language in which
-both agreement and proper use of relative pronouns are respected to be
-expressible in our formalism and on top of that, we would like this
-language to be easily expressed in terms of the two grammars for the
-individual constraints. We now have a requirement on the formalism in
-which we would write such grammars, we want a formalism where the set of
-all the ``useful'' definable languages is closed on intersection. If the
-object languages of ACGs are indeed closed on intersection, then our
-extension is just a convenience for expressing intersection more
-easily. However, if they are not, we believe our extension adds a
-critical level of expressivity needed for simple and modular
-wide-coverage grammars.
-
-\subsubsection{On the Decidability of Graphical ACGs}
-
-We turn to the problem of parsing with graphical ACGs. Parsing with ACGs
-is a problem on the limits of decidability. When we consider ACGs whose
-abstract signatures only contain types of order 2 or less, parsing is
-possible in polynomial type for both linear \cite{salvati2005problemes}
-and non-linear \cite{kanazawa2007parsing} type systems (such as the one
-we introduced in~\ref{sec:acg}). If the order of the abstract signature
-exceeds 2, the membership problem for linear ACGs becomes decidable if
-and only if the satisfiability problem in the multiplicative exponential
-fragment linear logic is decidable \cite{de2004vector}, which is an open
-problem. On the other hand, by extending the type system to include
-dependent types, parsing can become undecidable too since type checking
-in such a system is itself undecidable.
-
-Our intent here is to show that by extending the formalism, we have not
-made the problem of parsing with the grammar more difficult, at least in
-terms of decidability if not in terms of complexity. Let us imagine a
-graphical ACG where every edge (two signatures, distinguished types and
-a lexicon) is an ACG for which there is an effective parsing algorithm
-and all type checking is decidable in all the signatures. Then we can
-build a simple effective parsing algorithm for any of the extrinsic
-languages defined in the graphical ACG.
-
-We will be testing whether the term $\gamma$ belongs to the extrinsic
-language of some node $v$. Our algorithm first checks whether $\gamma$
-belongs to the intrinsic language of $v$ by checking its type. If it
-does, it proceeds to verify that it is also in the extrinsic language by
-finding antecedents in all of its abstract predecessors $u_i$, if there
-are any. We can use the presupposed parsing algorithm for the ACG
-$\mathopen{<} \Sigma_{u_i}, \Sigma_v, \mathcal{L}_{(u_i,v)},
-S_{u_i}\mathclose{>}$ to find all the abstract terms $\alpha_{i,j}$
-having type $S_{u_i}$ and being antecedents to $\gamma$ w.r.t. the
-lexicon $\mathcal{L}_{(u_i,v)}$. Such terms belong to the intrinsic
-language of $u_i$ but we are interested only in antecedents coming from
-the extrinsic language. Therefore, what we have to do is to recursively
-apply our parsing algorithm to all of the antecedent-candidates
-$\{\alpha_{i,j} \mid \forall j\}$ until we find one which belongs to the
-extrinsic language of $u_i$. Since the graph is acyclic, we are
-guaranteed that this recursive traversal will terminate. We repeat this
-for all the predecessors $u_i$ and if we succeed to find antecedents in
-all of them, we have confirmed the membership of $\gamma$ in
-$\mathcal{E}_{\mathcal{G}}(v)$ and also built its abstract
-antecedent(s).
-
-\subsubsection{On Grammar Engineering with Graphical ACGs}
-
-%% \subsubsection{On Alternative Interpretations of Graphical ACGs}
-%
-% Mention the example with the levels of linguistic representations
-% linked by many-to-many transductions and explain how a different
-% theory of graphs defining languages might be suitable to that
-% architecture.
-%
-% Also point out the weirdness of the diamond shape.

### 2013-06-17 front-page.tex
Last night shift

\begin{document}

\begin{titlepage}
\begin{center}

\newcommand{\HRule}{\includegraphics[width=\textwidth]{images/hrule.png}}

%\includegraphics[width=0.9\textwidth]{images/UL_UFR_MI.jpg}\\[2cm]

\textsc{\LARGE Master Sciences Cognitives et Applications\\Sp\'ecialit\'e Traitement Automatique du Langage}\\[1.5cm]

\textsc{\Large Master Thesis}\\[0.65cm]

\HRule \\[0.6cm]
{ \huge \bfseries Towards a Wide-Coverage Grammar: Graphical Abstract Categorial Grammars }\\[0.4cm]

\HRule \\[1.75cm]

\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
Ji\v{r}\'{i} \textsc{Mar\v{s}\'{i}k}\\
~
\end{flushleft}
\end{minipage}
\begin{minipage}{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
Maxime \textsc{Amblard} \\
~
\end{flushright}
\end{minipage}
~\\[2.1cm]
%\includegraphics[width=0.35\textwidth]{images/logoLORIA.jpg}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\includegraphics[width=0.35\textwidth]{images/tonequipe.png}

\end{center}
\end{titlepage}

%\maketitle

\newpage
~~ \pagenumbering{gobble}
\newpage
\pagenumbering{arabic}
\setcounter{page}{3}

### 2013-06-17 gacg.tex
Last night shift

--- 2013-06-16 the-last-part.tex
@@
-\chapter{The Last Part}
-
-In this chapter, we will attempt to incorporate accounts of several
-linguistic phenomena in a single framework, highlight the modularity
-issues this brings up and propose some solutions to them.
-
-\section{Negation}
-\label{sec:negation}
-
-We will start with an account of the paired grammatical words mechanism
-for negation in French. Recall the EPTDs we have seen
-in~\ref{ssec:frigram}, repeated here on Figure~\ref{fig:ig-neg-rep}.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.25]{ig-neg.pdf}
-  \caption{\label{fig:ig-neg-rep} Frigram EPTDs for a negative
-    determiner and the paired grammatical word $ne$.}
-\end{figure}
-
-At a basic level, $aucun$ functions like any other determiner and $ne$
-as any verb modifier. However, we want to encode the constraint that
-whenever a verb is modified by $ne$, the modified verb will demand that
-one of its arguments contains a negative determiner. Since we need to
-make a distinction between phrases that contain negative determiners and
-those that do not, we will need to provide different types for both (two
-terms with the same type are indistinguishable w.r.t. syntactic
-composition in a type-logical grammar such as ACGs). Our type for
-$aucun$ will thus end up being $N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S)
-\limp S)$ instead of the usual $N \limp ((NP \limp S) \limp
-S)$.\footnote{Strictly following the EPTD for $aucun$ would lead us to
-  give it a type $DET\_NEG{=}F$ and then give the two types
-  $DET\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S)$ and $DET\_NEG{=}T
-  \limp ((NP\_NEG{=}T \limp S) \limp S)$ to nouns. In this
-  demonstration, we prefer to align our examples with the treatment
-  which is customary to ACG research.} The type for $le$ will need to be
-able handle cases where its argument noun either already contains a
-negative determiner or not.
-
-\begin{align*}
-N_{aucun} &: N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S) \limp S) \\
-N_{le_1} &: N\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S) \\
-N_{le_2} &: N\_NEG{=}T \limp ((NP\_NEG{=}T \limp S) \limp S) \\
-\end{align*}
-
-The type of $ne$ will be more verbose, as it will consume a verb and
-transform its valency so that it demands that exactly one of its $NP$
-arguments contains a negative determiner.
-
-\begin{align*}
-N_{ne_{tv_1}} &: (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}F \limp S) \\
-N_{ne_{tv_2}} &: (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}F \limp NP\_NEG{=}T \limp S)
-\end{align*}
-
-Furthermore, we could add the case when both the subject and the object
-contain negative determiners, such as in~(\ref{ex:aucun2}). This case is
-currently not covered by Frigram (though it could be), but is considered
-grammatical by French speakers. Finally, in a complete grammar, $ne$
-would have to provide types capable of transforming all the other
-syntactic valencies.
-
-\begin{exe}
-  \ex \label{ex:aucun2} Aucune fourmi n'aime aucun tatou.
-\end{exe}
-
-\begin{align*}
-N_{ne_{tv_3}} &: (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}T \limp S)
-\end{align*}
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.25]{parse-aucun.pdf}
-  \caption{\label{fig:parse-aucun} One of the parse trees assigned to
-    sentence (\ref{ex:aucun-obj}) by Frigram/Leopar.}
-\end{figure}
-
-However, this is not the only change we have to effect on our grammar in
-order to properly handle the paired grammatical words for
-negation. Consider the example sentence (\ref{ex:aucun-obj}) and the
-parse tree assigned to it by Frigram on Figure~\ref{fig:parse-aucun}. It
-is not enough that the embedded noun phrase $aucun\ tatou$ has a type
-that tells us it contains a negative determiner. We need this bit of
-internal information also at the level of the enclosing noun phrase
-$l'odeur\ d'aucun\ tatou$ in which the former noun phrase is embedded
-since this is a property of the enclosing noun phrase that is crucial to
-its syntactic combinatorics. This presupposes the existence of a
-mechanism for propagating this kind of information up the parse tree. In
-the case of our ACG, this means that all the functions which modify
-phrases must be aware whether or not their argument contains a negative
-determiner and to convey this information in its result type as
-well. Moreover, since the negative determiner can also come from a
-complement of a prepositional phrase, the types of prepositions will
-need to take into account the presence of negative determiners in both
-their complements and the phrases they modify.
-
-\begin{align*}
-N_{que_1} &: (NP\_NEG{=}F \limp S) \limp N\_NEG{=}F \limp N\_NEG{=}F \\
-N_{que_2} &: (NP\_NEG{=}F \limp S) \limp N\_NEG{=}T \limp N\_NEG{=}T \\
-N_{de_1} &: NP\_NEG{=}F \limp N\_NEG{=}F \limp N\_NEG{=}F \\
-N_{de_2} &: NP\_NEG{=}F \limp N\_NEG{=}T \limp N\_NEG{=}T \\
-N_{de_3} &: NP\_NEG{=}T \limp N\_NEG{=}F \limp N\_NEG{=}T
-\end{align*}
-
-If we suppose the presence of the appropriate lexical items typed as in
-the examples of~\ref{sec:acg}, we have a type system in which we can
-type the syntactic terms corresponding to the sentences
-(\ref{ex:good-neg-double}), (\ref{ex:good-neg-embed}) and
-(\ref{ex:good-neg-rel}) while making it impossible to type the
-structures corresponding to sentences (\ref{ex:bad-neg-noneg}) and
-({\ref{ex:bad-neg-rel}}).
-
-In sentence (\ref{ex:bad-neg-noneg}), we have no constant
-$N_{ne_{tv_?}}$ for $ne$ which would yield a verb capable of consuming
-two noun phrases that contain no negative determiners.
-
-In sentence (\ref{ex:bad-neg-rel}), there are multiple problems with
-typing. First, $N_{chasse}$ requires its first argument to be of type
-$NP\_NEG{=}F$, a noun phrase with no negative determiner. This means
-that $\lambda^{\circ} x.\ N_{chasse}\ x\ y$ has type $NP\_NEG{=}F \limp
-S$ which makes it an invalid argument for the function
-$N_{aucun}\ N_{loup}$, which has type $(NP\_NEG{=}T \limp S) \limp
-S$. Even if we were able to assign it a type compatible with
-$N_{que_1}$, the final relative clause would map $N_{tatou}$ to another
-term of type $N\_NEG{=}F$, which has no ``free'' negative
-determiners. $N_{le_1}$ will map this to a term typed $(NP\_NEG{=}F
-\limp S) \limp S$. Finally, this term is not capable of taking
-$(\lambda^{\circ} x.\ N_{ne_{iv}}\ N_{court}\ x)$ as its argument, since
-its type is $NP\_NEG{=}T \limp S$.
-
-\begin{exe}
-  \ex \label{ex:good-neg-double} Aucune fourmi n'aime aucun tatou. \\
-      $(N_{aucune}\ N_{fourmi})\ (\lambda^{\circ} x.\ (N_{aucun}\ N_{tatou})\ (\lambda^{\circ} y.\ N_{ne_{tv_3}}\ N_{aime}\ x\ y))$
-  \ex \label{ex:good-neg-embed} Jean n'aime l'odeur d'aucun tatou. \\
-      $(N_{le_2}\ (N_{de_3}\ (N_{aucun}\ N_{tatou})\ N_{odeur}))\ (\lambda^{\circ} y.\ N_{ne_{tv_2}}\ N_{aime}\ N_{Jean}\ y)$
-  \ex \label{ex:good-neg-rel} Le tatou qu'aucun loup ne chasse court. \\
-      $(N_{le_1}\ (N_{que_1}\ (\lambda^{\circ} y.\ (N_{aucun}\ N_{loup})\ (\lambda^{\circ} x.\ N_{ne_{tv_1}}\ N_{chasse}\ x\ y))\ N_{tatou}))\ (\lambda^{\circ} x.\ N_{court}\ x)$
-  \ex * \label{ex:bad-neg-noneg} Jean n'aime le tatou. \\
-      * $(N_{le_1}\ N_{tatou})\ (\lambda^{\circ} y.\ N_{ne_{tv_?}}\ N_{aime}\ N_{Jean}\ y)$
-  \ex * \label{ex:bad-neg-rel} Le tatou qu'aucun loup chasse ne court. \\
-      * $(N_{le_1}\ (N_{que}\ (\lambda^{\circ} y.\ (N_{aucun}\ N_{loup})\ (\lambda^{\circ} x.\ N_{chasse}\ x\ y))\ N_{tatou}))\ (\lambda^{\circ} x.\ N_{ne_{iv}}\ N_{court}\ x)$
-\end{exe}
-
-There are two things that we would like to highlight about this
-grammatical treatment. First off, we had to refine our types to make
-them convey more than one piece of information in a single type
-(e.g. something is a noun phrase and at the same time it is something
-that contains a negative determiner). Second, a property that we wish to
-express in the type of a term may be due to one of its subconstituents
-and it is therefore necessary to encode the propagation of this property
-up the parse tree in the types of all the intervening operators. In our
-case, it meant that handling negative determiners and the paired
-grammatical word for negation required us to not only define and type
-constants for these two categories, but to also change the types of noun
-modifiers ($que\ ...$) and prepositions ($de$).
-
-\section{Extraction}
-\label{sec:extraction}
-
-We will now turn to another phenomenon, extraction in relative clauses,
-focus on one of its properties, the fact that subjects can be extracted
-using $qui$ only from the relative clause itself whereas objects can be
-extracted using $que$ from embedded clauses contained therein, and
-present an implementation of this constraint by way of types in an
-ACG. The solution that we will show was published in
-\cite{pogodalla2012controlling}. Our presentation is slightly adapted to
-conform to the patterns seen in previous examples and it does not use
-dependent types as they are not essential to this specific constraint.
-
-The key idea in the treatment presented below is to give a special type
-to extracted constituents. We will thus have two types for noun phrases:
-$NP\_VAR{=}T$, which stands for ``empty'' noun phrases introduced by
-extraction, and $NP\_VAR{=}F$, which stands for the other noun phrases,
-i.e.  proper names and nouns with determiners. The grammar gives no
-constant capable of constructing a value of type $NP\_VAR{=}T$. The only
-way to obtain a term with this type is to abstract over a variable
-having the type. Then, the types of relative pronouns end up being
-$(NP\_VAR{=}T \limp S) \limp N \limp N$, ensuring that the incomplete
-clause given as the first argument is a $\lambda$-abstraction whose
-argument has type $NP\_VAR{=}T$.
-
-Once we have this kind of information in the type system, we can start
-distinguishing clauses containing rooted extraction, where a subject or
-object is extracted directly from the relative clause, embedded
-extraction, where an object is extracted from a clause embedded inside
-the relative clause, or no extraction. Types of functions which produce
-values of type $S$ will be sensitive to the presence of extracted
-variables in their noun phrase and clausal arguments.
-
-\begin{align*}
-E_{dort_1} &: NP\_VAR{=}F \limp S\_EXT{=}NO \\
-E_{dort_2} &: NP\_VAR{=}T \limp S\_EXT{=}ROOT \\
-E_{aime_1} &: NP\_VAR{=}F \limp NP\_VAR{=}F \limp S\_EXT{=}NO \\
-E_{aime_2} &: NP\_VAR{=}T \limp NP\_VAR{=}F \limp S\_EXT{=}ROOT \\
-E_{aime_3} &: NP\_VAR{=}F \limp NP\_VAR{=}T \limp S\_EXT{=}ROOT \\
-E_{dit\ que_1} &: NP\_VAR{=}F \limp S\_EXT{=}NO \limp S\_EXT{=}NO \\
-E_{dit\ que_2} &: NP\_VAR{=}T \limp S\_EXT{=}NO \limp S\_EXT{=}ROOT \\
-E_{dit\ que_3} &: NP\_VAR{=}F \limp S\_EXT{=}ROOT \limp S\_EXT{=}EMB \\
-E_{dit\ que_4} &: NP\_VAR{=}F \limp S\_EXT{=}EMB \limp S\_EXT{=}EMB
-\end{align*}
-
-Now all of the infrastructure is in place for us to express the type of
-the relative pronouns $qui$ and $que$.
-
-\begin{align*}
-E_{qui} &: (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N \\
-E_{que_1} &: (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N \\
-E_{que_2} &: (NP\_VAR{=}T \limp S\_EXT{=}EMB) \limp N \limp N
-\end{align*}
-
-The above types ensure that the abstracted variables corresponding to
-traces of extracted constituents will be given the type $NP\_VAR{=}T$,
-that $qui$ can only be used to extract constituents from root positions
-while $que$ allows for extraction from both root and embedded
-positions.\footnote{NB: The constraint that $qui$ can only extract
-  subjects and that $que$ can only extract objects is not enforced
-  here. Likewise, this fragment does not handle multiple extraction.}
-
-We also note that the splitting of the clause type $S$ into three finer
-types $S\_EXT={NO}$, $S\_EXT{=}ROOT$ and $S\_EXT{=}EMB$ means that types
-that work with $S$ regardless of its extraction status will have to
-provide alternatives for all the possible variants.
-
-\begin{align*}
-E_{le_1} &: N \limp ((NP\_VAR{=}F \limp S\_EXT{=}NO) \limp S\_EXT{=}NO) \\
-E_{le_2} &: N \limp ((NP\_VAR{=}F \limp S\_EXT{=}ROOT) \limp S\_EXT{=}ROOT) \\
-E_{le_3} &: N \limp ((NP\_VAR{=}F \limp S\_EXT{=}EMB) \limp S\_EXT{=}EMB)
-\end{align*}
-
-While we have introduced the notion of an $NP\_VAR{=}T$ for describing
-traces, we can cover another element of our running fragment, the
-preposition $de$, and enforce the constraint that $qui$/$que$ cannot
-extract prepositional complements by providing only a single type for
-$de$ that accepts only unmoved noun phrases as complements.
-
-\begin{align*}
-E_{de} &: NP\_VAR{=}F \limp N \limp N
-\end{align*}
-
-Given this type system, we can assign the type $S\_EXT{=}NO$ to the term
-expressing the syntactic structure of (\ref{ex:good-ext}) but we cannot
-a find a typable term for sentence (\ref{ex:bad-ext}), where $qui$ is
-used to extract a subject from a clause embedded in a relative clause.
-The problem with sentence (\ref{ex:bad-ext}) is that the type of the
-relative clause is $NP\_VAR{=}T \limp S\_EXT{=}EMB$ which is not a valid
-argument for any constant representing the relative pronoun
-$qui$. Typing the variable $t$ with $NP\_VAR{=}F$ instead of
-$NP\_VAR{=}T$ and using $E_{aime_1}$ instead of $E_{aime_2}$ would not
-help, since the resulting relative clause would then have type
-$NP\_VAR{=}F \limp S\_EXT{=}NO$, which would not be admitted by any
-relative pronoun.
-
-\begin{exe}
-  \ex * \label{ex:bad-ext} Le tatou qui Jean dit que \_ aime Marie dort. \\
-      * $(E_{le_1}\ (E_{qui_?}\ (\lambda^{\circ} t.\ E_{dit\ que_3}\ E_{Jean}\ (E_{aime_2}\ t\ E_{Marie}))\ E_{tatou}))\ (\lambda^{\circ} x.\ E_{dort_1}\ x)$
-  \ex \label{ex:good-ext} Le tatou que Jean dit que Marie aime \_ dort. \\
-      $(E_{le_1}\ (E_{que_2}\ (\lambda^{\circ} t.\ E_{dit\ que_3}\ E_{Jean}\ (E_{aime_3}\ E_{Marie}\ t))\ E_{tatou}))\ (\lambda^{\circ} x.\ E_{dort_1}\ x)$
-\end{exe}
-
-As was the case with our treatment of negation, we refined our familiar
-types by tacking on new kinds of information and defined rules for how
-these pieces of information percolate up the parse tree all the way to
-the topmost level still pertinent for ensuring grammaticality.
-
-\section{Agreement}
-\label{sec:agreement}
-
-Finally, we will look at another example of implementing a linguistic
-constraint in ACGs by covering the phenomenon of number
-agreement.
-
-Number agreement motivates a straightforward refinement of our types,
-where all noun and noun phrase types carry number information.
-
-\begin{align*}
-A_{Marie} &: NP\_NUM{=}SG \\
-A_{tatou} &: N\_NUM{=}SG \\
-A_{tatous} &: N\_NUM{=}PL \\
-A_{le} &: N\_NUM{=}SG \limp ((NP\_NUM{=}SG \limp S) \limp S) \\
-A_{les} &: N\_NUM{=}PL \limp ((NP\_NUM{=}PL \limp S) \limp S) \\
-A_{dort} &: NP\_NUM{=}SG \limp S \\
-A_{dorment} &: NP\_NUM{=}PL \limp S \\
-A_{aime_1} &: NP\_NUM{=}SG \limp NP\_NUM{=}SG \limp S \\
-A_{aime_2} &: NP\_NUM{=}SG \limp NP\_NUM{=}PL \limp S \\
-A_{aiment_1} &: NP\_NUM{=}PL \limp NP\_NUM{=}SG \limp S \\
-A_{aiment_2} &: NP\_NUM{=}PL \limp NP\_NUM{=}PL \limp S \\
-A_{qui_1} &: (NP\_NUM{=}SG \limp S) \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
-A_{qui_2} &: (NP\_NUM{=}PL \limp S) \limp N\_NUM{=}PL \limp N\_NUM{=}PL \\
-A_{de_1} &: NP\_NUM{=}SG \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
-A_{de_2} &: NP\_NUM{=}PL \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
-A_{de_3} &: NP\_NUM{=}SG \limp N\_NUM{=}PL \limp N\_NUM{=}PL \\
-A_{de_4} &: NP\_NUM{=}PL \limp N\_Pl \limp N\_NUM{=}PL
-\end{align*}
-
-Given the above signature, we can assign the type $S$ to the syntactic
-term of sentence (\ref{ex:good-agr}) but not to the one of sentence
-(\ref{ex:bad-agr}). In the latter sentence, the type of the variable $t$
-has to be $NP\_NUM{=}SG$ since it is used as an argument to $A_{dort}$. This
-means that the relative clause $dort$ has type $NP\_NUM{=}SG \limp
-S$. However, there is no type for $qui$ which would allow us to modify
-the plural noun $A_{tatous}$ of type $N\_NUM{=}PL$ by a relative clause with a
-missing singular noun phrase, which has the type $NP\_NUM{=}SG \limp S$.
-
-\begin{exe}
-  \ex \label{ex:good-agr} Les tatous de Marie dorment. \\
-      $(A_{les}\ (A_{de}\ A_{Marie}\ A_{tatous}))\ (\lambda^{\circ} x.\ A_{dorment}\ x)$
-  \ex * \label{ex:bad-agr} Marie aime les tatous qui dort. \\
-      * $(A_{les}\ (A_{qui_?}\ (\lambda^{\circ} t.\ A_{dort}\ t)\ A_{tatous}))\ (\lambda^{\circ} y.\ A_{aime_2}\ A_{Marie}\ y)$
-\end{exe}
-
-Once again, we have expressed a linguistic constraint by refining our
-types, augmenting them with a new bit of information, we constrained the
-arguments of functions using these new types and we have described how
-this new type information propagates from the given arguments to the
-produced values.
-
-\section{Putting It All Together}
-\label{sec:together}
-
-In this section, we will see what it takes to enforce all of the
-constraints introduced in the preceding sections (negation, extraction
-and agreement) at the same time, which will give us some idea on how to
-go about building a grammar which handles a wide variety of such
-phenomena.
-
-In a grammar that handles negation, extraction and agreement, our types
-will need to carry all of the refinements we introduced before. To
-recapitulate, $NP$ will be parameterized by whether or not it contains a
-negative determiner ($\_\textcolor{red}{NEG{=}T}$ and
-$\_\textcolor{red}{NEG{=}F}$ respectively), whether or not it is a trace
-from an extraction ($\_\textcolor{green}{VAR{=}T}$ and
-$\_\textcolor{green}{VAR{=}F}$) and its number
-($\_\textcolor{blue}{NUM{=}SG}$ or $\_\textcolor{blue}{NUM{=}PL}$). For
-$N$, we have the same refinements except for $\textcolor{green}{VAR}$
-since extraction only moves $NP$s. Clauses, of type $S$, will be
-discriminated based on the level of extraction within them
-($\_\textcolor{green}{EXT{=}NO}$, $\_\textcolor{green}{EXT{=}ROOT}$ and
-$\_\textcolor{green}{EXT{=}EMB}$).
-
-The individual types will have to respect all of the constraints at the
-same time, each one being a complete specification of a possible
-situation (complete w.r.t. the features listed above). To see and
-appreciate what this means in practice, we give the types for the
-preposition $de$:
-
-\begin{align*}
-C_{de_1} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
-(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \\
-C_{de_2} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
-(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \\
-C_{de_3} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp
-(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
-C_{de_4} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp
-(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
-C_{de_5} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
-(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
-C_{de_6} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
-(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
-C_{de_7} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
-(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \\
-C_{de_8} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
-(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \\
-C_{de_9} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp
-(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
-C_{de_{10}} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp
-(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
-C_{de_{11}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
-(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
-C_{de_{12}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
-(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
-\end{align*}
-
-This highlights two important problems with the current approach. One is
-the untenable complexity. Even though the three mechanisms are
-completely independent of each other, here we are forced to refer to all
-of them in every type. It is no longer easy to glance at the types
-assigned to $de$ and see how it behaves with respect to, for example,
-the mechanism of negation (compare with the type assignments we gave to
-$de$ in previous subsections). Another cause of unreadability of this
-signature is due to the sheer number of types on display, which leads us
-to our second issue.
-
-The signature demonstrates an exponential growth in the number of typed
-constants included. In cases where the behaviors of a wordform across
-the different mechanisms are independent, the complete type signature
-will need to provide a type for every combination of situations in all
-of the mechanisms. Assuming that every mechanism will assign on average
-more than one type to every wordform, the number of the wordform's types
-in the final signature will be exponential w.r.t. the number of
-different mechanisms handled.
-
-We can try making the enumeration of these types easier for the grammar
-author by introducing metavariables for the values of the features. This
-means that we could write the following 3 templates instead of the 12
-types above:
-
-$$
-\forall \textcolor{blue}{x}, \textcolor{blue}{y} \in \{\textcolor{blue}{SG}, \textcolor{blue}{PL}\}
-$$
-\begin{align*}
-C_{de_{1(\textcolor{blue}{x}, \textcolor{blue}{y})}} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \\
-C_{de_{2(\textcolor{blue}{x}, \textcolor{blue}{y})}} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y}) \\
-C_{de_{3(\textcolor{blue}{x}, \textcolor{blue}{y})}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y})
-\end{align*}
-
-This is getting more concise and easier to comprehend. We can take this
-approach further by not only quantifying over variables but also
-computing some of the values using functions:
-
-$$
-\forall \textcolor{blue}{x}, \textcolor{blue}{y} \in\{\textcolor{blue}{SG}, \textcolor{blue}{PL}\}, \forall \textcolor{red}{p}, \textcolor{red}{q} \in \{\textcolor{red}{F}, \textcolor{red}{T}\}, \textcolor{red}{p \land q} \neq \textcolor{red}{T}
-$$
-\begin{align*}
-C_{de(\textcolor{blue}{x}, \textcolor{blue}{y}, \textcolor{red}{p}, \textcolor{red}{q})} &: (NP\_\textcolor{red}{NEG{=}p}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}q}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}(p \lor q)}\_\textcolor{blue}{NUM{=}y})
-\end{align*}
-
-The above is vastly more concise and manages to convey the intent more
-clearly (the intent being that it cannot be the case that there are
-negative determiners in both the complement noun phrase and the modified
-noun and that the value of the result's \textcolor{red}{NEG} feature is
-always the disjunction of the \textcolor{red}{NEG} features of the two
-arguments).
-
-As of right now, we have introduced this kind of rule only as
-meta-notation. This means that the real underlying grammar that will end
-up being used by some algorithm still suffers from the same exponential
-cardinality of the signature. Fortunately, there exists a well-studied
-construction in type theory that lets us assign to terms generic types
-such as the one defined in the template above. This construction is the
-\emph{dependent product} and its utility in ACGs has already been
-noted.\footnote{See \cite{de2007two} for its introduction into the
-  domain of ACGs, \cite{de2007type} for a small example grammar, which
-  handles the interaction between number and gender agreement and
-  coordination, and finally see \cite{pogodalla2012controlling} for
-  using dependent types to handle several constraints regarding
-  extraction (including the one we covered in~\ref{sec:extraction}).}
-
-While extending the type system by including dependent types is
-definitely useful and it helps to solve the problem of having our
-signatures grow exponentially in size, it does not solve our problem
-with the complexity of the new grammar. Granted, the new meta-type
-written in the dependent style is more readable and easier to reason
-about than the enumeration of 12 basic types we presented at the
-beginning of this section, however, the result is still complex in
-that it forces us to describe the behavior of three independent
-mechanisms in one place. This one place happens to be a wordform which
-has seemingly little connection to these mechanisms, the preposition
-$de$. We can therefore imagine that it will participate in many other
-phenomena and that the type we assign to it will grow beyond our
-abilities to reason about it effectively.
-
-Furthermore, the current notation does not even make it obvious that
-these are independent mechanisms. One has to examine the type to make
-sure that there is no interaction. In the rest of this chapter, we will
-devote ourselves to investigating ways of disentangling these.
-
-\section{Intersections of Types}
-
-Our goal is to split the types into smaller parts that only talk about
-features belonging to one of the three mechanisms covered above. This
-means that we would like to take some phrase about which we know a lot
-of disparate pieces of information (presence of negative determiner,
-trace status, number\ldots), having type
-e.g. $NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}$,
-and use it as an argument for some function which only cares about one
-channel of information and would have a type like
-$NP\_\textcolor{red}{NEG{=}T} \limp N\_\textcolor{red}{NEG{=}F} \limp
-N\_\textcolor{red}{NEG{=}T}$. To make a type such as
-$NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}$
-be a legal argument to a function which expects an argument of type
-$NP\_\textcolor{red}{NEG{=}T}$, we will need to introduce some notion of
-\emph{subtyping}.
-
-The types we have been showing up to now propose a very straightforward
-notion of subtyping. All of the types can be seen as conjunctions of
-properties where $\_$ (underscore) is the conjunction operator. A type
-$\tau$ can then be seen as a subtype of type $\sigma$ whenever the
-conjuncts of $\tau$ form a superset of the conjuncts of $\sigma$. This
-is not unlike the structural typing of records in OCaml (CITE SOME OCAML
-SPEC HERE), where a record type is a subtype of another record type if
-the former provides, among others, all the fields provided by the latter
-and gives them the same type. In our case, one type is a subtype of
-another if the former gives, among others, all the properties given by
-the latter. This notion of subtyping can be expressed by adding the rule
-of inference in Figure~\ref{fig:type-superset} to our type
-system.\footnote{In the rule \ldots (talk about $\land$ instead of $\_$
-  (underscore) here...)}
-
-\begin{figure}
-  \begin{prooftree}
-    \AxiomC{$\Gamma_l; \Gamma_i \vdash_\Sigma M : A_1 \land \ldots A_{i-1} \land A_i \land A_{i+1} \ldots \land A_n$}
-    \RightLabel{(subtype)}
-    \UnaryInfC{$\Gamma_l; \Gamma_i \vdash_\Sigma M : A_1 \land \ldots A_{i-1} \land A_{i+1} \ldots \land A_n$}
-  \end{prooftree}
-  \caption{\label{fig:type-superset} A rule of inference giving us
-    subtyping for conjunctions of atomic types. For readability, we use
-    $\land$ instead of $\_$ (underscore) to designate the conjunction of
-    properties.}
-\end{figure}
-
-(Talk about whether or not this is a conservative extension...)
-
-With the new subtyping rule, we can
-
-$$
-S_{tatou} : N \land NEG{=}F \land NUM{=}SG
-$$
-$$
-S_{le_{NEG_1}} : (N \land NEG{=}F) \limp (((NP \land NEG{=}F) \limp S) \limp S)
-$$
-$$
-S_{le_{NEG_2}} : (N \land NEG{=}T) \limp (((NP \land NEG{=}T) \limp S) \limp S)
-$$
-$$
-S_{le_{NUM}} : (N \land NUM{=}SG) \limp (((NP \land NUM{=}SG) \limp S) \limp S)
-$$
-$$
-S_{le_{VAR}} : N \limp (((NP \land VAR{=}F) \limp S) \limp S)
-$$
-
-\ldots
+\chapter{Graphical Abstract Categorial Grammars}
+\label{chap:gacg}
+
+In this chapter, we will introduce a generalization of abstract
+categorial grammars that gracefully solves the modularity issues that we
+saw in the previous chapter.
 
@@
 
-We will turn to another solution to the problem of complex grammars that
-try to implement multiple independent constraints. This solution will be
-based on a generalization of ACGs, but in order to introduce this
-generalization, we will first need to present the manner in which ACGs
-and their compositions are formulated in ACG literature.
+Before we proceed to define our extension of ACGs, we will first need to
+present the manner in which ACGs and their compositions are formulated
+in ACG literature.
 
@@
 of ACGs which will help us overcome the modularity issues that have been
-the theme of this chapter.\footnote{Incidentally, this was not the
-  reason for the way we presented ACGs in~\ref{sec:acg}. We chose this
-  style as it grouped together the information in a more natural way. We
-  approach signatures as descriptions of domains (strings, syntax,
-  semantics), each one being associated with a distinguished type
-  telling us which objects the domain has set out to describe (strings,
-  sentences, truth values). Having distinguished types directly attached
-  to signatures allows us to omit them from the ACG quadruples,
-  $\mathopen{<}\Sigma_A, \Sigma_O, \mathcal{L},
+the theme of Chapter~\ref{chap:constraints}.\footnote{Incidentally, this
+  was not the reason for the way we presented ACGs in~\ref{sec:acg}. We
+  chose this style as it grouped together the information in a more
+  natural way. We approach signatures as descriptions of domains
+  (strings, syntax, semantics), each one being associated with a
+  distinguished type telling us which objects the domain has set out to
+  describe (strings, sentences, truth values). Having distinguished
+  types directly attached to signatures allows us to omit them from the
+  ACG quadruples, $\mathopen{<}\Sigma_A, \Sigma_O, \mathcal{L},
   \cancel{S}\mathclose{>}$. However, this only leaves us with triples
@@
 signature ($SyntSem$) from whose terms the semantic representation can
-be derived by a function ($\mathcal{L}_{InterSem}$). This approach is
+be derived by a function ($\mathcal{L}_{interSem}$). This approach is
 studied more deeply in \cite{pogodalla2007generalizing}.
@@
 signature we introduce to handle the second constraint must also satisfy
-the first constraint, since mapping them by $\mathcal{L}_{Constr_2}$
+the first constraint, since mapping them by $\mathcal{L}_{constr_2}$
 must yield a well-typed term in the signature corresponding to the first
@@
 it directly to the syntactic signature ($Synt$) by taking the
-composition of the lexicons in between ($\mathcal{L}_{Constr_1} \circ
-\mathcal{L}_{Constr_2}$).
+composition of the lexicons in between ($\mathcal{L}_{constr_1} \circ
+\mathcal{L}_{constr_2}$).
 
@@
 
-\section{Graphical Abstract Categorial Grammars}
+\section{Definitions}
 
 In this section, we distill the reasoning about generalizing ACG
-diagrams to graphs in a definition of an extension for ACGs. We then
-examine its key properties and implications.
+diagrams to graphs in a definition of an extension for ACGs.
 
@@
 
+\section{General Remarks on Graphical ACGs}
+
+In this section, we then examine the key properties and implications of
+graphical ACGs.
+
 \subsection{On the Expressivity of Graphical ACGs}
@@
 \begin{observation}
-  The set of intrinsic languages definable by graphical ACGs is the same
-  as the set of abstract languages definable by ACGs.
+  The set of intrinsic languages definable by graphical ACGs is by
+  definition the same as the set of abstract languages definable by
+  ACGs.
 \end{observation}
@@
 abstract signatures only contain types of order 2 or less, parsing is
-possible in polynomial type for both linear \cite{salvati2005problemes}
+possible in polynomial time for both linear \cite{salvati2005problemes}
 and non-linear \cite{kanazawa2007parsing} type systems (such as the one
@@
 exactly those that have been presented in the first three sections of
-this chapter. The lexicons for the edges connecting these constraint
-nodes to $Synt$ all behave according to the scheme below:
+Chapter~\ref{chap:constraints}. The lexicons for the edges connecting
+these constraint nodes to $Synt$ all behave according to the scheme
+below:
 
@@
     \caption{{\label{fig:puzzle-solution} The graph of a grammar
-        enforcing all of the constraints discussed in this chapter.}}
+        enforcing all of the constraints discussed in
+        Chapter~\ref{chap:constraints}.}}
   \end{subfigure}
@@
 
-\subsection{On Alternative Interpretations of Graphical ACGs}
+\section{On Alternative Interpretations of Graphical ACGs}
 
@@
 
-    That is an interesting question indeed, whose resolution is left as
-    an exercise for the reader.
-
-  \item[Pangraphical languages are not ``less decidable'' than object
-    languages.] \hfill \\
-
-    Presuming that type checking is decidable for all the signatures of
-    the graphical ACGs and that we can parse effectively for all the ACG
-    edges in the graph, we can reuse the parsing algorithm we introduced
-    in \ref{ssec:graphical-decidability}. We apply this algorithm to
-    enumerate all the possible extrinsic parses in a language
-    $\mathcal{E}_{\mathcal{G}}(v)$ and keep only those where for each
-    ancestor $u$ we have the same abstract term for all paths from $u$
-    to $v$.
+    That is an interesting question indeed, one whose resolution is left
+    as a possible avenue for future work.
 \end{description}
-
-\subsection{On the Practical Consequences of Graphical ACGs}
-
-In this final subsection, we will enumerate some of the practical
-benefits and challenges of using graphical ACGs for implementing
-grammar-based systems.
-
-\begin{itemize}
-\item The lexicons that go from our constraint signatures ($Neg$, $Ext$
-  and $Agr$ in subsection \ref{ssec:graphical-engineering}) are all
-  trivial in the sense they map constants to constants. Finding an
-  antecedent to a term is then just a question of finding an antecedent
-  for every constant in the term (i.e. ``tagging'' the constants of the
-  term with antecedent constants). This means that there is an effective
-  parsing algorithm for verifying the constraints. Furthermore, this
-  algorithm could also be made efficient by introducing filtering
-  techniques \cite{guillaume2008toolchain} and statistical supertagging
-  \cite{moot2012categorial}.
-
-\item Constraints can introduce type extensions in their abstract
-  signatures and the added complexity of these extensions is then
-  contained within the constraint. Furthermore, the constraint signature
-  can be treated as a ``black box'' that verifies a constraint and in
-  the context of an implementation, it can be replaced with a
-  special-purpose hand-written procedure that does the same more
-  efficiently.
-
-\item The distributed nature of graphical ACGs has its own practical
-  applications. Parsing can be made more robust by being able to parse
-  sentences which violate some of the constraints. Parsing an
-  ungrammatical sentence and identifying the constraints it violates can
-  also serve as a basis for a syntax checker.
-
-\item If we decide to define all of our syntactic constraints in
-  separate signatures, we will end up with a lot of signatures which
-  look exactly like the syntactic signature but with some small
-  refinements (see the signatures in sections \ref{sec:negation},
-  \ref{sec:extraction} and \ref{sec:agreement}). Thus we will desire
-  some concise way of formulating these refinements so that we will not
-  need to replicate the entire syntactic signature for every
-  constraint.
-
-  We hope that the class of useful refinements is small and that these
-  constraint signatures could be systematically constructed from the
-  syntactic signature by applying a series of refining
-  transformations. This is an interesting engineering challenge awaiting
-  anyone willing to design a truly wide-coverage graphical ACG using
-  constraints.
-\end{itemize}

### 2013-06-17 ig.tex
Last night shift

@@
 
+\subsection{The Mechanisms of Interaction Grammars}
+
 We will illustrate the objects and mechanisms of interaction grammars on
-a simple example sentence.
+a simple example sentence. Since the largest and most-developed
+interaction grammar is the Frigram grammar of French, we will be using
+French example sentences and their analysis according to
+Frigram. Frigram will be discussed in more detail in
+subsection~\ref{ssec:frigram}.
 
@@
   \centering
-  \includegraphics[scale=0.25]{ig-parse.pdf}
+  \includegraphics[scale=0.25]{images/ig-parse.pdf}
   \caption{\label{fig:ig-parse} The syntactic tree of sentence
@@
 node being the root. Solid lines indicate immediate dominance with the
-higher node being the parent and the lower the child. The ordering of
-children is given by the arrows, which signify the immediate precedence
-relation between sister nodes.
+higher node being the parent and the lower node being the child. The
+ordering of children is given by the arrows, which signify the immediate
+precedence relation between sister nodes.
 
@@
 some set of algebraic operations (substitution and adjunction in case of
-TAG). Instead, \emph{polarized tree descriptions} (PTDs), the elementary
-structures of interaction grammars, impose constraints on the final
-structure and a structure is said to be generated by some PTDs if it is
-a minimal structure satisfying those constraints (it is a \emph{minimal
-  model}). This distinction separates TAG as a formalism in the
-generative-enumerative framework of syntax with IG as a formalism in the
-model-theoretic framework.
+TAGs). Instead, \emph{polarized tree descriptions} (PTDs), the
+elementary structures of interaction grammars, impose constraints on the
+final structure and a structure is said to be generated by some PTDs if
+it is a minimal structure satisfying those constraints (we say it is a
+\emph{minimal model}). This distinction separates TAGs as a formalism in
+the generative-enumerative framework of syntax from IGs as a formalism
+in the model-theoretic framework.
 
@@
   \centering
-  \includegraphics[scale=0.25]{ig-eptds.pdf}
+  \includegraphics[scale=0.25]{images/ig-eptds.pdf}
   \caption{\label{fig:ig-eptds} The elementary polarized tree
@@
 latter in the ordered tree (\emph{precedence}). Note that not all sister
-nodes in a PTD have to be linked with this precedence relation. See for
-example the nodes (1,1) and (1,5) in the EPTD of the clitic pronoun
-$la$.
+nodes in a PTD have to be linked with this precedence relation, the tree
+can be under-specified. See for example the nodes (1,1) and (1,5) in the
+EPTD of the clitic pronoun \emph{la}.
 
@@
 resource sensitivity of languages, while the virtual polarities
-($\textcolor{purple}{\sim}$) are used to for pattern matching against
-the context.
+($\textcolor{purple}{\sim}$) are used for pattern matching against the
+context.
 
@@
 (i.e. no positive, negative or virtual polarities). Whenever more than
-one node of the PTD is mapped onto a single node of the syntactic tree,
-the polarities of each feature are combined. The combination mechanism
-allows us to combine a positively polarized feature with a negatively
-polarized one to yield a saturated one. Virtual polarities can only be
-eliminated by combining them with a saturated polarity (they are
-analogous to the $=_{c}$ constraints of LFG \cite{kaplan1982lexical}).
+one node of the PTD is interpreted by the same node of the syntactic
+tree, the polarities of each feature are combined. The combination
+mechanism allows us to combine a positively polarized feature with a
+negatively polarized one to yield a saturated one. Virtual polarities
+can only be eliminated by combining them with a saturated polarity (they
+are analogous to the $=_{c}$ constraints of LFG
+\cite{kaplan1982lexical}).
 
 If we look at the PTDs of Figure~\ref{fig:ig-eptds}, we can see this
-polarity mechanism in action. The EPTD of $Jean$ has a positive $cat$
-feature in its root node saying that it provides one $np$, and a
+polarity mechanism in action. The EPTD of \emph{Jean} has a positive
+$cat$ feature in its root node saying that it provides one $np$, and a
 negative $funct$ feature saying that it expects some function. In the
-EPTD of $voit$, we have two nodes for the subject and object, both of
-them expecting an $np$ and providing them the $subj$ and $obj$
+EPTD of \emph{voit}, we have two nodes for the subject and the object,
+both of them expecting an $np$ and providing them the $subj$ and $obj$
 functions, respectively. The root of the EPTD then provides a complete
@@
 
-The clitic pronoun $la$ participates in the positive/negative resource
-management system as well, since using the clitic fills up the object
-slot in the valency of a verb. The EPTD of $la$ also uses virtual
-features heavily to select the right place where to hang the pronoun.
+The clitic pronoun \emph{la} participates in the positive/negative
+resource management system as well, since using the clitic fills up the
+object slot in the valency of a verb. The EPTD of \emph{la} also uses
+virtual features heavily to select the right place where to hang the
+pronoun in the resulting tree.
 
@@
 tree of Figure~\ref{fig:ig-parse}, we can look at the result of
-superposing the EPTDs of $la$ and $voit$ by merging the nodes (1,2) and
-(1,5) with (2,2) and (2,4) respectively on Figure~\ref{fig:ig-partial}.
+superposing the EPTDs of \emph{la} and \emph{voit} by merging the nodes
+(1,2) and (1,5) with (2,2) and (2,4) respectively on
+Figure~\ref{fig:ig-partial}.
 
@@
   \centering
-  \includegraphics[scale=0.25]{ig-partial.pdf}
+  \includegraphics[scale=0.25]{images/ig-partial.pdf}
   \caption{\label{fig:ig-partial} The result of merging the EPTDs of
-    $la$ and $voit$ by merging the nodes (1,5) and (1,5) with (2,2) and
-    (2,4), respectively.}
+    \emph{la} and \emph{voit} by merging the nodes (1,5) and (1,5) with
+    (2,2) and (2,4), respectively.}
 \end{figure}
 
-\subsection{Frigram, an interaction grammar}
+\subsection{Frigram, an Interaction Grammar}
 \label{ssec:frigram}
@@
 building our grammar on as well. Furthermore, as we will see
-in~\ref{ssec:link-ig-acg}, interaction grammars are closely linked with
+in~\ref{ssec:link-ig-acg}, interaction grammars are closely linked to
 abstract categorial grammars, which makes Frigram a suitable grammar to
@@
 
-In this chapter, we will introduce the metagrammatical structure of
+In this subsection, we will introduce the metagrammatical structure of
 Frigram and talk about how it uses the formalism of interaction grammars
 to solve several tricky linguistic phenomena.
+
+\subsubsection{Frigram as a Metagrammar}
 
@@
 of these EPTDs is connected to a specific wordform whose particular use
-it describes. Given the amount of wordforms that a useable grammar of
+it describes. Given the amount of wordforms that a usable grammar of
 French would need to cover, the number of EPTDs in our grammar could
@@
 differences between similar wordforms. For this purpose, Frilex was
-created. Frilex is a morpho-syntactic lexicon of French compiled from
-various other pre-existing lexicons of French. It is in effect a large
+created. Frilex is a morphosyntactic lexicon of French compiled from
+various other preexisting lexicons of French. It is in effect a large
 relation linking the wordforms of French to \emph{hypertags}, feature
@@
 With Frilex in place, Frigram can be defined as a set of unanchored
-EPTDs which are paired with feature structures which delimit the subset
+EPTDs which are paired with feature structures that delimit the subset
 of the Frilex items to which the EPTDs apply. This kind of simple
 ``metagrammar'' already saves us a lot of effort, the number of
-unanchored EPTDs defined by Frigram is somewhere around 4000 (TODO: Cite
-the Frig documentation somehow).
+unanchored EPTDs defined by Frigram is somewhere around
+4000\footnote{\url{http://wikilligramme.loria.fr/doku.php?id=frigram:frigram}}.
 
 However, defining some 4000 EPTDs manually still seems like a very
-tedious, error prone and hard to maintain process. Many of the EPTDs
-have to repeatedly describe common phenomena such as subject-predicate
-agreement or predicate-argument saturation. Similarly as in other
-software endeavors, it would be preferable to define these common
-patterns in some reusable module and compose EPTDs from these building
-blocks. This is where XMG steps in.
+tedious and error prone process that is likely to lead to a grammar that
+is hard to maintain. Many of the EPTDs have to repeatedly describe
+common phenomena such as subject-predicate agreement or
+predicate-argument saturation. Similarly as in other software endeavors,
+it would be preferable to define these common patterns in some reusable
+module and compose EPTDs from these building blocks. This is where XMG
+steps in \cite{duchier2012metagrammars}.
 
@@
 between what the grammar author writes and what ends up in the
-bottom-level interaction grammar. XMG provides a language that lets the
-grammar author define the lexical items of his grammar and to combine
-these definitions to yield new and more elaborate lexical items.
+bottom-level (interaction) grammar. XMG provides a language that lets
+the grammar author define the lexical items of his grammar and to
+combine these definitions to yield new and more elaborate lexical items.
 
@@
 chief tools that the grammar authors can use (and in the case of
-Frigram, have used) to succintly define their grammar.
+Frigram, have used) to succinctly define their grammar.
 
@@
 
+\subsubsection{Linguistic Description in Frigram}
+
 \begin{figure}
   \centering
-  \includegraphics[scale=0.25]{ig-neg.pdf}
-  \caption{\label{fig:ig-neg} EPTDs of $aucun$ and $ne$ demonstrating
-    the way polarities are used to model French negation.}
+  \includegraphics[scale=0.25]{images/ig-neg.pdf}
+  \caption{\label{fig:ig-neg} EPTDs of \emph{aucun} and \emph{ne}
+    demonstrating the way polarities are used to model French negation.}
 \end{figure}
@@
 particular, handle a tricky linguistic phenomenon in an interesting way.
-In French, negation is signalled by the particle $ne$ which must be
-accompanied by one of several designated determiners, pronouns or
-adverbs. The particle $ne$ assumes a position right before the inflected
-verb, but its partner, e.g. the determiner $aucun$ can appear in a rich
-variety of positions:
+In French, negation is signalled by the particle \emph{ne}, which must
+be accompanied by one of several designated determiners, pronouns or
+adverbs. The particle \emph{ne} assumes a position right before the
+inflected verb, but its partner, e.g. the determiner \emph{aucun} can
+appear in a rich variety of positions:
 
@@
 Sentences (\ref{ex:aucun-suj}) and (\ref{ex:aucun-obj}) demonstrate that
-the $aucun$ determiner can be used both in the subject and object
+the \emph{aucun} determiner can be used both in the subject and object
 positions, as a determiner of either one of the arguments directly or of
@@
 admissible to place the determiner in an embedded clause as it is in
-(\ref{ex:aucun-bad}). If the determiner $aucun$ is used in an embedded
-clause, it is the embedded clause itself that must be negated, as in
-sentence (\ref{ex:aucun-good}).
+(\ref{ex:aucun-bad}). If the determiner \emph{aucun} is used in an
+embedded clause, it is the embedded clause itself that must be negated,
+as in sentence (\ref{ex:aucun-good}).
 
@@
 the two wordforms which enable negation must always co-occur. The
-particle $ne$ provides a positive $neg$ feature to the clause in which
-it occurs and this positive polarity must be saturated by a negative
-polarity contributed by one of the possible partner words, such as
-$aucun$. This shows that the usefulness of positive and negative
+particle \emph{ne} provides a positive $neg$ feature to the clause in
+which it occurs and this positive polarity must be saturated by a
+negative polarity contributed by one of the possible partner words, such
+as \emph{aucun}. This shows that the usefulness of positive and negative
 polarities extends beyond the $cat$/$funct$ pair that we have already
@@
 
-Second, the determiner $aucun$ can occur deep, e.g. inside a complement
+Second, the determiner \emph{aucun} can occur deep, e.g. inside a complement
 of one of the arguments, but it cannot occur everywhere, as is
@@
 associate a feature structure with a dominance constraint which will
-force all the nodes that will span the distance between the two nodes
-linked by the dominance constraint to unify with that feature
-structure.
+force all of the nodes that span the distance between the two nodes
+linked by the dominance constraint to unify with that feature structure.
 
@@
 
-\subsection{The link with abstract categorial grammars}
+\subsection{The Link with Abstract Categorial Grammars}
 \label{ssec:link-ig-acg}
@@
 and of polarized tree descriptions in interaction grammars.
+
+\subsubsection{Perrier's Theorem}
 
@@
 will not cover the theorem deeply in our treatise, we will settle for
-knowing that the composition of the positive polarization operation $+$
-and the syntactic description producing operation $D$ maps an IILL
-formula into a syntactic description such that above-stated theorem
-holds. Furthermore, this mapping can be easily constructed since Perrier
-gives straightforward ways of computing the results of both operations.
+knowing that the composition $D \circ +$ of the positive polarization
+operation $+$ and the syntactic description producing operation $D$ maps
+an IILL formula into a syntactic description such that above-stated
+theorem holds. Furthermore, this mapping can be easily constructed since
+Perrier gives straightforward ways of computing the results of both
+operations.
 
-This theorem establishes a striking symmetry between the two
+Perrier's theorem establishes a striking symmetry between the two
 formalisms. Let us consider the case of parsing a sentence using a
@@
 In ACGs, we will try to take the typed constants and to use them to
-produce a term having some distinguished type $s$. Thanks to the
+produce a term having some distinguished type $S$. Thanks to the
 Curry-Howard correspondence, constructing a term having a given type is
@@
 our type system. What this means is that our parsing problem boils down
-to proving the IILL sequent $\tau_1, ..., \tau_n \vdash s$ (we omit the
+to proving the IILL sequent $\tau_1, ..., \tau_n \vdash S$ (we omit the
 inclusion of intuitionistic non-linear implications since they are not
-relevant for parsing sentences). The sentence thus becomes parseable by
+relevant for parsing sentences). The sentence thus becomes parsable by
 our grammar under the given lexical selection if and only if this
@@
 polarized tree description. Parsing is then just a matter of finding a
-model for this description. The sentence is therefore parseable by our
+model for this description. The sentence is therefore parsable by our
 grammar under the given lexical selection if and only if this
@@
 
-However, this approach is not sufficient. The $D \circ +$ transformation
-is far from a bijection.
+\subsubsection{Translating Tree Descriptions to Types}
+
+However, just inverting $D \circ +$ mapping to obtain types from tree
+descriptions is not sufficient. $D \circ +$, the transformation in
+Perrier's theorem, is far from a bijection.
 
@@
 we could simply choose one of the possible formulas which fit the
-syntactic description as our canonical representation.
+syntactic description as our canonical representation since the
+differences between the types would be inconsequential.
 
@@
   tree descriptions}, which is a proper subset of the set of all
-syntactic descriptions. Specifically, in an IILL tree descriptions, all
+syntactic descriptions. Specifically, in an IILL tree description, all
 immediate dominance constraints go from negative nodes to positive
@@
 still gives us a general plan for recovering IILL types. The strategy
-suggested by the description/formula connection given by the theorem
-yields a mapping $F$ from the under-specified trees to IILL formulas.
+suggested by the description/formula connection given by the definition
+of $D \circ +$ yields a mapping $F$ from the under-specified trees to
+IILL formulas.
 
@@
 
-This strategy corresponds to our intuitions about how should tree
-descriptions be expressed using IILL types. We will consider two very
-simple cases to demonstrate it.
+This strategy corresponds to our intuitions about how tree descriptions
+should be expressed using IILL types. We will consider two very simple
+cases to demonstrate it.
 
@@
 node of category $A$. This tree description can also be easily
-translated to TAG by turning the negative leaf nodes into substitution
-nodes. The conventional translation of TAG into categorial grammar then
-confirms the type proposed by our strategy.
+translated to TAGs by turning the negative leaf nodes into substitution
+nodes. The conventional translation of TAGs into categorial grammars
+then confirms the type proposed by our strategy.
 
@@
 
-The PTDs of IG include precedence constraints. We can see them as
+\subsubsection{IG Features Outside the Scope of Perrier's Theorem}
+
+The PTDs of IGs include precedence constraints. We can see them as
 serving two purposes, to both provide an ordering of the constituents
@@
 feature, $funct$), and ignore the others. The results and intuitions we
-have about intepreting tree descriptions as types do not offer an easy
+have about interpreting tree descriptions as types do not offer an easy
 way of modeling polarities across different dimensions, features, at the
-same time. This would mean that phenomena, such as the paired
-grammatical words for negation in French, which were handled using
-special polarized features will need new solutions in the framework of
-ACGs.
+same time. However, dropping the other polarized features would cost us
+the treatments of some phenomena, such as the paired grammatical words
+for negation in French, which were handled using the $neg$ feature.
 
@@
 would be to have a function of type $X \limp X$ where $X$ is the type of
-the consituent being modified. In TAG, we would have an adjunction tree
-with a root and a foot node of category $X$ and which pins the modifier
-on the $X$ phrase. This approach would work in interaction grammars as
-well but the formalism offers other solutions too. The designers of
-Frigram have opted for an alternate treatment in which modifying a
-constituent does not increase the depth of the parse tree. Modifiers are
-adjoined as sister nodes of the constituents they modify and it is
-through virtual features that IGs can ensure that a modifier is inserted
-only in the proper context.
+the constituent being modified. In TAGs, we would have an adjunction
+tree with a root and a foot node of category $X$ and which pins the
+modifier on the $X$ phrase. This approach would work in interaction
+grammars as well but the formalism offers other solutions too. The
+designers of Frigram have opted for an alternate treatment in which
+modifying a constituent does not increase the depth of the parse
+tree. Modifiers are adjoined as sister nodes of the constituents they
+modify and it is through virtual features that IGs can ensure that a
+modifier is inserted only in the proper context.
 
-See Figure~\ref{fig:ig-adj} for an example of an adjective modifier in
+See Figure~\ref{fig:ig-adj} for an example of a modifier adjective in
 Frigram. Nodes (1,3) and (1,2) select the noun to be modified and its
@@
   \centering
-  \includegraphics[scale=0.25]{ig-adj.pdf}
-  \caption{\label{fig:ig-adj} An EPTD for the adjective $grand$ in
+  \includegraphics[scale=0.25]{images/ig-adj.pdf}
+  \caption{\label{fig:ig-adj} An EPTD for the adjective \emph{grand} in
     Frigram.}
@@
 Understanding this pattern can give us a useful tool for reading
-EPTDs. Consider the EPTD of $la$ we saw in
+EPTDs. Consider the EPTD of \emph{la} we saw in
 Figure~\ref{fig:ig-eptds}. The EPTD can now be seen as applying this
 pattern twice at the same time to both cliticize the pronoun to the
-verbal kernel and to fill the object valency slot of the
-verb.
+verbal kernel and to fill the object valency slot of the verb.
 
@@
 (large) dominance relations. We have seen this before in the EPTD of the
-determiner $aucun$ (Figure~\ref{fig:ig-neg}), which had to reach up to
-the nearest containing clause and hang a negative $neg$ feature there.
-It is also present in the EPTD of the relative pronoun $que$
+determiner \emph{aucun} (Figure~\ref{fig:ig-neg}), which had to reach up
+to the nearest containing clause and hang a negative $neg$ feature
+there.  It is also present in the EPTD of the relative pronoun $que$
 (Figure~\ref{fig:ig-que}) which will bring us back to the translation
@@
   \centering
-  \includegraphics[scale=0.25]{ig-que.pdf}
+  \includegraphics[scale=0.25]{images/ig-que.pdf}
   \caption{\label{fig:ig-que} An EPTD of the relative pronoun $que$.}
@@
 represent a modification of some constituent, say a noun, then we get a
-type like $n \limp n$. However, the EPTD contains also positive and
+type like $n \limp n$. However, the EPTD also contains positive and
 negative $cat$ polarities. The node (2,0) has a negative $cat$ polarity
@@
 must be $decl$, but would also capture the information in the polarized
-$funct$ feature so that the relative pronoun $que$ is not permitted to
-extract subjects. There are still constraints that we would miss out on
-though, such as the fact that the subject (2,3) of the clause (2,0) must
-precede any embedded clauses (2,6)...(2,5), the fact that any such
+$funct$ feature so that the relative pronoun \emph{que} is not permitted
+to extract subjects. There are still constraints that we would miss out
+on though, such as the fact that the subject (2,3) of the clause (2,0)
+must precede any embedded clauses (2,6)...(2,5), the fact that any such
 embedded clause must serve either the $modal$ or the $obj$ function and
@@
 equivalent in ACGs. One example would be a feature of IGs that we have
-not even mentioned which allows us the grammar author to provide an
-\emph{exhaustive} list of children of a node in a PTD. Another kind of
-tricky constraint are the orange rectangles that we can see in
-Figure~\ref{fig:ig-que} which enforce that a constituent must be the
-leftmost or rightmost daughter of its parent. The problem with these is
-similar as with precedence relations. To model their impact on word
+not even mentioned which allows the grammar author to provide an
+\emph{exhaustive} list of children of a node in a PTD.
+
+Another kind of tricky constraint are the orange rectangles that we can
+see in Figure~\ref{fig:ig-que} which enforce that a constituent must be
+the leftmost or rightmost daughter of its parent. The problem with these
+is similar as with precedence relations. To model their impact on word
 order would be easy using an ACG lexicon, but to model their effect on
@@
 framework of linear logic and their fundamental principles of syntactic
-composition have been linked together by Perrier's theorem, we conclude
-that embedding Frigram into an ACG or constructing some automated
-technique for translating the Frigram EPTDs into ACG types and terms
-would not be feasible. We suspect that the incompatibility of these two
-formalisms is due to ACG being a formalism based in the
-generative-enumerative framework of syntax while IG is based in the
-model-theoretic approach. Perhaps we could model the operational view of
-IG, the combination of under-specified trees via superposition, by
-employing some of the other connectives of linear logic to model the
-fact that superposing PTDs usually involves many resources. However,
-this would lead us well out of the framework of ACGs in which we wish to
-conduct our research because of its elegant approach to discourse
-denotations.
+composition have been linked together by Perrier's theorem, we
+conjecture that straightforwardly embedding Frigram into an ACG or
+constructing some automated technique for directly translating the
+Frigram EPTDs into ACG types and terms would not be feasible. We suspect
+that the incompatibility of these two formalisms is due to ACGs being a
+formalism based in the generative-enumerative framework of syntax while
+IGs are based in the model-theoretic approach.
 

### 2013-06-17 implementation.tex
Last night shift

--- 2013-06-13 my-work-rewritten.tex
@@
-\chapter{Wide-coverage abstract categorial grammars}
+\chapter{Implementing Wide-Coverage Abstract Categorial Grammars}
+\label{chap:implementation}
 
@@
   refer to what is usually called a lexicon, in order not to cause
-  confusion with the lexicons of ACGs.} to produce the lexicalized ACG
-in a direct representation, ready to be loaded by the toolkit. This
-workflow would make development on the grammar tedious as every time we
-would change the metagrammar, we would have to regenerate the direct
-representation and then load it in the ACG toolkit. With extra effort,
-the former cost could be alleviated by devising a system for
-regenerating only parts of the grammars which will have changed since
-last time. However, we would still have to pay the price of loading the
-entire grammar into the toolkit before being able to interact with it
-again.
+  confusion with the lexicons of ACGs.} to produce the
+lexicalized\footnote{We will be using the term \emph{lexicalized} in the
+  context of ACGs to mean that the elements of the grammar (the typed
+  constants and their images given by lexicons) are generated by a
+  lexical database. This is not to be confused with the term used in
+  \cite{yoshinaka2005complexity}.} ACG in a direct representation, ready
+to be loaded by the toolkit. This workflow would make development on the
+grammar tedious as every time we would change the metagrammar, we would
+have to regenerate the direct representation and then load it in the ACG
+toolkit. With extra effort, the former cost could be alleviated by
+devising a system for regenerating only parts of the grammars which will
+have changed since last time. However, we would still have to pay the
+price of loading the entire grammar into the toolkit before being able
+to interact with it again.
 
@@
 
-In this chapter, we will spend time explaining the decisions which went
-into designing the system and the form it has now.
+In this chapter, we will spend some time explaining the design rationale
+underlying the system and the form it has now. The source code of the
+system in its current state can be perused at
+\url{https://github.com/jirkamarsik/acg-clj}.
 
-\section{The tools and techniques}
+\section{The Tools and Techniques}
 
@@
 By insisting on purity, I/O becomes more difficult as we would not allow
-the side-effectful pseudorelations of Prolog. Therefore, we will not
+the side-effectful pseudo-relations of Prolog. Therefore, we will not
 force ourselves to implement the entire program front-to-back using only
@@
 
-\subsection{The case for relational programming}
+\subsection{The Case for Relational Programming}
 
@@
 programming system for our implementation. First off, the problems which
-consiture the algorithmic core of our program are often mere textbook
+constitute the algorithmic core of our program are often mere textbook
 examples of how to use a relational programming system. This goes for
@@
 Finally, some of the structures that we will encounter in our domain are
-more directly modelled in mathematics as a relation instead of a
-function. This has the benefit that to implement a relation in
-relational programming, one just writes the relation, while if we were
-forced to encode everything in functions, we would have to manage this
-multiplicity of values manually. Examples of structures that have this
-kind of structures if Frilex, our lexical database, where a single
-wordform can be mapped to multiple hypertags. We will see another
-example of a relation within our domain in the next section and we
-will also discuss the advantages and drawbacks of modelling functions as
-relations.
+more directly modelled in mathematics as relations rather than
+functions. In a relational programming system, to implement a relation,
+one just writes a relation as it is, whereas if we were forced to encode
+everything in functions, we would have to simulate relational behavior
+manually. Frilex, our lexical database, is an example of a structure
+that has this relational shape since a single wordform can be mapped to
+multiple hypertags. We will see another example of a relation within our
+domain in \ref{ssec:def-sig} and we will also discuss the advantages and
+drawbacks of modelling functions using relations.
 
-\subsection{Choice of implementation language}
+\subsection{Choice of Implementation Language}
 
-There are two languages which play host to a substantial implementation
-of miniKanren (meaning one that has extensible unification and/or
-constraints and that includes support for nominal logic),
-Racket\footnote{\url{https://github.com/calvis/cKanren}} and
+There are two programming languages which play host to a substantial
+implementation of miniKanren (meaning one that has extensible
+unification and/or constraints and that includes support for nominal
+logic), Racket\footnote{\url{https://github.com/calvis/cKanren}} and
 Clojure\footnote{\url{https://github.com/clojure/core.logic}}. The two
@@
 no great significance. An appealing feature of Clojure is that it uses
-abstract collection types pervasively instead of tying the standard
-operations to specific structures such as lists and cons cells and that
-it follows other similarly enlightened design decisions. Racket on the
-other hand has had more time to mature, has better metaprogramming
+abstract collection types pervasively instead of coupling the standard
+operations to specific data structures such as lists and cons cells and
+that it follows other similarly enlightened design decisions. Racket on
+the other hand has had more time to mature, has better metaprogramming
 facilities and is more recognized in academia. In developing our system,
@@
 inspired the software community and the project has attracted
-contributors that submit issue reports and
-patches\footnote{\url{http://dev.clojure.org/jira/browse/LOGIC}}.
+contributors that submit numerous issue reports and
+patches.\footnote{\url{http://dev.clojure.org/jira/browse/LOGIC}}
 
-\section{Defining signatures and lexicons}
+\section{Defining Signatures and Lexicons}
 
-\subsection{Defining signatures}
+In this section, we will go over the representation of signatures and
+lexicons in our implementation and the basic facilities for defining
+them.
+
+\subsection{Defining Signatures}
+\label{ssec:def-sig}
 
@@
 provide the function $\tau$, which in its set-theoretical realization is
-just a set of pairs. In relational programming, a set can be easily
-encoded as a unary relation that answers the membership test. Therefore,
-our goal will be to construct to answer this relation and this relation
-will be our final representation of a signature.
+just a set of pairs. In relational programming, a specific set can be
+easily encoded as a unary relation so a unary will serve as our
+representation of a signature and defining a signature will be a matter
+of constructing this relation.
+
+\subsubsection{The Relationship Between the Lexical Database and a Signature}
 
@@
 and we want to arrive at a set of pairs of constants and their
-types. This mapping decomposed into smaller parts by considering the
-contribution of every part of the lexical database individually. What
-will be the right model for the link between some part of the lexical
-database and some part of our signature?
+types. This mapping can be decomposed into smaller parts by considering
+the contribution of every part of the lexical database
+individually. What will be the right model for the link between some
+part of the lexical database and some part of our signature?
 
@@
     lexical database is a relation that can assign more than one
-    distinct hypertag to a single wordform (e.g. with different part of
-    speech) and we would like to have constants of different types for
-    each of these hypertags.
+    distinct hypertag to a single wordform (e.g. a wordform with
+    different possible part of speech) and we would like to have
+    constants of different types for each of these hypertags.
   \item[Function of hypertag] Conversely, we might imagine that the
-    hypertags of the lexical database are mapped via a function to the
-    constants of the signature. Here we run into trouble when we
+    hypertags of the lexical database are mapped via some function to
+    the constants of the signature. Here we run into trouble when we
     consider variations in spelling or phonology. We might have two
     items in our lexical database that share the same hypertag but whose
-    wordforms still differ. We would like to have two distinct constants
-    in our signature then, so that when these two constants are mapped
-    by the surface lexicon, they will yield two different strings.
+    wordforms still differ. We would then like to have two distinct
+    constants in our signature as well, so that when these two constants
+    are mapped by the surface lexicon, they will yield two different
+    strings.
   \item[Function of lexical entry] Since neither the wordform nor the
@@
     in~\ref{ssec:acg-sem}, the constants of the semantic signature
-    $\Sigma_{Sem}$ contained predicates for the common nouns $man$ and
-    $woman$ and for the verb $loves$ and some common logical
-    furniture. However, it did not contain any constants which would
-    correspond to the determiners $every$ and $some$.
+    $\Sigma_{Sem}$ contained predicates for the common nouns \emph{man}
+    and \emph{woman} and for the verb \emph{loves} and some common
+    logical furniture. However, it did not contain any constants which
+    would correspond to the determiners \emph{every} and \emph{some} and
+    therefore there would be no constant that the function could assign
+    to these two wordforms.
   \item[Partial function of lexical entry] We could fix the issue above
@@
     lexical entry which represents e.g. the 3rd person singular present
-    tense form $loves$ of the transitive verb $love$. A legitimate
-    account of subject/object scope ambiguity would be to consider two
-    constants in our abstract syntactic signature that have distinct
-    images in some semantic lexicon, one giving the outer scope to the
-    subject and the other to the object. This means that in order to
-    define such a signature, we would need to link a single lexical
+    tense form \emph{loves} of the transitive verb \emph{love}. A
+    legitimate account of subject/object scope ambiguity would be to
+    consider two constants in our abstract syntactic signature that have
+    distinct images in some semantic lexicon, one giving the outer scope
+    to the subject and the other to the object. This means that in order
+    to define such a signature, we would need to link a single lexical
     entry in our database to more than one constant in the signature.
   \item[Relation between lexical entry and constant] Finally,
-    considering all the above common scenarios, the only mathematical
-    model general enough to handle all of the above cases is that of a
-    relation. In our system, this will correspond exactly to a relation
-    that will be provided by the grammar author.
+    considering all the above scenarios, the only mathematical model
+    general enough to handle all of these cases is a relation. In our
+    system, this will correspond exactly to a specific relation that
+    will be provided by the grammar author.
 \end{description}
 
-We have established that lexicons will be defined by a relation, a
+\subsubsection{Representing Constants and Defining Signatures}
+
+We have established that signatures will be defined by a relation, a
 relation between a lexical entry (a wordform and a hypertag) and a
-constant (an identity\footnote{The identity of a constant is simply what
-  distinguishes it from other constants of the same signature having the
-  same type.} and a type). The identity of a lexical constant also has a
-complex structure. We know that the lexical constants we define using
-the relation described above are always linked to a lexical entry. This
-lexical entry at least partially identifies the constant. For cases
-where we have more than one constant per lexical entry (such as the
-example of the two different scope readings of $loves$), we need some
-extra piece of information to specify which constant we are referring to
-(we call this extra piece of information a \emph{specifier}). This
-specifier can have any shape that the grammar author desires. We can see
-an example of a lexical constant in Figure~\ref{fig:lex-const}.
+constant (an identity\footnote{The identity of a constant is simply the
+  thing that distinguishes it from other constants of the same signature
+  having the same type.} and a type). The identity of a lexical constant
+also has a complex structure. We know that the lexical constants we
+define using the relation described above are always linked to a lexical
+entry. This lexical entry at least partially identifies the
+constant. For cases where we have more than one constant per lexical
+entry, such as in the example of the two different scope readings of
+\emph{loves}, we need some extra piece of information to specify exactly
+which of the two constants we are referring to (we call this extra piece
+of information a \emph{specifier}). This specifier can have any shape
+that the grammar author desires. We can see an example of a lexical
+constant in Figure~\ref{fig:lex-const}.
 
 \begin{figure}
-  % TODO: Center this.
   \centering
@@
 hypertag (that is, a single lexical entry) and outputs (possibly
-multiple) pairs of specifiers and types. Since we are working in a
-functional programming language that supports manipulation of
-higher-order functions, we can provide helper functions which construct
-these 4-ary relations for the grammar author in some simple cases. Our
-library thus provides functions like \texttt{unitypedr}, which expects a
-type and returns the 4-ary relation that assigns to every lexical item a
-single constant whose specifier is $nil$ and whose type is the supplied
-type, or \texttt{ht->typer}, which expects a mapping from hypertag
-patterns to types and returns the 4-ary relation which tries to map the
-hypertag against the patterns and yields constants with the
-corresponding types and $nil$ specifiers. When other patterns emerge,
-the grammar author is completely free to implement her new abstractions
-using the full power of the functional programming language. See
-Figure~\ref{fig:lex-sig-impl} for an example of implementing a
+multiple) pairs of specifiers and types.
+
+Since we are working in a functional programming language that supports
+manipulation of higher-order functions, we can provide helper functions
+which construct these 4-ary relations for the grammar author in some
+simple cases. Our library thus provides functions like
+\texttt{unitypedr}, which expects a type and returns the 4-ary relation
+that assigns to every lexical item a single constant whose specifier is
+$nil$ and whose type is the supplied type, or \texttt{ht->typer}, which
+expects a mapping from hypertag patterns to types and returns the 4-ary
+relation which tries to map the hypertag against the patterns and yields
+constants with the corresponding types and $nil$ specifiers. When other
+patterns emerge, the grammar author is completely free to implement her
+new abstractions using the full power of the functional programming
+language.
+
+See Figure~\ref{fig:lex-sig-impl} for an example of implementing a
 lexicalized version of the $\Sigma_{Synt}$ signature
@@
   \caption{\label{fig:lex-sig-impl} A lexicalized version of the
-    $\Sigma$(\_{Synt}) signature from~\ref{ssec:example-sig}.}
+    syntactic signature from~\ref{ssec:example-sig}.}
 \end{figure}
@@
 These constants are not associated to any lexical entry or specifier and
-therefore their representation differs. See
-Figure~\ref{fig:nonlex-const} for an example of an non-lexicalized
-constant from a semantic signature.
+therefore their representations must be different. See
+Figure~\ref{fig:nonlex-const} for an example of the representation we
+use for non-lexicalized constants.
 
@@
   \caption{\label{fig:nonlex-const} An example of the representation of
-    a non-lexicalized constant (the conjunction operator in a semantic
-    signature).}
+    a non-lexicalized constant (in this case, it is the conjunction
+    operator in a semantic signature).}
 \end{figure}
@@
 sufficient by themselves and will need to be complemented by other
-signatures which contain the lexical constants. Since we represent
+signatures which contain lexical constants. Since we represent
 signatures as unary relations, combining them to produce their unions or
-intersections is just a matter of taking their disjunctions or
-conjunctions, respectively. With this, we have enough to fully define
-the other signatures of section~\ref{sec:acg} in a lexicalized
-manner, see Figure~\ref{fig:example-sig-impl}.
+intersections is just a matter of taking their disjunctions
+(\texttt{ors}) or conjunctions (\texttt{ands}), respectively.
 
@@
 \end{verbatim}
-  \caption{\label{fig:nonlex-sig-impl} A signature of non-lexical
-    constants belonging to $\Sigma$(\_{Sem})
-    from~\ref{ssec:acg-sem}. The single arrow \texttt{->} corresponds
-    to linear implication while the double arrow \texttt{=>} corresponds
-    to intuitionistic implication.}
+  \caption{\label{fig:nonlex-sig-impl} A signature of the non-lexical
+    semantic constants belonging to the semantic signature
+    of~\ref{ssec:acg-sem}. The single arrow \texttt{->} corresponds to
+    linear implication while the double arrow \texttt{=>} corresponds to
+    intuitionistic implication.}
 \end{figure}
+
+With this, we have enough to fully define all the signatures of
+section~\ref{sec:acg} in a lexicalized manner, see
+Figure~\ref{fig:example-sig-impl}.
 
@@
 \end{verbatim}
-  \caption{\label{fig:example-sig-impl} The definition of the
+  \caption{\label{fig:example-sig-impl} The definitions of the
     lexicalized versions of the example signatures of
@@
 
-\subsection{Defining lexicons}
+\subsection{Defining Lexicons}
 
@@
 \tau\mathclose{>}$ to just the type assignment function $\tau$, we can
-similarly reduce lexicons from the pairs $\mathopen{<}F, G\mathclose{>}$
-to only $G$, the part which operates on terms. $F$, or at least its
-relevant subset, can be easily retrieved by putting the types of the
-abstract constants side-by-side with the types of their object terms to
-get a sufficient subset of $\hat{F}$ and finding $F$, the mapping from
-atomic abstract types whose homomorphic extension includes $\hat{F}$.
+reduce lexicons from the pairs $\mathopen{<}F, G\mathclose{>}$ to just
+$G$, the part which operates on terms. $F$, or at least its relevant
+subset, can be easily retrieved from $G$. We take the pairs of abstract
+constants and object terms that $G$ is composed of, we take their types
+and we arrive at pairs of abstract-level and object-level types. These
+pairs form a subset of $\hat{F}$. $F$ (or at least its relevant subset),
+which only maps the atomic abstract types, can be inferred from this
+subset of $\hat{F}$.
 
-In our relational framework, a mapping is most directly represented as a
-binary relation, which is the representation that we have chosen. In the
-previous subsection, we talked a lot about how individual entries in
-the lexical database generate the items of a signature. Solving the
-problem for signatures also solved it for lexicons, since their
-structure is very similar: a signature is a set of pairs of constants
-and types, a lexicon is a set of pairs of constants and
-terms. Therefore, a lexical entry contributes the object terms for the
-constants which it generated in the abstract signature.
+In our relational framework, a mapping can be represented most directly
+as a binary relation and that is the representation that we use in our
+system. In the previous subsection, we talked a lot about how individual
+entries in the lexical database generate the items of a
+signature. Solving the problem for signatures also solved it for
+lexicons, since a lexicon contains exactly one object term for every
+constant in its abstract signature. Therefore, a pair of an abstract
+constant and an object term belonging to some lexicon is generated by
+the same lexical entry that generated the abstract constant.
 
@@
 as to make the lexical entry directly accessible to the lexicon
-implementation. This way, the lexicon author can inspect the hypertag
-and the wordform of the lexical entry, the specifier of the concrete
-constant that it is mapping right now and even the type she assigned to
-the constant in the abstract signature to allow her to choose the
-correct object term. Furthermore, since lexicons are still
-relations/sets, we can compose them together by disjunction. See
-Figure~\ref{fig:lex-impl} for the implementation of the lexicalized
-versions of the $\mathcal{L}_{Syntax}$ and $\mathcal{L}_{Sem}$ lexicons.
+implementation. Since the lexicon is directly implemented by the grammar
+author as a binary relation between constants and terms, the grammar
+author can inspect the abstract constant and directly access the
+hypertag and wordform of the lexical entry that generated it, the
+specifier that was assigned to it and the type of the constant and then
+express the object term assigned to the constant using any combination
+of these.
+
+\subsubsection{Succinct Definitions of Lexicons}
+
+The above is already enough to define any lexicon. However, our toolkit
+presents convenience facilities that capture common forms of
+lexicons. With these, we will be able to define lexicalized versions of
+the lexicons of section~\ref{sec:acg} in a concise point-free
+style. Furthermore, the grammar author is not limited to the convenience
+facilities we have provided and is completely free to use the same
+programming language we have used to build on our abstractions and build
+her own right within her grammar definition.
 
@@
 \end{verbatim}
-  \caption{\label{fig:lex-impl} The implementation of the
-    $\mathcal{L}$(\_{Syntax}) and $\mathcal{L}$(\_{Sem}) lexicons from
-    section~\ref{sec:acg}.}
+  \caption{\label{fig:lex-impl} The definitions of the lexicalized
+    versions of the lexicons of section~\ref{sec:acg}.}
 \end{figure}
 
- NOTE: This is too verbose both to put on the paper and also to
- code. These examples demonstrate two common and related shapes of
- lexicons. The $\mathcal{L}$(\_{Syntax}) and the first half of the
- $\mathcal{L}$(\_{Sem}) are lexicalized, in the sense that the object
- term contains a lexical constant with the same lexical entry as the
- abstract constant.\footnote{This notion of ``lexicalized'' is quite
-   different to the one I use in the rest of the thesis so far (which
-   roughly means ``generated by a lexical database'') and also different
-   from the one used by Kanazawa (FIX!).} This means that before
- constructing the object term, we have to acquire the object constant
- that is contained in the term and then we just fit it in of the
- recipes, which we choose according to the hypertag. The second half of
- $\mathcal{L}$(\_{Sem}) simply assigns object terms based on the
- hypertag. A more concise and desirable implementation would allow us
- simply write the pairs of hypertags and object terms (recipes encoded
- as lambdas). My plan is to have exactly this, but I have not thought
- yet about how to split the various aspects into different reusable
- abstractions.
+We have the lexicon analogues to \texttt{nonlex-sigr} and
+\texttt{ht->typer}, they are called \texttt{nonlex-lexiconr} and
+\texttt{ht-lexiconr} and they work the same way as the signature
+versions but assign object terms instead of types.
 
-\section{Testing signatures and lexicons}
+When mapping lexical constants by a lexicon, we usually produce object
+terms which contain constants of the object signature that are generated
+by the same lexical entry as the abstract constant
+(e.g. $\mathcal{L}_{syntax}(\synt{love}) = \lambda^{\circ} x y.\ x +
+loves + y$). Managing this involves some tedious boilerplate that can be
+avoided by another convenience facility. First, we define a lexicon that
+produces the desired object term but with the object constant abstracted
+out (e.g. $\lambda^{\circ} c x y.\ x + c + y$). Then, we apply the
+\texttt{lexicalizer} function to the object signature and to this
+lexicon to get the desired lexicon which fills in the correct object
+constants.
+
+It is also useful to be able to easily insert the non-lexical constants
+of the object signature in the object terms we are giving when defining
+a lexicon. This service is provided by the anaphoric macro
+\texttt{with-sig-consts} which introduces all of the non-lexical
+constants of a signature into scope for easy use inside of lexicon
+definitions.
+
+Finally, since lexicons are again just sets of pairs, we can take their
+unions and intersections using disjunction (\texttt{orr}) and
+conjunction (\texttt{andr}).
+
+This gives us everything we need to define the lexicalized versions of
+the lexicons of section~\ref{sec:acg} in a concise and elegant
+manner. See Figure~\ref{fig:lex-impl} for the definitions.
+
+\section{Testing Signatures and Lexicons}

### 2013-06-17 motivation.tex
Switched to the report class, start using theorem envs
Last night shift

--- 2013-05-06 motivation.tex
@@
-\subsection{Motivation}
+\section{Motivation}
 
@@
 propositions have shown to have an important effect on anaphora and on
-the semantic content of the discourse.
+the semantic content of the discourse \cite{asher2003logics}.
 
@@
 (\ref{narxexp-nar-b}) occurred after the event described in
-(\ref{narxexp-nar-a}). In the latter case (\ref{narxexp-exp}), we can
-conversely infer that the events happened in the opposite order. We feel
-that a complete understanding of the two examples above entails
-inferring the correct temporal order and we would welcome a principled
-way to make this kind of decisions.
+(\ref{narxexp-nar-a}). In (\ref{narxexp-exp}), we can conversely infer
+that the events happened in the opposite order. We feel that a complete
+understanding of the two examples above presupposes correctly inferring
+the temporal order of the events described and we would thus welcome a
+principled way to handle these distinctions.
 
@@
   \ex \label{salmon} \begin{xlist}
-    \ex Max had a lovely evening last night.
-    \ex He had a great meal.
-    \ex \label{salmon-here} He ate salmon.
-    \ex He devoured lots of cheese.
-    \ex He then won a dancing competition.
+    \ex \label{salmon-evening} Max had a lovely evening last night.
+    \ex \label{salmon-meal} He had a great meal.
+    \ex \label{salmon-ate} He ate salmon.
+    \ex \label{salmon-cheese} He devoured lots of cheese.
+    \ex \label{salmon-competition} He then won a dancing competition.
   \end{xlist}
@@
 
-%% \sdrtree{
-%%  &                      & \LAB{Max had a lovely evening} \\
-%%           & \LAB{He had a great meal} & & \LAB{He won a dancing competition} \\
-%% \LAB{He ate salmon} &     & \LAB{He devoured cheese}
-%% }
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{diagrams/sdrt.pdf}
+  \caption{\label{fig:sdrt-discourse} The SDRT structure of the
+    discourse in \ref{salmon}.}
+\end{figure}
 
-The discourse in (\ref{salmon}) and its structure in (TODO) demonstrate
-another feature of discourse structures. First off, seeing the
-hiearchical structure of the discourse gives us important information
-about the granularity of the descriptions given in the individual
-propositions, information that could be useful for performing tasks such
-as text summarization.
+The discourse in (\ref{salmon}) and the representation of its structure
+on Figure~\ref{fig:sdrt-discourse} demonstrate another feature of
+discourse structures. First off, seeing the hierarchical structure of
+the discourse gives us important information about the granularity of
+the description employed in the individual propositions, information
+that could be useful for performing tasks such as text summarization.
 
@@
 coherently follow our excerpt. SDRT, \cite{asher2003logics}, the theory
-of discourse structure that we adhere to, would not license a discourse
-structure in which the new proposition connects to (\ref{salmon-here})
-as its \emph{Elaboration}. As a consequence, it states that the pronoun
-``it'' cannot have the salmon as its antecedent, which is a constraint
-that is not enforced by prior theories of discourse such as DRT.
+of discourse structure that we will adhere to, would not license a
+discourse structure in which the new proposition connects to
+(\ref{salmon-ate}) as its \emph{Elaboration}. As a consequence, it
+states that the pronoun ``it'' cannot have the salmon as its antecedent.
 
@@
 The grammatical formalism of our choice for this task will be the
-Abstract Categorial Grammars (ACG) \cite{de2001towards}. This framework
+abstract categorial grammars (ACGs) \cite{de2001towards}. This framework
 lets use lambda calculus to express the syntax-semantics interface in a
 fashion similar to Montague's. Furthermore, elegant techniques for using
-ACGs to implemenent DRT and SDRT using continuations have been
-discovered \cite{de2006towards} \cite{asher2011sdrt}
-\cite{asher2011montagovian}.
+ACGs to implement DRT and SDRT using continuations have been
+discovered (see \cite{de2006towards}, \cite{asher2011sdrt} and
+\cite{asher2011montagovian}).
 
@@
 
-Finally, our work will be supported by the existence of a corpus
+Finally, our work will be motivated by the existence of a corpus
 annotated with the rhetorical relations of SDRT, the Annodis corpus
-\cite{afantenos2012empirical}.
+\cite{afantenos2012empirical}, which would allow evaluation of the final
+grammar.

### 2013-06-17 outline.tex
Last night shift

\section{Outline}

In the rest of this chapter, we proceed to introduce two grammatical
formalisms that are of interest to our work, the formalism of abstract
categorial grammars and the formalism interaction grammars. We talk
about a formal connection between the two and its value with respect to
constructing our grammar.

In Chapter~\ref{chap:implementation}, we present our software
implementation of the ACG machinery which will enable us to write and
test grammars lexicalized by a lexicon.

In Chapter~\ref{chap:constraints}, we take the first steps into
designing a wide-coverage grammar by showing the treatment of several
disparate linguistic constraints in the framework of ACGs and the
challenges inherent in trying to combine them all in a single grammar.

Chapter~\ref{chap:gacg} builds on the problems highlighted in
Chapter~\ref{chap:constraints} and introduces a generalization of
abstract categorial grammars that is geared towards solving these
problems.

We conclude our findings with Chapter~\ref{chap:conclusion}, in which we
summarize our findings and point out potential directions for future
work.

### 2013-06-17 acg.tex
This is the printed version.

@@
 the string concatenation operator, $+$. $\tau_{String}$ assigns the type
-$\textsc{String}$ to all the wordforms and to $\epsilon$ and the type
+$\textsc{String}$ to all the wordforms and to $\epsilon$, and the type
 $\textsc{String} \limp \textsc{String} \limp
@@
   $\hat{F}$ and $\hat{G}$ to refer to the homomorphic extensions of $F$
-  and $G$ to $\mathcal{T}(A_1)$ and $\Lambda(\Sigma_1)$ respectively).
+  and $G$ to $\mathcal{T}(A_1)$ and $\Lambda(\Sigma_1)$
+  respectively). This property will be referred to as the
+  \emph{homomorphism property} of a lexicon.
 \end{itemize}
@@
 computation representing a phrase into a more specific object
-computation that computes the string representing the phrase.
+computation which calculates the string representation.
 
@@
 
-Given the above mappings, we can use the homomorphic properties of our
-type mapping to arrive at the object-level interpretations of complex
-abstract types, such as the type of predicates, $NP \limp S$, which
-becomes a unary string function, $\textsc{String} \limp
-\textsc{String}$, or the type of quantified noun phrases, $(NP \limp S)
-\limp S$, which becomes a higher-order string function,
-$(\textsc{String} \limp \textsc{String}) \limp \textsc{String}$.
+We can construct the homomorphic extension of the above type mapping
+which can give us the object-level interpretations of complex abstract
+types, such as the type of predicates, $NP \limp S$, which becomes a
+unary string function, $\textsc{String} \limp \textsc{String}$, or the
+type of quantified noun phrases, $(NP \limp S) \limp S$, which becomes a
+higher-order string function, $(\textsc{String} \limp \textsc{String})
+\limp \textsc{String}$.
 

### 2013-06-17 acknowledgments.tex
This is the printed version.

\vspace{2\baselineskip}

\topskip0pt
\vspace*{\fill}

\centerline{\textbf{Acknowledgments}}

\vspace{2\baselineskip}

I would like to thank my friends in Nancy for making my year at
Universit\'{e} de Lorraine a wonderful one.

\vspace{\baselineskip}

I would like to thank my supervisor for his enlightening guidance and
his dedication and attention to our work.

\vspace{\baselineskip}

I would like to thank my parents for always supporting me in the things
I have chosen to do.

\vspace{\baselineskip}

\vspace{\baselineskip}

I would like to thank Wes Anderson for \emph{Moonrise Kingdom}.

\vspace*{\fill}

### 2013-06-17 constraints.tex
This is the printed version.

@@
 
-TODO: Change this after splitting the chapters.
-
-In the rest of this chapter, we will devote ourselves to investigating
-ways of disentangling these.
-
-\section{Intersections of Types}
-
-Our goal is to split the types into smaller parts that only talk about
-features belonging to one of the three mechanisms covered above. This
-means that we would like to take some phrase about which we know a lot
-of disparate pieces of information (presence of negative determiner,
-trace status, number\ldots), having type
-e.g. $NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}$,
-and use it as an argument for some function which only cares about one
-channel of information and would have a type like
-$NP\_\textcolor{red}{NEG{=}T} \limp N\_\textcolor{red}{NEG{=}F} \limp
-N\_\textcolor{red}{NEG{=}T}$. To make a type such as
-$NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}$
-be a legal argument to a function which expects an argument of type
-$NP\_\textcolor{red}{NEG{=}T}$, we will need to introduce some notion of
-\emph{subtyping}.
-
-The types we have been showing up to now propose a very straightforward
-notion of subtyping. All of the types can be seen as conjunctions of
-properties where $\_$ (underscore) is the conjunction operator. A type
-$\tau$ can then be seen as a subtype of type $\sigma$ whenever the
-conjuncts of $\tau$ form a superset of the conjuncts of $\sigma$. This
-is not unlike the structural typing of records in OCaml (CITE SOME OCAML
-SPEC HERE), where a record type is a subtype of another record type if
-the former provides, among others, all the fields provided by the latter
-and gives them the same type. In our case, one type is a subtype of
-another if the former gives, among others, all the properties given by
-the latter. This notion of subtyping can be expressed by adding the rule
-of inference in Figure~\ref{fig:type-superset} to our type
-system.\footnote{In the rule \ldots (talk about $\land$ instead of $\_$
-  (underscore) here...)}
-
-\begin{figure}
-  \begin{prooftree}
-    \AxiomC{$\Gamma_l; \Gamma_i \vdash_\Sigma M : A_1 \land \ldots A_{i-1} \land A_i \land A_{i+1} \ldots \land A_n$}
-    \RightLabel{(subtype)}
-    \UnaryInfC{$\Gamma_l; \Gamma_i \vdash_\Sigma M : A_1 \land \ldots A_{i-1} \land A_{i+1} \ldots \land A_n$}
-  \end{prooftree}
-  \caption{\label{fig:type-superset} A rule of inference giving us
-    subtyping for conjunctions of atomic types. For readability, we use
-    $\land$ instead of $\_$ (underscore) to designate the conjunction of
-    properties.}
-\end{figure}
-
-(Talk about whether or not this is a conservative extension...)
-
-With the new subtyping rule, we can
-
-$$
-S_{tatou} : N \land NEG{=}F \land NUM{=}SG
-$$
-$$
-S_{le_{NEG_1}} : (N \land NEG{=}F) \limp (((NP \land NEG{=}F) \limp S) \limp S)
-$$
-$$
-S_{le_{NEG_2}} : (N \land NEG{=}T) \limp (((NP \land NEG{=}T) \limp S) \limp S)
-$$
-$$
-S_{le_{NUM}} : (N \land NUM{=}SG) \limp (((NP \land NUM{=}SG) \limp S) \limp S)
-$$
-$$
-S_{le_{VAR}} : N \limp (((NP \land VAR{=}F) \limp S) \limp S)
-$$
-
-\ldots
+In the next chapter, we will present our proposed solution to the
+problem of writing simple grammars that cover multiple linguistic
+constraints.

### 2013-06-17 gacg.tex
This is the printed version.

@@
 and $\mathcal{L}_{23}(A_2)$, images of the abstract languages given by
-the lexicons. Since the lexicons are homomorphic mappings w.r.t. the
-types, all elements of $\mathcal{L}_{12}(A_1)$ will have the type
+the lexicons. Because of the homomorphism property of lexicons, all
+elements of $\mathcal{L}_{12}(A_1)$ will have the type
 $\mathcal{L}_{12}(S_1) = S_2$ and will therefore also belong to $A_2$
@@
 distinguished type. To all the other nodes, we will assign the result of
-mapping their parent's language by the lexicon linking the two. By
-homomorphism of the lexicons, we can find distinguished types for all
-the non-root nodes. They are simply the images of their parents'
-distinguished types by the connecting lexicons. For all of the nodes, it
-is true that every term in the language assigned to the node has the
-node's distinguished type.
+mapping their parent's language by the lexicon linking the two. Thanks
+to the homomorphism property of the lexicons, we can find distinguished
+types for all the non-root nodes. They are simply the images of their
+parents' distinguished types by the connecting lexicons. For all of the
+nodes, it is true that every term in the language assigned to the node
+has the node's distinguished type.
 
@@
 of the language must have the root's distinguished type. This is a
-property that is also true in the non-root cases due to the homomorphic
-properties of lexicons. If we wanted to unify the definitions of the
-root language and the non-root languages, we would have to check that a
-term of the language has the distinguished type and that it has an
-antecedent in the parent, if it has any parent. A relational statement
-of this unification could look like this.
+property that is also true in the non-root cases due to the homomorphism
+property of lexicons. If we wanted to unify the definitions of the root
+language and the non-root languages, we would have to check that a term
+of the language has the distinguished type and that it has an antecedent
+in the parent, if it has any parent. A relational statement of this
+unification could look like this.
 
@@
 
+  \begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.2\textwidth}
+      \centering
+      \includegraphics[height=0.3\textheight]{diagrams/proof-object.pdf}
+      \caption{\label{fig:proof-object} The case for object languages.}
+    \end{subfigure}
+    \quad
+    \begin{subfigure}[b]{0.4\textwidth}
+      \centering
+      \includegraphics[height=0.3\textheight]{diagrams/proof-intersection.pdf}
+      \caption{\label{fig:proof-intersection} The case for intersections.}
+    \end{subfigure}
+    \quad
+    \begin{subfigure}[b]{0.3\textwidth}
+      \centering
+      \includegraphics[height=0.3\textheight]{diagrams/proof-lexicon.pdf}
+      \caption{\label{fig:proof-lexicon} The case for transformations by
+        a lexicon.}
+    \end{subfigure}
+    \caption{Figures demonstrating that extrinsic languages contain
+      object languages closed on transformation by a lexicon and
+      intersection.}
+  \end{figure}
+
   \begin{itemize}
@@
       and $\Sigma_O$ as signatures and $S$ and $\mathcal{L}(S)$ as
-      distinguished types respectively. Finally, we label the edge
+      distinguished types respectively . Finally, we label the edge
       $(A,O)$ with $\mathcal{L}$ and we get a graphical ACG $G'$ in
-      which $\mathcal{E}_{\mathcal{G'}}(O) = \mathcal{O}(G) = L$.
+      which $\mathcal{E}_{\mathcal{G'}}(O) = \mathcal{O}(G) = L$
+      (Figure~\ref{fig:proof-object}).
 
@@
       then add a new node $v$ such that $\mathcal{E}_{\mathcal{G}'}(v) =
-      L_1 \cap L_2$.
+      L_1 \cap L_2$ (Figure~\ref{fig:proof-intersection}).
 
@@
         \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
-        \mathcal{E}_{\mathcal{G}}(u_1) \cap
-        \mathcal{E}_{\mathcal{G}}(u_2) \\ &= L_1 \cap L_2
+        \mathcal{E}_{\mathcal{G}_1}(u_1) \cap
+        \mathcal{E}_{\mathcal{G}_2}(u_2) \\ &= L_1 \cap L_2
       \end{align*}
@@
       \mathcal{E}_{\mathcal{G}'}(u_2)$, that
-      $\mathcal{E}_{\mathcal{G}'}(u_1) = \mathcal{E}_{\mathcal{G}}(u_1)$
-      and $\mathcal{E}_{\mathcal{G}'}(u_2) =
-      \mathcal{E}_{\mathcal{G}}(u_2)$ and finally that
+      $\mathcal{E}_{\mathcal{G}'}(u_1) =
+      \mathcal{E}_{\mathcal{G}_1}(u_1)$ and
+      $\mathcal{E}_{\mathcal{G}'}(u_2) =
+      \mathcal{E}_{\mathcal{G}_2}(u_2)$ and finally that
       $\mathcal{E}_{\mathcal{G}}(u_1) = L_1$ and
@@
     $u$. To this graph, we conjoin the edge $(u,v)$ and label it with
-    $\mathcal{L}$. Now we verify that $\mathcal{E}_{\mathcal{G}'}(v) =
-    L$.
+    $\mathcal{L}$ (Figure~\ref{fig:proof-lexicon}).
+
+    Now we verify that $\mathcal{E}_{\mathcal{G}'}(v) = L$.
 

### 2013-06-17 ig.tex
This is the printed version.

@@
 We will briefly discuss one more grammar formalism and that is the
-formalism of Interaction Grammars \cite{guillaume2009interaction}.
+formalism of interaction grammars \cite{guillaume2009interaction}. This
+formalism is of particular interest to us since it is based on the same
+logical basis as ACGs and there is already a detailed wide-coverage
+grammar written in it.
+
 Interaction grammars are a grammatical formalism centered around the

### 2013-06-17 implementation.tex
This is the printed version.

@@
 practical functional programming language presents an empowering way of
-defining metagrammars. This allows the grammar designer to compose his
-grammar in any way he sees fit and to use the metaprogramming facilities
-of the underlying programming language to transcribe the grammar in a
-way that makes the most sense in her case.
+defining metagrammars. This allows the grammar designer to compose her
+grammar in any way she sees fit and to use the metaprogramming
+facilities of the underlying programming language to transcribe the
+grammar in a way that makes the most sense in her case.
 
@@
     system, this will correspond exactly to a specific relation that
-    will be provided by the grammar author.
+    will be provided by the grammar designer.
 \end{description}
@@
 of information a \emph{specifier}). This specifier can have any shape
-that the grammar author desires. We can see an example of a lexical
+that the grammar designer desires. We can see an example of a lexical
 constant in Figure~\ref{fig:lex-const}.
@@
 The system takes care of wiring up the relations and assembling the
-signature. The grammar author's duty is to provide a 4-ary relation that
-links the four parts of a lexical constant: the wordform, the hypertag,
-the specifier and the type. This relation can be thought of as a
-multi-valued ``function'' that receives as input a wordform and a
+signature. The grammar designer's duty is to provide a 4-ary relation
+that links the four parts of a lexical constant: the wordform, the
+hypertag, the specifier and the type. This relation can be thought of as
+a multi-valued ``function'' that receives as input a wordform and a
 hypertag (that is, a single lexical entry) and outputs (possibly
@@
 manipulation of higher-order functions, we can provide helper functions
-which construct these 4-ary relations for the grammar author in some
+which construct these 4-ary relations for the grammar designer in some
 simple cases. Our library thus provides functions like
@@
 constants with the corresponding types and $nil$ specifiers. When other
-patterns emerge, the grammar author is completely free to implement her
-new abstractions using the full power of the functional programming
+patterns emerge, the grammar designer is completely free to implement
+her new abstractions using the full power of the functional programming
 language.
@@
 \subsection{Defining Lexicons}
+\label{ssec:def-lex}
 
@@
 implementation. Since the lexicon is directly implemented by the grammar
-author as a binary relation between constants and terms, the grammar
-author can inspect the abstract constant and directly access the
+designer as a binary relation between constants and terms, the grammar
+designer can inspect the abstract constant and directly access the
 hypertag and wordform of the lexical entry that generated it, the
@@
 the lexicons of section~\ref{sec:acg} in a concise point-free
-style. Furthermore, the grammar author is not limited to the convenience
+style. Furthermore, the grammar designer is not limited to the convenience
 facilities we have provided and is completely free to use the same
@@
 
-\section{Testing Signatures and Lexicons}
+\section{Checking Signatures and Lexicons}
+
+Since we let the grammar designer define signatures and lexicons using
+arbitrary relations, she can also end up defining relations which cannot
+be interpreted as well-formed signatures or lexicons. For this reason,
+our toolkit provides a set of testing facilities which help check that
+the signatures and lexicons are well-defined.
+
+\subsection{Checking Properties of Large Structures}
+
+As opposed to the existing \textbf{ACG development toolkit}, we cannot
+easily enumerate all of the elements in a given signature or lexicon and
+check that they are all consistent. The size of the lexicalized grammar
+is too large to make this approach practical.
+
+We solve this problem by demanding that the grammar designer names some
+subset of the structure and testing is done only on that subset. In our
+case, the most practical solution has been to let the designer just list
+wordforms that generate a sufficient subset of the structure, i.e. the
+designer lists one wordform per every category she has handled in her
+grammar.
+
+A more comfortable solution, which would also be more technically
+involved, would be to have the system test the grammar on an
+automatically deduced selection of constants such that they are
+guaranteed to cover all the cases and code paths in the grammar
+designer's definitions. Implementing such a system could still be
+manageable given the small number of primitives the relational
+programming system is built upon.
+
+\subsection{Checking Signatures}
+\label{ssec:check-sig}
+
+The crucial property that we need to check for in signatures is that the
+relation between constant identities and their types is a function. This
+simply reduces to querying the system for the number of typed constants
+belonging to the signature and having some specific identity. If this
+number is one for all of our test cases, we have verified that (the
+tested subset of) the signature assigns exactly one type to every one of
+its constants.
+
+\subsection{Checking Lexicons}
+
+For lexicons, we check the three following properties:
+
+\begin{description}
+\item[The lexicon is a function,] assigning exactly one object term to
+  each abstract constant. The verification proceeds analogously to the
+  case of checking signatures (subsection \ref{ssec:check-sig}).
+
+\item[The lexicon assigns only well-typed object terms.] This is a
+  consequence of the homomorphism property of lexicons. We verify it by
+  checking whether we can infer some type for every object term assigned
+  to a constant in the test set.
+
+\item[The lexicon has the homomorphism property.] We use the process
+  that we described in subsection~\ref{ssec:def-lex} for inferring the
+  type mapping of a lexicon from its term mapping.
+
+  We first enumerate the pairs of abstract constants belonging to the
+  test set and the object terms assigned to them by the lexicon. We then
+  take the types of these pairs. Finally, we state, by using
+  unification, that this set of pairs is a subset of the homomorphic
+  extension of some mapping of atomic abstract types and let the
+  relational programming system find us that mapping. If it succeeds, we
+  have verified the homomorphism property of the lexicon.
+\end{description}
+
+The above technique of verifying the signature and lexicon definitions
+by automated tests is not only useful for checking their
+well-formedness. The grammar designer can go further and implement new
+tests which demonstrate grammar-specific properties, an example of which
+would be the $cat$-$funct$ principle in Frigram that imposes constraints
+on the relationship between the $cat$ and $funct$ features in the EPTDs
+of the grammar.

### 2013-06-17 motivation.tex
This is the printed version.

@@
 fashion similar to Montague's. Furthermore, elegant techniques for using
-ACGs to implement DRT and SDRT using continuations have been
-discovered (see \cite{de2006towards}, \cite{asher2011sdrt} and
-\cite{asher2011montagovian}).
+ACGs to implement DRT and SDRT using continuations have been discovered
+(see \cite{de2006towards}, \cite{asher2011sdrt},
+\cite{asher2011montagovian} and \cite{qian2011event}).
 

### 2013-06-17 gacg.tex
Fixing mistakes post-print

@@
 It is easy to see that the newly introduced pangraphical languages solve
-all of the counter-intuitive quirks of object languages that we
+all of the counter-intuitive quirks of extrinsic languages that we
 discussed in this subsection.

### 2013-06-23 slides/presentation.tex
First draft of the presentation (unfinished)

\documentclass{beamer}

\usepackage{cmll}

\hypersetup{pdfstartview={Fit}}
\def\limp {\mathbin{{-}\mkern-3.5mu{\circ}}}

\AtBeginSection[]
{
  \begin{frame}
    \frametitle{Outline}
    \tableofcontents[currentsection]
  \end{frame}
}

\begin{document}

\title[G-ACGs]{Towards a Wide-Coverage Grammar}
\subtitle{Graphical Abstract Categorial Grammars}
\author{Ji\v{r}\'{i} Mar\v{s}\'{i}k}
\institute[INRIA Loria]
{
  INRIA Loria
  \and
  UFR Math\'{e}matiques et Informatiques \\
  Universit\'{e} de Lorraine
}
\date[June 2013]{June 25, 2013}

\frame{\titlepage}

\begin{frame}
\frametitle{Outline}
\tableofcontents
\end{frame}

\section{Introduction}

\begin{frame}
  \frametitle{Towards a Wide-Coverage Grammar}

TODO: Mention discourse here.

  Requirements for building a wide-coverage grammar:
    \begin{itemize}
    \item Having linguistic resources to build upon
      \begin{itemize}
      \item Frilex, a formalism-independent lexical database
      \item Frigram, a French interaction grammar
      \end{itemize}
    \item Debugging the grammar
      \begin{itemize}
        \item A toolkit for formulating ACGs as simple metagrammars, for
          examining how they work and for checking their consistency.
      \end{itemize}
    \item Using a formalism that allows for a simple grammar
      \begin{itemize}
      \item The need to express multiple independent constraints led us
        to graphical abstact categorial grammars.
      \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{ACGs and IGs}

  \begin{itemize}
  \item ACGs are based on types (IILL formulas).
  \item IGs are based on polarized tree descriptons.
  \item Perrier has linked the two.

    \vspace{2 mm}
    \begin{columns}[c]
      \begin{column}{0.25\textwidth}\end{column}
      \begin{column}{0.2\textwidth}
        \includegraphics[height=0.3\textheight]{../diagrams/ptd.pdf}
      \end{column}
      \begin{column}{0.1\textwidth}$\to$\end{column}
      \begin{column}{0.2\textwidth}
        $$(NP \limp S) \limp N \limp N$$
        $$N \limp (NP \limp S) \limp N$$
      \end{column}
      \begin{column}{0.25\textwidth}\end{column}
    \end{columns}
  \item The IG of Frigram use different constraints in tree
    descriptions.
    \begin{itemize}
    \item The deductive syntactic model of IGs is built on a wider range
      of linear logic primitives ($\otimes$, $\with$).
    \end{itemize}
  \end{itemize}
\end{frame}

\section{ACG Preliminaries}

\frame{}

\section{Why Graphical ACGs?}

\frame{}

\section{Incompatible Intuitions in G-ACGs}

\begin{frame}
  \frametitle{Two Intuitions for G-ACGs}

  \begin{block}{Algebraic Perspective}
    \begin{itemize}
    \item Every node is a language.
    \item An arrow means that a language is produced from another by
      mapping through a lexicon.
    \item Two arrows converging on a node mean an intersection of
      languages.
    \end{itemize}
  \end{block}
  \begin{block}{Generative Perspective}
    \begin{itemize}
    \item Every node is a term.
    \item An arrow means that a term is produced from another by
      applying a lexicon.
    \item Two arrows converging on a node mean that the two (abstract)
      terms yield the same (object) term.
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Algebraic Intuition}

  \begin{columns}[c]
    \begin{column}{0.7\textwidth}
      \begin{itemize}
      \item Every node is a language.
      \item An arrow means that a language is produced from another by
        mapping through a \textcolor{green}{lexicon}.
      \item Two arrows converging on a node mean an
        \textcolor{red}{intersection} of languages.
      \end{itemize}

$$
\mathcal{I}_{\mathcal{G}}(v) = \{t \in \Lambda(\Sigma_v)
\mid\ \vdash_{\Sigma_v} t : S_v\}
$$
$$
\mathcal{E}_{\mathcal{G}}(v) = \mathcal{I}_{\mathcal{G}}(v) \cap
\textcolor{red}{\bigcap_{(u,v) \in E}}
\textcolor{green}{\mathcal{L}_{(u,v)}}(\mathcal{E}_{\mathcal{G}}(u)).
$$
    \end{column}
    \begin{column}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../diagrams/algebraic-perspective.pdf}
    \end{column}
  \end{columns}

\end{frame}

\begin{frame}
  \frametitle{Two Problems with the Algebraic View}

  \begin{columns}[t]
    \begin{column}{0.5\textwidth}
      \includegraphics[height=0.6\textheight]{../diagrams/both-patterns.pdf}

      How do we get the languages of meanings expressible only by
      syntactically correct sentences?
    \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[height=0.6\textheight]{../diagrams/diamond-grammar.pdf}

      How do we express the language of $Bottom$ terms whose $Left$ and
      $Right$ antecedent are generated by the same $Top$ antecedent?
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Generative Intuition}

  \begin{columns}[c]
    \begin{column}{0.7\textwidth}
      \begin{itemize}
      \item \textcolor{red}{Every node is a term.}
      \item An arrow means that a term is produced from another by
        applying a \textcolor{green}{lexicon}.
      \item Two arrows converging on a node mean that the two (abstract)
        terms yield \textcolor{red}{the same} (object) term.
      \end{itemize}

      \vspace{3 mm}

      $t$ belongs to $\mathcal{P}_{\mathcal{G}}(u)$ iff \\$\exists T :
      \textcolor{red}{V(G) \to \Lambda(\Sigma_u)}$ such that:
      \begin{itemize}
      \item $T_u = t$.
      \item For all $v \in V(G)$, $\vdash_{\Sigma_v} T_v : S_v$.
      \item For all $(v,w) \in E(G)$,
        $\textcolor{green}{\mathcal{L}_{(v,w)}(T_v) = T_w}$.
      \end{itemize}
    \end{column}
    \begin{column}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../diagrams/generative-perspective.pdf}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Only Different Points of View?}

  \begin{itemize}
  \item $\forall \, \mathcal{G}, u$ where $\mathcal{G}$ is arborescent,
    we have $\mathcal{E}_{\mathcal{G}}(u) = \mathcal{P}_{\mathcal{G}}(u)
    = \mathcal{O}_{\mathcal{G}}(u)$.
  \item However, $\exists \, \mathcal{G}, u$ such that
    $\mathcal{E}_{\mathcal{G}}(u) \neq \mathcal{P}_{\mathcal{G}}(u)$.
  \item The two different ``perspectives'' lead to different languages.
    $\implies$ They are not interchangeable points of view.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Properties of G-ACG languages}

  \begin{itemize}
  \item The three language families we defined above are\ldots
    \begin{itemize}
    \item {\ldots}increasingly constraining, $\forall \, \mathcal{G},
      u. \; \mathcal{I}_{\mathcal{G}}(u) \supseteq
      \mathcal{E}_{\mathcal{G}}(u) \supseteq
      \mathcal{P}_{\mathcal{G}}(u)$.
    \item {\ldots}increasingly expressive, $\mathcal{I} \subseteq
      \mathcal{E} \subseteq \mathcal{P}$.
    \end{itemize}
  \item Intrinsic languages are just the abstract languages of ACGs.
    \begin{itemize}
    \item $\mathcal{I} = \mathcal{A}$
    \end{itemize}
  \item Extrinsic languages are simple extensions of ACGs that add
    intersections.
    \begin{itemize}
    \item $\mathcal{E} = \mathcal{O}^{\mathcal{L}{\cap}}$
    \end{itemize}
  \item Pangraphical languages are faithful to the generative
    perspective and subsume the previous classes of languages.
    \begin{itemize}
    \item They solve both the example problems we saw before
      (\emph{meanings of grammatical sentences} and the \emph{diamond}).
    \end{itemize}
  \end{itemize}
\end{frame}

\section{Conclusion}

\begin{frame}
  \frametitle{Summary of Work Done}

  \begin{itemize}
  \item Direct embedding of IGs into ACGs based on Perrier's theorem
    conjectured not to be practical.
  \item Created a toolkit to serve as basis for more explorations or
    implementations.
  \item Formalized graphical abstract categorial grammars.
    \begin{itemize}
    \item Helps solve multiple constraints elegantly.
    \item Facilitates discussion and inquiry into
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Future Directions}

  \begin{itemize}
  \item Make a closer examination of the relationship between extrinsic
    and pangraphical languages in G-ACGs.
  \item Determine whether ACG object languages are closed on intersection
    or not.
  \item Write grammars using the G-ACG constraint architecture. TODO:
    More specific.
  \item Try to express the multiple independent polarized features of IGs
    using G-ACGs.
  \end{itemize}
TODO: Discourse perspective.
\end{frame}

\end{document}

### 2013-06-24 slides/presentation.tex
Before mock-presentation

@@
 \documentclass{beamer}
-
-\usepackage{cmll}
 
@@
 
-TODO: Mention discourse here.
+  We want to build discourse representations in a compositional manner
+  by using ACGs. Therefore, we need a wide-coverage ACG grammar.
+
+  \vspace{3 mm}
 
@@
     \begin{columns}[c]
-      \begin{column}{0.25\textwidth}\end{column}
-      \begin{column}{0.2\textwidth}
-        \includegraphics[height=0.3\textheight]{../diagrams/ptd.pdf}
+      \begin{column}{0.1\textwidth}\end{column}
+      \begin{column}{0.4\textwidth}
+        $$(NP \limp S) \limp N \limp N$$
+        $$N \limp (NP \limp S) \limp N$$
       \end{column}
@@
       \begin{column}{0.2\textwidth}
-        $$(NP \limp S) \limp N \limp N$$
-        $$N \limp (NP \limp S) \limp N$$
+        \includegraphics[height=0.3\textheight]{../diagrams/ptd.pdf}
       \end{column}
@@
     \item The deductive syntactic model of IGs is built on a wider range
-      of linear logic primitives ($\otimes$, $\with$).
+      of linear logic primitives ($\otimes$, $\&$).
     \end{itemize}
@@
 
-\frame{}
+\newcommand{\synt}[1]{C_{\textrm{#1}}}
+
+\begin{frame}
+  \begin{block}{Abstract languages}
+    Languages of (abstract) terms constrained by the constants they are
+    built on and their types. Defined by a \textcolor{red}{signature}
+    and a \textcolor{green}{type}.
+
+    \begin{columns}[t]
+      \begin{column}{0.5\textwidth}
+        \begin{align*}
+          \synt{tatous} &: \textcolor{red}{N} \\
+          \synt{Laura} &: \textcolor{red}{NP}
+        \end{align*}
+      \end{column}
+      \begin{column}{0.5\textwidth}
+        \begin{align*}
+          \synt{les} &: \textcolor{red}{N \limp NP} \\
+          \synt{aime} &: \textcolor{red}{NP \limp NP \limp S}
+        \end{align*}
+      \end{column}
+    \end{columns}
+
+    $$\synt{aime}\ \synt{Laura}\ (\synt{les}\ \synt{tatous}) :
+    \textcolor{green}{S}$$
+  \end{block}
+
+  \begin{block}{Object languages}
+    Languages of (object) terms which are obtained by replacing the
+    constants in an existing (abstract) term using a \textcolor{blue}{lexicon}.
+
+    \begin{columns}[t]
+      \begin{column}{0.5\textwidth}
+        \begin{align*}
+          \mathcal{L}(\synt{tatous}) &= \textcolor{blue}{tatous} \\
+          \mathcal{L}(\synt{Laura}) &= \textcolor{blue}{Laura}
+        \end{align*}
+      \end{column}
+      \begin{column}{0.5\textwidth}
+        \begin{align*}
+          \mathcal{L}(\synt{les}) &= \textcolor{blue}{\lambda^{\circ} x.\ les + x} \\
+          \mathcal{L}(\synt{aime}) &= \textcolor{blue}{\lambda^{\circ} x y.\ x + aime + y}
+        \end{align*}
+      \end{column}
+    \end{columns}
+
+    $$\mathcal{L}(\synt{aime}\ \synt{Laura}\ (\synt{les}\ \synt{tatous}))
+    =_{\beta} Laura + aime + les + tatous$$
+  \end{block}
+\end{frame}
+
+\begin{frame}
+  \frametitle{ACG Diagrams}
+
+  \begin{columns}[c]
+    \begin{column}{0.5\textwidth}
+      \begin{itemize}
+      \item Two different lexicons interpret the same abstract terms.
+      \item Instead of referring to the two ACGs $\mathopen{<}
+        \Sigma_{Synt}, \Sigma_{String}, \mathcal{L}_{syntax}, S
+        \mathclose{>}$ and $\mathopen{<} \Sigma_{Synt}, \Sigma_{Sem},
+        \mathcal{L}_{sem}, S \mathclose{>}$, we refer to this single
+        diagram.
+      \end{itemize}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \includegraphics[width=\textwidth]{../diagrams/double-acg.pdf}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{ACG Patterns}
+
+  \begin{columns}[c]
+    \begin{column}{0.5\textwidth}
+      \begin{block}{Semantic ambiguities}
+        \vspace{2 mm}
+        \begin{center}
+          \includegraphics[height=0.7\textheight]{../diagrams/parallel-over-serial.pdf}
+        \end{center}
+      \end{block}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \begin{block}{Syntactic constraint}
+        \vspace{2 mm}
+        \begin{center}
+          \includegraphics[height=0.7\textheight]{../diagrams/serial-over-parallel.pdf}
+        \end{center}
+      \end{block}
+    \end{column}
+  \end{columns}
+\end{frame}
 
@@
 
-\frame{}
+\begin{frame}
+  \frametitle{Composing ACG Patterns I}
+
+  In order to stay in the scope of ACGs, our diagrams must be shaped as
+  arborescences (at most one parent per node).
+
+  \begin{columns}[c]
+    \begin{column}{0.5\textwidth}
+      \begin{center}
+        \includegraphics[height=0.7\textheight]{../diagrams/fail1.pdf}
+      \end{center}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \begin{center}
+        \includegraphics[height=0.7\textheight]{../diagrams/fail2.pdf}
+      \end{center}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Multiple Constraints}
+
+  Implementing multiple constraints in a single signature is
+  \textbf{difficult}.
+
+  Agreement:
+    \begin{align*}
+      A_{aime_1} &: NP\_\textcolor{blue}{NUM{=}SG} \limp NP\_\textcolor{blue}{NUM{=}SG} \limp S \\
+      A_{aime_2} &: NP\_\textcolor{blue}{NUM{=}SG} \limp NP\_\textcolor{blue}{NUM{=}PL} \limp S
+    \end{align*}
+
+  Extraction:
+    \begin{align*}
+      E_{aime_1} &: NP\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}NO} \\
+      E_{aime_2} &: NP\_\textcolor{green}{VAR{=}T} \limp NP\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
+      E_{aime_3} &: NP\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{green}{VAR{=}T} \limp S\_\textcolor{green}{EXT{=}ROOT}
+    \end{align*}
+
+  Combined:
+  \begin{small}
+  \begin{align*}
+      C_{aime_1} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}NO} \\
+      C_{aime_2} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}T} \limp NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
+      C_{aime_3} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}T} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
+      C_{aime_4} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}PL}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}NO} \\
+      C_{aime_5} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}T} \limp NP\_\textcolor{blue}{NUM{=}PL}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
+      C_{aime_6} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}PL}\_\textcolor{green}{VAR{=}T} \limp S\_\textcolor{green}{EXT{=}ROOT}
+  \end{align*}
+
+  \end{small}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Composing ACG Patterns II}
+
+  We will need to be able to specify constraints independently if we are
+  to write a manageably simple wide-coverage grammar.
+
+  \begin{columns}[c]
+    \begin{column}{0.5\textwidth}
+      \begin{center}
+        \includegraphics[height=0.7\textheight]{../diagrams/constr-fail-1.pdf}
+      \end{center}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \begin{center}
+        \includegraphics[height=0.7\textheight]{../diagrams/constr-fail-2.pdf}
+      \end{center}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Composing ACG Patterns III}
+
+  Therefore, we would like to able to do something like this (i.e., not
+  an arborescence).
+
+  \begin{columns}[c]
+    \begin{column}{0.6\textwidth}
+      \begin{center}
+        \includegraphics[height=0.7\textheight]{../diagrams/final.pdf}
+      \end{center}
+    \end{column}
+    \begin{column}{0.4\textwidth}
+      How do we assign languages to nodes of the graph? \\
+      $\implies$ Graphical Abstract Categorial Grammars (G-ACGs)
+
+      \vspace{5 mm}
+
+      Formally, we have:
+        \begin{itemize}
+        \item A DAG.
+        \item Nodes labelled with:
+          \begin{itemize}
+          \item Signatures.
+          \item Distinguished types.
+          \end{itemize}
+        \item Edges labelled with:
+          \begin{itemize}
+          \item Lexicons.
+          \end{itemize}
+        \end{itemize}
+    \end{column}
+  \end{columns}
+\end{frame}
 
@@
       $t$ belongs to $\mathcal{P}_{\mathcal{G}}(u)$ iff \\$\exists T :
-      \textcolor{red}{V(G) \to \Lambda(\Sigma_u)}$ such that:
+      \textcolor{red}{V(G) \to \Lambda}$ such that:
       \begin{itemize}
@@
     \item Helps solve multiple constraints elegantly.
-    \item Facilitates discussion and inquiry into
+    \item Facilitates discussion and inquiry into compositions of ACGs.
     \end{itemize}
@@
     or not.
-  \item Write grammars using the G-ACG constraint architecture. TODO:
-    More specific.
+  \item Build a wide-coverage G-ACG and study the patterns used in the
+    graphical architecture.
   \item Try to express the multiple independent polarized features of IGs
     using G-ACGs.
+  \item Assess the applicability of G-ACG architectures at other levels
+    of description, e.g. discourse structure.
   \end{itemize}
-TODO: Discourse perspective.
 \end{frame}

### 2013-06-24 slides/presentation.tex
Typos

@@
       \item The need to express multiple independent constraints led us
-        to graphical abstact categorial grammars.
+        to graphical abstract categorial grammars.
       \end{itemize}
@@
   \item ACGs are based on types (IILL formulas).
-  \item IGs are based on polarized tree descriptons.
+  \item IGs are based on polarized tree descriptions.
   \item Perrier has linked the two.

### 2013-06-24 slides/presentation.tex
New colors for presentation

@@
 \documentclass{beamer}
+
+\usepackage{xcolor}
+
+\definecolor{darkgreen}{rgb}{0., 0.6, 0.}
+\definecolor{darkgold}{rgb}{8., 0.6, 0.}
 
@@
     built on and their types. Defined by a \textcolor{red}{signature}
-    and a \textcolor{green}{type}.
+    and a \textcolor{darkgreen}{type}.
 
@@
     $$\synt{aime}\ \synt{Laura}\ (\synt{les}\ \synt{tatous}) :
-    \textcolor{green}{S}$$
+    \textcolor{darkgreen}{S}$$
   \end{block}
@@
     Languages of (object) terms which are obtained by replacing the
-    constants in an existing (abstract) term using a \textcolor{blue}{lexicon}.
+    constants in an existing (abstract) term using a \textcolor{darkgold}{lexicon}.
 
@@
         \begin{align*}
-          \mathcal{L}(\synt{tatous}) &= \textcolor{blue}{tatous} \\
-          \mathcal{L}(\synt{Laura}) &= \textcolor{blue}{Laura}
+          \mathcal{L}(\synt{tatous}) &= \textcolor{darkgold}{tatous} \\
+          \mathcal{L}(\synt{Laura}) &= \textcolor{darkgold}{Laura}
         \end{align*}
@@
         \begin{align*}
-          \mathcal{L}(\synt{les}) &= \textcolor{blue}{\lambda^{\circ} x.\ les + x} \\
-          \mathcal{L}(\synt{aime}) &= \textcolor{blue}{\lambda^{\circ} x y.\ x + aime + y}
+          \mathcal{L}(\synt{les}) &= \textcolor{darkgold}{\lambda^{\circ} x.\ les + x} \\
+          \mathcal{L}(\synt{aime}) &= \textcolor{darkgold}{\lambda^{\circ} x y.\ x + aime + y}
         \end{align*}
@@
     \begin{align*}
-      A_{aime_1} &: NP\_\textcolor{blue}{NUM{=}SG} \limp NP\_\textcolor{blue}{NUM{=}SG} \limp S \\
-      A_{aime_2} &: NP\_\textcolor{blue}{NUM{=}SG} \limp NP\_\textcolor{blue}{NUM{=}PL} \limp S
+      A_{aime_1} &: NP\_\textcolor{darkgold}{NUM{=}SG} \limp NP\_\textcolor{darkgold}{NUM{=}SG} \limp S \\
+      A_{aime_2} &: NP\_\textcolor{darkgold}{NUM{=}SG} \limp NP\_\textcolor{darkgold}{NUM{=}PL} \limp S
     \end{align*}
@@
     \begin{align*}
-      E_{aime_1} &: NP\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}NO} \\
-      E_{aime_2} &: NP\_\textcolor{green}{VAR{=}T} \limp NP\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
-      E_{aime_3} &: NP\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{green}{VAR{=}T} \limp S\_\textcolor{green}{EXT{=}ROOT}
+      E_{aime_1} &: NP\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
+      E_{aime_2} &: NP\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
+      E_{aime_3} &: NP\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT}
     \end{align*}
@@
   \begin{align*}
-      C_{aime_1} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}NO} \\
-      C_{aime_2} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}T} \limp NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
-      C_{aime_3} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}T} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
-      C_{aime_4} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}PL}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}NO} \\
-      C_{aime_5} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}T} \limp NP\_\textcolor{blue}{NUM{=}PL}\_\textcolor{green}{VAR{=}F} \limp S\_\textcolor{green}{EXT{=}ROOT} \\
-      C_{aime_6} &: NP\_\textcolor{blue}{NUM{=}SG}\_\textcolor{green}{VAR{=}F} \limp NP\_\textcolor{blue}{NUM{=}PL}\_\textcolor{green}{VAR{=}T} \limp S\_\textcolor{green}{EXT{=}ROOT}
+      C_{aime_1} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
+      C_{aime_2} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
+      C_{aime_3} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
+      C_{aime_4} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
+      C_{aime_5} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
+      C_{aime_6} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT}
   \end{align*}
@@
       \item An arrow means that a language is produced from another by
-        mapping through a \textcolor{green}{lexicon}.
+        mapping through a \textcolor{darkgreen}{lexicon}.
       \item Two arrows converging on a node mean an
@@
 \textcolor{red}{\bigcap_{(u,v) \in E}}
-\textcolor{green}{\mathcal{L}_{(u,v)}}(\mathcal{E}_{\mathcal{G}}(u)).
+\textcolor{darkgreen}{\mathcal{L}_{(u,v)}}(\mathcal{E}_{\mathcal{G}}(u)).
 $$
@@
       \item An arrow means that a term is produced from another by
-        applying a \textcolor{green}{lexicon}.
+        applying a \textcolor{darkgreen}{lexicon}.
       \item Two arrows converging on a node mean that the two (abstract)
@@
       \item For all $(v,w) \in E(G)$,
-        $\textcolor{green}{\mathcal{L}_{(v,w)}(T_v) = T_w}$.
+        $\textcolor{darkgreen}{\mathcal{L}_{(v,w)}(T_v) = T_w}$.
       \end{itemize}

### 2013-06-24 slides/presentation.tex
Frame numbers
Moved multiple constraints to annex
Descriptions for schema examples

@@
     \tableofcontents[currentsection]
+    \addtocounter{framenumber}{-1}
   \end{frame}
 }
+
+\setbeamertemplate{navigation symbols}{}
+\setbeamertemplate{footline}
+  {\hfill {\normalsize \insertframenumber{}/\inserttotalframenumber{}}}
 
@@
 
-\frame{\titlepage}
+\frame{\titlepage \setcounter{framenumber}{1}}
 
@@
   \begin{columns}[c]
-    \begin{column}{0.5\textwidth}
+    \begin{column}{0.25\textwidth}
+      {\small Syntax-semantics interface deals with syntactic
+        constraints.}
+
+      $\implies$ Complex.
+    \end{column}
+    \begin{column}{0.25\textwidth}
       \begin{center}
@@
     \end{column}
-    \begin{column}{0.5\textwidth}
+    \begin{column}{0.25\textwidth}
+      {\small Syntactic constraints deal with the syntax-semantics
+        interface.}
+
+      $\implies$ Complex.
+    \end{column}
+    \begin{column}{0.25\textwidth}
       \begin{center}
@@
   \end{columns}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Multiple Constraints}
-
-  Implementing multiple constraints in a single signature is
-  \textbf{difficult}.
-
-  Agreement:
-    \begin{align*}
-      A_{aime_1} &: NP\_\textcolor{darkgold}{NUM{=}SG} \limp NP\_\textcolor{darkgold}{NUM{=}SG} \limp S \\
-      A_{aime_2} &: NP\_\textcolor{darkgold}{NUM{=}SG} \limp NP\_\textcolor{darkgold}{NUM{=}PL} \limp S
-    \end{align*}
-
-  Extraction:
-    \begin{align*}
-      E_{aime_1} &: NP\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
-      E_{aime_2} &: NP\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
-      E_{aime_3} &: NP\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT}
-    \end{align*}
-
-  Combined:
-  \begin{small}
-  \begin{align*}
-      C_{aime_1} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
-      C_{aime_2} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
-      C_{aime_3} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
-      C_{aime_4} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
-      C_{aime_5} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
-      C_{aime_6} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT}
-  \end{align*}
-
-  \end{small}
 \end{frame}
@@
   \begin{columns}[c]
-    \begin{column}{0.5\textwidth}
+    \begin{column}{0.25\textwidth}
+      Second constraint reimplements first constraint.
+
+      $\implies$ Pointless.
+    \end{column}
+    \begin{column}{0.25\textwidth}
       \begin{center}
@@
     \end{column}
-    \begin{column}{0.5\textwidth}
+    \begin{column}{0.25\textwidth}
+      Third constraint reimplements first and second constraint.
+
+      $\implies$ Pointless.
+    \end{column}
+    \begin{column}{0.25\textwidth}
       \begin{center}
@@
 
+\appendix
+\newcounter{finalframe}
+\setcounter{finalframe}{\value{framenumber}}
+
+\begin{frame}
+  \frametitle{Multiple Constraints}
+
+  Implementing multiple constraints in a single signature is
+  \textbf{difficult}.
+
+  Agreement:
+    \begin{align*}
+      A_{aime_1} &: NP\_\textcolor{darkgold}{NUM{=}SG} \limp NP\_\textcolor{darkgold}{NUM{=}SG} \limp S \\
+      A_{aime_2} &: NP\_\textcolor{darkgold}{NUM{=}SG} \limp NP\_\textcolor{darkgold}{NUM{=}PL} \limp S
+    \end{align*}
+
+  Extraction:
+    \begin{align*}
+      E_{aime_1} &: NP\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
+      E_{aime_2} &: NP\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
+      E_{aime_3} &: NP\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT}
+    \end{align*}
+
+  Combined:
+  \begin{small}
+  \begin{align*}
+      C_{aime_1} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
+      C_{aime_2} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
+      C_{aime_3} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
+      C_{aime_4} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
+      C_{aime_5} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
+      C_{aime_6} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT}
+  \end{align*}
+  \end{small}
+\end{frame}
+
+\setcounter{framenumber}{\value{finalframe}}
+
 \end{document}

### 2013-06-24 slides/presentation.tex
Pointless -> Ineffective
Final sitdown with Maxime over the slides

@@
 \definecolor{darkgold}{rgb}{8., 0.6, 0.}
+\definecolor{limegreen}{RGB}{175, 255, 175}
+\definecolor{limeblue}{RGB}{200, 200, 235}
 
@@
 
-  We want to build discourse representations in a compositional manner
-  by using ACGs. Therefore, we need a wide-coverage ACG grammar.
+  \textbf{Objective:} Build discourse representations using ACGs: \\a
+  wide-coverage ACG grammar.
 
-  \vspace{3 mm}
+  \vfill
 
-  Requirements for building a wide-coverage grammar:
   \begin{itemize}
@@
     \end{itemize}
+    \vfill
   \item Debugging the grammar
@@
     \end{itemize}
+    \vfill
   \item Using a formalism that allows for a simple grammar
     \begin{itemize}
-      \item The need to express multiple independent constraints led us
-        to graphical abstract categorial grammars.
+    \item Express multiple independent constraints \\ $\implies$ graphical abstract categorial grammars.
     \end{itemize}
@@
   \item IGs are based on polarized tree descriptions.
-  \item Perrier has linked the two.
+    \vfill
+  \item[$\Rightarrow$] Perrier has linked the two.
 
@@
     \end{columns}
-  \item The IG of Frigram use different constraints in tree
-    descriptions.
+    \vfill
+  \item IG of Frigram use different constraints in tree descriptions.
     \begin{itemize}
@@
   \begin{block}{Abstract languages}
-    Languages of (abstract) terms constrained by the constants they are
-    built on and their types. Defined by a \textcolor{red}{signature}
-    and a \textcolor{darkgreen}{type}.
+    %% Languages of (abstract) terms constrained by the constants they are
+    %% built on and their types. Defined by a \textcolor{red}{signature}
+    %% and a \textcolor{darkgreen}{type}.
+
+    Lambda terms constrained by the types of constants. \\ Defined by a
+    \textcolor{red}{signature} and a \textcolor{darkgreen}{type}.
 
@@
         \mathclose{>}$ and $\mathopen{<} \Sigma_{Synt}, \Sigma_{Sem},
-        \mathcal{L}_{sem}, S \mathclose{>}$, we refer to this single
+        \mathcal{L}_{sem}, S \mathclose{>}$, refer to this single
         diagram.
@@
 
-  In order to stay in the scope of ACGs, our diagrams must be shaped as
-  arborescences (at most one parent per node).
+  ACG diagrams are arborescences.
 
@@
 
-      $\implies$ Complex.
+      \onslide<2->{$\implies$ Complex.}
     \end{column}
@@
 
-      $\implies$ Complex.
+      \onslide<3->{$\implies$ Complex.}
     \end{column}
@@
 
-  We will need to be able to specify constraints independently if we are
-  to write a manageably simple wide-coverage grammar.
+  Writing a simple wide-coverage grammar implies specifying constraints
+  independently.
 
@@
 
-      $\implies$ Pointless.
+      \onslide<2->{$\implies$ Ineffective.}
     \end{column}
@@
 
-      $\implies$ Pointless.
+      \onslide<3->{$\implies$ Ineffective.}
     \end{column}
@@
 \begin{frame}
-  \frametitle{Composing ACG Patterns III}
-
-  Therefore, we would like to able to do something like this (i.e., not
-  an arborescence).
+  \frametitle{Graphical Abstract Categorial Grammars (G-ACGs)}
 
@@
     \end{column}
+
+    \pause
+
     \begin{column}{0.4\textwidth}
-      How do we assign languages to nodes of the graph? \\
-      $\implies$ Graphical Abstract Categorial Grammars (G-ACGs)
-
-      \vspace{5 mm}
-
-      Formally, we have:
         \begin{itemize}
@@
 
-\section{Incompatible Intuitions in G-ACGs}
+\section{Semantics of G-ACGs Nodes}
 
@@
 
+  \setbeamercolor{block title}{bg=limeblue}
+  \setbeamercolor{block body}{bg=limeblue}
   \begin{block}{Algebraic Perspective}
     \begin{itemize}
-    \item Every node is a language.
-    \item An arrow means that a language is produced from another by
-      mapping through a lexicon.
-    \item Two arrows converging on a node mean an intersection of
-      languages.
+    \item Nodes as \emph{languages}.
+    \item Edges define \emph{languages} in terms of other
+      \emph{languages} using lexicons and intersections.
     \end{itemize}
   \end{block}
+  \setbeamercolor{block title}{bg=limegreen}
+  \setbeamercolor{block body}{bg=limegreen}
   \begin{block}{Generative Perspective}
     \begin{itemize}
-    \item Every node is a term.
-    \item An arrow means that a term is produced from another by
-      applying a lexicon.
-    \item Two arrows converging on a node mean that the two (abstract)
-      terms yield the same (object) term.
+    \item Nodes as \emph{terms}.
+    \item Edges describe how \emph{terms} are obtained from other
+      \emph{terms} by applying lexicons.
     \end{itemize}
@@
     \begin{column}{0.7\textwidth}
+      \setbeamercolor{block title}{bg=limeblue}
+      \setbeamercolor{block body}{bg=limeblue}
+      \begin{block}{}
         \begin{itemize}
@@
         \end{itemize}
+      \end{block}
 
@@
     \begin{column}{0.5\textwidth}
+      \begin{center}
         \includegraphics[height=0.6\textheight]{../diagrams/both-patterns.pdf}
+      \end{center}
 
-      How do we get the languages of meanings expressible only by
+      How do we get the language of meanings expressible only by
       syntactically correct sentences?
@@
     \begin{column}{0.5\textwidth}
+      \begin{center}
         \includegraphics[height=0.6\textheight]{../diagrams/diamond-grammar.pdf}
+      \end{center}
 
@@
     \begin{column}{0.7\textwidth}
+      \setbeamercolor{block title}{bg=limegreen}
+      \setbeamercolor{block body}{bg=limegreen}
+      \begin{block}{}
         \begin{itemize}
-      \item \textcolor{red}{Every node is a term.}
+        \item Every node is a \textcolor{red}{term}.
         \item An arrow means that a term is produced from another by
@@
 \begin{frame}
-  \frametitle{Only Different Points of View?}
+  \frametitle{Algebraic and Generative Relations}
 
   \begin{itemize}
-  \item $\forall \, \mathcal{G}, u$ where $\mathcal{G}$ is arborescent,
-    we have $\mathcal{E}_{\mathcal{G}}(u) = \mathcal{P}_{\mathcal{G}}(u)
-    = \mathcal{O}_{\mathcal{G}}(u)$.
+  \item $\forall \, \mathcal{G}, u$ where $\mathcal{G}$ is arborescent:
+    $$\mathcal{E}_{\mathcal{G}}(u) = \mathcal{P}_{\mathcal{G}}(u)
+    = \mathcal{O}_{\mathcal{G}}(u)$$
+    \pause
   \item However, $\exists \, \mathcal{G}, u$ such that
-    $\mathcal{E}_{\mathcal{G}}(u) \neq \mathcal{P}_{\mathcal{G}}(u)$.
-  \item The two different ``perspectives'' lead to different languages.
-    $\implies$ They are not interchangeable points of view.
+    $$\mathcal{E}_{\mathcal{G}}(u) \neq \mathcal{P}_{\mathcal{G}}(u)$$
+    \vfill
+    \pause
+  \item[$\Rightarrow$] They are not interchangeable points of view.
   \end{itemize}
@@
     \end{itemize}
+  \vfill
   \item Intrinsic languages are just the abstract languages of ACGs.
@@
     \end{itemize}
+  \vfill
   \item Extrinsic languages are simple extensions of ACGs that add
@@
     \end{itemize}
+  \vfill
   \item Pangraphical languages are faithful to the generative
@@
     conjectured not to be practical.
+  \vfill
   \item Created a toolkit to serve as basis for more explorations or
     implementations.
+  \vfill
   \item Formalized graphical abstract categorial grammars.
@@
   \begin{itemize}
-  \item Make a closer examination of the relationship between extrinsic
-    and pangraphical languages in G-ACGs.
-  \item Determine whether ACG object languages are closed on intersection
-    or not.
-  \item Build a wide-coverage G-ACG and study the patterns used in the
-    graphical architecture.
-  \item Try to express the multiple independent polarized features of IGs
-    using G-ACGs.
-  \item Assess the applicability of G-ACG architectures at other levels
-    of description, e.g. discourse structure.
+  \item Are extrinsic languages as expressive as pangraphical
+    ones? $$\mathcal{P} \subseteq \mathcal{E}$$
+  \item Are ACG object languages closed on
+    intersection? $$\mathcal{O}^{\mathcal{L}{\cap}} \subseteq \mathcal{O}^{\mathcal{L}}$$
+  \item Express the multiple polarized features of IGs in G-ACGs.
+    \vfill
+  \item Build and study a wide-coverage G-ACG.
+    \vfill
+  \item Apply G-ACGs to discourse.
   \end{itemize}

### 2013-06-26 gacg.tex
Fixing something...???
Fixed the assumptions in the decid. of G-ACGs

@@
 \mathcal{E}_{\mathcal{G}}(v) = \mathcal{I}_{\mathcal{G}}(v) \cap
-\bigcap_{(u,v) \in E} \mathcal{L}_{(u,v)}(\mathcal{E}_{\mathcal{G}}(u)).
+\bigcap_{(u,v) \in E} \mathcal{L}_{(u,v)}(\mathcal{E}_{\mathcal{G}}(u))
 $$
@@
 a lexicon) is an ACG for which there is an effective parsing algorithm
-and type checking is decidable in all the signatures of the graphical
-ACG. Then we can prove the existence of a simple effective parsing
-algorithm for any of the extrinsic languages defined in the graphical
-ACG.
+that lets us find all the abstract terms for a given object term in
+finite time and where type checking is decidable in all the signatures
+of the graphical ACG. Then we can prove the existence of a simple
+effective parsing algorithm for any of the extrinsic languages defined
+in the graphical ACG.
 

### 2013-06-26 ig.tex
Made small changes in IG (precedence vs types)

@@
 interaction grammar is the Frigram grammar of French, we will be using
-French example sentences and their analysis according to
+French example sentences and their analyses according to
 Frigram. Frigram will be discussed in more detail in
@@
   \caption{\label{fig:ig-partial} The result of merging the EPTDs of
-    \emph{la} and \emph{voit} by merging the nodes (1,5) and (1,5) with
+    \emph{la} and \emph{voit} by merging the nodes (1,2) and (1,5) with
     (2,2) and (2,4), respectively.}
@@
 
-This example demonstrates two points of interest to us.
+We can see the EPTDs of both \emph{aucun} and \emph{ne} on
+Figure~\ref{fig:ig-neg}. This example demonstrates two points of
+interest to us.
 
@@
 operation $+$ and the syntactic description producing operation $D$ maps
-an IILL formula into a syntactic description such that above-stated
+an IILL formula into a syntactic description such that the above-stated
 theorem holds. Furthermore, this mapping can be easily constructed since
@@
 formalisms. Let us consider the case of parsing a sentence using a
-lexicalized grammar in both formalism. First, the sentence is tokenized
+lexicalized grammar in both formalisms. First, the sentence is tokenized
 into wordforms and for each wordform, a lexical item is selected, be it
@@
 
-In ACGs, we will try to take the typed constants and to use them to
-produce a term having some distinguished type $S$. Thanks to the
-Curry-Howard correspondence, constructing a term having a given type is
-nothing more than proving that term's type as a formula in the logic of
-our type system. What this means is that our parsing problem boils down
-to proving the IILL sequent $\tau_1, ..., \tau_n \vdash S$ (we omit the
+In ACGs, we will try to take the typed constants and use them to produce
+a term having some distinguished type $S$. Thanks to the Curry-Howard
+correspondence, constructing a term having a given type is nothing more
+than proving that term's type as a formula in the logic of our type
+system. What this means is that our parsing problem boils down to
+proving the IILL sequent $\tau_1, ..., \tau_n \vdash S$ (we omit the
 inclusion of intuitionistic non-linear implications since they are not
@@
 sequent is provable and furthermore, the proof of the sequent is our
-desired syntactic structure in a disguise.
+desired syntactic structure in disguise.
 
@@
 
-A bigger problem is posed by the fact that $D \circ +$ is not
+Secondly, a bigger problem is posed by the fact that $D \circ +$ is not
 surjective. Perrier \cite{perrier2001intuitionistic} characterized the
@@
 precedence constraints, to establish further restrictions on syntactic
-composition, would be lost, as the only restricting factor on syntactic
-composition in ACGs is the type system.
+composition, would be lost.
 
@@
 However, this translation is very lossy in terms of its discriminatory
-power. Many of the constraints which were forced by the original EPTD
+power. Many of the constraints which were enforced by the original EPTD
 are now ignored. We can imagine improving our strategy by using the
@@
 must precede any embedded clauses (2,6)...(2,5), the fact that any such
-embedded clause must serve either the $modal$ or the $obj$ function and
-the fact that the embedded clause (2,5) must have a subject (2,4).
+embedded clauses must serve either the $modal$ or the $obj$ functions
+and the fact that the embedded clause (2,5) must have a subject (2,4).
 

### 2013-06-26 front-page.tex
Fixed links by omitting PDFjam from the pipeline

@@
+\documentclass{report}
+
+\usepackage{graphicx}
+
 \begin{document}
@@
 
-%\includegraphics[width=0.9\textwidth]{images/UL_UFR_MI.jpg}\\[2cm]
+\includegraphics[width=0.9\textwidth]{images/UL_UFR_MI.jpg}\\[1.5cm]
 
-\textsc{\LARGE Master Sciences Cognitives et Applications\\Sp\'ecialit\'e Traitement Automatique du Langage}\\[1.5cm]
+\textsc{\LARGE Master Sciences Cognitives et Applications\\[0.5cm]Sp\'ecialit\'e Traitement Automatique du Langage}\\[1.5cm]
 
@@
 
-\HRule \\[1.75cm]
+\HRule \\[0.5cm]
+
+June 25, 2013 \\[0.75cm]
 
@@
 \end{minipage}
-~\\[2.1cm]
-%\includegraphics[width=0.35\textwidth]{images/logoLORIA.jpg}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\includegraphics[width=0.35\textwidth]{images/tonequipe.png}
+~\\[0.6cm]
+\includegraphics[width=0.35\textwidth]{images/logoLORIA.jpg}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\includegraphics[height=0.10\textheight]{images/Calligramme.png}
 
@@
 
-\newpage
-~~ \pagenumbering{gobble}
-\newpage
-\pagenumbering{arabic}
-\setcounter{page}{3}
+%% \newpage
+%% ~~ \pagenumbering{gobble}
+%% \newpage
+%% \pagenumbering{arabic}
+%% \setcounter{page}{3}
+
+\end{document}

### 2013-06-26 acg.tex
Proofreading for publication

@@
   \item $A$ is a (finite) set of atomic types
-  \item $\tau$ is the type associating mapping from $C$ to $\mathcal{T}(A)$,
-    the set of types built over $A$
+  \item $\tau$ is the type-associating mapping from $C$ to
+    $\mathcal{T}(A)$, the set of types built over $A$
 \end{itemize}
@@
 \item $F$ is a mapping from $A_1$ to $\mathcal{T}(A_2)$ which links the
-  abstract-level types with the object-level types that they realized
-  as.
+  abstract-level types with the object-level types that they are
+  realized as.
 \item $F$ and $G$ are compatible, meaning that for any $c \in C_1$, we
@@
   x.\ (\synt{some}\ \synt{woman})\ (\lambda^{\circ}
-  y.\ \synt{love}\ x\ y))$ \\ Finally, the sentence (type $s$) is mapped
+  y.\ \synt{love}\ x\ y))$ \\ Finally, the sentence (type $S$) is mapped
   into a simple $\textsc{String}$, $\mathcal{L}_{syntax}(t_5) =_{\beta}

### 2013-06-26 constraints.tex
Proofreading for publication

@@
 At a basic level, \emph{aucun} functions like any other determiner and
-\emph{ne} as any verb modifier. However, we want to encode the
+\emph{ne} as any other verb modifier. However, we want to encode the
 constraint that whenever a verb is modified by \emph{ne}, the modified
@@
 being\footnote{Strictly following the EPTD for \emph{aucun} would lead
-  us to give it a type $DET\_NEG{=}F$ and then give the two types
+  us to give it a type $DET\_NEG{=}T$ and then give the two types
   $DET\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S)$ and $DET\_NEG{=}T
@@
   demonstration, we prefer to align our examples with the treatment
-  which is customary to ACG research.} $$N\_NEG{=}F \limp ((NP\_NEG{=}T
+  which is customary in ACG research.} $$N\_NEG{=}F \limp ((NP\_NEG{=}T
 \limp S) \limp S)$$ instead of the usual $$N \limp ((NP \limp S) \limp
-S)$$ The type for \emph{le} will need to be able handle cases where its
-argument noun either already contains a negative determiner or not.
+S)$$ The type for \emph{le} will need to be able to handle cases where
+its argument noun either already contains a negative determiner or not.
 
@@
 of internal information also at the level of the enclosing noun phrase
-\emph{l'odeur d'aucun tatou} in which the former noun phrase is embedded
-since this is a property of the enclosing noun phrase that is crucial to
-its syntactic combinatorics. This presupposes the existence of a
-mechanism for propagating this kind of information up the parse tree. In
-the case of our ACG, this means that all the functions which modify
-phrases must be aware whether or not their argument contains a negative
-determiner and to convey this information in its result type as
+\emph{l'odeur d'aucun tatou}, in which the former noun phrase is
+embedded, since this is a property of the enclosing noun phrase that is
+crucial to its syntactic combinatorics. This presupposes the existence
+of a mechanism for propagating this kind of information up the parse
+tree. In the case of our ACG, this means that all the functions which
+modify phrases must be aware whether or not their argument contains a
+negative determiner and to convey this information in its result type as
 well. Moreover, since the negative determiner can also come from a
@@
 case, it meant that handling negative determiners and the paired
-grammatical word for negation required us to not only define and type
+grammatical words for negation required us to not only define and type
 constants for these two categories, but to also change the types of noun
@@
 We also note that the splitting of the clause type $S$ into three finer
-types $S\_EXT={NO}$, $S\_EXT{=}ROOT$ and $S\_EXT{=}EMB$ means that types
+types $S\_EXT{=}NO$, $S\_EXT{=}ROOT$ and $S\_EXT{=}EMB$ means that types
 that work with $S$ regardless of its extraction status will have to

### 2013-06-26 gacg.tex
Proofreading for publication

@@
 constrained not only by the types of the signature they are built upon
-but also by the type system of the abstract signature.
+but also by the type system of the abstract signature and the lexicon
+connecting the two.
 
@@
 lexicons. Our definition only introduced the requisite machinery, but
-delayed the delimitation of the way languages can be defined. Leaving
+delayed the delimitation of the ways languages can be defined. Leaving
 this question open will let us answer it now by providing an extension
@@
 representations of forms and one for the semantic representations of
-meaning), which would mean having at least two different ACGs which are
+meanings), which would mean having at least two different ACGs which are
 somehow linked together (e.g. by sharing the same abstract
@@
 
-First, we can show that our interpretation is sound meaning that any
+First, we can show that our interpretation is sound, meaning that any
 arborescence adorned with signatures on the nodes, compatible lexicons
@@
 
-In this section, we then examine the key properties and implications of
+In this section, we examine the key properties and implications of
 graphical ACGs.
@@
       intersections and lexicons thanks to the induction
-      hypothesis). Therefore even $\mathcal{E}_{\mathcal{G}}(v_n)$ can
+      hypothesis). Therefore, even $\mathcal{E}_{\mathcal{G}}(v_n)$ can
       be constructed from object languages using only intersections and
@@
   Now we have to prove the converse. We do so by showing that extrinsic
-  languages contain object languages and are closed under intersection
-  and transformation by a lexicon.
+  languages contain object languages and are closed on intersection and
+  transformation by a lexicon.
 
@@
   as the set of object languages definable by ACGs if and only if the
-  set of object languages definable by ACGs is closed under
-  intersection.
+  set of object languages definable by ACGs is closed on intersection.
 \end{corollary}
@@
 
-We turn to the problem of parsing with graphical ACGs. Parsing with ACGs
-is a problem on the limits of decidability. When we consider ACGs whose
-abstract signatures only contain types of order 2 or less, parsing is
-possible in polynomial time for both linear \cite{salvati2005problemes}
-and non-linear \cite{kanazawa2007parsing} type systems (such as the one
-we introduced in~\ref{sec:acg}). If the order of the abstract signature
-exceeds 2, the membership problem for linear ACGs becomes decidable if
-and only if the satisfiability problem in the multiplicative exponential
-fragment of linear logic is decidable \cite{de2004vector}, which is an
-open problem. On the other hand, by extending the type system to include
-dependent types, parsing can become undecidable too since in such a
-system type checking itself is undecidable.
+We now turn to the problem of parsing with graphical ACGs. Parsing with
+ACGs is a problem on the limits of decidability. When we consider ACGs
+whose abstract signatures only contain types of order 2 or less, parsing
+is possible in polynomial time for both linear
+\cite{salvati2005problemes} and non-linear \cite{kanazawa2007parsing}
+type systems (such as the one we introduced in~\ref{sec:acg}). If the
+order of the abstract signature exceeds 2, the membership problem for
+linear ACGs becomes decidable if and only if the satisfiability problem
+in the multiplicative exponential fragment of linear logic is decidable
+\cite{de2004vector}, which is an open problem. On the other hand, by
+extending the type system to include dependent types, parsing can become
+undecidable too since in such a system type checking itself is
+undecidable.
 
@@
 \item The diamond grammar of Figure~\ref{fig:diamond-grammar} is also no
-  longer problematic since the labeling of nodes with terms guarantees
+  longer problematic since the labeling of nodes with terms guarantee
   that there is always a single term per node that can generate the term
@@
 
-  \item[$\mathcal{P}_{\mathcal{G}}(u) \subseteq
-    \mathcal{E}_{\mathcal{G}}(u) \subseteq
-    \mathcal{I}_{\mathcal{G}}(u)$] \hfill \\
+  \item[The three languages are increasingly constraining:
+    $\mathcal{P}_{\mathcal{G}}(u) \subseteq \mathcal{E}_{\mathcal{G}}(u)
+    \subseteq \mathcal{I}_{\mathcal{G}}(u)$] \hfill \\
 

### 2013-06-26 implementation.tex
Proofreading for publication

@@
 limitations which make it not sufficient for our needs. The foremost of
-those is that the signatures and lexicons are all explicitly realized in
+these is that the signatures and lexicons are all explicitly realized in
 (primary) memory and the toolkit expects them to be given directly in a
@@
 grammar, to reduce the time it takes to both generate the direct
-representation of the grammar and the time it takes to load it in the
-toolkit. This would have made developing the grammar already
-tenable.
+representation of the grammar and the time it takes to load it into the
+toolkit. This would have made developing the grammar already tenable.
 
@@
 with $\beta$-normalization are provided as relations which can be used
-interactively or as part of a larger program. En example of such a
-larger program might be a routine to check the validity of an ACG or to
-test some of its properties.
+interactively or as part of larger programs. En example of such a larger
+program might be a routine to check the validity of an ACG or to test
+some of its properties.
 
@@
 generating new variables and $\alpha$-converting) is a motivational
-example for introducing \emph{nominal logic programming}. To reach
-feature parity with the ACG toolkit, one would only need to implement
-the mapping of a lexicon over a term, which is a simple functor.
+example for introducing \emph{nominal logic programming}. Finally, to
+reach feature parity with the ACG toolkit, one would only need to
+implement the mapping of a lexicon over a term, which is a simple
+functor.
 
@@
 
-The third system was XMG, a language for writing metagrammars, which is
-another goal of our system, the one which distinguishes it from the
-original ACG toolkit. We believe that using a declarative programming
-system such as miniKanren with the ability to extend it using a
-practical functional programming language presents an empowering way of
-defining metagrammars. This allows the grammar designer to compose her
-grammar in any way she sees fit and to use the metaprogramming
-facilities of the underlying programming language to transcribe the
-grammar in a way that makes the most sense in her case.
+The third system at the conference was XMG, a language for writing
+metagrammars, which is another goal of our system, one which
+distinguishes it from the original ACG toolkit. We believe that using a
+declarative programming system such as miniKanren with the ability to
+extend it using a practical functional programming language presents an
+empowering way of defining metagrammars. This allows the grammar
+designer to compose her grammar in any way she sees fit and to use the
+metaprogramming facilities of the underlying programming language to
+transcribe the grammar in a way that makes the most sense in her case.
 
@@
 operations to specific data structures such as lists and cons cells and
-that it follows other similarly enlightened design decisions. Racket on
-the other hand has had more time to mature, has better metaprogramming
+that it follows other similarly enlightened design decisions. Racket, on
+the other hand, has had more time to mature, has better metaprogramming
 facilities and is more recognized in academia. In developing our system,
@@
 In~\ref{ssec:sig}, we formally presented signatures as triples
-$\mathopen{<}A, C, \tau\mathclose{>}$. The set $C$ can be easily
-reconstructed as the domain of $\tau$ and the set $A$ is just the set of
-all atomic types occurring in the range of $\tau$. Therefore, to define
-the signature $\mathopen{<}A, C, \tau\mathclose{>}$, it is enough to
-provide the function $\tau$, which in its set-theoretical realization is
-just a set of pairs. In relational programming, a specific set can be
-easily encoded as a unary relation so a unary will serve as our
-representation of a signature and defining a signature will be a matter
-of constructing this relation.
+$\mathopen{<}A, C, \tau\mathclose{>}$. However, the set $C$ can be
+always reconstructed as the domain of $\tau$ and the set $A$ is just the
+set of all the atomic types occurring in the range of $\tau$. Therefore,
+to define the signature $\mathopen{<}A, C, \tau\mathclose{>}$, it is
+enough to provide the function $\tau$, which in its set-theoretical
+realization is just a set of pairs. In relational programming, a
+specific set can be easily encoded as a unary relation so a unary
+relation will serve as our representation of a signature and defining a
+signature will be a matter of constructing this relation.
 
@@
 The system takes care of wiring up the relations and assembling the
-signature. The grammar designer's duty is to provide a 4-ary relation
-that links the four parts of a lexical constant: the wordform, the
-hypertag, the specifier and the type. This relation can be thought of as
-a multi-valued ``function'' that receives as input a wordform and a
-hypertag (that is, a single lexical entry) and outputs (possibly
-multiple) pairs of specifiers and types.
+signature that yields these structures. The grammar designer's duty is
+to provide a 4-ary relation that links the four parts of a lexical
+constant: the wordform, the hypertag, the specifier and the type. This
+relation can be thought of as a multi-valued ``function'' that receives
+as input a wordform and a hypertag (that is, a single lexical entry) and
+outputs (possibly multiple) pairs of specifiers and types.
 
@@
 manipulation of higher-order functions, we can provide helper functions
-which construct these 4-ary relations for the grammar designer in some
+which construct the signature relations for the grammar designer in some
 simple cases. Our library thus provides functions like
-\texttt{unitypedr}, which expects a type and returns the 4-ary relation
-that assigns to every lexical item a single constant whose specifier is
-$nil$ and whose type is the supplied type, or \texttt{ht->typer}, which
-expects a mapping from hypertag patterns to types and returns the 4-ary
-relation which tries to map the hypertag against the patterns and yields
-constants with the corresponding types and $nil$ specifiers. When other
-patterns emerge, the grammar designer is completely free to implement
-her new abstractions using the full power of the functional programming
-language.
+\texttt{unitypedr}, which expects a type and returns a signature that
+assigns to every lexical item a single constant whose specifier is $nil$
+and whose type is the supplied type, or \texttt{ht->typer}, which
+expects a mapping from hypertag patterns to types and returns a
+signature which tries to map the hypertags of lexical items against the
+patterns and yields constants with the corresponding types and $nil$
+specifiers. When other patterns emerge, the grammar designer is
+completely free to implement her new abstractions using the full power
+of the functional programming language.
 
@@
 connectives of the $\Sigma_{Sem}$ signature in~\ref{ssec:acg-sem}.
-These constants are not associated to any lexical entry or specifier and
-therefore their representations must be different. See
+These constants are not associated to any lexical entry and therefore
+their representations must be different. See
 Figure~\ref{fig:nonlex-const} for an example of the representation we
@@
 Signatures of non-lexical constants can be simply defined by a mapping
-from the constant names to their types as
-in~\ref{fig:nonlex-sig-impl}. Such signatures will generally not be
+from the constant names to their types as in
+Figure~\ref{fig:nonlex-sig-impl}. Such signatures will generally not be
 sufficient by themselves and will need to be complemented by other
@@
 
-With this, we have enough to fully define all the signatures of
+With this, we have enough tools to fully define all the signatures of
 section~\ref{sec:acg} in a lexicalized manner, see
@@
 
-In our relational framework, a mapping can be represented most directly
-as a binary relation and that is the representation that we use in our
-system. In the previous subsection, we talked a lot about how individual
-entries in the lexical database generate the items of a
+In the relational framework, a mapping can be represented most directly
+as a binary relation and that is the representation of lexicons that we
+use in our system. In the previous subsection, we talked a lot about how
+individual entries in the lexical database generate the items of a
 signature. Solving the problem for signatures also solved it for
@@
 constant and an object term belonging to some lexicon is generated by
-the same lexical entry that generated the abstract constant.
+the same lexical entry that generated the abstract constant and its
+type.
 
@@
 as to make the lexical entry directly accessible to the lexicon
-implementation. Since the lexicon is directly implemented by the grammar
-designer as a binary relation between constants and terms, the grammar
-designer can inspect the abstract constant and directly access the
-hypertag and wordform of the lexical entry that generated it, the
-specifier that was assigned to it and the type of the constant and then
-express the object term assigned to the constant using any combination
-of these.
+implementation. Since the lexicon is implemented by the grammar designer
+as a binary relation between constants and terms, the grammar designer
+can inspect the abstract constant and directly access the hypertag and
+wordform of the lexical entry that generated it, the specifier that was
+assigned to it and the type of the constant, and then express the object
+term assigned to the constant using any combination of these.
 
@@
 the lexicons of section~\ref{sec:acg} in a concise point-free
-style. Furthermore, the grammar designer is not limited to the convenience
-facilities we have provided and is completely free to use the same
-programming language we have used to build on our abstractions and build
-her own right within her grammar definition.
+style. Furthermore, the grammar designer is not limited to the
+convenience facilities we have provided and is completely free to use
+the same programming language we have used to build her own abstractions
+on top of ours right within her grammar definition.
 
@@
 produces the desired object term but with the object constant abstracted
-out (e.g. $\lambda^{\circ} c x y.\ x + c + y$). Then, we apply the
+out (e.g. $\lambda^{\circ} c x y.\ x + c + y$). Then, we can apply the
 \texttt{lexicalizer} function to the object signature and to this
@@
 It is also useful to be able to easily insert the non-lexical constants
-of the object signature in the object terms we are giving when defining
-a lexicon. This service is provided by the anaphoric macro
+of the object signature into the object terms we are giving when
+defining a lexicon. This service is provided by the anaphoric macro
 \texttt{with-sig-consts} which introduces all of the non-lexical
@@
 
-Finally, since lexicons are again just sets of pairs, we can take their
+Finally, since lexicons are again just relations, we can take their
 unions and intersections using disjunction (\texttt{orr}) and
-conjunction (\texttt{andr}).
+conjunction (\texttt{andr}), respectively.
 
@@
 The crucial property that we need to check for in signatures is that the
-relation between constant identities and their types is a function. This
-simply reduces to querying the system for the number of typed constants
-belonging to the signature and having some specific identity. If this
-number is one for all of our test cases, we have verified that (the
-tested subset of) the signature assigns exactly one type to every one of
-its constants.
+relation between the identities of constants and their types is a
+function. This simply reduces to querying the system for the number of
+typed constants belonging to the signature and having a specific
+identity. If this number is one for all the identities in our test set,
+we have verified that (the tested subset of) the signature assigns
+exactly one type to every one of its constants.
 

### 2013-06-26 outline.tex
Proofreading for publication

@@
 formalisms that are of interest to our work, the formalism of abstract
-categorial grammars and the formalism interaction grammars. We talk
+categorial grammars and the formalism of interaction grammars. We talk
 about a formal connection between the two and its value with respect to
@@
 implementation of the ACG machinery which will enable us to write and
-test grammars lexicalized by a lexicon.
+test grammars lexicalized by Frilex.
 
@@
 
-We conclude our findings with Chapter~\ref{chap:conclusion}, in which we
+We conclude our report with Chapter~\ref{chap:conclusion}, in which we
 summarize our findings and point out potential directions for future

### 2013-07-02 abstract.tex
Typo in abstract

@@
 \begin{abstract}
-We present work whose ultimate goal is the creation of wide-coverage
-abstract categorial grammars (ACGs) that could be used to automatically
+We present work whose ultimate goal is the creation of a wide-coverage
+abstract categorial grammar (ACG) that could be used to automatically
 build discourse-level representations. In our work, we advance towards

### 2013-07-02 gacg.tex
Typography changes near end of GACGs

@@
   \item[Extrinsic languages can be defined similarly to pangraphical
-    languages] \hfill \\
+    languages] \hfill
 
@@
 
+  \vspace{0.5cm}
+
   \item[The three languages are increasingly constraining:
     $\mathcal{P}_{\mathcal{G}}(u) \subseteq \mathcal{E}_{\mathcal{G}}(u)
-    \subseteq \mathcal{I}_{\mathcal{G}}(u)$] \hfill \\
+    \subseteq \mathcal{I}_{\mathcal{G}}(u)$] \hfill
 
@@
 
+  \vspace{0.5cm}
+
   \item[On arborescences, $\mathcal{P}_{\mathcal{G}}(u) =
-    \mathcal{E}_{\mathcal{G}}(u)$.] \hfill \\
+    \mathcal{E}_{\mathcal{G}}(u)$.] \hfill
 
@@
 
+  \vspace{0.5cm}
+
   \item[The set of extrinsic languages is included in the set of
-    pangraphical languages.] \hfill \\
+    pangraphical languages.] \hfill
 
@@
 
+  \vspace{0.5cm}
+
   \item[Is the set of pangraphical languages included in the set of
-    extrinsic languages?] \hfill \\
+    extrinsic languages?] \hfill
 

### 2013-07-02 resume.tex
Wrote the resume in both languages

\newpage

\pagenumbering{gobble}

\begin{description}[style=nextline]
\item[Résumé]

Le but final de notre travail est la création d'une grammaire
catégorielle abstraite (ACG) à large échelle qu'on pourra utiliser pour
construire les représentations au niveau de discours d'une manière
automatique.

On commence par examiner les ressources linguistiques qui existent déjà,
en particulier la grammaire d'interaction Frigram et son lexique Frilex,
et on examine leur utilité pour la construction d'une ACG à large
échelle. Ensuite, on présente une implémentation des mécanismes ACG qui
nous permet d'expérimenter avec des grammaires lexicalisées par
Frilex. Enfin, on considère les problèmes posés par le traitement des
plusieurs contraintes linguistiques dans une seule ACG et on propose une
généralisation du formalisme: les grammaires catégorielles abstraites
graphiques.

\item[Mots-clés]

Grammaires catégorielles abstraites (ACG), Génie des grammaires,
Grammaires d'interaction, Metagrammaires, Programmation relationnelle,
Formalismes grammaticaux, Grammaires formelles, Linguistique
computationnelle, Lambda-calcul, Théorie des types
\end{description}

\vfill

\begin{description}[style=nextline]
\item[Summary]

We present work whose ultimate goal is the creation of a wide-coverage
abstract categorial grammar (ACG) that could be used to automatically
build discourse-level representations.

We first examine existing language resources, in particular the Frigram
interaction grammar and its lexicon Frilex, and assess their utility to
building a wide-coverage ACG. We then present our implementation of the
ACG machinery which allows us to experiment with grammars lexicalized by
Frilex. Finally, we consider the challenge of integrating the treatment
of disparate linguistic constraints in a single ACG and propose a
generalization of the formalism: graphical abstract categorial
grammars.

\item[Keywords]

Abstract Categorial Grammars (ACG), Grammar Engineering, Interaction
Grammars, Metagrammars, Relational Programming, Grammatical Formalisms,
Formal Grammars, Computational Linguistics, Lambda Calculus, Type Theory
\end{description}

### 2013-07-03 resume.tex
Resume after review

@@
 catégorielle abstraite (ACG) à large échelle qu'on pourra utiliser pour
-construire les représentations au niveau de discours d'une manière
+construire les représentations au niveau du discours de manière
 automatique.
 
-On commence par examiner les ressources linguistiques qui existent déjà,
-en particulier la grammaire d'interaction Frigram et son lexique Frilex,
-et on examine leur utilité pour la construction d'une ACG à large
-échelle. Ensuite, on présente une implémentation des mécanismes ACG qui
-nous permet d'expérimenter avec des grammaires lexicalisées par
-Frilex. Enfin, on considère les problèmes posés par le traitement des
-plusieurs contraintes linguistiques dans une seule ACG et on propose une
-généralisation du formalisme: les grammaires catégorielles abstraites
-graphiques.
+Nous commençons par examiner les ressources linguistiques qui existent
+déjà, en particulier la grammaire d'interaction Frigram et son lexique
+Frilex. Nous examinons ensuite leur utilité pour la construction d'une
+ACG à large échelle. Puis, nous présentons une implémentation des
+mécanismes ACG qui nous permet d'expérimenter des grammaires
+lexicalisées à partir de Frilex. Enfin, nous considérons les problèmes
+posés par le traitement de plusieurs contraintes linguistiques dans une
+unique ACG et nous proposons une généralisation du formalisme : les
+grammaires catégorielles abstraites graphiques.
 
@@
 
-Grammaires catégorielles abstraites (ACG), Génie des grammaires,
+Grammaires catégorielles abstraites (ACG), Développement de grammaires,
 Grammaires d'interaction, Metagrammaires, Programmation relationnelle,

### 2013-07-08 lenls/abstract.tex
Framework for LENLS abstract
First draft of LENLS abstract

--- 2013-05-06 thesis.tex
@@
-\documentclass{article}
+\documentclass[twocolumn]{article}
 
-%% \usepackage{sdrt}
+\usepackage[utf8]{inputenc}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{xcolor}
+\usepackage{graphicx}
 \usepackage{gb4e}
 \usepackage{bussproofs}
+\usepackage{hyperref}
+\usepackage[labelformat=simple]{subcaption}
+\renewcommand\thesubfigure{(\alph{subfigure})}
+\usepackage{cancel}
+\usepackage[nottoc,numbib]{tocbibind}
+\usepackage{pdfpages}
+\usepackage{enumitem}
+\usepackage{titling}
+
+\setlength{\droptitle}{-13em}
 
@@
 
+\newtheorem*{observation}{Observation}
+\newtheorem*{theorem}{Theorem}
+\newtheorem*{corollary}{Corollary}
+
 \begin{document}
 
-\author{Ji\v{r}\'{i} Mar\v{s}\'{i}k}
-\title{Master Thesis}
+\title{Towards a Wide-Coverage Grammar: \\ Graphical Abstract Categorial Grammars}
+\author{Jiří Maršík \and Maxime Amblard}
 \maketitle
 
-\section{Preliminaries}
+\begin{abstract}
+  We present work whose ultimate goal is the creation of a wide-coverage
+  abstract categorial grammar (ACG) that could be used to automatically
+  build discourse-level representations. We focus on the challenge of
+  integrating the treatment of disparate linguistic constraints in a
+  single ACG and propose a generalization of the formalism: graphical
+  abstract categorial grammars.
+\end{abstract}
 
-\input{motivation}
+\section{Motivation}
 
-\input{acg}
+Abstract categorial grammars (ACGs) have shown to be a viable formalism
+for elegantly encoding the dynamic nature of discourse. Proposals based
+on continuation semantics \cite{de2006towards} have tackled topics such
+as event anaphora \cite{qian2011event}, SDRT discourse structure
+\cite{asher2011sdrt} and modal accessibility constraints
+\cite{asher2011montagovian}. However, all of these treatments only
+consider tiny fragments of languages. We are interested in building a
+wide-coverage grammar which integrates and reconciles the existing
+formal treatments of discourse and allows us to study their interactions
+and to build discourse representations automatically.
+
+In this paper we focus on the problem of constructing a wide-coverage
+ACG. We identify a problem that comes up when one tries to enforce
+multiple constraints/dependencies in a single ACG and we propose a
+generalization of the formalism which allows for a natural solution to
+the problem.
+
+\begin{figure*}[t]
+  \centering
+  \begin{subfigure}[b]{0.25\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{../diagrams/double-acg.pdf}
+    \caption{\label{fig:acg-comp-basic} Connecting form with meaning.}
+  \end{subfigure}
+  \qquad
+  \begin{subfigure}[b]{0.25\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{../diagrams/serial-over-parallel.pdf}
+    \caption{\label{fig:acg-comp-constr} Adding a constraint.}
+  \end{subfigure}
+  \qquad
+  \begin{subfigure}[b]{0.25\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{../diagrams/parallel-over-serial.pdf}
+    \caption{\label{fig:acg-comp-sem} Distinguishing syntactic and
+      semantic ambiguities..}
+  \end{subfigure}
+  \caption{\label{fig:acg-comp} Diagrams of systems of ACGs.}
+\end{figure*}
+
+\section{The Problem of Multiple Constraints}
+\label{sec:constraints}
+
+We will consider several linguistic constraints that have been given
+formal treatments in grammatical formalisms.
+
+In French, negation is signalled by prepending the particle \emph{ne} to
+the negated verb in conjunction with using a properly placed
+accompanying word, such as a negative determiner in one of the verb's
+arguments. This phenomenon has been elegantly formalized in the Frigram
+interaction grammar \cite{perrier2007french}.
+
+\begin{exe}
+  \ex \label{ex:aucun-shallow} Jean \textbf{ne} parle à \textbf{aucun} collègue. \\
+      (Jean speaks to no colleague.)
+  \ex \label{ex:aucun-deep-obj} Jean \textbf{ne} parle à la femme d'\textbf{aucun} collègue. \\
+      (Jean speaks to the wife of no colleague.)
+  \ex \label{ex:aucun-deep-subj} \textbf{Aucun} collègue de Jean \textbf{ne} parle à sa femme. \\
+      (No colleague of Jean's speaks to his wife.)
+\end{exe}
+
+We see here that the negative determiner \emph{aucun} can be present in
+the subject or the object of the negated verb and it can modify the
+argument directly or be attached to one of its complements. Furthermore,
+we note that omitting either the word \emph{ne} or the word \emph{aucun}
+while keeping the other produces a sentence which is considered
+ungrammatical.
+
+This difference in syntactic behavior between noun phrases that contain
+a negative determiner and those that do not has implications for our
+grammar. Since two terms having an identical type in an ACG signature
+can be freely interchanged in any sentence, we are forced to assign two
+different types to these two different kinds of noun phrases.
+
+% NOTE: We could use a single constant for the two paired words.
+% NOTE: We could use a higher-order type for the negative determiner
+% which demands the negative particle itself.
+
+This leads us to a grammar in which we subdivide the atomic types $N$
+and $NP$ into subtypes that reflect whether or not they contain a
+negative determiner inside. Types of the other constants, such as the
+preposition \emph{de} seen in (\ref{ex:aucun-deep-obj}) and
+(\ref{ex:aucun-deep-subj}), will constrain their arguments to have
+compatible features on their types and will propagate the information
+carried in the features to its result type, e.g.:
+
+\begin{align*}
+N_{de_1} &: NP\_NEG{=}F \limp N\_NEG{=}F \limp N\_NEG{=}F \\
+N_{de_2} &: NP\_NEG{=}F \limp N\_NEG{=}T \limp N\_NEG{=}T \\
+N_{de_3} &: NP\_NEG{=}T \limp N\_NEG{=}F \limp N\_NEG{=}T
+\end{align*}
+
+Enforcing other constraints leads us to subdivide our ``atomic'' types
+even further (e.g. the authors of \cite{pogodalla2012controlling} add
+features to the $S$ and $NP$ types to implement constraints about
+extraction). Other phenomena, such as agreement on morphological
+properties like number, gender, person or case, intuitively lead us to
+make our types even more specific.
+
+If we were to use this approach to write a grammar that enforces
+multiple constraints at the same time, we would end up with complicated
+types like the one below which provide complete specifications of the
+various possible situations.
+
+\begin{align*}
+C_{de_{11}} :\ &(NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \\
+&\limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \\
+&\limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
+\end{align*}
+
+This creates two problems. Firstly, the number of such types grows
+exponentially with the number of features added. This can be fixed by
+introducing dependent types into the type system as in
+\cite{de2007type}. Secondly, and perhaps more importantly, we have lost
+the clarity and simplicity of describing the individual phenomena
+independently. This makes adding new constraints more and more
+cumbersome and reduces our capability to reason about our grammar. It no
+longer becomes apparent which language mechanisms behave independently
+and their interactions are hard to track. We firmly believe that
+simplicity is a fundamental requirement for constructing a large and
+robust grammar and our proposal aims to reclaim that simplicity.
+
+\section{Systems of ACGs}
+\label{sec:acg-comp}
+
+An ACG is defined in terms of two higher-order signatures, sets of typed
+constants on which we can form well-typed lambda terms, and a
+translating function from terms of one of the signatures to terms of the
+other (the former signature being called the \emph{abstract} signature,
+the latter being called the \emph{object} signature and the function
+being called a \emph{lexicon}). We are then interested in the terms of
+the abstract signature having some distinguished type and their images
+through the lexicon (these two sets of terms are called the
+\emph{abstract language} and the \emph{object language}, respectively).
+
+When describing language, we are usually interested in systems of more
+than one ACG and we use diagrams such as the ones on
+Figure~\ref{fig:acg-comp} to explain them.
+
+In Figure~\ref{fig:acg-comp-basic}, we have abstract syntactic terms
+from the signature $\Sigma_{Synt}$ yielding the phonological strings and
+semantic representations using two different lexicons. In
+Figure~\ref{fig:acg-comp-constr}, we extend this pattern by adding a new
+abstract signature which narrows the language of syntactic terms and
+implements some linguistic constraint. Finally, in
+Figure~\ref{fig:acg-comp-sem}, we split the abstract signature
+$\Sigma_{Synt}$ of Figure~\ref{fig:acg-comp-basic} into two signatures,
+$\Sigma_{Syntax}$ and $\Sigma_{SyntSem}$, in order to model syntactic
+and semantic ambiguities in different grammars.
+
+In our grammar, we would like to combine both of the patterns on
+Figures~\ref{fig:acg-comp-constr} and \ref{fig:acg-comp-sem} and
+furthermore, to use the pattern of Figure~\ref{fig:acg-comp-constr}
+multiple times to implement multiple constraints independently. However,
+these combinations are not possible in the framework of ACGs without
+introducing complexity such as by contaminating the syntax-semantics
+interface by the implementation details of the syntactic layer.
+
+We would like to have a system which would be characterized by a diagram
+like the one on Figure~\ref{fig:gacg}. However, ACG diagrams are limited
+to arborescences and we are obliged to generalize the formalism in order
+to get the expected interpretation of Figure~\ref{fig:gacg}.
+
+\section{Graphical ACGs}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.25\textwidth,height=0.2\textheight]{../diagrams/final.pdf}
+  \caption{{\label{fig:gacg} A graphical ACG that implements two
+      independent syntactic constraints and distinguishes syntactic and
+      semantic ambiguities.}}
+\end{figure}
+
+We define a \emph{graphical abstract categorial grammar} as a directed
+acyclic graph (DAG) whose nodes are labeled with signatures (and
+distinguished types) and whose edges are labeled with lexicons, a
+mathematical reification of an ACG diagram generalized from
+arborescences to DAGs. We then search for an appropriate semantics for
+these structures, i.e. how to determine what languages are defined by a
+graphical ACG.
+
+We first follow the paradigm in which we associate languages with the
+nodes of the graph while the edges tell us how these languages are
+defined in terms of each other. We argue that intersection of languages
+is the suitable meaning of more than two edges arriving at the same node
+based both on the simplicity of the resulting definitions and on our
+expectations about the desired semantics. This leads us to the following
+definitions of \emph{intrinsic} and \emph{extrinsic} languages
+associated with some node $v$ in a graphical ACG $\mathcal{G}$:
+
+$$
+\mathcal{I}_{\mathcal{G}}(v) = \{t \in \Lambda(\Sigma_v)
+\mid\ \vdash_{\Sigma_v} t : S_v\}
+$$
+$$
+\mathcal{E}_{\mathcal{G}}(v) = \mathcal{I}_{\mathcal{G}}(v) \cap
+\bigcap_{(u,v) \in E} \mathcal{L}_{(u,v)}(\mathcal{E}_{\mathcal{G}}(u))
+$$
+
+The intrinsic language is just the set of terms built on the node's
+signature and having the node's distinguished type. The extrinsic
+language is established by taking the extrinsic languages of its
+predecessors, mapping them through a lexicon and taking their
+intersection, or just taking the node's intrinsic language if it has no
+predecessors.
+
+We examine the relationship between the languages defined by ACGs and
+graphical ACGs (G-ACGs). Intrinsic languages correspond exactly to
+abstract languages and therefore the sets of languages definable by both
+are equal.
+
+$$
+\mathcal{I} = \mathcal{A}
+$$
+
+G-ACG extrinsic languages correspond to ACG object languages with
+intersection. More formally, we can say that ACG object languages are
+just ACG abstract languages closed on transformation by a lexicon and we
+show that G-ACG extrinsic languages are just ACG abstract languages
+closed on transformation by a lexicon and intersection.
+
+$$
+\mathcal{O} = \mathcal{A}^{\mathcal{L}}
+$$
+$$
+\mathcal{E} = \mathcal{A}^{\mathcal{L}{\cap}}
+$$
+
+We then argue that closure on intersection is an intuitive property of a
+grammatical formalism that allows independent combination of linguistic
+constraints. Furthermore, from the above we can see that object
+languages are as expressive as extrinsic languages iff object languages
+are closed on intersection, which is conjectured to be false.
+
+However, extrinsic languages turn out to give counter-intuitive
+interpretations to some graphical ACGs and so after examining the
+algebraic paradigm which gave us the intrinsic and extrinsic languages,
+we consider an alternative paradigm, which we call generative.
+
+In this paradigm, we identify the nodes of the graph with terms and the
+edges with statements that one term is mapped into another using a
+lexicon. This leads us to the definition of the \emph{pangraphical}
+language of a node $u$ in a G-ACG $\mathcal{G}$.
+
+A term $t$ belongs to $\mathcal{P}_{\mathcal{G}}(u)$ whenever there
+exists a labeling $T$ of the nodes of the graph such that:
+
+\begin{itemize}
+  \item $T_u = t$.
+  \item For all $v \in V(G)$, $\vdash_{\Sigma_v} T_v : S_v$.
+  \item For all $(v,w) \in E(G)$, $\mathcal{L}_{(v,w)}(T_v) = T_w$.
+\end{itemize}
+
+The two paradigms under which we can interpret ACGs are equivalent in
+the context of plain ACGs and arborescent diagrams and the two metaphors
+(nodes as languages and nodes as terms) can be used interchangeably. We
+back this up formally by showing that for every node $u$ in every
+arborescent G-ACG $\mathcal{G}$, we have $\mathcal{E}_{\mathcal{G}}(u) =
+\mathcal{P}_{\mathcal{G}}(u)$. However, as we start to consider
+non-arborescent graphs, we find, interestingly, that the two paradigms
+diverge.
+
+The newly defined pangraphical languages solve the problem of extrinsic
+languages giving us counter-intuitive interpretations for some specific
+G-ACGs. We show by construction that pangraphical languages are as
+expressive as extrinsic languages and so the kinds of languages we have
+defined here are increasingly more expressive, $\mathcal{I} \subseteq
+\mathcal{E} \subseteq \mathcal{P}$. We also show that
+$\mathcal{I}_{\mathcal{G}}(u) \supseteq \mathcal{E}_{\mathcal{G}}(u)
+\supseteq \mathcal{P}_{\mathcal{G}}(u)$, meaning that the language
+definitions are more and more constraining and that the extrinsic
+languages went in the right direction as pangraphical languages, but not
+as far.
+
+Finally, we construct a G-ACG which integrates the French negation
+constraint discussed in Section~\ref{sec:constraints}, the constraints
+on extraction introduced in \cite{pogodalla2012controlling} and a
+constraint handling gender agreement in a single grammar
+specification. The negation constraint is implemented exactly as was
+described in Section~\ref{sec:constraints} and the extraction
+constraints are taken as is from \cite{pogodalla2012controlling}. The
+underlying syntactic signature is defined without any concern for the
+additional constraints and the syntax-semantics signature is kept simple
+as well.
+
+In the end, this lets us define the syntax in a clean way using the
+idiomatic style of categorial grammars (simple atomic types like $N$,
+$NP$ and $S$) and then define the constraints themselves the same way
+they are defined in papers which try to formalize them individually
+(such as the case with \cite{pogodalla2012controlling}). We can freely
+combine both the constraint pattern and the syntax-semantics ambiguities
+pattern from Section~\ref{sec:acg-comp}.
 
 \bibliographystyle{plain}
-\bibliography{biblio}
+\bibliography{../biblio}
 

### 2013-07-13 lenls/abstract.tex
LENLS: Changed the title
First rereading of the LENLS abstract

@@
 
-\title{Towards a Wide-Coverage Grammar: \\ Graphical Abstract Categorial Grammars}
+\title{An Integration of Multiple Constraints in ACG}
 \author{Jiří Maršík \and Maxime Amblard}
@@
 translating function from terms of one of the signatures to terms of the
-other (the former signature being called the \emph{abstract} signature,
-the latter being called the \emph{object} signature and the function
+other (the former signature being called the \emph{abstract signature},
+the latter being called the \emph{object signature} and the function
 being called a \emph{lexicon}). We are then interested in the terms of
@@
 
-When describing language, we are usually interested in systems of more
-than one ACG and we use diagrams such as the ones on
-Figure~\ref{fig:acg-comp} to explain them.
+When describing languages, we are usually interested in systems of more
+than one ACG and diagrams such as the ones on Figure~\ref{fig:acg-comp}
+have been used to explain them.
 
-In Figure~\ref{fig:acg-comp-basic}, we have abstract syntactic terms
-from the signature $\Sigma_{Synt}$ yielding the phonological strings and
-semantic representations using two different lexicons. In
-Figure~\ref{fig:acg-comp-constr}, we extend this pattern by adding a new
-abstract signature which narrows the language of syntactic terms and
-implements some linguistic constraint. Finally, in
-Figure~\ref{fig:acg-comp-sem}, we split the abstract signature
-$\Sigma_{Synt}$ of Figure~\ref{fig:acg-comp-basic} into two signatures,
+In Figure~\ref{fig:acg-comp-basic}, abstract syntactic terms from the
+signature $\Sigma_{Synt}$ yield phonological strings and semantic
+representations using two different lexicons. In
+Figure~\ref{fig:acg-comp-constr}, this pattern is extended by adding a
+new abstract signature which narrows the language of syntactic terms and
+implements some linguistic constraint
+\cite{pogodalla2012controlling}. Finally, in
+Figure~\ref{fig:acg-comp-sem}, the abstract signature $\Sigma_{Synt}$ of
+Figure~\ref{fig:acg-comp-basic} is split into two signatures,
 $\Sigma_{Syntax}$ and $\Sigma_{SyntSem}$, in order to model syntactic
-and semantic ambiguities in different grammars.
+and semantic ambiguities in different grammars
+\cite{pogodalla2007generalizing}.
 
@@
 these combinations are not possible in the framework of ACGs without
-introducing complexity such as by contaminating the syntax-semantics
+introducing complexity such as contaminating the syntax-semantics
 interface by the implementation details of the syntactic layer.
@@
 acyclic graph (DAG) whose nodes are labeled with signatures (and
-distinguished types) and whose edges are labeled with lexicons, a
-mathematical reification of an ACG diagram generalized from
-arborescences to DAGs. We then search for an appropriate semantics for
-these structures, i.e. how to determine what languages are defined by a
-graphical ACG.
+distinguished types) and whose edges are labeled with lexicons, in other
+words just a mathematical reification of an ACG diagram that has also
+been generalized from arborescences to DAGs. We then search for an
+appropriate semantics for these structures, i.e. how to determine what
+languages are defined by a graphical ACG.
 
-We first follow the paradigm in which we associate languages with the
-nodes of the graph while the edges tell us how these languages are
-defined in terms of each other. We argue that intersection of languages
-is the suitable meaning of more than two edges arriving at the same node
-based both on the simplicity of the resulting definitions and on our
+We first follow a paradigm in which nodes of the diagrams are
+interpreted as languages with the edges telling us how these languages
+are defined in terms of each other. A single arrow leading to a language
+means that the target language is produced from the source by mapping it
+through a lexicon. We then argue that the suitable meaning of two or
+more edges arriving at the same node is intersection of languages based
+both on the simplicity of the resulting definitions and on our
 expectations about the desired semantics. This leads us to the following
@@
 language is established by taking the extrinsic languages of its
-predecessors, mapping them through a lexicon and taking their
+predecessors, mapping them through lexicons and taking their
 intersection, or just taking the node's intrinsic language if it has no
@@
 
-We examine the relationship between the languages defined by ACGs and
-graphical ACGs (G-ACGs). Intrinsic languages correspond exactly to
+We then examine the relationship between the languages defined by ACGs
+and graphical ACGs (G-ACGs). Intrinsic languages correspond exactly to
 abstract languages and therefore the sets of languages definable by both
@@
 G-ACG extrinsic languages correspond to ACG object languages with
-intersection. More formally, we can say that ACG object languages are
-just ACG abstract languages closed on transformation by a lexicon and we
-show that G-ACG extrinsic languages are just ACG abstract languages
-closed on transformation by a lexicon and intersection.
+intersection. More formally, it can be said that ACG object languages
+are just ACG abstract languages closed on transformation by a lexicon
+and we then show that G-ACG extrinsic languages are just ACG abstract
+languages closed on transformation by a lexicon and intersection.
 
@@
 However, extrinsic languages turn out to give counter-intuitive
-interpretations to some graphical ACGs and so after examining the
-algebraic paradigm which gave us the intrinsic and extrinsic languages,
-we consider an alternative paradigm, which we call generative.
+interpretations to some graphical ACGs and we turn to another paradigm.
 
-In this paradigm, we identify the nodes of the graph with terms and the
-edges with statements that one term is mapped into another using a
+In the new paradigm, we interpret the nodes of the graph as terms and
+the edges with statements that one term is mapped into another using a
 lexicon. This leads us to the definition of the \emph{pangraphical}
@@
 The two paradigms under which we can interpret ACGs are equivalent in
-the context of plain ACGs and arborescent diagrams and the two metaphors
-(nodes as languages and nodes as terms) can be used interchangeably. We
-back this up formally by showing that for every node $u$ in every
-arborescent G-ACG $\mathcal{G}$, we have $\mathcal{E}_{\mathcal{G}}(u) =
-\mathcal{P}_{\mathcal{G}}(u)$. However, as we start to consider
-non-arborescent graphs, we find, interestingly, that the two paradigms
-diverge.
+the context of plain ACGs whose diagrams are arborescent and the two
+metaphors (nodes as languages and nodes as terms) can be used
+interchangeably. We back this up formally by showing that for every node
+$u$ in every arborescent G-ACG $\mathcal{G}$, we have
+$\mathcal{E}_{\mathcal{G}}(u) = \mathcal{P}_{\mathcal{G}}(u)$. However,
+as we start to consider non-arborescent graphs, we find, interestingly,
+that the two paradigms diverge (i.e. $\exists \mathcal{G},
+u.\ \mathcal{E}_{\mathcal{G}}(u) \neq \mathcal{P}_{\mathcal{G}}(u)$).
 
@@
 definitions are more and more constraining and that the extrinsic
-languages went in the right direction as pangraphical languages, but not
-as far.
+languages went in the same direction as pangraphical languages when
+narrowing the elements of $\mathcal{I}_{\mathcal{G}}(u)$, but not as
+far.
 
@@
 on extraction introduced in \cite{pogodalla2012controlling} and a
-constraint handling gender agreement in a single grammar
-specification. The negation constraint is implemented exactly as was
-described in Section~\ref{sec:constraints} and the extraction
-constraints are taken as is from \cite{pogodalla2012controlling}. The
-underlying syntactic signature is defined without any concern for the
-additional constraints and the syntax-semantics signature is kept simple
-as well.
+constraint handling agreement in a single grammar specification. The
+negation constraint is implemented exactly as was described in
+Section~\ref{sec:constraints} and the extraction constraints are taken
+``as is'' from \cite{pogodalla2012controlling}. The underlying syntactic
+signature is defined without any concern for the additional constraints
+and the syntax-semantics signature is kept simple as well.
 
@@
 idiomatic style of categorial grammars (simple atomic types like $N$,
-$NP$ and $S$) and then define the constraints themselves the same way
+$NP$ and $S$) and then define the constraints themselves the same way as
 they are defined in papers which try to formalize them individually
-(such as the case with \cite{pogodalla2012controlling}). We can freely
-combine both the constraint pattern and the syntax-semantics ambiguities
-pattern from Section~\ref{sec:acg-comp}.
+(such as the case with \cite{pogodalla2012controlling}). We can also
+freely combine both the constraint pattern and the syntax-semantics
+ambiguities pattern from Section~\ref{sec:acg-comp}.
 

### 2013-07-16 lenls/abstract.tex
Swapped Systems of ACGs with Multiple Ling Cons

@@
     \caption{\label{fig:acg-comp-sem} Distinguishing syntactic and
-      semantic ambiguities..}
+      semantic ambiguities.}
   \end{subfigure}
@@
 \end{figure*}
-
-\section{The Problem of Multiple Constraints}
-\label{sec:constraints}
-
-We will consider several linguistic constraints that have been given
-formal treatments in grammatical formalisms.
-
-In French, negation is signalled by prepending the particle \emph{ne} to
-the negated verb in conjunction with using a properly placed
-accompanying word, such as a negative determiner in one of the verb's
-arguments. This phenomenon has been elegantly formalized in the Frigram
-interaction grammar \cite{perrier2007french}.
-
-\begin{exe}
-  \ex \label{ex:aucun-shallow} Jean \textbf{ne} parle à \textbf{aucun} collègue. \\
-      (Jean speaks to no colleague.)
-  \ex \label{ex:aucun-deep-obj} Jean \textbf{ne} parle à la femme d'\textbf{aucun} collègue. \\
-      (Jean speaks to the wife of no colleague.)
-  \ex \label{ex:aucun-deep-subj} \textbf{Aucun} collègue de Jean \textbf{ne} parle à sa femme. \\
-      (No colleague of Jean's speaks to his wife.)
-\end{exe}
-
-We see here that the negative determiner \emph{aucun} can be present in
-the subject or the object of the negated verb and it can modify the
-argument directly or be attached to one of its complements. Furthermore,
-we note that omitting either the word \emph{ne} or the word \emph{aucun}
-while keeping the other produces a sentence which is considered
-ungrammatical.
-
-This difference in syntactic behavior between noun phrases that contain
-a negative determiner and those that do not has implications for our
-grammar. Since two terms having an identical type in an ACG signature
-can be freely interchanged in any sentence, we are forced to assign two
-different types to these two different kinds of noun phrases.
-
-% NOTE: We could use a single constant for the two paired words.
-% NOTE: We could use a higher-order type for the negative determiner
-% which demands the negative particle itself.
-
-This leads us to a grammar in which we subdivide the atomic types $N$
-and $NP$ into subtypes that reflect whether or not they contain a
-negative determiner inside. Types of the other constants, such as the
-preposition \emph{de} seen in (\ref{ex:aucun-deep-obj}) and
-(\ref{ex:aucun-deep-subj}), will constrain their arguments to have
-compatible features on their types and will propagate the information
-carried in the features to its result type, e.g.:
-
-\begin{align*}
-N_{de_1} &: NP\_NEG{=}F \limp N\_NEG{=}F \limp N\_NEG{=}F \\
-N_{de_2} &: NP\_NEG{=}F \limp N\_NEG{=}T \limp N\_NEG{=}T \\
-N_{de_3} &: NP\_NEG{=}T \limp N\_NEG{=}F \limp N\_NEG{=}T
-\end{align*}
-
-Enforcing other constraints leads us to subdivide our ``atomic'' types
-even further (e.g. the authors of \cite{pogodalla2012controlling} add
-features to the $S$ and $NP$ types to implement constraints about
-extraction). Other phenomena, such as agreement on morphological
-properties like number, gender, person or case, intuitively lead us to
-make our types even more specific.
-
-If we were to use this approach to write a grammar that enforces
-multiple constraints at the same time, we would end up with complicated
-types like the one below which provide complete specifications of the
-various possible situations.
-
-\begin{align*}
-C_{de_{11}} :\ &(NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \\
-&\limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \\
-&\limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
-\end{align*}
-
-This creates two problems. Firstly, the number of such types grows
-exponentially with the number of features added. This can be fixed by
-introducing dependent types into the type system as in
-\cite{de2007type}. Secondly, and perhaps more importantly, we have lost
-the clarity and simplicity of describing the individual phenomena
-independently. This makes adding new constraints more and more
-cumbersome and reduces our capability to reason about our grammar. It no
-longer becomes apparent which language mechanisms behave independently
-and their interactions are hard to track. We firmly believe that
-simplicity is a fundamental requirement for constructing a large and
-robust grammar and our proposal aims to reclaim that simplicity.
 
@@
 
-In our grammar, we would like to combine both of the patterns on
-Figures~\ref{fig:acg-comp-constr} and \ref{fig:acg-comp-sem} and
-furthermore, to use the pattern of Figure~\ref{fig:acg-comp-constr}
-multiple times to implement multiple constraints independently. However,
-these combinations are not possible in the framework of ACGs without
-introducing complexity such as contaminating the syntax-semantics
-interface by the implementation details of the syntactic layer.
+\section{The Problem of Multiple Constraints}
+\label{sec:constraints}
 
-We would like to have a system which would be characterized by a diagram
-like the one on Figure~\ref{fig:gacg}. However, ACG diagrams are limited
-to arborescences and we are obliged to generalize the formalism in order
-to get the expected interpretation of Figure~\ref{fig:gacg}.
+We will consider several linguistic constraints that have been given
+formal treatments in grammatical formalisms.
+
+In French, negation is signalled by prepending the particle \emph{ne} to
+the negated verb in conjunction with using a properly placed
+accompanying word, such as a negative determiner in one of the verb's
+arguments. This phenomenon has been elegantly formalized in the Frigram
+interaction grammar \cite{perrier2007french}.
+
+\begin{exe}
+  \ex \label{ex:aucun-shallow} Jean \textbf{ne} parle à \textbf{aucun} collègue. \\
+      (Jean speaks to no colleague.)
+  \ex \label{ex:aucun-deep-obj} Jean \textbf{ne} parle à la femme d'\textbf{aucun} collègue. \\
+      (Jean speaks to the wife of no colleague.)
+  \ex \label{ex:aucun-deep-subj} \textbf{Aucun} collègue de Jean \textbf{ne} parle à sa femme. \\
+      (No colleague of Jean's speaks to his wife.)
+\end{exe}
+
+We see here that the negative determiner \emph{aucun} can be present in
+the subject or the object of the negated verb and it can modify the
+argument directly or be attached to one of its complements. Furthermore,
+we note that omitting either the word \emph{ne} or the word \emph{aucun}
+while keeping the other produces a sentence which is considered
+ungrammatical.
+
+This difference in syntactic behavior between noun phrases that contain a
+negative determiner and those that do not has implications for our
+grammar. Since two terms that have an identical type in an ACG signature can
+be freely interchanged in any sentence, we are forced to assign two different
+types to these two different kinds of noun phrases.
+
+% NOTE: We could use a single constant for the two paired words.
+% NOTE: We could use a higher-order type for the negative determiner
+% which demands the negative particle itself.
+
+This leads us to a grammar in which we subdivide the atomic types $N$
+and $NP$ into subtypes that reflect whether or not they contain a
+negative determiner inside. Types of the other constants, such as the
+preposition \emph{de} seen in (\ref{ex:aucun-deep-obj}) and
+(\ref{ex:aucun-deep-subj}), will constrain their arguments to have
+compatible features on their types and will propagate the information
+carried in the features to its result type, e.g.:
+
+\begin{align*}
+N_{de_1} &: NP\_NEG{=}F \limp N\_NEG{=}F \limp N\_NEG{=}F \\
+N_{de_2} &: NP\_NEG{=}F \limp N\_NEG{=}T \limp N\_NEG{=}T \\
+N_{de_3} &: NP\_NEG{=}T \limp N\_NEG{=}F \limp N\_NEG{=}T
+\end{align*}
+
+Enforcing other constraints leads us to subdivide our ``atomic'' types
+even further (e.g. the authors of \cite{pogodalla2012controlling} add
+features to the $S$ and $NP$ types to implement constraints about
+extraction). Other phenomena, such as agreement on morphological
+properties like number, gender, person or case, intuitively lead us to
+make our types even more specific.
+
+If we were to use this approach to write a grammar that enforces multiple
+constraints at the same time, we would end up with complicated types, like the
+one below, which provide complete specifications of the various possible
+situations.
+
+\begin{align*}
+C_{de_{11}} :\ &(NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \\
+&\limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \\
+&\limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
+\end{align*}
+
+This creates two problems. Firstly, the number of such types grows
+exponentially with the number of features added. This can be fixed by
+introducing dependent types into the type system as in
+\cite{de2007type}. However, secondly, and perhaps more importantly, we have
+lost the clarity and simplicity of describing the individual phenomena
+independently. This makes adding new constraints more and more cumbersome and
+reduces our capability to reason about our grammar. It no longer becomes
+apparent which language mechanisms behave independently and their interactions
+are hard to track. We firmly believe that simplicity is a fundamental
+requirement for constructing a large and robust grammar and our proposal aims
+to reclaim that simplicity.
+
+In our grammar, we would like to combine several constraints
+(Figure~\ref{fig:acg-comp-constr}) and possibly to also separate the syntactic
+ambiguities from the purely semantic ones (Figure~\ref{fig:acg-comp-sem}).
+However, these combinations are not possible in the framework of ACGs without
+introducing complexity by solving all the constraints in one single type
+signature (the consequences of we saw above) or contaminating the
+syntax-semantics interface by the implementation details of the syntactic
+layer.
+
+We would like to have a system which would be characterized by a diagram like
+the one on Figure~\ref{fig:gacg}. However, ACG diagrams are limited to
+arborescences and we are obliged to generalize the formalism in order to get
+the expected interpretation of Figure~\ref{fig:gacg}.
 

### 2013-07-16 lenls/abstract.tex
First draft of a Conclusion

@@
 
+\section{Conclusion}
+
+We have considered the problem of building a wide-coverage ACG, specifically
+the problem of expressing a multitude of linguistic constraints. We have
+examined previous techniques and found no satisfying solution. We have thus
+provided an extension of the ACG formalism to solve the problem and justified
+the need for the increased expressivity.
+
 \bibliographystyle{plain}

### 2013-07-16 lenls/abstract.tex
Some small refactoring...

@@
 G-ACG extrinsic languages correspond to ACG object languages with
-intersection. More formally, it can be said that ACG object languages
-are just ACG abstract languages closed on transformation by a lexicon
-and we then show that G-ACG extrinsic languages are just ACG abstract
-languages closed on transformation by a lexicon and intersection.
+intersection. More formally, it can be said that ACG object languages are just
+ACG abstract languages closed on transformation by a lexicon, and we then show
+that G-ACG extrinsic languages are just ACG abstract languages closed on
+transformation by a lexicon and intersection.
 
@@
 
-However, extrinsic languages turn out to give counter-intuitive
-interpretations to some graphical ACGs and we turn to another paradigm.
+However, the extrinsic languages we have defined above turn out to give
+counter-intuitive interpretations to some graphical ACGs and we revisit the
+question in an another paradigm.
 
-In the new paradigm, we interpret the nodes of the graph as terms and
-the edges with statements that one term is mapped into another using a
-lexicon. This leads us to the definition of the \emph{pangraphical}
-language of a node $u$ in a G-ACG $\mathcal{G}$.
+In the new paradigm, we interpret the nodes of the graph as terms and the
+edges as statements that one term is mapped into another using a lexicon. This
+leads us to the definition of the \emph{pangraphical} language of a node $u$
+in a G-ACG $\mathcal{G}$.
 
@@
 signature is defined without any concern for the additional constraints
-and the syntax-semantics signature is kept simple as well.
+and the syntax-semantics interface is kept simple as well.
 

### 2013-07-17 lenls/abstract.tex
LENLS: More feedback and back to 4p

@@
 
-\setlength{\droptitle}{-13em}
+\setlength{\droptitle}{-14em}
+\title{Integration of Multiple Constraints in ACG}
+\date{\vspace{-5em}}
 
@@
 
-\title{An Integration of Multiple Constraints in ACG}
-\author{Jiří Maršík \and Maxime Amblard}
 \maketitle
 
-\begin{abstract}
-  We present work whose ultimate goal is the creation of a wide-coverage
-  abstract categorial grammar (ACG) that could be used to automatically
-  build discourse-level representations. We focus on the challenge of
-  integrating the treatment of disparate linguistic constraints in a
-  single ACG and propose a generalization of the formalism: graphical
-  abstract categorial grammars.
-\end{abstract}
+This proposal is a first step towards a wide-coverage Abstract Categorial
+Grammar (ACG) that could be used to automatically build discourse-level
+representations. We focus on the challenge of integrating the treatment of
+disparate linguistic constraints in a single ACG and propose a generalization
+of the formalism: Graphical Abstract Categorial Grammars.
 
@@
 
-Abstract categorial grammars (ACGs) have shown to be a viable formalism
-for elegantly encoding the dynamic nature of discourse. Proposals based
-on continuation semantics \cite{de2006towards} have tackled topics such
-as event anaphora \cite{qian2011event}, SDRT discourse structure
-\cite{asher2011sdrt} and modal accessibility constraints
-\cite{asher2011montagovian}. However, all of these treatments only
-consider tiny fragments of languages. We are interested in building a
-wide-coverage grammar which integrates and reconciles the existing
-formal treatments of discourse and allows us to study their interactions
-and to build discourse representations automatically.
+Abstract Categorial Grammars (ACGs) have shown to be a viable formalism for
+elegantly encoding the dynamic nature of discourse. Proposals based on
+continuation semantics \cite{de2006towards} have tackled topics such as event
+anaphora \cite{qian2011event}, SDRT discourse structure \cite{asher2011sdrt}
+and modal accessibility constraints \cite{asher2011montagovian}. However, all
+of these treatments only consider tiny fragments of languages. We are
+interested in building a wide-coverage grammar which integrates and reconciles
+the existing formal treatments of discourse and allows us to study their
+interactions and to build discourse representations automatically.
 
-In this paper we focus on the problem of constructing a wide-coverage
-ACG. We identify a problem that comes up when one tries to enforce
-multiple constraints/dependencies in a single ACG and we propose a
-generalization of the formalism which allows for a natural solution to
-the problem.
+In this paper we focus on the problem of constructing a wide-coverage ACG. We
+identify a problem that comes up when one tries to enforce multiple
+constraints/dependencies in a single ACG and we propose a generalization of
+the formalism which resolves to the issue.
 
@@
 introducing dependent types into the type system as in
-\cite{de2007type}. However, secondly, and perhaps more importantly, we have
-lost the clarity and simplicity of describing the individual phenomena
-independently. This makes adding new constraints more and more cumbersome and
-reduces our capability to reason about our grammar. It no longer becomes
-apparent which language mechanisms behave independently and their interactions
-are hard to track. We firmly believe that simplicity is a fundamental
-requirement for constructing a large and robust grammar and our proposal aims
-to reclaim that simplicity.
+\cite{de2007type}. However, while this allows us to abstract over the
+combinations of feature values and write our grammar down concisely, it does
+not take away the complexity. The treatments of the various linguistic
+phenomena are all expressed in the same types making it hard to see whether
+they are independent or not. Since the treatments cannot be considered in
+isolation, reasoning about the entire grammar becomes difficult and so does
+enhancing it with more constraints. This is a fatal problem for a grammar
+which strives to cover a wide range of linguistic facts. We firmly believe
+that simplicity is a fundamental requirement for constructing a large and
+robust grammar and our proposal aims to reclaim that simplicity.
 
@@
 ambiguities from the purely semantic ones (Figure~\ref{fig:acg-comp-sem}).
-However, these combinations are not possible in the framework of ACGs without
-introducing complexity by solving all the constraints in one single type
-signature (the consequences of we saw above) or contaminating the
-syntax-semantics interface by the implementation details of the syntactic
-layer.
+However, trying to mix these strategies in the ACG framework forces us to
+solve all the constraints in a single type signature or contaminate the
+syntax-semantics interface with the implementation details of the syntactic
+layer, both of which introduce incidental complexity we want to avoid.
 
 We would like to have a system which would be characterized by a diagram like
-the one on Figure~\ref{fig:gacg}. However, ACG diagrams are limited to
-arborescences and we are obliged to generalize the formalism in order to get
+the one on Figure~\ref{fig:gacg} where the constraint signatures delimit the
+legal syntactic structures independently of each other and without interfering
+with the syntax-semantics interface. However, ACG diagrams are limited to
+arborescences and we are obliged to generalize them in order to get
 the expected interpretation of Figure~\ref{fig:gacg}.
@@
   \includegraphics[width=0.25\textwidth,height=0.2\textheight]{../diagrams/final.pdf}
-  \caption{{\label{fig:gacg} A graphical ACG that implements two
+  \caption{\label{fig:gacg} A graphical ACG that implements two
       independent syntactic constraints and distinguishes syntactic and
-      semantic ambiguities.}}
+      semantic ambiguities.}
 \end{figure}
 
-We define a \emph{graphical abstract categorial grammar} as a directed
-acyclic graph (DAG) whose nodes are labeled with signatures (and
-distinguished types) and whose edges are labeled with lexicons, in other
-words just a mathematical reification of an ACG diagram that has also
-been generalized from arborescences to DAGs. We then search for an
-appropriate semantics for these structures, i.e. how to determine what
-languages are defined by a graphical ACG.
+We define a \emph{graphical abstract categorial grammar} as a directed acyclic
+graph (DAG) whose nodes are labeled with signatures (and distinguished types)
+and whose edges are labeled with lexicons, in other words a mathematical
+reification of an ACG diagram that has been generalized from
+arborescences to DAGs. We then search for an appropriate semantics for these
+structures, i.e. how to determine what languages are defined by a graphical
+ACG.
+
+\subsection{Nodes as Languages}
 
@@
 G-ACG extrinsic languages correspond to ACG object languages with
-intersection. More formally, it can be said that ACG object languages are just
-ACG abstract languages closed on transformation by a lexicon, and we then show
-that G-ACG extrinsic languages are just ACG abstract languages closed on
-transformation by a lexicon and intersection.
+intersection. More formally, while ACG object languages are ACG abstract
+languages closed on transformation by a lexicon, G-ACG extrinsic languages are
+ACG abstract languages closed on transformation by a lexicon and intersection.
 
@@
 grammatical formalism that allows independent combination of linguistic
-constraints. Furthermore, from the above we can see that object
-languages are as expressive as extrinsic languages iff object languages
-are closed on intersection, which is conjectured to be false.
+constraints. Furthermore, from the above we can see that object languages are
+as expressive as extrinsic languages iff object languages are closed on
+intersection, which is conjectured to be false.
 
@@
 
+\subsection{Nodes as Terms}
+
 In the new paradigm, we interpret the nodes of the graph as terms and the
 edges as statements that one term is mapped into another using a lexicon. This
-leads us to the definition of the \emph{pangraphical} language of a node $u$
-in a G-ACG $\mathcal{G}$.
+leads us to the definition of the \emph{pangraphical}\footnote{As opposed to
+  extrinsic languages, which are constrained only by their predecessors in the
+  graph, pangraphical languages are constrained by the entire graph.} language
+of a node $u$ in a G-ACG $\mathcal{G}$.
 
@@
 
-Finally, we construct a G-ACG which integrates the French negation
-constraint discussed in Section~\ref{sec:constraints}, the constraints
-on extraction introduced in \cite{pogodalla2012controlling} and a
-constraint handling agreement in a single grammar specification. The
-negation constraint is implemented exactly as was described in
-Section~\ref{sec:constraints} and the extraction constraints are taken
-``as is'' from \cite{pogodalla2012controlling}. The underlying syntactic
-signature is defined without any concern for the additional constraints
-and the syntax-semantics interface is kept simple as well.
-
-In the end, this lets us define the syntax in a clean way using the
-idiomatic style of categorial grammars (simple atomic types like $N$,
-$NP$ and $S$) and then define the constraints themselves the same way as
-they are defined in papers which try to formalize them individually
-(such as the case with \cite{pogodalla2012controlling}). We can also
-freely combine both the constraint pattern and the syntax-semantics
-ambiguities pattern from Section~\ref{sec:acg-comp}.
+Finally, we construct a G-ACG which integrates the French negation constraint
+discussed in Section~\ref{sec:constraints}, the constraints on extraction
+introduced in \cite{pogodalla2012controlling} and a constraint handling
+agreement in a single grammar specification. The constraints are implemented
+exactly as they were originally described. The underlying syntactic signature
+is defined without any concern for the additional constraints and the
+syntax-semantics interface is kept simple as well.
 
@@
 provided an extension of the ACG formalism to solve the problem and justified
-the need for the increased expressivity.
+the need for the increased expressivity. This embedding of syntactic
+constraints will be used to define a syntax-semantics interface and then to
+build discourse structures.
+
+In the end, this lets us define the syntax in a clean way using the
+idiomatic style of categorial grammars (simple atomic types like $N$,
+$NP$ and $S$) and then define the constraints themselves the same way as
+they are defined in papers which try to formalize them individually
+(such as the case with \cite{pogodalla2012controlling}).
 

### 2013-07-24 lenls/abstract.tex
LENLS: Final proof-read of the abstract

@@
 
-In this paper we focus on the problem of constructing a wide-coverage ACG. We
+In this paper, we focus on the problem of constructing a wide-coverage ACG. We
 identify a problem that comes up when one tries to enforce multiple
 constraints/dependencies in a single ACG and we propose a generalization of
-the formalism which resolves to the issue.
+the formalism which resolves the issue.
 
@@
 
-When describing languages, we are usually interested in systems of more
-than one ACG and diagrams such as the ones on Figure~\ref{fig:acg-comp}
-have been used to explain them.
+When describing languages, we are usually interested in systems of more than
+one ACG, and diagrams such as the ones on Figure~\ref{fig:acg-comp} have been
+used to explain them.
 
@@
 
-In French, negation is signalled by prepending the particle \emph{ne} to
-the negated verb in conjunction with using a properly placed
-accompanying word, such as a negative determiner in one of the verb's
-arguments. This phenomenon has been elegantly formalized in the Frigram
-interaction grammar \cite{perrier2007french}.
+In French, negation is signalled by prepending the particle \emph{ne} to the
+negated verb in conjunction with using a properly placed accompanying word,
+such as a negative determiner, in one of the verb's arguments. This phenomenon
+has been elegantly formalized in the Frigram interaction grammar
+\cite{perrier2007french}.
 
@@
 
-The two paradigms under which we can interpret ACGs are equivalent in
-the context of plain ACGs whose diagrams are arborescent and the two
+The two paradigms under which we can interpret ACGs are equivalent in the
+context of plain ACGs whose diagrams are arborescent and hence the two
 metaphors (nodes as languages and nodes as terms) can be used
-interchangeably. We back this up formally by showing that for every node
-$u$ in every arborescent G-ACG $\mathcal{G}$, we have
-$\mathcal{E}_{\mathcal{G}}(u) = \mathcal{P}_{\mathcal{G}}(u)$. However,
-as we start to consider non-arborescent graphs, we find, interestingly,
-that the two paradigms diverge (i.e. $\exists \mathcal{G},
-u.\ \mathcal{E}_{\mathcal{G}}(u) \neq \mathcal{P}_{\mathcal{G}}(u)$).
+interchangeably. We back this up formally by showing that for every node $u$
+in every arborescent G-ACG $\mathcal{G}$, we have
+$\mathcal{E}_{\mathcal{G}}(u) = \mathcal{P}_{\mathcal{G}}(u)$. However, as we
+start to consider non-arborescent graphs, we find, interestingly, that the two
+paradigms diverge (i.e. $\exists \mathcal{G}, u.\ \mathcal{E}_{\mathcal{G}}(u)
+\neq \mathcal{P}_{\mathcal{G}}(u)$).
 
@@
 
-In the end, this lets us define the syntax in a clean way using the
-idiomatic style of categorial grammars (simple atomic types like $N$,
-$NP$ and $S$) and then define the constraints themselves the same way as
-they are defined in papers which try to formalize them individually
-(such as the case with \cite{pogodalla2012controlling}).
+In the end, this lets us define the syntax in a clean way using the idiomatic
+style of categorial grammars (simple atomic types like $N$, $NP$ and $S$) and
+then define the constraints themselves the same way as they are defined in
+papers which try to formalize them individually (such as is the case with
+\cite{pogodalla2012controlling}).
 

### 2013-09-12 lenls/paper.tex
Converted LENLS abstract to LNCS paper class

--- 2013-07-24 lenls/abstract.tex
@@
-\documentclass[twocolumn]{article}
+% This is LLNCS.DEM the demonstration file of
+% the LaTeX macro package from Springer-Verlag
+% for Lecture Notes in Computer Science,
+% version 2.4 for LaTeX2e as of 16. April 2010
+%
+\documentclass{llncs}
+%
+\usepackage{makeidx}
+
+\usepackage{caption}
+\usepackage[labelformat=simple]{subcaption}
+\renewcommand\thesubfigure{(\alph{subfigure})}
 
@@
 \usepackage{amsmath}
-\usepackage{amsthm}
+%\usepackage{amsthm}
 \usepackage{xcolor}
@@
 \usepackage{hyperref}
-\usepackage[labelformat=simple]{subcaption}
-\renewcommand\thesubfigure{(\alph{subfigure})}
 \usepackage{cancel}
-\usepackage[nottoc,numbib]{tocbibind}
 \usepackage{pdfpages}
 \usepackage{enumitem}
-\usepackage{titling}
-
-\setlength{\droptitle}{-14em}
-\title{Integration of Multiple Constraints in ACG}
-\date{\vspace{-5em}}
 
@@
 
-\newtheorem*{observation}{Observation}
-\newtheorem*{theorem}{Theorem}
-\newtheorem*{corollary}{Corollary}
+%
+\begin{document}
+%
+\frontmatter          % for the preliminaries
+%
+\pagestyle{headings}  % switches on printing of running heads
 
-\begin{document}
+\mainmatter              % start of the contributions
+%
+\title{Integration of Multiple Constraints in ACG}
+%
+\titlerunning{Constraints in ACG}  % abbreviated title (for running head)
+%                                    also used for the TOC unless
+%                                    \toctitle is used
+%
+\author{Ji\v{r}\'i Mar\v{s}\'ik\inst{1} \and Maxime Amblard\inst{2}}
+%
+\authorrunning{Ji\v{r}\'i Mar\v{s}\'ik et al.} % abbreviated author list (for running head)
+%
+%%%% list of authors for the TOC (use if author list has to be modified)
+\tocauthor{Ji\v{r}\'i Mar\v{s}\'ik and Maxime Amblard}
+%
+\institute{Princeton University, Princeton NJ 08544, USA,\\
+\email{I.Ekeland@princeton.edu},\\ WWW home page:
+\texttt{http://users/\homedir iekeland/web/welcome.html}
+\and
+Universit\'{e} de Paris-Sud,
+Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
+F-91405 Orsay Cedex, France}
 
-\maketitle
+\maketitle              % typeset the title of the contribution
 
+\begin{abstract}
 This proposal is a first step towards a wide-coverage Abstract Categorial
@@
 of the formalism: Graphical Abstract Categorial Grammars.
+\keywords{abstract categorial grammars, grammar engineering, grammatical formalisms, formal grammars, computational linguistics}
+\end{abstract}
 
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
   \centering
@@
   \caption{\label{fig:acg-comp} Diagrams of systems of ACGs.}
-\end{figure*}
+\end{figure}
 
@@
   \centering
-  \includegraphics[width=0.25\textwidth,height=0.2\textheight]{../diagrams/final.pdf}
+  \includegraphics[height=0.3\textheight]{../diagrams/final.pdf}
   \caption{\label{fig:gacg} A graphical ACG that implements two
@@
 
-\bibliographystyle{plain}
+%
+% ---- Bibliography ----
+%
+
+\bibliographystyle{splncs}
 \bibliography{../biblio}

### 2013-09-12 lenls/paper.tex
LENLS: Affiliations

@@
 %
-\author{Ji\v{r}\'i Mar\v{s}\'ik\inst{1} \and Maxime Amblard\inst{2}}
+\author{Jiří Maršík\inst{1,2} \and Maxime Amblard\inst{2}}
 %
-\authorrunning{Ji\v{r}\'i Mar\v{s}\'ik et al.} % abbreviated author list (for running head)
+\authorrunning{Maršík et al.} % abbreviated author list (for running head)
 %
 %%%% list of authors for the TOC (use if author list has to be modified)
-\tocauthor{Ji\v{r}\'i Mar\v{s}\'ik and Maxime Amblard}
+\tocauthor{Jiří Maršík and Maxime Amblard}
 %
-\institute{Princeton University, Princeton NJ 08544, USA,\\
-\email{I.Ekeland@princeton.edu},\\ WWW home page:
-\texttt{http://users/\homedir iekeland/web/welcome.html}
+\institute{Université de Lorraine, LORIA, Nancy, 54000, France\\
+\email{\{jiri.marsik, maxime.amblard\}@loria.fr},\\
 \and
-Universit\'{e} de Paris-Sud,
-Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
-F-91405 Orsay Cedex, France}
+INRIA, Nancy, 54000, France}
 

### 2013-09-18 lenls/paper.tex
Elucidating the notation in sec:constraints

@@
 
-Enforcing other constraints leads us to subdivide our ``atomic'' types
-even further (e.g. the authors of \cite{pogodalla2012controlling} add
-features to the $S$ and $NP$ types to implement constraints about
-extraction). Other phenomena, such as agreement on morphological
-properties like number, gender, person or case, intuitively lead us to
-make our types even more specific.
+In the above, we elaborate the types $NP$ and $N$ with features ($\_NEG{=}F$
+and $\_NEG{=}T$) and we give the different types for the preposition \emph{de}
+in the fragment for negation ($N$). The types accept an $NP$ and an $N$ as
+arguments with any combination of values for the feature $NEG$, except for the
+case when both the prepositional $NP$ and the $N$ being modified both contain
+free negative determiners (i.e. there is no $N_{de_4} : NP\_NEG{=}T \limp
+N\_NEG{=}T \limp N\_NEG{=}\ldots$). This encodes a constraint that there can
+be only one free negative determiner per phrase (free as in not hidden inside
+an embedded clause). Besides this constraint, the types also dictate how the
+information about negative determiners should propagate from the argument
+constituent to the complex constituent (in this case, it is simple
+disjunction).
+
+Enforcing other constraints leads us to subdividing our ``atomic'' types even
+further (e.g. the authors of \cite{pogodalla2012controlling} add features to
+the $S$ and $NP$ types to implement constraints about extraction). Other
+phenomena, such as agreement on morphological properties like number, gender,
+person or case, intuitively lead us to make our types even more specific.
 
@@
 one below, which provide complete specifications of the various possible
-situations.
+situations (in this hypothetical grammar ($C$), the preposition \emph{de} has
+12 different types).
 
@@
 In our grammar, we would like to combine several constraints
-(Figure~\ref{fig:acg-comp-constr}) and possibly to also separate the syntactic
-ambiguities from the purely semantic ones (Figure~\ref{fig:acg-comp-sem}).
-However, trying to mix these strategies in the ACG framework forces us to
-solve all the constraints in a single type signature or contaminate the
-syntax-semantics interface with the implementation details of the syntactic
-layer, both of which introduce incidental complexity we want to avoid.
+(Figure~\ref{fig:acg-comp-constr}), and possibly to also separate the
+syntactic ambiguities from the purely semantic ones
+(Figure~\ref{fig:acg-comp-sem}).  However, trying to mix these strategies in
+the ACG framework forces us to solve all the constraints in a single type
+signature or contaminate the syntax-semantics interface with the
+implementation details of the syntactic layer, both of which introduce
+incidental complexity we want to avoid.
 

### 2013-09-18 lenls/paper.tex
Expanded section on G-ACGs

@@
 
-We first follow a paradigm in which nodes of the diagrams are
-interpreted as languages with the edges telling us how these languages
-are defined in terms of each other. A single arrow leading to a language
-means that the target language is produced from the source by mapping it
-through a lexicon. We then argue that the suitable meaning of two or
-more edges arriving at the same node is intersection of languages based
-both on the simplicity of the resulting definitions and on our
-expectations about the desired semantics. This leads us to the following
-definitions of \emph{intrinsic} and \emph{extrinsic} languages
-associated with some node $v$ in a graphical ACG $\mathcal{G}$:
+We first follow a paradigm in which nodes of the diagrams are interpreted as
+languages with the edges telling us how these languages are defined in terms
+of each other. A single arrow leading to a language means that the target
+language is produced from the source by mapping it through a lexicon. We now
+argue that the suitable meaning of two or more edges arriving at the same node
+is intersection of languages based both on the simplicity of the resulting
+definitions and on our expectations about the desired semantics.
+
+In an ACG diagram, a node with no inbound edges stands for an abstract
+language. This language is defined as the set of terms having the correct
+type. If a node has an inbound edge, and therefore a parent, then the elements
+of its language are also obliged to have an antecedent in the parent
+language. It is a small step to go from this definition to the following: the
+language of a node is the set of terms having the correct type and an
+antecedent in the language of any of its parent nodes. This correctly
+characterizes the current use of ACG diagrams, recognizing abstract languages
+as a special case of object languages. Furthermore, this restatement
+generalizes to DAGs and gives us the desired semantics for implementing
+multiple constraints: intersection.
+
+We can formalize the above definitions by introducing the notions of
+\emph{intrinsic} and \emph{extrinsic} languages associated with some node $v$
+in a graphical ACG $\mathcal{G}$:
 
@@
 languages closed on transformation by a lexicon, G-ACG extrinsic languages are
-ACG abstract languages closed on transformation by a lexicon and intersection.
+ACG abstract languages closed on transformation by a lexicon and
+intersection\footnote{Can be show by induction on the topological ordering of
+  any given graphical ACG.}.
 
@@
 
-We then argue that closure on intersection is an intuitive property of a
-grammatical formalism that allows independent combination of linguistic
-constraints. Furthermore, from the above we can see that object languages are
+If we want our grammatical framework to be modular w.r.t. the different
+linguistic constraints it encodes, we essentially need
+intersection\footnote{If I have a grammar that enforces $A$ and a grammar that
+  enforces $B$, then I want to have access to the grammar of the language
+  where both $A$ and $B$ are enforced.}. Not only would we want a framework in
+which intersection is possible, we would also like it to be an easy
+operation. Graphical ACGs make intersections trivial to encode and as for
+expressive power, we know, from the equations above, that object languages are
 as expressive as extrinsic languages iff object languages are closed on
-intersection, which is conjectured to be false.
+intersection (which is, at this moment, conjectured to be false). This means
+that graphical ACGs are either a conservative extension providing a more
+convenient way of expressing intersection or are extending ACGs by adding
+only intersection, which enables constraint-based composition of grammars.
 
-However, the extrinsic languages we have defined above turn out to give
-counter-intuitive interpretations to some graphical ACGs and we revisit the
-question in an another paradigm.
+\begin{figure}
+  \centering
+  \includegraphics[height=0.3\textheight]{../diagrams/diamond-grammar.pdf}
+  \caption{\label{fig:diamond} The diamond-shaped G-ACG $\mathcal{D}$.}
+\end{figure}
+
+While the interpretation of G-ACGs given above does have some nice properties,
+it fails to predict the intuitive understanding of more complex ACG
+diagrams. This is most visible in the diamond-shaped grammar on
+Figure~\ref{fig:diamond}. If we take a term from
+$\mathcal{E}_{\mathcal{D}}(Bottom)$, we know it has antecedents both in
+$\mathcal{E}_{\mathcal{D}}(Left)$ and
+$\mathcal{E}_{\mathcal{D}}(Right)$. However, these two do not have to share an
+antecedent in $\mathcal{E}_{\mathcal{D}}(Top)$. This contradicts the
+generative story one might imagine in which a single term from $Top$ generates
+terms in $Left$ and $Right$ which finally generate a term in $Bottom$.
+
+We can observe another peculiarity on a more practical example. Consider the
+G-ACG in Figure~\ref{fig:gacg}. In classical ACGs, one can always take an
+element in a signature, like $Sem$, and by transduction find its corresponding
+element in an another signature, like $Syntax$. However, in an G-ACG such as
+this it is possible that the $Syntax$ term we obtain by transduction does not
+belong to $\mathcal{E}_{\mathcal{G}}(Syntax)$ because it lacks antecedents in
+either $Constr_1$ or $Constr_2$. This means that the notion of an extrinsic
+language is not capable to give us the set of all meaning terms in $Sem$ which
+actually correspond to syntactically correct sentences in this G-ACG.
+
+The above characteristics motivated us to explore alternative interpretations
+of G-ACGs. We will present one such alternative now, which exchanges the
+language-algebraic point of view for a generative one.
 
@@
 
-The two paradigms under which we can interpret ACGs are equivalent in the
-context of plain ACGs whose diagrams are arborescent and hence the two
-metaphors (nodes as languages and nodes as terms) can be used
-interchangeably. We back this up formally by showing that for every node $u$
-in every arborescent G-ACG $\mathcal{G}$, we have
-$\mathcal{E}_{\mathcal{G}}(u) = \mathcal{P}_{\mathcal{G}}(u)$. However, as we
-start to consider non-arborescent graphs, we find, interestingly, that the two
-paradigms diverge (i.e. $\exists \mathcal{G}, u.\ \mathcal{E}_{\mathcal{G}}(u)
-\neq \mathcal{P}_{\mathcal{G}}(u)$).
+If we compare this new interpretation of G-ACGs to the former one, we find out
+that in the case when the graph of the grammar is an arborescence they are
+actually equivalent. This means that in classical ACGs, where all the diagrams
+are arborescences, the two metaphors (nodes as languages and nodes as terms)
+can be and are used interchangeably. However, as we start to consider
+non-arborescent graphs, we find, interestingly, that the two paradigms diverge
+(i.e. $\exists \mathcal{G}, u.\ \mathcal{E}_{\mathcal{G}}(u) \neq
+\mathcal{P}_{\mathcal{G}}(u)$).
 
@@
 languages giving us counter-intuitive interpretations for some specific
-G-ACGs. We show by construction that pangraphical languages are as
-expressive as extrinsic languages and so the kinds of languages we have
-defined here are increasingly more expressive, $\mathcal{I} \subseteq
-\mathcal{E} \subseteq \mathcal{P}$. We also show that
-$\mathcal{I}_{\mathcal{G}}(u) \supseteq \mathcal{E}_{\mathcal{G}}(u)
-\supseteq \mathcal{P}_{\mathcal{G}}(u)$, meaning that the language
-definitions are more and more constraining and that the extrinsic
-languages went in the same direction as pangraphical languages when
-narrowing the elements of $\mathcal{I}_{\mathcal{G}}(u)$, but not as
-far.
+G-ACGs. Specifically, the members of $\mathcal{P}_{\mathcal{D}}(Bottom)$ in
+the diamond grammar have a single antecedent in
+$\mathcal{P}_{\mathcal{D}}(Top)$ and the language
+$\mathcal{P}_{\mathcal{G}}(Sem)$ (from Figure~\ref{fig:gacg}) contains only
+meanings expressible by syntactically correct sentences.
+
+Pangraphical languages turn out to be at least as expressive as the extrinsic
+languages\footnote{This might come as no surprise given the extended domain of
+  constraints compared to extrinsic languages (constrained by the entire graph
+  as opposed to just the node's predecessors).}. The proof is carried out by
+transforming a G-ACG such that a particular node will have the same
+pangraphical language in the latter G-ACG as the extrinsic language it had in
+the former. The nodes of the newly constructed G-ACG correspond to paths in
+the former one terminating in the node in question\footnote{This means that
+  when the G-ACG is $\mathcal{D}$, the diamond grammar, and the node in
+  question is $Bottom$, then we will have two different nodes for the two
+  paths $[Top, Left, Bottom]$ and $[Top, Right, Bottom]$.}. Thus the labelling
+of nodes with terms which certifies a given term's presence in the
+pangraphical language of the new G-ACG also serves as a proof of its presence
+in the extrinsic language of the old G-ACG.
+
+This gives us the following ladder of expressivity
+
+$$
+\mathcal{I} \subseteq \mathcal{E} \subseteq \mathcal{P}
+$$
+
+which can be complemented with the following series
+
+$$
+\mathcal{I}_{\mathcal{G}}(u) \supseteq \mathcal{E}_{\mathcal{G}}(u) \supseteq
+\mathcal{P}_{\mathcal{G}}(u)
+$$
+
+\section{Example Grammar}
 

### 2013-09-20 lenls/paper.tex
Expanding the Motivation and ACG sections

@@
 %
-\author{Jiří Maršík\inst{1,2} \and Maxime Amblard\inst{2}}
+\author{Jiří Maršík\inst{1,2} \and Maxime Amblard\inst{1,2}}
 %
@@
 
-Abstract Categorial Grammars (ACGs) have shown to be a viable formalism for
-elegantly encoding the dynamic nature of discourse. Proposals based on
-continuation semantics \cite{de2006towards} have tackled topics such as event
-anaphora \cite{qian2011event}, SDRT discourse structure \cite{asher2011sdrt}
-and modal accessibility constraints \cite{asher2011montagovian}. However, all
-of these treatments only consider tiny fragments of languages. We are
-interested in building a wide-coverage grammar which integrates and reconciles
-the existing formal treatments of discourse and allows us to study their
-interactions and to build discourse representations automatically.
+Abstract Categorial Grammars (ACGs) \cite{de2001towards} have shown to be a
+viable formalism for elegantly encoding the dynamic nature of
+discourse. Proposals based on continuation semantics \cite{de2006towards} have
+tackled topics such as event anaphora \cite{qian2011event}, SDRT discourse
+structure \cite{asher2011sdrt} and modal accessibility constraints
+\cite{asher2011montagovian}. However, all of these treatments only consider
+tiny fragments of languages. We are interested in building a wide-coverage
+grammar which integrates and reconciles the existing formal treatments of
+discourse and allows us to study their interactions and to build discourse
+representations automatically.
 
-In this paper, we focus on the problem of constructing a wide-coverage ACG. We
-identify a problem that comes up when one tries to enforce multiple
-constraints/dependencies in a single ACG and we propose a generalization of
-the formalism which resolves the issue.
+A presupposed condition to actually using an ACG to describe discourse in a
+large scope is to have a large scale ACG grammar in the first place. The work
+presented here was motivated by this quest for a wide-coverage ACG grammar, as
+seen from a point of view of language as a system of largely orthogonal
+constraints. Encoding ancillary constraints in a type system can make the
+system hard to read and understand. If multiple constraints written in the
+style of \cite{pogodalla2012controlling} are to be enforced in the same
+grammar, we advocate for extending the formalism to prevent the incidental
+complexity that would otherwise appear.
+
+We begin with a short review of Abstract Categorial Grammars in which we
+highlight the ways in which Abstract Categorial Grammars facilitate
+decomposition of linguistic description. We follow that by a closer look at an
+example linguistic constraint and how its treatment might look like in a
+categorial grammar, which motivates our proposed extension. We then present
+the Graphical Abstract Categorial Grammars and discuss some of its formal
+properties. We then finish with an illustration of our approach which
+incorporates linguistic contribution from multiple sources into a single
+coherent and simple grammar and discuss the limitations and challenges of
+adopting this approach.
+
+\section{Abstract Categorial Grammars}
+\label{sec:acg-comp}
+
+We first review the grammatical framework in which we conduct our
+work. Abstract Categorial Grammars are built upon two mathematical structures,
+\emph{(higher-order) signatures} and \emph{lexicons}.
+
+\subsection{Higher-Order Signatures}
+\label{ssec:sig}
+
+A \textbf{higher-order signature} is a set of elements that we call
+\emph{constants}, each of which is associated with a type. Formally, it
+is defined as a triple $\Sigma = \mathopen{<}A, C, \tau\mathclose{>}$,
+where:
+\begin{itemize}
+  \item $C$ is the (finite) set of constants
+  \item $A$ is a (finite) set of atomic types
+  \item $\tau$ is the type-associating mapping from $C$ to
+    $\mathcal{T}(A)$, the set of types built over $A$
+\end{itemize}
+
+In our case, $\mathcal{T}(A)$ is the implicative fragment of linear and
+intuitionistic logic with $A$ being the atomic propositions. This means
+that $\mathcal{T}(A)$ contains all the $a \in A$ and all the $\alpha \limp
+\beta$ and $\alpha \to \beta$ for $\alpha, \beta \in \mathcal{T}(A)$.
+
+A signature $\Sigma = \mathopen{<}A, C, \tau\mathclose{>}$, by itself, already
+lets us define an interesting set of structures, that is the set
+$\Lambda(\Sigma)$ of \emph{well-typed lambda terms} built upon the signature
+$\Sigma$. To make this structure even more useful, we often focus ourselves
+only on terms that have a specific \emph{distinguished type}. Using this
+notion of a signature of typed constants and some distinguished type, we can
+describe languages of, e.g. tree-like (and by extension string-like), lambda
+terms.
+
+\subsection{Lexicons}
+
+The idea of a signature is coupled with the one of \textbf{lexicon},
+which is a mapping between two different signatures (mapping the
+constants of one into well-typed terms of the other). Formally speaking,
+a lexicon $\mathcal{L}$ from a signature $\Sigma_1 = \mathopen{<}A_1,
+C_1, \tau_1\mathclose{>}$ (which we call the abstract signature) to a
+signature $\Sigma_2 = \mathopen{<}A_2, C_2, \tau_2\mathclose{>}$ (which
+we call the object signature) is a pair $\mathopen{<}F, G\mathclose{>}$
+such that:
+
+\begin{itemize}
+\item $G$ is a mapping from $C_1$ to $\Lambda(\Sigma_2)$ assigning to
+  every constant of the abstract signature a term in the object
+  signature, which can be understood as its
+  interpretation/implementation/realization.
+\item $F$ is a mapping from $A_1$ to $\mathcal{T}(A_2)$ which links the
+  abstract-level types with the object-level types that they are
+  realized as.
+\item $F$ and $G$ are compatible, meaning that for any $c \in C_1$, we
+  have $\vdash_{\Sigma_2} G(c) : \hat{F}(\tau_1(c))$ (we will be using
+  $\hat{F}$ and $\hat{G}$ to refer to the homomorphic extensions of $F$
+  and $G$ to $\mathcal{T}(A_1)$ and $\Lambda(\Sigma_1)$
+  respectively).
+\end{itemize}
+
+Now we have enough machinery in hand to describe how ACGs define
+languages. Given some signature and a distinguished type, we can talk of the
+\emph{abstract language} which is the set of well-typed terms built on that
+signature having that distinguished type. If we then also consider a lexicon
+and its abstract signature, then the language we get by mapping the abstract
+language of the signature through that lexicon is called an \emph{object
+  language}.
+
+An ACG for us will then be just a collection of signatures with their
+distinguished types and the lexicons connecting these signatures, which we
+will write down in diagrams such as the ones in Figure~\ref{fig:acg-comp}.
+
+The most common pattern will have us using two object signatures for the
+surface forms of utterances (strings) and their logical forms (logical
+propositions) and an abstract signature which is connected to both of the
+object signatures via lexicons (as we can see in
+Figure~\ref{fig:acg-comp-basic}). Parsing is then just a matter of inverting
+the surface lexicon to get the abstract term and then applying to it the
+logical lexicon. Generation is symmetric, we simply invert the logical lexicon
+and apply the surface lexicon.
+
+\subsection{Systems of ACGs}
 
@@
     \includegraphics[height=0.2\textheight]{../diagrams/serial-over-parallel.pdf}
-    \caption{\label{fig:acg-comp-constr} Adding a constraint.}
+    \caption{\label{fig:acg-comp-constr} Adding a constraint (Constraint).}
   \end{subfigure}
@@
     \caption{\label{fig:acg-comp-sem} Distinguishing syntactic and
-      semantic ambiguities.}
+      semantic ambiguities (Elaboration).}
   \end{subfigure}
@@
 
-\section{Systems of ACGs}
-\label{sec:acg-comp}
+Besides the pattern in Figure~\ref{fig:acg-comp-basic}, there have been other
+techniques of decomposing linguistic description in the ACG literature. Here
+we will focus on two that are relevant to our work.
 
-An ACG is defined in terms of two higher-order signatures, sets of typed
-constants on which we can form well-typed lambda terms, and a
-translating function from terms of one of the signatures to terms of the
-other (the former signature being called the \emph{abstract signature},
-the latter being called the \emph{object signature} and the function
-being called a \emph{lexicon}). We are then interested in the terms of
-the abstract signature having some distinguished type and their images
-through the lexicon (these two sets of terms are called the
-\emph{abstract language} and the \emph{object language}, respectively).
+The first constitutes the Constraint pattern used throughout
+\cite{pogodalla2012controlling}. Imagine we want our grammar to enforce a
+particular linguistic constraint, say that tensed clauses form islands for
+quantifier scope, and suppose we want to proceed by refining our types so as
+to make linguistic structures violating the constraint untypable. We could
+dive in to our grammar and rewrite it to use the new types. However, we can
+also keep the existing grammar intact and complement it with a signature and a
+lexicon that translates that signature to the abstract-most signature of our
+existing grammar. Usually, the new signature will resemble the prior
+abstract-most one, albeit with more fine-grained types necessary to express
+the constraint in question.
 
-When describing languages, we are usually interested in systems of more than
-one ACG, and diagrams such as the ones on Figure~\ref{fig:acg-comp} have been
-used to explain them.
+This pattern is practical for examining linguistic constraints one-by-one and
+demonstrating that the formalism is capable of expressing them
+individually. However, this pattern does not serve us much in the context of
+ACGs when we try to build a grammar which incorporates many of these at the
+same time. This is due to the fact that after grafting on the first
+constraint, its added signature now becomes the abstract-most signature in the
+grammar. Any other constraint that would be added using the Constraint pattern
+would have to be translated into this new signature. This would mean that the
+type systems of newly added constraints would have to re-enforce all of the
+constraints introduced before, therefore nullifying almost any benefit of the
+pattern.
 
-In Figure~\ref{fig:acg-comp-basic}, abstract syntactic terms from the
-signature $\Sigma_{Synt}$ yield phonological strings and semantic
-representations using two different lexicons. In
-Figure~\ref{fig:acg-comp-constr}, this pattern is extended by adding a
-new abstract signature which narrows the language of syntactic terms and
-implements some linguistic constraint
-\cite{pogodalla2012controlling}. Finally, in
-Figure~\ref{fig:acg-comp-sem}, the abstract signature $\Sigma_{Synt}$ of
-Figure~\ref{fig:acg-comp-basic} is split into two signatures,
-$\Sigma_{Syntax}$ and $\Sigma_{SyntSem}$, in order to model syntactic
-and semantic ambiguities in different grammars
-\cite{pogodalla2007generalizing}.
+Another pattern we will briefly mention is the Elaboration pattern in which an
+edge in the diagram (a lexicon) is expanded into two edges and an intermediate
+node (signature). We can illustrate this pattern on Tree Adjoining Grammars
+\cite{degroote02}. Suppose we have a signature describing TAG derivation trees
+and a lexicon which maps them to their yield. We might want to make this
+grammar richer, and arguably more synoptic, by elaborating the translation
+from derivation trees to yields by introducing derived trees and describe
+separately how both derivation trees produce derived trees and how derived
+trees produce the yields. This way, we can make emerge a potentially useful
+linguistic structure.
+
+This pattern was used in \cite{pogodalla2007generalizing} to elaborate the
+translation from the common Montague-style \cite{montague1973proper}
+higher-order signature to its yield by introducing lower-order syntactic
+structures more reminiscent of TAG or CFG trees. This helps make a clear
+distinction between ambiguities caused by syntax and by semantics by
+segregating the purely syntactic description in the $Syntax$ signature from
+the issues of operator scope that have to be solved in the $SyntSem$ signature
+(Figure~\ref{fig:acg-comp-sem}). However, in classical ACGs, we cannot exploit
+this pattern as much as we would like. If we were to use this grammar as a
+kernel on top of which we would like to express a syntactic constraint, we
+would not be able to attach the constraint's signature directly to the
+$Syntax$ node but we would instead need to attach it to
+$SyntSem$\footnote{This is due to the fact that ACG diagrams are always
+  arborescences, in which the root node represents the abstract language from
+  whose terms are generated terms of all the object languages.}. This means
+that the syntactic constraint would end up being expressed by strengthening
+the type system of $SyntSem$, which only adds unnecessary complexity.
 

### 2013-09-21 lenls/paper.tex
LENLS: Wrote up the example grammar

@@
 \begin{align*}
-N_{de_1} &: NP\_NEG{=}F \limp N\_NEG{=}F \limp N\_NEG{=}F \\
-N_{de_2} &: NP\_NEG{=}F \limp N\_NEG{=}T \limp N\_NEG{=}T \\
-N_{de_3} &: NP\_NEG{=}T \limp N\_NEG{=}F \limp N\_NEG{=}T
+N_{de_1} &: NP_{NEG{=}F} \limp N_{NEG{=}F} \limp N_{NEG{=}F} \\
+N_{de_2} &: NP_{NEG{=}F} \limp N_{NEG{=}T} \limp N_{NEG{=}T} \\
+N_{de_3} &: NP_{NEG{=}T} \limp N_{NEG{=}F} \limp N_{NEG{=}T}
 \end{align*}
 
-In the above, we elaborate the types $NP$ and $N$ with features ($\_NEG{=}F$
-and $\_NEG{=}T$) and we give the different types for the preposition \emph{de}
+In the above, we elaborate the types $NP$ and $N$ with features ($_{NEG{=}F}$
+and $_{NEG{=}T}$) and we give the different types for the preposition \emph{de}
 in the fragment for negation ($N$). The types accept an $NP$ and an $N$ as
@@
 case when both the prepositional $NP$ and the $N$ being modified both contain
-free negative determiners (i.e. there is no $N_{de_4} : NP\_NEG{=}T \limp
-N\_NEG{=}T \limp N\_NEG{=}\ldots$). This encodes a constraint that there can
+free negative determiners (i.e. there is no $N_{de_4} : NP_{NEG{=}T} \limp
+N_{NEG{=}T} \limp N\_NEG{=}\ldots$). This encodes a constraint that there can
 be only one free negative determiner per phrase (free as in not hidden inside
@@
 
-\begin{align*}
-C_{de_{11}} :\ &(NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \\
-&\limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \\
-&\limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
-\end{align*}
+$$
+C_{de_{11}} :\ NP_{\textcolor{red}{NEG{=}T}, \textcolor{green}{VAR{=}F}, \textcolor{blue}{NUM{=}PL}}
+\limp N_{\textcolor{red}{NEG{=}F},\textcolor{blue}{NUM{=}SG}}
+\limp N_{\textcolor{red}{NEG{=}T},\textcolor{blue}{NUM{=}SG}}
+$$
 
@@
 
-\section{Example Grammar}
+\section{Illustration}
 
-Finally, we construct a G-ACG which integrates the French negation constraint
-discussed in Section~\ref{sec:constraints}, the constraints on extraction
-introduced in \cite{pogodalla2012controlling} and a constraint handling
-agreement in a single grammar specification. The constraints are implemented
-exactly as they were originally described. The underlying syntactic signature
-is defined without any concern for the additional constraints and the
-syntax-semantics interface is kept simple as well.
+In this final section, we assemble a G-ACG which integrates the French
+negation constraint discussed in Section~\ref{sec:constraints}, the
+constraints on extraction introduced in \cite{pogodalla2012controlling} and a
+constraint handling agreement in a single grammar specification to demonstrate
+our approach.
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{../diagrams/single-cont.pdf}
+    \caption{\label{fig:single-cont} An extraction constraint }
+  \end{subfigure}
+  \qquad
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{../diagrams/merged-conts.pdf}
+    \caption{\label{fig:merged-conts} G-ACG merging the three extraction
+      constraints.}
+  \end{subfigure}
+  \caption{\label{fig:conts} Combining the constraints on extraction.}
+\end{figure}
+
+We start with the extraction constraints. \cite{pogodalla2012controlling}
+describes three different constraints on extraction, all of them expressed
+using the Constraint pattern (Figure~\ref{fig:single-cont}). We can take the
+union of these three G-ACGs to get the G-ACG on Figure~\ref{fig:merged-conts}.
+Next, we can incorporate knowledge from another paper,
+\cite{pogodalla2007generalizing}, by splitting the $\mathcal{L}_{Syn}$ lexicon
+and introducing a new intermediate signature
+(Figure~\ref{fig:conts-elab}). The new signature, $Syntax$, deals purely with
+syntax without any issues of operator scope and its functions have lower-order
+types. We will finally add our contribution, the signatures $Neg$ and $Agr$
+and the lexicons $\mathcal{L}_{Neg}$ and $\mathcal{L}_{Agr}$ which implement
+the negation phenomenon we discussed before and some simple notion of
+agreement, respectively.
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{../diagrams/conts-elab.pdf}
+    \caption{\label{fig:conts-elab} Elaborating the lexicon from syntax to
+      strings. }
+  \end{subfigure}
+  \qquad
+  \begin{subfigure}[b]{0.4\textwidth}
+    \centering
+    \includegraphics[height=0.2\textheight]{../diagrams/final-synth.pdf}
+    \caption{\label{fig:final-synth} Adding our own syntactic constraints,
+      $Neg$ and $Agr$.}
+  \end{subfigure}
+  \caption{\label{fig:final} Building the example grammar up to its final
+    state.}
+\end{figure}
+
+We notice that while the three constraints introduced in
+\cite{pogodalla2012controlling} are expressed in terms of the $Syn$ signature,
+our constraints $Agr$ and $Neg$ use a different interface, the $Syntax$
+signature, to constrain the languages being defined. The $Cont_i$ constraints
+are not actually constraining syntax itself, but the syntax-semantics
+interface. This is in some cases a deliberate decision, since e.g. $Cont_1$ is
+the constraint enforcing that quantifier scope must not escape outside of
+tensed clauses. The $Syntax$ signature is not equipped to talk about operator
+scope and so the constraint is expressed above the $Syn$ signature. On the
+other hand, constraints $Cont_2$ and $Cont_3$ from
+\cite{pogodalla2012controlling} are purely syntactic. However, we see it as a
+perk of our approach that we can ignore that and import the constraint
+wholesale without having to adapt them.
+
+\subsection{The $Neg$ Signature}
+
+We will now look under the covers of the two signatures implementing the
+constraints, starting with negation and the $Neg$ signature.
+
+As was said in Section~\ref{sec:constraints}, we will proceed by splitting the
+atomic types $N$ and $NP$ into $N_{NEG{=}F}$ and $N_{NEG{=}T}$, and
+$NP_{NEG{=}F}$ and $NP_{NEG{=}T}$, respectively. We will use these new types to
+distinguish whether an $N$ or $NP$ phrase contains a negative determiner that
+can pair up with a negated verb. Where before we had a single type, we will
+now have possibly multiple to account for the different values of the
+arguments' $NEG$ features. Here are some representative examples:
+
+\begin{align*}
+N_{aucun} &: N_{NEG{=}F} \limp NP_{NEG{=}T} \\
+N_{le_1} &: N_{NEG{=}F} \limp NP_{NEG{=}F} \\
+N_{le_2} &: N_{NEG{=}T} \limp NP_{NEG{=}T} \\
+N_{ne^{tv}_1} &: (NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S) \limp (NP_{NEG{=}T} \limp NP_{NEG{=}F} \limp S) \\
+N_{ne^{tv}_2} &: (NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S) \limp (NP_{NEG{=}F}
+\limp NP_{NEG{=}T} \limp S) \\
+N_{ne^{tv}_3} &: (NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S) \limp (NP_{NEG{=}T}
+\limp NP_{NEG{=}T} \limp S) \\
+N_{aime} &: NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S \\
+N_{que_1} &: (NP_{NEG{=}F} \limp S) \limp N_{NEG{=}F} \limp N_{NEG{=}F} \\
+N_{que_2} &: (NP_{NEG{=}F} \limp S) \limp N_{NEG{=}T} \limp N_{NEG{=}T} \\
+\end{align*}
+
+We see that the negative determiner \emph{aucun} can only attach to phrases
+which do not already contain a free negative determiner and that the resulting
+phrase is marked as having a free negative determiner. On the other hand, the
+determiner \emph{le} does not have any interaction and is still able to take
+any $N$ phrase, preserving any free negative determiners inside. The type of
+the negative particle \emph{ne} modifies a verb (in the example above,
+$N_{ne^{tv}_i}$, a transitive verb) and produces a new verb that makes sure
+that at least one of its arguments contains a negative determiner. Verbs are
+not agnostic about the $NEG$ feature. By default, they specifically demand it
+to be false, since a negative determiner cannot be coupled with a non-negated
+verb. Finally, the types of the relativizer \emph{que} tell us two other
+things: the $NP$ trace is considered as containing no negative determiners,
+and a relative clause does not care about or alter the presence of free
+negative determiners in the $N$ it modifies.
+
+The lexicon which translates from this signature to $Syntax$ is a
+straightforward relabeling. Its items can be characterized by the following
+schema:
+
+$$
+\mathcal{L}_{Neg}(N_{wordform_{i}}) = C_{wordform}
+$$
+
+where $C_{wordform}$ is the constant in $Syntax$ corresponding to the
+$wordform$.
+
+\subsection{The $Agr$ Signature}
+
+The $Agr$ signature will be constructed using the same strategy as $Neg$. $N$
+and $NP$ will be subdivided into $N_{NUM{=}SG}$, $N_{NUM{=}PL}$,
+$NP_{NUM{=}SG}$ and $NP_{NUM{=}PL}$ (we only treat number agreement in this
+example). Here are some example types:
+
+\begin{align*}
+A_{tatou} &: N_{NUM{=}SG} \\
+A_{tatous} &: N_{NUM{=}PL} \\
+A_{le} &: N_{NUM{=}SG} \limp NP_{NUM{=}SG} \\
+A_{les} &: N_{NUM{=}PL} \limp NP_{NUM{=}PL} \\
+A_{aime_1} &: NP_{NUM{=}SG} \limp NP_{NUM{=}SG} \limp S \\
+A_{aime_2} &: NP_{NUM{=}SG} \limp NP_{NUM{=}PL} \limp S \\
+A_{qui_1} &: (NP_{NUM{=}SG} \limp S) \limp N_{NUM{=}SG} \limp N_{NUM{=}SG} \\
+A_{qui_2} &: (NP_{NUM{=}PL} \limp S) \limp N_{NUM{=}PL} \limp N_{NUM{=}PL}
+\end{align*}
+
+The types for nouns and determiners are quite predictable. For the transitive
+verb \emph{aime}, we see that we need a second type to account for the fact
+that \emph{aime} does not care about the number of the object. Finally, the
+relativizer \emph{qui} enforces that the number of the trace $NP$ in the
+relative clause must be the same as the number of the $N$ being modified.
+
+The lexicon $\mathcal{L}_{Agr}$ follows exactly the same schema as that of
+$\mathcal{L}_{Neg}$.
+
+\subsection{Discussion}
 

### 2013-09-22 lenls/paper.tex
LENLS: Wrote Limits of Modularity

@@
 union of these three G-ACGs to get the G-ACG on Figure~\ref{fig:merged-conts}.
-Next, we can incorporate knowledge from another paper,
-\cite{pogodalla2007generalizing}, by splitting the $\mathcal{L}_{Syn}$ lexicon
-and introducing a new intermediate signature
-(Figure~\ref{fig:conts-elab}). The new signature, $Syntax$, deals purely with
-syntax without any issues of operator scope and its functions have lower-order
-types. We will finally add our contribution, the signatures $Neg$ and $Agr$
-and the lexicons $\mathcal{L}_{Neg}$ and $\mathcal{L}_{Agr}$ which implement
-the negation phenomenon we discussed before and some simple notion of
-agreement, respectively.
+Next, we can incorporate knowledge from \cite{pogodalla2007generalizing}, by
+splitting the $\mathcal{L}_{Syn}$ lexicon and introducing a new intermediate
+signature (Figure~\ref{fig:conts-elab}). The new signature, $Syntax$, deals
+purely with syntax without any issues of operator scope and its functions have
+lower-order types than those in $Syn$. We will finally add our contribution,
+the signatures $Neg$ and $Agr$ and the lexicons $\mathcal{L}_{Neg}$ and
+$\mathcal{L}_{Agr}$ which implement the constraint on negation we discussed
+before and some simple notion of agreement, respectively.
 
@@
 other hand, constraints $Cont_2$ and $Cont_3$ from
-\cite{pogodalla2012controlling} are purely syntactic. However, we see it as a
-perk of our approach that we can ignore that and import the constraint
-wholesale without having to adapt them.
+\cite{pogodalla2012controlling} are purely syntactic and therefore could be
+more easily expressed in terms of $Syntax$ instead of $Syn$. However, we see
+it as a perk of our approach that we can ignore that and import the
+constraints wholesale without having to adapt them.
 
@@
 atomic types $N$ and $NP$ into $N_{NEG{=}F}$ and $N_{NEG{=}T}$, and
-$NP_{NEG{=}F}$ and $NP_{NEG{=}T}$, respectively. We will use these new types to
-distinguish whether an $N$ or $NP$ phrase contains a negative determiner that
-can pair up with a negated verb. Where before we had a single type, we will
-now have possibly multiple to account for the different values of the
-arguments' $NEG$ features. Here are some representative examples:
+$NP_{NEG{=}F}$ and $NP_{NEG{=}T}$, respectively. We will use these new types
+to distinguish whether an $N$ or $NP$ phrase contains a negative determiner
+that can pair up with a negated verb. Where before we had a single type, we
+will now have possibly multiple types to account for the different values of
+the arguments' $NEG$ features. Here are some representative examples:
 
@@
 phrase is marked as having a free negative determiner. On the other hand, the
-determiner \emph{le} does not have any interaction and is still able to take
-any $N$ phrase, preserving any free negative determiners inside. The type of
-the negative particle \emph{ne} modifies a verb (in the example above,
-$N_{ne^{tv}_i}$, a transitive verb) and produces a new verb that makes sure
-that at least one of its arguments contains a negative determiner. Verbs are
-not agnostic about the $NEG$ feature. By default, they specifically demand it
-to be false, since a negative determiner cannot be coupled with a non-negated
-verb. Finally, the types of the relativizer \emph{que} tell us two other
-things: the $NP$ trace is considered as containing no negative determiners,
-and a relative clause does not care about or alter the presence of free
-negative determiners in the $N$ it modifies.
+determiner \emph{le} does not have any interaction with the new $NEG$ feature
+and is still able to take any $N$ phrase, preserving any free negative
+determiners inside. The type of the negative particle \emph{ne} modifies a
+verb (in the example above, $N_{ne^{tv}_i}$, a transitive verb) and produces a
+new verb that makes sure that at least one of its arguments contains a
+negative determiner. Verbs are not agnostic about the $NEG$ feature. By
+default, they specifically demand it to be false, since a negative determiner
+cannot be coupled with a non-negated verb. Finally, the types of the
+relativizer \emph{que} tell us two other things: the $NP$ trace is considered
+as containing no negative determiners, and a relative clause does not care
+about or alter the presence of free negative determiners in the $N$ it
+modifies.
 
@@
 
-\subsection{Discussion}
+\subsection{Limits of Modularity}
+
+We have decomposed our grammar into 9 signatures and 8 lexicons. Besides
+making the grammar more readable and easily navigable, it also makes the
+grammar more attractive from a computational point of view. The complexity of
+parsing the highest abstract-most structures and verifying all the constraints
+is decomposed into smaller subtasks. A string can be parsed into its
+representation at the $Syntax$ level fairly easily (compared to parsing it at
+the higher levels) since $Syntax$ employs types of lower orders than the more
+abstract signatures. Finding the syntactic structure for a string is thus
+feasible. We then proceed to verifying the syntactic constraints. Here we can
+observe that the lexicons $\mathcal{L}_{Neg}$ and $\mathcal{L}_{Agr}$ are mere
+relabelings and therefore parsing (i.e. constraint verification) is trivial
+(i.e. decidable). Now in order to determine the meaning of our sentence, we
+have to continue parsing by finding an antecedent in $Syn$. This involves
+higher-order types used to encode the various scope-bearing
+operators. Finally, we can opt in to verify the high-level syntax-semantics
+constraints which employ dependent types. However, the lexicons in this case
+resemble relabelings and a semi-decidable procedure could be used to solve the
+problem. Having the grammar decomposed like this opens it up for computational
+processing by letting us engage in different tactics at different stages
+instead of having to parse a single hidden structure which enforces
+grammaticality w.r.t. every constraint. G-ACGs share these parks also with
+classical ACGs but they let use this composition in another dimension, which
+facilitates the writing of multiple constraints.
+
+However, it is important to keep in mind that the composability demonstrated
+above is limited only to the composition of constraints. If we would like to
+extend our language with, e.g., a new lexical category, then every signature
+and every lexicon would have to be extended to cover this new category. It is
+also important to point out that the constraint signatures that we have been
+composing are not light-weight objects. They are usually defined by taking an
+existing signature that describes the language at a relevant level and by
+strengthening (usually all) the types of that signature to enforce the
+constraint\footnote{This makes the expressivity of model-theoretic syntactic
+  formalisms formalisms such as Interaction Grammars, in which adding the
+  lexical items for \emph{ne} and \emph{aucun} along with enforcing the $Neg$
+  constraint takes only two tree descriptions \cite{perrier2007french}, even
+  more striking when compared to our approach.}. However, most of these type
+alterations are mundane and we believe that this is fertile ground for
+applying metagrammars.
 

### 2013-09-22 lenls/paper.tex
LENLS: Finishing the Conclusion

@@
 the need for the increased expressivity. This embedding of syntactic
-constraints will be used to define a syntax-semantics interface and then to
-build discourse structures.
+constraints will contribute to an effort to define a syntax-semantics
+interface and later to build discourse structures.
 
-In the end, this lets us define the syntax in a clean way using the idiomatic
-style of categorial grammars (simple atomic types like $N$, $NP$ and $S$) and
-then define the constraints themselves the same way as they are defined in
-papers which try to formalize them individually (such as is the case with
+In the end, our approach us define the syntax in a clean way using the
+idiomatic style of categorial grammars (simple atomic types like $N$, $NP$ and
+$S$) and then define the constraints themselves the same way as they are
+defined in ACG research (such as is the case with
 \cite{pogodalla2012controlling}).
+
+Interesting avenues of future work include: digging deeper into the
+language-theoretical properties of G-ACGs (Are ACG object languages closed on
+intersection? Are pangraphical languages expressible as extrinsic languages?),
+searching for a metagrammatical description of constraint signatures to ease
+work on G-ACG grammars and finally building large-scale grammars to verify the
+utility of our approach.
 

### 2013-09-22 lenls/paper.tex
LENLS: Fixed the affiliations
LENLS: Implemented style remarks

@@
 %
-\author{Jiří Maršík\inst{1,2} \and Maxime Amblard\inst{1,2}}
+\author{Jiří Maršík\inst{1,2,3} \and Maxime Amblard\inst{1,2,3}}
 %
@@
 %
-\institute{Université de Lorraine, LORIA, Nancy, 54000, France\\
+\institute{Université de Lorraine, Laboratoire lorrain de recherche en informatique et ses applications, UMR 7503, Vandoeuvre-lès-Nancy, 54500, France\\
 \email{\{jiri.marsik, maxime.amblard\}@loria.fr},\\
 \and
-INRIA, Nancy, 54000, France}
+INRIA, Villers-lès-Nancy, 54600, France\\
+\and
+CNRS, Loria, Vandoeuvre-lès-Nancy, UMR 7503, 54500, France}
 
@@
 
-We begin with a short review of Abstract Categorial Grammars in which we
+We begin with a short review of Abstract Categorial Grammars where we
 highlight the ways in which Abstract Categorial Grammars facilitate
 decomposition of linguistic description. We follow that by a closer look at an
-example linguistic constraint and how its treatment might look like in a
-categorial grammar, which motivates our proposed extension. We then present
-the Graphical Abstract Categorial Grammars and discuss some of its formal
-properties. We then finish with an illustration of our approach which
-incorporates linguistic contribution from multiple sources into a single
-coherent and simple grammar and discuss the limitations and challenges of
-adopting this approach.
+example of a linguistic constraint. This discussion motivates our proposed
+extension. We then present the Graphical Abstract Categorial Grammars and
+consider some of its formal properties. We then finish with an illustration of
+our approach which incorporates linguistic contribution from multiple sources
+into a single coherent and simple grammar and examine the limitations and
+challenges of adopting this approach.
 
@@
 We first review the grammatical framework in which we conduct our
-work. Abstract Categorial Grammars are built upon two mathematical structures,
+work. Abstract Categorial Grammars are built upon two mathematical structures:
 \emph{(higher-order) signatures} and \emph{lexicons}.
@@
 
-The most common pattern will have us using two object signatures for the
-surface forms of utterances (strings) and their logical forms (logical
-propositions) and an abstract signature which is connected to both of the
-object signatures via lexicons (as we can see in
+The most common pattern will have us using two object signatures,
+$\Sigma_{String}$ and $\Sigma_{Sem}$, for the surface forms of utterances
+(strings) and their logical forms (logical propositions) and an abstract
+signature, $\Sigma_{Synt}$, which is connected to both of the object
+signatures via lexicons (as we can see in
 Figure~\ref{fig:acg-comp-basic}). Parsing is then just a matter of inverting
@@
 to make linguistic structures violating the constraint untypable. We could
-dive in to our grammar and rewrite it to use the new types. However, we can
+dive into our grammar and rewrite it to use the new types. However, we can
 also keep the existing grammar intact and complement it with a signature and a

### 2013-09-22 lenls/paper.tex
LENLS: Live feedback from Maxime

@@
 \institute{Université de Lorraine, Laboratoire lorrain de recherche en informatique et ses applications, UMR 7503, Vandoeuvre-lès-Nancy, 54500, France\\
-\email{\{jiri.marsik, maxime.amblard\}@loria.fr},\\
 \and
@@
 \and
-CNRS, Loria, Vandoeuvre-lès-Nancy, UMR 7503, 54500, France}
+CNRS, Loria, Vandoeuvre-lès-Nancy, UMR 7503, 54500, France\\
+\email{\{jiri.marsik, maxime.amblard\}@loria.fr}}
 
@@
 \end{abstract}
-
-\section{Motivation}
 
@@
 
+\vspace{-4ex}
+
 \begin{align*}
@@
 \end{align*}
+
+\vspace{-1ex}
 
@@
 one below, which provide complete specifications of the various possible
-situations (in this hypothetical grammar ($C$), the preposition \emph{de} has
-12 different types).
+situations (in this grammar ($C$), the preposition \emph{de} has 12 different
+types).
+
+\vspace{-4ex}
 
@@
 $$
+
+\vspace{-1ex}
 
@@
 arborescences and we are obliged to generalize them in order to get
-the expected interpretation of Figure~\ref{fig:gacg}.
+the expected interpretation.
 
@@
 generative story one might imagine in which a single term from $Top$ generates
-terms in $Left$ and $Right$ which finally generate a term in $Bottom$.
+terms in $Left$ and $Right$ which generate a term in $Bottom$.
 
@@
 
+\vspace{-3ex}
+
 We see that the negative determiner \emph{aucun} can only attach to phrases
@@
 
+\vspace{-3ex}
+
 \begin{align*}
@@
 A_{aime_1} &: NP_{NUM{=}SG} \limp NP_{NUM{=}SG} \limp S \\
-A_{aime_2} &: NP_{NUM{=}SG} \limp NP_{NUM{=}PL} \limp S \\
+A_{aime_2} &: NP_{NUM{=}SG} \limp NP_{NUM{=}PL} \limp S
+\end{align*}
+\begin{align*}
 A_{qui_1} &: (NP_{NUM{=}SG} \limp S) \limp N_{NUM{=}SG} \limp N_{NUM{=}SG} \\
@@
 constraint\footnote{This makes the expressivity of model-theoretic syntactic
-  formalisms formalisms such as Interaction Grammars, in which adding the
-  lexical items for \emph{ne} and \emph{aucun} along with enforcing the $Neg$
-  constraint takes only two tree descriptions \cite{perrier2007french}, even
-  more striking when compared to our approach.}. However, most of these type
+  formalisms such as Interaction Grammars, in which adding the lexical items
+  for \emph{ne} and \emph{aucun} along with enforcing the $Neg$ constraint
+  takes only two tree descriptions \cite{perrier2007french}, even more
+  striking when compared to our approach.}. However, most of these type
 alterations are mundane and we believe that this is fertile ground for
@@
 
-In the end, our approach us define the syntax in a clean way using the
+In the end, our approach lets us define the syntax in a clean way using the
 idiomatic style of categorial grammars (simple atomic types like $N$, $NP$ and
@@
 language-theoretical properties of G-ACGs (Are ACG object languages closed on
-intersection? Are pangraphical languages expressible as extrinsic languages?),
-searching for a metagrammatical description of constraint signatures to ease
-work on G-ACG grammars and finally building large-scale grammars to verify the
-utility of our approach.
+intersection? Are extrinsic languages as expressive as pangraphical
+languages?), searching for a metagrammatical description of constraint
+signatures to ease work on G-ACG grammars and finally building large-scale
+grammars to verify the usability of our approach.
 

### 2013-09-22 lenls/paper.tex
LENLS: Final proofreading

@@
 grammar, we advocate for extending the formalism to prevent the incidental
-complexity that would otherwise appear.
+complexity that would otherwise emerge.
 
@@
 consider some of its formal properties. We then finish with an illustration of
-our approach which incorporates linguistic contribution from multiple sources
+our approach which incorporates linguistic contributions from multiple sources
 into a single coherent and simple grammar and examine the limitations and
@@
 languages. Given some signature and a distinguished type, we can talk of the
-\emph{abstract language} which is the set of well-typed terms built on that
-signature having that distinguished type. If we then also consider a lexicon
-and its abstract signature, then the language we get by mapping the abstract
-language of the signature through that lexicon is called an \emph{object
-  language}.
+\emph{abstract language}, which is the set of well-typed terms built on the
+signature and having the distinguished type. If we then also consider a
+lexicon and its abstract signature, then the language we get by mapping the
+abstract language of the signature through the lexicon is called an
+\emph{object language}.
 
@@
 lexicon that translates that signature to the abstract-most signature of our
-existing grammar. Usually, the new signature will resemble the prior
-abstract-most one, albeit with more fine-grained types necessary to express
-the constraint in question.
+existing grammar (Figure~\ref{fig:acg-comp-constr}). Usually, the new
+signature will resemble the prior abstract-most one, albeit with more
+fine-grained types necessary to express the constraint in question.
 
@@
   arborescences, in which the root node represents the abstract language from
-  whose terms are generated terms of all the object languages.}. This means
-that the syntactic constraint would end up being expressed by strengthening
-the type system of $SyntSem$, which only adds unnecessary complexity.
+  whose terms are generated the terms of all the object languages.}. This
+means that the syntactic constraint would end up being expressed in terms of
+the types of the syntax-semantics interface ($SyntSem$) rather than just pure
+syntax ($Syntax$), which only adds unnecessary complexity.
 
@@
 
-In the above, we elaborate the types $NP$ and $N$ with features ($_{NEG{=}F}$
-and $_{NEG{=}T}$) and we give the different types for the preposition \emph{de}
+In the above, we elaborate the types $NP$ and $N$ with features ($NEG{=}F$
+and $NEG{=}T$) and we give the different types for the preposition \emph{de}
 in the fragment for negation ($N$). The types accept an $NP$ and an $N$ as
@@
 free negative determiners (i.e. there is no $N_{de_4} : NP_{NEG{=}T} \limp
-N_{NEG{=}T} \limp N\_NEG{=}\ldots$). This encodes a constraint that there can
+N_{NEG{=}T} \limp N_{NEG{=}}\ldots$). This encodes a constraint that there can
 be only one free negative determiner per phrase (free as in not hidden inside
@@
 graph (DAG) whose nodes are labeled with signatures (and distinguished types)
-and whose edges are labeled with lexicons, in other words a mathematical
-reification of an ACG diagram that has been generalized from
-arborescences to DAGs. We then search for an appropriate semantics for these
-structures, i.e. how to determine what languages are defined by a graphical
-ACG.
+and whose edges are labeled with lexicons, in other words, a mathematical
+reification of an ACG diagram that has been generalized from arborescences to
+DAGs. We then search for an appropriate semantics for these structures,
+i.e. how to determine what languages are defined by a graphical ACG.
 
@@
 ACG abstract languages closed on transformation by a lexicon and
-intersection\footnote{Can be show by induction on the topological ordering of
+intersection\footnote{Can be shown by induction on the topological ordering of
   any given graphical ACG.}.
@@
 element in a signature, like $Sem$, and by transduction find its corresponding
-element in an another signature, like $Syntax$. However, in an G-ACG such as
-this it is possible that the $Syntax$ term we obtain by transduction does not
+element in an another signature, like $Syntax$. However, in a G-ACG such as
+this it, is possible that the $Syntax$ term we obtain by transduction does not
 belong to $\mathcal{E}_{\mathcal{G}}(Syntax)$ because it lacks antecedents in
@@
 If we compare this new interpretation of G-ACGs to the former one, we find out
-that in the case when the graph of the grammar is an arborescence they are
+that in the case when the graph of the grammar is an arborescence, they are
 actually equivalent. This means that in classical ACGs, where all the diagrams
 are arborescences, the two metaphors (nodes as languages and nodes as terms)
-can be and are used interchangeably. However, as we start to consider
+can be, and are, used interchangeably. However, as we start to consider
 non-arborescent graphs, we find, interestingly, that the two paradigms diverge
@@
 pangraphical language of the new G-ACG also serves as a proof of its presence
-in the extrinsic language of the old G-ACG.
+in the extrinsic language of the old G-ACG and vice versa.
 
@@
 $\mathcal{L}_{Agr}$ which implement the constraint on negation we discussed
-before and some simple notion of agreement, respectively.
+before and some simple notion of agreement, respectively
+(Figure~\ref{fig:final-synth}).
 
@@
 instead of having to parse a single hidden structure which enforces
-grammaticality w.r.t. every constraint. G-ACGs share these parks also with
+grammaticality w.r.t. every constraint. G-ACGs share these perks also with
 classical ACGs but they let use this composition in another dimension, which
@@
 languages?), searching for a metagrammatical description of constraint
-signatures to ease work on G-ACG grammars and finally building large-scale
+signatures to ease work on G-ACG grammars, and finally building large-scale
 grammars to verify the usability of our approach.

### 2013-09-22 lenls/paper.tex
LENLS: Final remark from Maxime

@@
 element in an another signature, like $Syntax$. However, in a G-ACG such as
-this it, is possible that the $Syntax$ term we obtain by transduction does not
+this, it is possible that the $Syntax$ term we obtain by transduction does not
 belong to $\mathcal{E}_{\mathcal{G}}(Syntax)$ because it lacks antecedents in

### 2013-10-21 lenls/slides.tex
LENLS: First usable version of slides

--- 2013-06-24 slides/presentation.tex
@@
 
+\usepackage[utf8]{inputenc}
 \usepackage{xcolor}
@@
 
-\title[G-ACGs]{Towards a Wide-Coverage Grammar}
+\title[G-ACGs]{Integration of Multiple Constraints in ACG}
 \subtitle{Graphical Abstract Categorial Grammars}
-\author{Ji\v{r}\'{i} Mar\v{s}\'{i}k}
-\institute[INRIA Loria]
-{
-  INRIA Loria
-  \and
-  UFR Math\'{e}matiques et Informatiques \\
-  Universit\'{e} de Lorraine
-}
-\date[June 2013]{June 25, 2013}
+%
+\author{Jiří Maršík \and Maxime Amblard}
+%
+\institute{Université de Lorraine, Laboratoire lorrain de recherche en informatique et ses applications, UMR 7503, Vandoeuvre-lès-Nancy, 54500, France\\
+INRIA, Villers-lès-Nancy, 54600, France\\
+CNRS, Loria, Vandoeuvre-lès-Nancy, UMR 7503, 54500, France\\
+\texttt{\{jiri.marsik, maxime.amblard\}@loria.fr}}
+
+\date[October 2013]{October 28, 2013}
 
@@
 \begin{frame}
-  \frametitle{Towards a Wide-Coverage Grammar}
-
-  \textbf{Objective:} Build discourse representations using ACGs: \\a
-  wide-coverage ACG grammar.
-
-  \vfill
+  \frametitle{Introduction}
 
   \begin{itemize}
-  \item Having linguistic resources to build upon
-    \begin{itemize}
-    \item Frilex, a formalism-independent lexical database
-    \item Frigram, a French interaction grammar
-    \end{itemize}
+    \item Abstract Categorial Grammars provide interesting tools for managing
+      discourse and dynamic semantics. \vfill
+    \item Refining grammars to enforce numerous constraints breeds complexity.
       \vfill
-  \item Debugging the grammar
-    \begin{itemize}
-    \item A toolkit for formulating ACGs as simple metagrammars, for
-      examining how they work and for checking their consistency.
-    \end{itemize}
-    \vfill
-  \item Using a formalism that allows for a simple grammar
-    \begin{itemize}
-    \item Express multiple independent constraints \\ $\implies$ graphical abstract categorial grammars.
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{ACGs and IGs}
-
-  \begin{itemize}
-  \item ACGs are based on types (IILL formulas).
-  \item IGs are based on polarized tree descriptions.
-    \vfill
-  \item[$\Rightarrow$] Perrier has linked the two.
-
-    \vspace{2 mm}
-    \begin{columns}[c]
-      \begin{column}{0.1\textwidth}\end{column}
-      \begin{column}{0.4\textwidth}
-        $$(NP \limp S) \limp N \limp N$$
-        $$N \limp (NP \limp S) \limp N$$
-      \end{column}
-      \begin{column}{0.1\textwidth}$\to$\end{column}
-      \begin{column}{0.2\textwidth}
-        \includegraphics[height=0.3\textheight]{../diagrams/ptd.pdf}
-      \end{column}
-      \begin{column}{0.25\textwidth}\end{column}
-    \end{columns}
-    \vfill
-  \item IG of Frigram use different constraints in tree descriptions.
-    \begin{itemize}
-    \item The deductive syntactic model of IGs is built on a wider range
-      of linear logic primitives ($\otimes$, $\&$).
-    \end{itemize}
+    \item We formalize an extension of ACGs which lets us use the ACG
+      apparatus and maintain independent constraints. \vfill
+    \item Our goal is a large-scale grammar usable for discussing discourse.
   \end{itemize}
@@
   \begin{block}{Abstract languages}
-    %% Languages of (abstract) terms constrained by the constants they are
-    %% built on and their types. Defined by a \textcolor{red}{signature}
-    %% and a \textcolor{darkgreen}{type}.
-
     Lambda terms constrained by the types of constants. \\ Defined by a
-    \textcolor{red}{signature} and a \textcolor{darkgreen}{type}.
+    \textcolor{red}{signature} and a \textcolor{darkgreen}{distinguished
+      type}.
 
@@
 \begin{frame}
+  \frametitle{Multiple Constraints}
+
+  Implementing multiple constraints in a single signature is
+  \textbf{difficult}.
+
+  Agreement:
+    \begin{align*}
+      A_{aime_1} &: NP_{\textcolor{darkgold}{NUM{=}SG}} \limp NP_{\textcolor{darkgold}{NUM{=}SG}} \limp S \\
+      A_{aime_2} &: NP_{\textcolor{darkgold}{NUM{=}SG}} \limp NP_{\textcolor{darkgold}{NUM{=}PL}} \limp S
+    \end{align*}
+
+  Extraction:
+    \begin{align*}
+      E_{aime_1} &: NP_{\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}NO}} \\
+      E_{aime_2} &: NP_{\textcolor{darkgreen}{VAR{=}T}} \limp NP_{\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}} \\
+      E_{aime_3} &: NP_{\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgreen}{VAR{=}T}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}
+}    \end{align*}
+
+  Combined:
+  \begin{small}
+  \begin{align*}
+      C_{aime_1} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}NO}} \\
+      C_{aime_2} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}T}} \limp NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}} \\
+      C_{aime_3} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}T}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}} \\
+      C_{aime_4} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgold}{NUM{=}PL},\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}NO}} \\
+      C_{aime_5} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}T}} \limp NP_{\textcolor{darkgold}{NUM{=}PL},\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}} \\
+      C_{aime_6} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgold}{NUM{=}PL},\textcolor{darkgreen}{VAR{=}T}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}
+}  \end{align*}
+  \end{small}
+\end{frame}
+
+\begin{frame}
   \frametitle{Graphical Abstract Categorial Grammars (G-ACGs)}
@@
 
-\section{Semantics of G-ACGs Nodes}
+\section{Semantics of G-ACG Nodes}
 
@@
   \item Intrinsic languages are just the abstract languages of ACGs.
-    \begin{itemize}
-    \item $\mathcal{I} = \mathcal{A}$
-    \end{itemize}
-  \vfill
+    $$\mathcal{I} = \mathcal{A}$$
   \item Extrinsic languages are simple extensions of ACGs that add
     intersections.
-    \begin{itemize}
-    \item $\mathcal{E} = \mathcal{O}^{\mathcal{L}{\cap}}$
-    \end{itemize}
+    $$\mathcal{O} = \mathcal{A}^{\mathcal{L}}$$
+    $$\mathcal{E} = \mathcal{A}^{\mathcal{L}{\cap}}$$
   \vfill
@@
 
+\section{Illustration}
+
+\begin{frame}
+  \frametitle{Basic Setup}
+  \begin{center}
+    \includegraphics[height=0.8\textheight]{../diagrams/illustration0.pdf}
+  \end{center}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Single Constraint}
+  \begin{center}
+    \includegraphics[height=0.8\textheight]{../diagrams/illustration1.pdf}
+  \end{center}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Multiple Constraints}
+  \begin{center}
+    \includegraphics[height=0.8\textheight]{../diagrams/illustration2.pdf}
+  \end{center}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Exposing Another Signature}
+  \begin{center}
+    \includegraphics[height=0.8\textheight]{../diagrams/illustration3.pdf}
+  \end{center}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Constraints on Multiple Signatures}
+  \begin{center}
+    \includegraphics[height=0.8\textheight]{../diagrams/illustration4.pdf}
+  \end{center}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Reflection}
+  \begin{columns}[c]
+    \begin{column}{0.5\textwidth}
+      \begin{center}
+        \includegraphics[height=0.8\textheight]{../diagrams/illustration5.pdf}
+      \end{center}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \begin{itemize}
+        \item 9 signatures
+        \item 8 lexicons
+        \item Constraints decomplected from each other
+        \item Constraints on different signatures
+        \item High level of redundancy
+      \end{itemize}
+    \end{column}
+  \end{columns}
+\end{frame}
+
 \section{Conclusion}
@@
 \begin{frame}
-  \frametitle{Summary of Work Done}
+  \frametitle{Recapitulation}
 
   \begin{itemize}
-  \item Direct embedding of IGs into ACGs based on Perrier's theorem
-    conjectured not to be practical.
+  \item We reinterpreted ACGs by treating the diagrams as primary constructs
+    instead of as artifacts.
   \vfill
-  \item Created a toolkit to serve as basis for more explorations or
-    implementations.
+  \pause
+  \item This lets us handle non-arborescent diagrams in a principled manner.
   \vfill
-  \item Formalized graphical abstract categorial grammars.
-    \begin{itemize}
-    \item Helps solve multiple constraints elegantly.
-    \item Facilitates discussion and inquiry into compositions of ACGs.
-    \end{itemize}
+  \pause
+  \item Non-arborescence enables intersection and therefore conjunction of
+    independent constraints.
+  \vfill
+  \pause
+  \item We presented the kind of grammar we envisioned, built by collecting
+    fragments described in existing research.
   \end{itemize}
@@
     intersection? $$\mathcal{O}^{\mathcal{L}{\cap}} \subseteq \mathcal{O}^{\mathcal{L}}$$
-  \item Express the multiple polarized features of IGs in G-ACGs.
+  \item Is there a metagrammatical approach that makes refining the types of a
+    signature more concise?
   \vfill
-  \item Build and study a wide-coverage G-ACG.
-    \vfill
-  \item Apply G-ACGs to discourse.
+  \item Is this enough to build a satisfactory grammar of a real language?
   \end{itemize}
@@
 
-\appendix
-\newcounter{finalframe}
-\setcounter{finalframe}{\value{framenumber}}
-
-\begin{frame}
-  \frametitle{Multiple Constraints}
-
-  Implementing multiple constraints in a single signature is
-  \textbf{difficult}.
-
-  Agreement:
-    \begin{align*}
-      A_{aime_1} &: NP\_\textcolor{darkgold}{NUM{=}SG} \limp NP\_\textcolor{darkgold}{NUM{=}SG} \limp S \\
-      A_{aime_2} &: NP\_\textcolor{darkgold}{NUM{=}SG} \limp NP\_\textcolor{darkgold}{NUM{=}PL} \limp S
-    \end{align*}
-
-  Extraction:
-    \begin{align*}
-      E_{aime_1} &: NP\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
-      E_{aime_2} &: NP\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
-      E_{aime_3} &: NP\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT}
-    \end{align*}
-
-  Combined:
-  \begin{small}
-  \begin{align*}
-      C_{aime_1} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
-      C_{aime_2} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
-      C_{aime_3} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
-      C_{aime_4} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
-      C_{aime_5} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
-      C_{aime_6} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT}
-  \end{align*}
-  \end{small}
-\end{frame}
-
-\setcounter{framenumber}{\value{finalframe}}
-
 \end{document}

### 2013-12-25 lenls/slides.tex
LENLS slides, backlog commit

@@
 \usepackage{xcolor}
+\usepackage{pgfpages}
 
@@
 
+\setbeameroption{show notes on second screen=right}
+%\setbeamertemplate{note page}[plain]
+
 \begin{document}
@@
 CNRS, Loria, Vandoeuvre-lès-Nancy, UMR 7503, 54500, France\\
+$\,$  \\
 \texttt{\{jiri.marsik, maxime.amblard\}@loria.fr}}
@@
   \begin{itemize}
-    \item Abstract Categorial Grammars provide interesting tools for managing
-      discourse and dynamic semantics. \vfill
-    \item Refining grammars to enforce numerous constraints breeds complexity.
-      \vfill
-    \item We formalize an extension of ACGs which lets us use the ACG
-      apparatus and maintain independent constraints. \vfill
-    \item Our goal is a large-scale grammar usable for discussing discourse.
+    \item Long-term goal: large-scale grammar usable for discourse \vfill
+    \item Abstract Categorial Grammars: a tool for discourse and dynamic
+      semantics \vfill
+    \item Refining grammars to enforce numerous constraints \\ $\Rightarrow$
+      complexity \vfill
+    \item Extension of ACGs for independent constraints
   \end{itemize}
@@
 
+\note[itemize]{\item In our team, we are interested in formal treatments of
+  discourse-level phenomena.
+
+\item Our tool of choice is the formalism of Abstract Categorial
+  Grammars. ACGs provide us with interesting tools to handle discourse and
+  dynamic semantics with continuations allowing us to capture dynamic meaning
+  and the generality of the lambda calculus allowing us to introduce our own
+  constructs such as discourse structures.
+
+\item However, during our work when trying to use ACGs, we observed that it
+  proved difficult to try to enforce multiple independent constraints within a
+  single grammar without the grammar becoming too complex\ldots
+
+\item \ldots and it is this inadequacy that prompted the extension of the
+  formalism that is the focus of our contribution.
+
+\item 0:34}
+
 \section{ACG Preliminaries}
+
+\note{I will now give a short introduction to the ACG formalism.}
 
@@
   \begin{block}{Abstract languages}
-    Lambda terms constrained by the types of constants. \\ Defined by a
-    \textcolor{red}{signature} and a \textcolor{darkgreen}{distinguished
-      type}.
+    Defined by a \textcolor{red}{signature} and a
+    \textcolor{darkgreen}{distinguished type}.
 
@@
   \begin{block}{Object languages}
-    Languages of (object) terms which are obtained by replacing the
-    constants in an existing (abstract) term using a \textcolor{darkgold}{lexicon}.
+    A projection of an abstract language with a \textcolor{darkgold}{lexicon}.
 
@@
 
+\note[itemize]{\item Abstract Categorial Grammars introduce two families of
+  languages: the so-called abstract languages and object languages.
+
+\item Abstract languages are sets of lambda terms built from some signature of
+  typed constants where every term of the language is well-typed and has a
+  specific distinguished type. Here we have an example signature giving some
+  familiar types to French words and a term of type S which corresponds to the
+  structure of a French sentence.
+
+\item Object languages are then derived from abstract languages by replacing
+  all the constants in their terms with specific substitutes given by a
+  lexicon. Here we have such a lexicon that when applied to the example
+  sentence structure, it will yield a term that beta-reduces to the French
+  sentence ``Laura aime les tatous''.
+
+\item An ACG is then usually represented as a quadruple consisting of an
+  abstract signature, a distinguished abstract type, an object signature and a
+  lexicon connecting the two signatures.
+
+\item 0:35}
+
 \begin{frame}
@@
 
-  \begin{columns}[c]
-    \begin{column}{0.5\textwidth}
-      \begin{itemize}
-      \item Two different lexicons interpret the same abstract terms.
-      \item Instead of referring to the two ACGs $\mathopen{<}
-        \Sigma_{Synt}, \Sigma_{String}, \mathcal{L}_{syntax}, S
-        \mathclose{>}$ and $\mathopen{<} \Sigma_{Synt}, \Sigma_{Sem},
-        \mathcal{L}_{sem}, S \mathclose{>}$, refer to this single
-        diagram.
-      \end{itemize}
-    \end{column}
-    \begin{column}{0.5\textwidth}
-      \includegraphics[width=\textwidth]{../diagrams/double-acg.pdf}
-    \end{column}
-  \end{columns}
+  \begin{center}
+    \includegraphics[height=0.8\textheight]{../diagrams/double-acg.pdf}
+  \end{center}
 \end{frame}
+
+\note[itemize]{\item Outside of language-theoretical explorations, considering
+  single ACGs is seldom useful. In this typical example, we have two ACGs that
+  share a single abstract language of sentence structures that is mapped via
+  two different lexicons to phonological strings and semantic formulae.
+
+\item Instead of talking about two quadruples that share some elements, we
+  will prefer the more eloquent diagrams.
+
+\item 0:20}
 
@@
      \begin{column}{0.5\textwidth}
-      \begin{block}{Semantic ambiguities}
+      \begin{block}{Constraint}
         \vspace{2 mm}
         \begin{center}
-          \includegraphics[height=0.7\textheight]{../diagrams/parallel-over-serial.pdf}
+          \includegraphics[height=0.7\textheight]{../diagrams/serial-over-parallel-slides.pdf}
         \end{center}
@@
     \begin{column}{0.5\textwidth}
-      \begin{block}{Syntactic constraint}
+      \begin{block}{Elaboration}
         \vspace{2 mm}
         \begin{center}
-          \includegraphics[height=0.7\textheight]{../diagrams/serial-over-parallel.pdf}
+          \includegraphics[height=0.7\textheight]{../diagrams/parallel-over-serial-slides.pdf}
         \end{center}
@@
 
+\note[itemize]{\item Besides the diagram we just saw, there are other
+  established patterns in ACGs.
+
+\item The Constraint pattern lets us retroactively restrain the members of
+  previously defined languages. Whereas before, any sentence structure that
+  had the distinguished type was considered valid, with the added constraint,
+  we only consider those that can be generated from the newly added signature
+  and lexicon. The new signature usually mimics the old one but refines its
+  types so that certain terms will no longer be admissible.
+
+\item The Elaboration pattern serves a different purpose. It lets us decompose
+  an involved translation into two smaller translations and expose the
+  intermediate product. In this case, we split the signature of sentence
+  structures into a more superficial analytic layer which accounts only for
+  syntactic ambiguities and a deeper tectogrammatical layer which goes
+  further, into the semantic ambiguities of scope.
+
+\item 0:45}
+
 \section{Why Graphical ACGs?}
+
+\note{Now that we are familiar with ACGs, I will proceed to show the problems
+  that motivated our proposed extension.}
 
@@
 
-  ACG diagrams are arborescences.
-
   \begin{columns}[c]
-    \begin{column}{0.25\textwidth}
-      {\small Syntax-semantics interface deals with syntactic
-        constraints.}
-
-      \onslide<2->{$\implies$ Complex.}
-    \end{column}
-    \begin{column}{0.25\textwidth}
+    \begin{column}{0.5\textwidth}
       \begin{center}
@@
     \end{column}
-    \begin{column}{0.25\textwidth}
-      {\small Syntactic constraints deal with the syntax-semantics
-        interface.}
-
-      \onslide<3->{$\implies$ Complex.}
-    \end{column}
-    \begin{column}{0.25\textwidth}
+    \begin{column}{0.5\textwidth}
       \begin{center}
@@
 
+\note[itemize]{{\scriptsize
+\item Our problems revolve around the fact that all the diagrams drawn for
+  ACGs are arborescences. There is always a single abstract language and all
+  the other languages are established by projecting that abstract language
+  with some composition of lexicons. This limits us when we try to combine the
+  two patterns we just saw in a single grammar.
+
+\item Consider the situation where we want to distinguish the analytic syntax
+  and tectogrammatical levels of sentence structure, while adding a constraint
+  on the analytic level. One way we can try to do this is to have the
+  constraint signature constrain the analytic syntax level directly. But then
+  the tectogrammatical level, which provides the link between syntax and
+  semantics, has to deal with the constraint signature by having to generate
+  correct members of its language. Thus we have an unwarranted interaction
+  between internal details of syntax and the syntax-semantics interface.
+
+}{\scriptsize
+\item Another solution would be to keep the syntax-semantics interface pure by
+  having it still connect just the analytic syntax with the
+  semantics. However, then the constraint has to be implemented on top of all
+  that, meaning that the constraint now has to deal with matters of
+  tectogrammatics and the syntax-semantics interface. Again, we arrive at
+  accidental complexity, this time with the syntax-semantics interface
+  contaminating the syntactic constraint.
+
+\item 0:54}}
+
 \begin{frame}
@@
 
-  Writing a simple wide-coverage grammar implies specifying constraints
-  independently.
-
   \begin{columns}[c]
-    \begin{column}{0.25\textwidth}
-      Second constraint reimplements first constraint.
-
-      \onslide<2->{$\implies$ Ineffective.}
-    \end{column}
-    \begin{column}{0.25\textwidth}
+    \begin{column}{0.5\textwidth}
       \begin{center}
@@
     \end{column}
-    \begin{column}{0.25\textwidth}
-      Third constraint reimplements first and second constraint.
-
-      \onslide<3->{$\implies$ Ineffective.}
-    \end{column}
-    \begin{column}{0.25\textwidth}
+    \begin{column}{0.5\textwidth}
       \begin{center}
@@
 
+\note[itemize]{{\footnotesize \item When writing a wide-coverage grammar, we
+    will be especially interested in implementing many different constraints
+    and we would like to try to factor them out, maybe by using the Constraint
+    pattern. However, this proves to be difficult thanks to the homomorphism
+    property of lexicons, which lies at the heart of ACGs. The consequence of
+    the homomorphism property that's interesting to us is that the image of a
+    well-typed term with a lexicon must be another well-typed term.
+
+\item This means that when we try to attach a second constraint in the example
+  on the left, its types have to essentially ensure that all the licensed
+  structures not only satisfy the new constraint, but also the existing one.
+
+\item This problem is exacerbated by iteration. When adding a third
+  constraint, we are obliged to engineer its signature such that the licensed
+  terms conform to all three constraints. In the end, there is not much
+  redeeming value to stacking constraints like this and we end up implementing
+  multiple constraints in a single signature which, as we'll see, can get very
+  complex.
+
+\item 0:34}}
+
 \begin{frame}
   \frametitle{Multiple Constraints}
-
-  Implementing multiple constraints in a single signature is
-  \textbf{difficult}.
-
   Agreement:
@@
 
+\note[itemize]{\item Here we have an example of trying to combine the
+  treatment of two independent constraints in a single signature. In the first
+  fragment, we see a few types refined to handle agreement in number. In the
+  second, we have the same wordform in a refinement which enforces a
+  constraint on extraction and relative clauses. Finally, in the third block,
+  we have a slice of a signature which enforces both constraints at the same
+  time.
+
+\item While the individual constraint signatures are reasonably
+  understandable, the combined signature grows exponentially with the number
+  of features and it becomes difficult to see whether constraints are
+  independent of each other or not.
+
+\item 0:29}
+
 \begin{frame}
@@
       \begin{center}
-        \includegraphics[height=0.7\textheight]{../diagrams/final.pdf}
+        \includegraphics[height=0.7\textheight]{../diagrams/final-slides.pdf}
       \end{center}
@@
 
-    \pause
-
     \begin{column}{0.4\textwidth}
         \begin{itemize}
-        \item A DAG.
-        \item Nodes labelled with:
+        \item DAG
+        \item Nodes:
           \begin{itemize}
-          \item Signatures.
-          \item Distinguished types.
+          \item Signatures
+          \item Distinguished types
           \end{itemize}
-        \item Edges labelled with:
+        \item Edges:
           \begin{itemize}
-          \item Lexicons.
+          \item Lexicons
           \end{itemize}
@@
 
-\section{Semantics of G-ACG Nodes}
+\note[itemize]{\item The framework that we would like to work in would ideally
+  let us build a grammar like this. Individual constraints are laid
+  side-by-side without any cumulative complexity and outside of the scope of
+  the syntax-semantics interface. With no directed paths leading between the
+  constraint or tectogrammatics signatures, there are no direct dependencies
+  between signatures which handle different issues in our grammar.
+
+\item What we have here is the basic notation of our Graphical Abstract
+  Categorial Grammars: a directed acyclic graph whose nodes are labelled with
+  signatures and distinguished types and whose edges are labelled with
+  lexicons. Having introduced the notation, we are now obliged to answer the
+  question of how this notation establishes languages.
+
+\item 0:36}
+
+\section{Semantics of G-ACG}
 
@@
     \begin{itemize}
-    \item Nodes as \emph{languages}.
-    \item Edges define \emph{languages} in terms of other
-      \emph{languages} using lexicons and intersections.
+    \item Nodes as \emph{languages}
     \end{itemize}
@@
     \begin{itemize}
-    \item Nodes as \emph{terms}.
-    \item Edges describe how \emph{terms} are obtained from other
-      \emph{terms} by applying lexicons.
+    \item Nodes as \emph{terms}
     \end{itemize}
@@
 \end{frame}
+
+\note[itemize]{\item We will present two interpretations of the G-ACG
+  notation.
+
+\item In the first, which we dubbed ``algebraic'', nodes are identified with
+  languages and edges tell us how languages are defined in terms of each other
+  using lexicons and intersections
+
+\item In the second interpretation, the ``generative'' one, nodes are seen as
+  standing for individual terms and edges are seen as statements that applying
+  a particular lexicon to one term produces another. The languages are then
+  established through the notion of a model.
+
+\item 0:31}
 
@@
         \begin{itemize}
-        \item Every node is a language.
-        \item An arrow means that a language is produced from another by
-          mapping through a \textcolor{darkgreen}{lexicon}.
-        \item Two arrows converging on a node mean an
-          \textcolor{red}{intersection} of languages.
+        \item Node = Language
+        \item Edge = Projection with a \textcolor{darkgreen}{lexicon}
+        \item Two incoming edges = \textcolor{red}{Intersection} of languages
         \end{itemize}
@@
 
+\note[itemize]{\item In the algebraic paradigm, we associate a language with
+  every node of the G-ACG. A single incoming edge tells us that the target
+  language is the projection of the source language with the edge's
+  lexicon. When multiple incoming edges meet, we take the intersection of the
+  projections.
+
+\item Formally, this is realized by establishing intrinsic languages for every
+  node which correspond exactly to the abstract languages of ACGs; they are
+  simply sets of terms having some distinguished type. This then helps us
+  inductively build the extrinsic languages, which are defined as the
+  intersection of the lexical projections of the node's predecessors'
+  extrinsic languages, or the node's intrinsic language if it has no
+  predecessors.
+
+\item 0:43}
+
 \begin{frame}
@@
       \begin{center}
-        \includegraphics[height=0.6\textheight]{../diagrams/both-patterns.pdf}
+        \includegraphics[height=0.6\textheight]{../diagrams/both-patterns-slides.pdf}
       \end{center}
@@
       How do we express the language of $Bottom$ terms whose $Left$ and
-      $Right$ antecedent are generated by the same $Top$ antecedent?
+      $Right$ antecedents are generated by the same $Top$ antecedent?
     \end{column}
@@
 \end{frame}
+
+\note[itemize]{{\scriptsize \item The algebraic interpretation that we
+    developed can be counter-intuitive in certain situations. First, consider
+    the combination of a syntactic constraint that is independent of the
+    syntax-semantics interface, as can be seen in the diagram on the
+    left. When we take a semantic formula belonging to the extrinsic language
+    of the node $Sem$, we can take its antecedent in the syntax-semantics
+    interface $SyntSem$ and project it to find the syntactic counterparts of
+    our semantic formula. However, we have no guarantee that these
+    counterparts will satisfy the syntactic constraint, in other words, they
+    might lack an antecedent in the $Constr$ signature. Thus, it becomes
+    difficult to establish the language of meanings which are expressible only
+    by syntactically correct sentences since the extrinsic language of $Sem$
+    is not constrained by the $Constr$ signature and thus overgenerates in
+    this case.
+
+\item Another counter-intuitive example is the diamond grammar on the
+  right. If we take a member of the $Bottom$ extrinsic language, it must have,
+  by definition, antecedents in the $Left$ and $Right$ extrinsic
+  languages. These two antecedents then must have their own antecedents in the
+  $Top$ extrinsic language. However, they don't have to be the same
+  antecedent as the single shared node on the diagram might suggest.
+
+\item 0:58}}
 
@@
         \begin{itemize}
-        \item Every node is a \textcolor{red}{term}.
-        \item An arrow means that a term is produced from another by
-          applying a \textcolor{darkgreen}{lexicon}.
-        \item Two arrows converging on a node mean that the two (abstract)
-          terms yield \textcolor{red}{the same} (object) term.
+        \item Node = \textcolor{red}{Term}
+        \item Edge = Image with a \textcolor{darkgreen}{lexicon}
+        \item Two incoming edges = \textcolor{red}{Equality} of images
         \end{itemize}
@@
 
-      $t$ belongs to $\mathcal{P}_{\mathcal{G}}(u)$ iff \\$\exists T :
-      \textcolor{red}{V(G) \to \Lambda}$ such that:
+      $t \in \mathcal{P}_{\mathcal{G}}(u)$ iff $\exists T :
+      \textcolor{red}{V(G) \to \Lambda}$ s.t.:
       \begin{itemize}
       \item $T_u = t$.
-      \item For all $v \in V(G)$, $\vdash_{\Sigma_v} T_v : S_v$.
-      \item For all $(v,w) \in E(G)$,
+      \item $\forall v \in V(G)$, $\vdash_{\Sigma_v} T_v : S_v$.
+      \item $\forall (v,w) \in E(G)$,
         $\textcolor{darkgreen}{\mathcal{L}_{(v,w)}(T_v) = T_w}$.
@@
 
+\note[itemize]{\item We now turn to another interpretation of G-ACGs, one that
+  we called the ``generative'' interpretation, though ``logical'' or
+  ``relational'' are also apt names.
+
+\item The nodes of the graph stand for terms, that is, they are term variables
+  that range over the nodes' intrinsic languages. An edge is a statement
+  saying that the pair of terms must belong to the edge's lexicon, which is to
+  say that applying the lexicon to the source term must yield the target term.
+
+\item We then consider a model of this graphical formula to be a labelling
+  which assigns terms to the vertices such that both the domain restrictions
+  and the edge statements are satisfied. From this we establish the
+  pangraphical language of a node as the set of all the possible terms that
+  can label the node in a valid model.
+
+\item 0:44}
+
 \begin{frame}
   \frametitle{Algebraic and Generative Relations}
+
+  \note{We have presented two interpretations of G-ACGs based on two different
+    intuitive readings of ACG diagrams, one where nodes are thought of as
+    languages and one where they stand for individual terms. The result shown
+    here validates these two ways of reasoning about diagrams in the context
+    of classical ACGs since they lead to the same languages as the canonical
+    definition of an object language.
+
+    0:25}
 
@@
   \pause
-  \item However, $\exists \, \mathcal{G}, u$ such that
+  \item However, $\exists \, \mathcal{G}, u$ such that:
     $$\mathcal{E}_{\mathcal{G}}(u) \neq \mathcal{P}_{\mathcal{G}}(u)$$
-    \vfill
-    \pause
-  \item[$\Rightarrow$] They are not interchangeable points of view.
   \end{itemize}
@@
 
+\note{However, as soon as we turn to G-ACGs which are not arborescent, the
+  difference between these two perspectives becomes substantial since the two
+  definitions we elaborated on before lead to different languages.
+
+  0:14}
+
 \begin{frame}
-  \frametitle{Properties of G-ACG languages}
+  \frametitle{Properties of G-ACG Languages}
 
-  \begin{itemize}
-  \item The three language families we defined above are\ldots
-    \begin{itemize}
-    \item {\ldots}increasingly constraining, $\forall \, \mathcal{G},
-      u. \; \mathcal{I}_{\mathcal{G}}(u) \supseteq
-      \mathcal{E}_{\mathcal{G}}(u) \supseteq
-      \mathcal{P}_{\mathcal{G}}(u)$.
-    \item {\ldots}increasingly expressive, $\mathcal{I} \subseteq
-      \mathcal{E} \subseteq \mathcal{P}$.
-    \end{itemize}
-  \vfill
-  \item Intrinsic languages are just the abstract languages of ACGs.
+  $$\forall \, \mathcal{G}, u. \; \mathcal{I}_{\mathcal{G}}(u) \supseteq
+  \mathcal{E}_{\mathcal{G}}(u) \supseteq \mathcal{P}_{\mathcal{G}}(u)$$
+
+  $$\mathcal{I} \subseteq \mathcal{E} \subseteq \mathcal{P}$$
+
   $$\mathcal{I} = \mathcal{A}$$
-  \item Extrinsic languages are simple extensions of ACGs that add
-    intersections.
+
   $$\mathcal{O} = \mathcal{A}^{\mathcal{L}}$$
+
   $$\mathcal{E} = \mathcal{A}^{\mathcal{L}{\cap}}$$
-  \vfill
-  \item Pangraphical languages are faithful to the generative
-    perspective and subsume the previous classes of languages.
-    \begin{itemize}
-    \item They solve both the example problems we saw before
-      (\emph{meanings of grammatical sentences} and the \emph{diamond}).
-    \end{itemize}
-  \end{itemize}
 \end{frame}
 
+\note[itemize]{{\scriptsize \item We now proceed to give some basic properties
+    of the language families we have just introduced.
+
+\item The three language families that we have just defined (intrinsic,
+  extrinsic and pangraphical) were increasingly restrictive. The intrinsic
+  languages were constrained only by the node's signature and distinguished
+  type, whereas the extrinsic languages were constrained also by the node's
+  predecessors and the pangraphical languages were constrained by the entire
+  graph.
+
+\item We can prove that any extrinsic language is also pangraphical and
+  therefore we get that the three language families we introduced were also
+  increasingly expressive.
+
+\item Intrinsic languages were useful to establish extrinsic and pangraphical
+  languages. They are exactly the same thing as the abstract languages of
+  ACGs.
+
+\item Extrinsic languages, even though they can give counter-intuitive
+  interpretations in some cases, are still very interesting. It can be shown
+  that they extend the expressive power of object languages precisely by
+  introducing closure on intersection and nothing more.
+
+}{\tiny
+\item We don't have a similar result for pangraphical languages but we find
+  them useful since they more closely capture our intuitions about ACG
+  diagrams and so, for example, they resolve the two problems we saw with
+  interpreting G-ACGs as extrinsic languages.
+
+\item 1:01}}
+
 \section{Illustration}
+
+\note{We now illustrate the kind of grammar for which we introduced this
+  extension.}
 
@@
 
+\note{We start with the canonical example of a tectogrammatical language that
+  is projected into phenogrammatical strings and semantic formulae.
+
+  0:09}
+
 \begin{frame}
+  \addtocounter{framenumber}{-1}
   \frametitle{Single Constraint}
@@
 
+\note{We then add one of the three constraints on extraction developed by
+  Sylvain Pogodalla and Florent Pompigne in their 2010 paper.
+
+  0:08}
+
 \begin{frame}
+  \addtocounter{framenumber}{-1}
   \frametitle{Multiple Constraints}
@@
 
+\note{Then we add the other two constraints the same way. It is here that we
+  leave the scope of what classical ACGs let us do.
+
+  0:09}
+
 \begin{frame}
+  \addtocounter{framenumber}{-1}
   \frametitle{Exposing Another Signature}
@@
 
+\note{Next, we proceed by elaborating the syntactic lexicon, again by
+  literally copying existing work from one of Sylvain Pogodalla's papers. This
+  exposes a new signature of lower-order syntactic structures which are more
+  comfortable to work with.
+
+  0:14}
+
 \begin{frame}
+  \addtocounter{framenumber}{-1}
   \frametitle{Constraints on Multiple Signatures}
@@
 \end{frame}
+
+\note{We then exploit this new syntactic signature by adding constraints. We
+  refine some types to simulate Guy Perrier's treatment of negative
+  determiners in French. We also add a constraint for enforcing agreement of
+  number.
+
+  0:13}
 
@@
         \item 8 lexicons
+        \item High level of redundancy
         \item Constraints decomplected from each other
         \item Constraints on different signatures
-        \item High level of redundancy
       \end{itemize}
@@
 \end{frame}
+
+\note{One thing to remark about this grammar is its size. Having a different
+  signature for each constraint means that there is quite a lot of redundancy
+  since we are repeatedly refining the same signature in different ways. On
+  the other hand, we have segregated all of our constraints into individual
+  signatures, keeping them simple. Furthermore, the constraints are attached
+  to different nodes of the G-ACG, making the introduction of new constraints
+  flexible in how deeply they want to get involved with the rest of the
+  grammar.
+
+  0:21}
 
@@
   \begin{itemize}
-  \item We reinterpreted ACGs by treating the diagrams as primary constructs
-    instead of as artifacts.
+  \item ACG diagrams as mathematical objects
   \vfill
-  \pause
-  \item This lets us handle non-arborescent diagrams in a principled manner.
+  \item Non-arborescence enables intersection
   \vfill
-  \pause
-  \item Non-arborescence enables intersection and therefore conjunction of
-    independent constraints.
+  \item Intersection enables conjunction of independent constraints
   \vfill
-  \pause
-  \item We presented the kind of grammar we envisioned, built by collecting
-    fragments described in existing research.
+  \item Large-scale grammar: collection of fragments
   \end{itemize}
 \end{frame}
+
+\note[itemize]{\item We started by reinterpreting ACGs by treating their
+  diagrams as primary defining constructs instead of as artifacts used only
+  for illustration.
+
+\item This let us explore the semantics of non-arborescent diagrams
+  directly. Through our explorations, we have arrived at a formalism which
+  facilitates intersection.
+
+\item Intersection turned out to be a key feature for the arbitrary
+  conjunction of independent constraints, which is useful in crafting detailed
+  grammars.
+
+\item We illustrated our approach by building a grammar which enforces a small
+  but diverse range of constraints and is broken up into manageable
+  chunks. This modularity was showcased by assembling the grammar in large
+  part by simply copying existing work without any modification.
+
+\item 0:35}
 
@@
   \begin{itemize}
+  \item Are ACG object languages closed on
+    intersection? $$\mathcal{O}^{\mathcal{L}{\cap}} \subseteq \mathcal{O}^{\mathcal{L}}$$
   \item Are extrinsic languages as expressive as pangraphical
     ones? $$\mathcal{P} \subseteq \mathcal{E}$$
-  \item Are ACG object languages closed on
-    intersection? $$\mathcal{O}^{\mathcal{L}{\cap}} \subseteq \mathcal{O}^{\mathcal{L}}$$
   \item Is there a metagrammatical approach that makes refining the types of a
@@
 
+\note[itemize]{{\scriptsize \item Are ACG object languages closed on
+    intersection? If this were true, then our proposal would be just a
+    conservative extension which makes conjunction of constraints easier. If
+    this were false, then our proposal is necessary to make the languages
+    definable by the formalism closed on intersection, which we see as
+    critical for conjoining independent constraints.
+
+\item We like that we can exactly describe the set of extrinsic languages
+  w.r.t. ACG object languages. However, we still prefer pangraphical languages
+  as we find them more intuitive and easier to understand. We already know
+  that extrinsic languages are pangraphical. Proving the converse would be one
+  way of characterizing the set of pangraphical languages.
+
+\item A constraint signature that is created by refining the types shares a
+  lot of similarities with the original signature. Likewise, constraint
+  signatures developed using this techniques are alike. This makes us think
+  that it should be possible to find some kind of meta-notation that will be
+  able to convey new constraints more concisely than refining an entire
+  signature by hand.
+
+\item Finally, we are also curious to see how far we can get in building G-ACG
+  grammars before we run into another problem that will warrant extending the
+  formalism in yet another direction.
+
+\item 1:07}}
+
 \end{document}

### 2013-12-25 sjtu/body/abstract.tex
SJTU thesis version

%%==================================================
%% abstract.tex for SJTU Master Thesis
%% based on CASthesis
%% modified by wei.jianwen@gmail.com
%% version: 0.3a
%% Encoding: UTF-8
%% last update: Dec 5th, 2010
%%==================================================

\begin{abstract}

本研究的最终目的在于创建一个具有广覆盖面的抽象范畴语法（Abstract Category Grammar - ACG）。该语法可以用于自动构建自然语言段落层的表达式。为达到这一目标，我们成功地对创建广覆盖面抽象范畴语法的必要基础条件进行了实现。

在该论文中，我们首先调查了现存的语言资源，尤其是基于交互语法的Frigram以及它的语料库：Frilex，并评估了其作为构建广覆盖面抽象范畴语法的实用性。然后，我们介绍了对于范畴语法功能的具体实现，这项实现可以使用经Frilex词汇化了的语法进行测试。再后，我们考虑尝试使用单一的抽象范畴语法来整合不同语言限定条件，并提出了之前系统框架的一个推广，即图形化抽样范畴语法。最后，我们探索了一些图形化抽样范畴语法的形式特性，以此作为论文的总结。

  \keywords{\large 抽样范畴语法 \quad 语法工程 \quad 交互语法 \quad 元语法
    \quad 关系化编程 \quad 语法形式 \quad 形式语法 \quad 计算语言 \quad Lambda
    算子 \quad 类型论}
\end{abstract}

\begin{englishabstract}

We present work whose ultimate goal is the creation of a wide-coverage
abstract categorial grammar (ACG) that could be used to automatically build
discourse-level representations. In our work, we advance towards that goal by
laying down the foundations necessary for building wide-coverage ACGs.

We first examine existing language resources, in particular the Frigram
interaction grammar and its lexicon Frilex, and assess their utility to
building a wide-coverage ACG. We then present our implementation of the ACG
machinery which allows us to experiment with grammars lexicalized by
Frilex. Finally, we consider the challenge of integrating the treatment of
disparate linguistic constraints in a single ACG and propose a generalization
of the formalism: graphical abstract categorial grammars. The report concludes
with an exploration of some of the formal properties of graphical ACGs.

  \englishkeywords{\large Abstract Categorial Grammars (ACG), Grammar
    Engineering, Interaction Grammars, Metagrammars, Relational Programming,
    Grammatical Formalisms, Formal Grammars, Computational Linguistics, Lambda
    Calculus, Type Theory}
\end{englishabstract}

### 2013-12-25 sjtu/body/ack.tex
SJTU thesis version

%%==================================================
%% thanks.tex for SJTU Master Thesis
%% based on CASthesis
%% modified by wei.jianwen@gmail.com
%% version: 0.3a
%% Encoding: UTF-8
%% last update: Dec 5th, 2010
%%==================================================

\begin{thanks}

  I thank the European Union and the Erasmus Mundus Language and Communication
  Technologies Master program for funding my studies.

  I thank my supervisors, Maxime Amblard in Nancy and Zhao Hai in Shanghai,
  for their leadership and assistance.

  I thank my colleague Sai Qian for helping me with the Chinese translations.

  I thank my friends in Shanghai and Nancy for two wonderful years.

  I thank my parents in supporting me in everything I chose to do.

  Thank you!

\end{thanks}

### 2013-12-25 sjtu/constraints.tex
SJTU thesis version

--- 2013-06-26 constraints.tex
@@
 \chapter{Treatment of Multiple Linguistic Constraints}
+\chaptermark{Multiple Constraints}
 \label{chap:constraints}
@@
 A_{dorment} &: NP\_NUM{=}PL \limp S \\
+\end{align*}
+\begin{align*}
 A_{aime_1} &: NP\_NUM{=}SG \limp NP\_NUM{=}SG \limp S \\
 A_{aime_2} &: NP\_NUM{=}SG \limp NP\_NUM{=}PL \limp S \\
-\end{align*}
-\begin{align*}
 A_{aiment_1} &: NP\_NUM{=}PL \limp NP\_NUM{=}SG \limp S \\
@@
 
+{\centering
+{\small
 \begin{align*}
@@
 (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
-C_{de_{11}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
-(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
-C_{de_{12}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
-(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
+C_{de_{11}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
+C_{de_{12}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
 \end{align*}
+}
+}
 
@@
 $$
+\vspace{-1cm}
+{\small
 \begin{align*}
@@
 \end{align*}
+}
 
@@
 \begin{align*}
-C_{de(\textcolor{blue}{x}, \textcolor{blue}{y}, \textcolor{red}{p}, \textcolor{red}{q})} &: (NP\_\textcolor{red}{NEG{=}p}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}q}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}(p \lor q)}\_\textcolor{blue}{NUM{=}y})
+C_{de(\textcolor{blue}{x}, \textcolor{blue}{y}, \textcolor{red}{p},
+  \textcolor{red}{q})} :
+(NP\_\textcolor{red}{NEG{=}p}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x})
+&\limp (N\_\textcolor{red}{NEG{=}q}\_\textcolor{blue}{NUM{=}y}) \\
+&\limp (N\_\textcolor{red}{NEG{=}(p \lor q)}\_\textcolor{blue}{NUM{=}y})
 \end{align*}

### 2013-12-25 sjtu/temptest.tex
SJTU thesis version

%========================================
%% Minimun Test File
%% Encoding: UTF-8
%% HOW TO RUN: xelatex test.tex
%% 正确的PDF编译结果：
% 显示4幅图像，分别是EPS，PDF，PNG，JPG格式
% 显示"中文测试"和英文内容
% 嵌入AdobeSongStd和NimbusRomNo9字体（通过PDF阅读器的“属性”查看）
% 英文字符正确连字(ligature)
% textsc正确工作
%========================================

\documentclass[12pt,a4paper,UTF8,adobefonts]{ctexbook}

% Font Setting
\usepackage{fontspec}
\RequirePackage{metalogo}
\setmainfont{TeXGyreTermes} % an SC-capable OTF font that looks like Times
\renewcommand\labelitemi{\ensuremath{\bullet}} % 原来定义为 \textbullet

% Math Tools
\RequirePackage{mathtools, amsfonts, mathrsfs}

% Graph
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{ccaption}
% 设置图形文件的搜索路径
\graphicspath{{figure/}{figures/}{logo/}{logos/}{graph/}{graphs}}
% 如果插入的图片没有指定扩展名，那么依次搜索下面的扩展名所对应的文件
\DeclareGraphicsExtensions{.pdf,.eps,.png,.jpg,.jpeg}
% 重定义\nobreakspace命令
\DeclareRobustCommand\nobreakspace{\leavevmode\nobreak\ }

\begin{document}

中文测试 TeX Gyre Termes ff fi fl ffi \textsc{PostScript} % 检查连字

\begin{displaymath}
  E=mc^2 \quad \mathscr{ABCDE abcde 1234} \quad \mathfrak{ABCDE abcde 1234} \quad \mathbb{ABCDE abcde 1234}
\end{displaymath}

\begin{figure}
  \centering
  \subfigure[EPS Figure]{
    \label{fig:demo:a} %% label for first subfigure
    \includegraphics[width=0.3\textwidth]{chap2/testeps}}
  \hspace{1in}
  \subfigure[PDF Figure]{
    \label{fig:demo:b} %% label for second subfigure
    \includegraphics[angle=-90,origin=br,width=0.3\textwidth]{chap2/testpdf}}
  \\ % newline
  \subfigure[PNG Figure]{
    \label{fig:demo:c} %% label for third subfigure
    \includegraphics[width=0.3\textwidth]{chap2/testpng}}
  \hspace{1in}
  \subfigure[JPG Figure]{
    \label{fig:demo:d} %% label for fourth subfigure
    \includegraphics[width=0.3\textwidth]{chap2/testjpg}}
  \bicaption[fig:demo]{插入图像}{插入图像的例子}{Fig}{A demo}
\end{figure}

\begin{itemize}
\item Hello
\item World
\end{itemize}

\section{一段测试文字}

这这是一个最简单的\LaTeX测试文件。这是一个最简单的\LaTeX测试文件。这是一个最简单的\LaTeX测试文件。这是一个最简单的\LaTeX测试文件。这是一个最简单的\LaTeX测试文件。这是一个最简单的\LaTeX测试文件。这是一个最简单的\LaTeX测试文件。这是一个最简单的\LaTeX测试文件。

这这是一个最简单的 \LaTeX 测试文件。这是一个最简单的 \LaTeX 测试文件。这是一个最简单的 \LaTeX 测试文件。这是一个最简单的 \LaTeX 测试文件。这是一个最简单的 \LaTeX 测试文件。这是一个最简单的 \LaTeX 测试文件。这是一个最简单的 \LaTeX 测试文件。这是一个最简单的 \LaTeX 测试文件。

 上海交通大学是我国历史最悠久的高等学府之一，是教育部直属、教育部与上海市共建的全国重点大学，是国家 “七五”、“八五”重点建设和“211工程”、“985工程”的首批建设高校。经过115年的不懈努力，上海交通大学已经成为一所“综合性、研究型、国际化”的国内一流、国际知名大学，并正在向世界一流大学稳步迈进。

\end{document}

### 2013-12-25 slides/sjtu-presentation.tex
Slides for SJTU defense

--- 2013-10-21 lenls/slides.tex
@@
 
-\title[G-ACGs]{Integration of Multiple Constraints in ACG}
+\title[G-ACGs]{Towards a Wide-Coverage Grammar}
 \subtitle{Graphical Abstract Categorial Grammars}
-%
-\author{Jiří Maršík \and Maxime Amblard}
-%
-\institute{Université de Lorraine, Laboratoire lorrain de recherche en informatique et ses applications, UMR 7503, Vandoeuvre-lès-Nancy, 54500, France\\
-INRIA, Villers-lès-Nancy, 54600, France\\
-CNRS, Loria, Vandoeuvre-lès-Nancy, UMR 7503, 54500, France\\
-\texttt{\{jiri.marsik, maxime.amblard\}@loria.fr}}
-
-\date[October 2013]{October 28, 2013}
+\author{Jiří Maršík}
+\institute[INRIA Loria]
+{
+  Department of Computer Science \& Engineering \\
+  Shanghai Jiao Tong University
+  \and
+  INRIA Loria
+}
+\date[December 2013]{December 26, 2013}
 
@@
 \begin{frame}
-  \frametitle{Introduction}
+  \frametitle{Towards a Wide-Coverage Grammar}
+
+  \textbf{Objective:} Build discourse representations using ACGs: \\a
+  wide-coverage ACG grammar.
+
+  \vfill
 
   \begin{itemize}
-    \item Abstract Categorial Grammars provide interesting tools for managing
-      discourse and dynamic semantics. \vfill
-    \item Refining grammars to enforce numerous constraints breeds complexity.
+  \item Having linguistic resources to build upon
+    \begin{itemize}
+    \item Frilex, a formalism-independent lexical database
+    \item Frigram, a French interaction grammar
+    \end{itemize}
     \vfill
-    \item We formalize an extension of ACGs which lets us use the ACG
-      apparatus and maintain independent constraints. \vfill
-    \item Our goal is a large-scale grammar usable for discussing discourse.
+  \item Debugging the grammar
+    \begin{itemize}
+    \item A toolkit for formulating ACGs as simple metagrammars, for
+      examining how they work and for checking their consistency.
+    \end{itemize}
+    \vfill
+  \item Using a formalism that allows for a simple grammar
+    \begin{itemize}
+    \item Express multiple independent constraints \\ $\implies$ graphical abstract categorial grammars.
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{ACGs and IGs}
+
+  \begin{itemize}
+  \item ACGs are based on types (IILL formulas).
+  \item IGs are based on polarized tree descriptions.
+    \vfill
+  \item[$\Rightarrow$] Perrier has linked the two.
+
+    \vspace{2 mm}
+    \begin{columns}[c]
+      \begin{column}{0.1\textwidth}\end{column}
+      \begin{column}{0.4\textwidth}
+        $$(NP \limp S) \limp N \limp N$$
+        $$N \limp (NP \limp S) \limp N$$
+      \end{column}
+      \begin{column}{0.1\textwidth}$\to$\end{column}
+      \begin{column}{0.2\textwidth}
+        \includegraphics[height=0.3\textheight]{../diagrams/ptd.pdf}
+      \end{column}
+      \begin{column}{0.25\textwidth}\end{column}
+    \end{columns}
+    \vfill
+  \item IG of Frigram use different constraints in tree descriptions.
+    \begin{itemize}
+    \item The deductive syntactic model of IGs is built on a wider range
+      of linear logic primitives ($\otimes$, $\&$).
+    \end{itemize}
   \end{itemize}
@@
 
-  \begin{columns}[c]
-    \begin{column}{0.5\textwidth}
-      \begin{itemize}
-      \item Two different lexicons interpret the same abstract terms.
-      \item Instead of referring to the two ACGs $\mathopen{<}
-        \Sigma_{Synt}, \Sigma_{String}, \mathcal{L}_{syntax}, S
-        \mathclose{>}$ and $\mathopen{<} \Sigma_{Synt}, \Sigma_{Sem},
-        \mathcal{L}_{sem}, S \mathclose{>}$, refer to this single
-        diagram.
-      \end{itemize}
-    \end{column}
-    \begin{column}{0.5\textwidth}
-      \includegraphics[width=\textwidth]{../diagrams/double-acg.pdf}
-    \end{column}
-  \end{columns}
+  \begin{center}
+    \includegraphics[height=0.8\textheight]{../diagrams/double-acg.pdf}
+  \end{center}
 \end{frame}
@@
   \frametitle{Multiple Constraints}
-
-  Implementing multiple constraints in a single signature is
-  \textbf{difficult}.
-
   Agreement:
@@
 
+  \note{We have presented two interpretations of G-ACGs based on two different
+    intuitive readings of ACG diagrams, one where nodes are thought of as
+    languages and one where they stand for individual terms. The result shown
+    here validates these two ways of reasoning about diagrams in the context
+    of classical ACGs since they lead to the same languages as the canonical
+    definition of an object language.
+
+    0:25}
+
   \begin{itemize}
@@
 \begin{frame}
-  \frametitle{Properties of G-ACG languages}
+  \frametitle{Properties of G-ACG Languages}
 
-  \begin{itemize}
-  \item The three language families we defined above are\ldots
-    \begin{itemize}
-    \item {\ldots}increasingly constraining, $\forall \, \mathcal{G},
-      u. \; \mathcal{I}_{\mathcal{G}}(u) \supseteq
-      \mathcal{E}_{\mathcal{G}}(u) \supseteq
-      \mathcal{P}_{\mathcal{G}}(u)$.
-    \item {\ldots}increasingly expressive, $\mathcal{I} \subseteq
-      \mathcal{E} \subseteq \mathcal{P}$.
-    \end{itemize}
-  \vfill
-  \item Intrinsic languages are just the abstract languages of ACGs.
+  $$\forall \, \mathcal{G}, u. \; \mathcal{I}_{\mathcal{G}}(u) \supseteq
+  \mathcal{E}_{\mathcal{G}}(u) \supseteq \mathcal{P}_{\mathcal{G}}(u)$$
+
+  $$\mathcal{I} \subseteq \mathcal{E} \subseteq \mathcal{P}$$
+
   $$\mathcal{I} = \mathcal{A}$$
-  \item Extrinsic languages are simple extensions of ACGs that add
-    intersections.
+
   $$\mathcal{O} = \mathcal{A}^{\mathcal{L}}$$
+
   $$\mathcal{E} = \mathcal{A}^{\mathcal{L}{\cap}}$$
-  \vfill
-  \item Pangraphical languages are faithful to the generative
-    perspective and subsume the previous classes of languages.
-    \begin{itemize}
-    \item They solve both the example problems we saw before
-      (\emph{meanings of grammatical sentences} and the \emph{diamond}).
-    \end{itemize}
-  \end{itemize}
 \end{frame}
@@
 \section{Illustration}
+
+\note{We now illustrate the kind of grammar for which we introduced this
+  extension.}
 
@@
 \begin{frame}
+  \addtocounter{framenumber}{-1}
   \frametitle{Multiple Constraints}
@@
 \begin{frame}
+  \addtocounter{framenumber}{-1}
   \frametitle{Exposing Another Signature}
@@
 \begin{frame}
+  \addtocounter{framenumber}{-1}
   \frametitle{Constraints on Multiple Signatures}
@@
 \begin{frame}
-  \frametitle{Recapitulation}
+  \frametitle{Summary of Work Done}
 
   \begin{itemize}
-  \item We reinterpreted ACGs by treating the diagrams as primary constructs
-    instead of as artifacts.
+  \item Formalized graphical abstract categorial grammars.
+    \begin{itemize}
+      \item ACG diagrams as mathematical objects
+      \item Non-arborescence enables intersection
+      \item Intersection enables conjunction of independent constraints
+      \item Large-scale grammar: collection of fragments
+    \end{itemize}
   \vfill
-  \pause
-  \item This lets us handle non-arborescent diagrams in a principled manner.
+  \item Created a toolkit to serve as basis for more explorations or
+    implementations.
   \vfill
-  \pause
-  \item Non-arborescence enables intersection and therefore conjunction of
-    independent constraints.
-  \vfill
-  \pause
-  \item We presented the kind of grammar we envisioned, built by collecting
-    fragments described in existing research.
+  \item Direct embedding of IGs into ACGs based on Perrier's theorem
+    conjectured not to be practical.
   \end{itemize}
@@
   \begin{itemize}
+  \item Are ACG object languages closed on
+    intersection? $$\mathcal{O}^{\mathcal{L}{\cap}} \subseteq \mathcal{O}^{\mathcal{L}}$$
   \item Are extrinsic languages as expressive as pangraphical
     ones? $$\mathcal{P} \subseteq \mathcal{E}$$
-  \item Are ACG object languages closed on
-    intersection? $$\mathcal{O}^{\mathcal{L}{\cap}} \subseteq \mathcal{O}^{\mathcal{L}}$$
+  \item Express the multiple polarized features of IGs in G-ACGs.
+  \vfill
   \item Is there a metagrammatical approach that makes refining the types of a

### 2014-02-17 lenls/post-proc.tex
Forked LENLS paper for post-proc
POST: Minor revisions, pangraphical models

--- 2013-09-22 lenls/paper.tex
@@
-% This is LLNCS.DEM the demonstration file of
-% the LaTeX macro package from Springer-Verlag
-% for Lecture Notes in Computer Science,
-% version 2.4 for LaTeX2e as of 16. April 2010
-%
 \documentclass{llncs}
@@
 \usepackage{amsmath}
-%\usepackage{amsthm}
 \usepackage{xcolor}
@@
 presented here was motivated by this quest for a wide-coverage ACG grammar, as
-seen from a point of view of language as a system of largely orthogonal
+seen from the point of view of language as a system of largely orthogonal
 constraints. Encoding ancillary constraints in a type system can make the
@@
 free negative determiners (i.e. there is no $N_{de_4} : NP_{NEG{=}T} \limp
-N_{NEG{=}T} \limp N_{NEG{=}}\ldots$). This encodes a constraint that there can
+N_{NEG{=}T} \limp N_{NEG{=}?}$). This encodes a constraint that there can
 be only one free negative determiner per phrase (free as in not hidden inside
@@
 
-The intrinsic language is just the set of terms built on the node's
-signature and having the node's distinguished type. The extrinsic
-language is established by taking the extrinsic languages of its
-predecessors, mapping them through lexicons and taking their
-intersection, or just taking the node's intrinsic language if it has no
-predecessors.
+The intrinsic language is just the set of terms built on the node's signature
+and having the node's distinguished type. The extrinsic language is
+established by taking the extrinsic languages of its predecessors, mapping
+them through the lexicons and taking their intersection, or just taking the
+node's intrinsic language if it has no predecessors.
 
@@
 of G-ACGs. We will present one such alternative now, which exchanges the
-language-algebraic point of view for a generative one.
+language-algebraic point of view for a term-relational one.
 
@@
 
-In the new paradigm, we interpret the nodes of the graph as terms and the
-edges as statements that one term is mapped into another using a lexicon. This
-leads us to the definition of the \emph{pangraphical}\footnote{As opposed to
-  extrinsic languages, which are constrained only by their predecessors in the
-  graph, pangraphical languages are constrained by the entire graph.} language
-of a node $u$ in a G-ACG $\mathcal{G}$.
+In the new paradigm, we interpret the graph as a logical formula. The nodes
+will be seen as variables whose domains are their intrinsic languages,
+i.e. each node $v$ as a variable will range over the terms built from its
+signature $\Sigma_v$ and having its distinguished type $S_v$. Edges $(v,w)$
+will be statements saying that the image of the term $v$ through the lexicon
+$\mathcal{L}_{(v,w)}$ is exactly the term $w$.
 
-A term $t$ belongs to $\mathcal{P}_{\mathcal{G}}(u)$ whenever there
-exists a labeling $T$ of the nodes of the graph such that:
+We can think of \emph{models} for this graphical propositional logic as
+assignments of terms to variables, i.e. labelings of the nodes of the graph
+with terms. A model $T$ is \emph{valid} when it satisfies the following
+conditions:
 
 \begin{itemize}
-  \item $T_u = t$.
   \item For all $v \in V(G)$, $\vdash_{\Sigma_v} T_v : S_v$.
@@
 
+A term $t$ belongs to the \emph{pangraphical}\footnote{As opposed to extrinsic
+  languages, which are constrained only by their predecessors in the graph,
+  pangraphical languages are constrained by the entire graph, hence the name.}
+language $\mathcal{P}_{\mathcal{G}}(u)$ if there exists a valid model $T$ such
+that $T(u) = t$.
+
 If we compare this new interpretation of G-ACGs to the former one, we find out
-that in the case when the graph of the grammar is an arborescence, they are
+that in the cases when the graph of the grammar is an arborescence, they are
 actually equivalent. This means that in classical ACGs, where all the diagrams
@@
 purely with syntax without any issues of operator scope and its functions have
-lower-order types than those in $Syn$. We will finally add our contribution,
-the signatures $Neg$ and $Agr$ and the lexicons $\mathcal{L}_{Neg}$ and
-$\mathcal{L}_{Agr}$ which implement the constraint on negation we discussed
-before and some simple notion of agreement, respectively
+types of a lower order than those in $Syn$. We will finally add our
+contribution, the signatures $Neg$ and $Agr$ and the lexicons
+$\mathcal{L}_{Neg}$ and $\mathcal{L}_{Agr}$ which implement the constraint on
+negation we discussed before and some simple notion of agreement, respectively
 (Figure~\ref{fig:final-synth}).
@@
 operators. Finally, we can opt in to verify the high-level syntax-semantics
-constraints which employ dependent types. However, the lexicons in this case
-resemble relabelings and a semi-decidable procedure could be used to solve the
+constraints which employ dependent types
+\cite{pogodalla2012controlling}. However, the lexicons in this case resemble
+relabelings and a semi-decidable procedure could be used to solve the
 problem. Having the grammar decomposed like this opens it up for computational

### 2014-02-18 lenls/post-proc.tex
POST: Final remarks from Maxime, flipped sigs in story diagrams

@@
 ACG abstract languages closed on transformation by a lexicon and
-intersection\footnote{Can be shown by induction on the topological ordering of
-  any given graphical ACG.}.
+intersection.
 
@@
 $$
+
+\begin{proof}
+  We prove $\mathcal{E} = \mathcal{A}^{\mathcal{L}{\cap}}$, or equivalently
+  $\mathcal{E} = \mathcal{I}^{\mathcal{L}{\cap}}$, by proving $\mathcal{E}
+  \subseteq \mathcal{I}^{\mathcal{L}{\cap}}$ and
+  $\mathcal{I}^{\mathcal{L}{\cap}} \subseteq \mathcal{E}$.
+
+  We begin with $\mathcal{E} \subseteq \mathcal{I}^{\mathcal{L}{\cap}}$. Let
+  us consider any graphical ACG $\mathcal{G}$ and some topological ordering
+  $v_1, \ldots, v_n$ of its nodes $V(G)$ (an ordering such that more abstract
+  nodes always precede less abstract ones). We proceed by induction on the
+  topological ordering and prove that $\mathcal{E}_{\mathcal{G}}(v_i) \in
+  \mathcal{I}^{\mathcal{L}{\cap}}$ for all $v_i$ in $V(G)$:
+
+  \begin{itemize}
+  \item For $v_1$, the case is trivial. $v_1$ has no predecessors and so
+    $\mathcal{E}_{\mathcal{G}}(v_1) = \mathcal{I}_{\mathcal{G}}(v_1) \in
+    \mathcal{I}$.
+  \item For any other node $v_n$, we inspect the definition of
+    $\mathcal{E}_{\mathcal{G}}(v_n)$.
+
+    $$\mathcal{E}_{\mathcal{G}}(v_n) = \mathcal{I}_{\mathcal{G}}(v_n) \cap
+    \bigcap_{(u,v_n) \in E(G)}
+    \mathcal{L}_{(u,v_n)}(\mathcal{E}_{\mathcal{G}}(u))$$
+
+    By induction hypothesis, $\mathcal{E}_{\mathcal{G}}(u) \in
+    \mathcal{I}^{\mathcal{L}{\cap}}$ for all the predecessors $u$ of
+    $v_n$. Since, $\mathcal{I}^{\mathcal{L}{\cap}}$ is closed on projection by
+    a lexicon, we have that
+    $\mathcal{L}_{(u,v_n)}(\mathcal{E}_{\mathcal{G}}(u)) \in
+    \mathcal{I}^{\mathcal{L}{\cap}}$ as well. Finally, by the closure of
+    $\mathcal{I}^{\mathcal{L}{\cap}}$ on intersection, we have that their
+    intersection $\bigcap_{(u,v_n) \in E(G)}
+    \mathcal{L}_{(u,v_n)}(\mathcal{E}_{\mathcal{G}}(u))$ also belongs to
+    $\mathcal{I}^{\mathcal{L}{\cap}}$. Since $\mathcal{I}_{\mathcal{G}}(v_n)$
+    belongs to $\mathcal{I}^{\mathcal{L}{\cap}}$ too, we know that their
+    intersection, $\mathcal{E}_{\mathcal{G}}(v_n)$ belongs to
+    $\mathcal{I}^{\mathcal{L}{\cap}}$.
+  \end{itemize}
+
+  Now we prove the converse, that is that $\mathcal{I}^{\mathcal{L}{\cap}}
+  \subseteq \mathcal{E}$. We do so by proving that $\mathcal{I} \subseteq
+  \mathcal{E}$ and that $\mathcal{E}$ is closed on projection through a
+  lexicon and intersection.
+
+  \begin{description}
+    \item[Inclusion of intrinsic languages] Let us consider the intrinsic
+      language $\mathcal{I}_{\mathcal{G}}(u)$ of some node $u$ in some G-ACG
+      $\mathcal{G}$. We can produce a G-ACG $\mathcal{G}_u$, whose graph
+      contains only the node $u$ labeled with the same signature and
+      distinguished type as in $\mathcal{G}$. Since $u$ has no predecessors in
+      $\mathcal{G}_u$, it will be the case that
+      $\mathcal{E}_{\mathcal{G}_u}(u) = \mathcal{I}_{\mathcal{G}_u}(u) =
+      \mathcal{I}_{\mathcal{G}}(u)$ and therefore the intrinsic language
+      $\mathcal{I}_{\mathcal{G}}(u)$ is also an extrinsic language
+      $\mathcal{E}_{\mathcal{G}_u}(u)$.
+
+    \vspace{5mm}
+
+    \item[Closure on projection through a lexicon] Let $L_1$ and $L_2$ be two
+      extrinsic languages defined by the nodes $u_1$ and $u_2$ in some
+      graphical ACGs $\mathcal{G}_1$ and $\mathcal{G}_2$, respectively. We can
+      construct a new graphical ACG $\mathcal{G}'$ by taking the union of the
+      two ACGs (union of graphs (union of vertices and edges) and unions of
+      the labelings).\footnote{Assuming, without loss of generality, that the
+        sets of vertices of the two graphs are disjoint.} We will then add a
+      new node $v$ such that $\mathcal{E}_{\mathcal{G}'}(v) = L_1 \cap L_2$.
+
+      The signature labeling $v$ will be the union of the two signatures
+      labeling $u_1$ and $u_2$, ensuring that $\forall u \in \{u_1,
+      u_2\}$, $\mathcal{I}_{\mathcal{G}'}(v) \supseteq
+      \mathcal{I}_{\mathcal{G}'}(u) \supseteq
+      \mathcal{E}_{\mathcal{G}'}(u)$. We can use either of the two
+      distinguished types that label $u_1$ or $u_2$ to label $v$ (if
+      they are different, then the intersection $L_1 \cap L_2$ is
+      trivially empty). Finally, we will add two edges leading from
+      $u_1$ and $u_2$ to $v$, both labeled with the identity lexicon. We
+      will now show that $\mathcal{E}_{\mathcal{G}'}(v) = L_1 \cap L_2$.
+
+      \begin{align*}
+        \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v)
+        \cap \mathcal{L}_{(u_1,v)}(\mathcal{E}_{\mathcal{G}'}(u_1)) \cap
+        \mathcal{L}_{(u_2,v)}(\mathcal{E}_{\mathcal{G}'}(u_2)) \\ &=
+        \mathcal{I}_{\mathcal{G}'}(v) \cap
+        \mathcal{E}_{\mathcal{G}'}(u_1) \cap
+        \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
+        \mathcal{E}_{\mathcal{G}'}(u_1) \cap
+        \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
+        \mathcal{E}_{\mathcal{G}_1}(u_1) \cap
+        \mathcal{E}_{\mathcal{G}_2}(u_2) \\ &= L_1 \cap L_2
+      \end{align*}
+
+      The simplifications use the facts that: the two new lexicons are
+      identities, that $\mathcal{I}_{\mathcal{G}'}(v) \supseteq
+      \mathcal{E}_{\mathcal{G}'}(u_1) \cap \mathcal{E}_{\mathcal{G}'}(u_2)$
+      (due to $\Sigma'_v = \Sigma'_{u_1} \cup \Sigma'_{u_2}$), that
+      $\mathcal{E}_{\mathcal{G}'}(u_1) = \mathcal{E}_{\mathcal{G}_1}(u_1)$ and
+      $\mathcal{E}_{\mathcal{G}'}(u_2) = \mathcal{E}_{\mathcal{G}_2}(u_2)$
+      (due to $u_1$ and $u_2$ having the same predecessors in $\mathcal{G}$
+      and $\mathcal{G}'$) and finally that $\mathcal{E}_{\mathcal{G}}(u_1) =
+      L_1$ and $\mathcal{E}_{\mathcal{G}}(u_2) = L_2$ (our starting
+      assumptions).
+
+  \vspace{5mm}
+
+  \item[Closure on intersection] Let $L \subseteq \Lambda(\Sigma_O)$ be a
+    language which is the result of mapping an extrinsic language $L'
+    \subseteq \Lambda(\Sigma_A)$ through the lexicon $\mathcal{L} : \Sigma_A
+    \to \Sigma_O$. We will show that $L$ is also an extrinsic language. Let
+    $u$ be the node that defines the extrinsic language $L$ in some graphical
+    ACG $\mathcal{G}$. We construct a new graphical ACG $\mathcal{G}'$ by
+    adding a new vertex $v$ to $\mathcal{G}$ labeled with the signature
+    $\Sigma_O$ and the distinguished type $\mathcal{L}(S)$ where $S$ is the
+    distinguished type labeling $u$. To this graph, we conjoin the edge
+    $(u,v)$ and label it with $\mathcal{L}$.
+
+    Now we verify that $\mathcal{E}_{\mathcal{G}'}(v) = L$.
+
+    \begin{align*}
+      \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v) \cap
+      \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
+      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
+      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}}(u)) \\
+      &= \mathcal{L}(L') \\
+      &= L
+    \end{align*}
+
+    The assumptions used here are in turn: that $\mathcal{I}_{\mathcal{G}'}(v)
+    \supseteq \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u))$ (due to the
+    homomorphism property of lexicon $\mathcal{L} : \Sigma_A \to \Sigma_O$),
+    that $\mathcal{E}_{\mathcal{G}'}(u) = \mathcal{E}_{\mathcal{G}}(u)$ (due
+    to $u$ having the same predecessors in $\mathcal{G}$ and $\mathcal{G}'$),
+    that $\mathcal{E}_{\mathcal{G}}(u) = L'$ and that $\mathcal{L}(L') = L$
+    (our starting assumptions).
+  \end{description}
+\hfill $\qed$
+\end{proof}
 
@@
   as opposed to just the node's predecessors).}. The proof is carried out by
-transforming a G-ACG such that a particular node will have the same
-pangraphical language in the latter G-ACG as the extrinsic language it had in
-the former. The nodes of the newly constructed G-ACG correspond to paths in
-the former one terminating in the node in question\footnote{This means that
-  when the G-ACG is $\mathcal{D}$, the diamond grammar, and the node in
-  question is $Bottom$, then we will have two different nodes for the two
-  paths $[Top, Left, Bottom]$ and $[Top, Right, Bottom]$.}. Thus the labelling
-of nodes with terms which certifies a given term's presence in the
-pangraphical language of the new G-ACG also serves as a proof of its presence
-in the extrinsic language of the old G-ACG and vice versa.
+transforming any G-ACG $\mathcal{G}$ where $L = \mathcal{E}_{\mathcal{G}}(u)$
+to a G-ACG $\mathcal{G'}$ where for some $v$, $\mathcal{P}_{\mathcal{G'}}(v) =
+L$. We do so by taking all the (finitely many) paths in $G$ that end in $u$ to
+be the vertices of $G'$. Each of the vertices in $G'$ (paths in $G$) will be
+labeled by the same signature and distinguished type as the first node of the
+corresponding path. Furthermore, there will be an edge $(p_1, p_2)$ in $G'$
+labeled with the lexicon $\mathcal{L}$ whenever the path $p_2$ is the maximal
+strict suffix of $p_1$ such that $p_1 = e; p_2$ where the edge $e$ is labeled
+with $\mathcal{L}$ in $G$.\footnote{This means that when the G-ACG is
+  $\mathcal{D}$, the diamond grammar, and the node in question is $Bottom$,
+  then we will have the following nodes in $\mathcal{D}'$: $[Top, Left,
+    Bottom]$, $[Top, Right, Bottom]$, $[Left, Bottom]$, $[Right, Bottom]$,
+  $[Bottom]$. Note that we have two nodes for (i.e. beginning with) $Top$,
+  which will let us admit two different antecedents in
+  $\mathcal{E}_{\mathcal{D}}(Top)$.} In this G-ACG $\mathcal{G'}$, valid
+models $T$ which assign $T([u]) = t$ are isomorphic to proofs of membership of
+$t$ in $\mathcal{E}_{\mathcal{G}}(u)$ and therefore
+$\mathcal{P}_{\mathcal{G'}}([u]) = \mathcal{E}_{\mathcal{G}}(u)$.
 
@@
     \centering
-    \includegraphics[height=0.2\textheight]{../diagrams/single-cont.pdf}
+    \includegraphics[height=0.2\textheight]{../diagrams/story-single-cont.pdf}
     \caption{\label{fig:single-cont} An extraction constraint }
@@
     \centering
-    \includegraphics[height=0.2\textheight]{../diagrams/merged-conts.pdf}
+    \includegraphics[height=0.2\textheight]{../diagrams/story-merged-conts.pdf}
     \caption{\label{fig:merged-conts} G-ACG merging the three extraction
@@
     \centering
-    \includegraphics[height=0.2\textheight]{../diagrams/conts-elab.pdf}
+    \includegraphics[height=0.2\textheight]{../diagrams/story-conts-elab.pdf}
     \caption{\label{fig:conts-elab} Elaborating the lexicon from syntax to
@@
     \centering
-    \includegraphics[height=0.2\textheight]{../diagrams/final-synth.pdf}
+    \includegraphics[height=0.2\textheight]{../diagrams/story-final-synth.pdf}
     \caption{\label{fig:final-synth} Adding our own syntactic constraints,
@@
 N_{aucun} &: N_{NEG{=}F} \limp NP_{NEG{=}T} \\
+\end{align*} % for page break purposes
+\begin{align*}
 N_{le_1} &: N_{NEG{=}F} \limp NP_{NEG{=}F} \\
@@
 
-\vspace{-3ex}
-
 \begin{align*}
@@
 A_{aime_1} &: NP_{NUM{=}SG} \limp NP_{NUM{=}SG} \limp S \\
-A_{aime_2} &: NP_{NUM{=}SG} \limp NP_{NUM{=}PL} \limp S
-\end{align*}
+A_{aime_2} &: NP_{NUM{=}SG} \limp NP_{NUM{=}PL} \limp S \\
+\end{align*} % for page break purposes
 \begin{align*}
 A_{qui_1} &: (NP_{NUM{=}SG} \limp S) \limp N_{NUM{=}SG} \limp N_{NUM{=}SG} \\
-A_{qui_2} &: (NP_{NUM{=}PL} \limp S) \limp N_{NUM{=}PL} \limp N_{NUM{=}PL}
+A_{qui_2} &: (NP_{NUM{=}PL} \limp S) \limp N_{NUM{=}PL} \limp N_{NUM{=}PL} \\
 \end{align*}

### 2014-02-21 lenls/post-proc.tex
POST: Swap sigs in figs, embed biblio

@@
     \centering
-    \includegraphics[height=0.2\textheight]{../diagrams/double-acg.pdf}
+    \includegraphics[height=0.2\textheight]{diagrams/double-acg.pdf}
     \caption{\label{fig:acg-comp-basic} Connecting form with meaning.}
@@
     \centering
-    \includegraphics[height=0.2\textheight]{../diagrams/serial-over-parallel.pdf}
+    \includegraphics[height=0.2\textheight]{diagrams/serial-over-parallel.pdf}
     \caption{\label{fig:acg-comp-constr} Adding a constraint (Constraint).}
@@
     \centering
-    \includegraphics[height=0.2\textheight]{../diagrams/parallel-over-serial.pdf}
+    \includegraphics[height=0.2\textheight]{diagrams/parallel-over-serial.pdf}
     \caption{\label{fig:acg-comp-sem} Distinguishing syntactic and
@@
   \centering
-  \includegraphics[height=0.3\textheight]{../diagrams/final.pdf}
+  \includegraphics[height=0.3\textheight]{diagrams/final.pdf}
   \caption{\label{fig:gacg} A graphical ACG that implements two
@@
   \centering
-  \includegraphics[height=0.3\textheight]{../diagrams/diamond-grammar.pdf}
+  \includegraphics[height=0.3\textheight]{diagrams/diamond-grammar.pdf}
   \caption{\label{fig:diamond} The diamond-shaped G-ACG $\mathcal{D}$.}
@@
     \centering
-    \includegraphics[height=0.2\textheight]{../diagrams/story-single-cont.pdf}
+    \includegraphics[height=0.2\textheight]{diagrams/story-single-cont.pdf}
     \caption{\label{fig:single-cont} An extraction constraint }
@@
     \centering
-    \includegraphics[height=0.2\textheight]{../diagrams/story-merged-conts.pdf}
+    \includegraphics[height=0.2\textheight]{diagrams/story-merged-conts.pdf}
     \caption{\label{fig:merged-conts} G-ACG merging the three extraction
@@
     \centering
-    \includegraphics[height=0.2\textheight]{../diagrams/story-conts-elab.pdf}
+    \includegraphics[height=0.2\textheight]{diagrams/story-conts-elab.pdf}
     \caption{\label{fig:conts-elab} Elaborating the lexicon from syntax to
@@
     \centering
-    \includegraphics[height=0.2\textheight]{../diagrams/story-final-synth.pdf}
+    \includegraphics[height=0.2\textheight]{diagrams/story-final-synth.pdf}
     \caption{\label{fig:final-synth} Adding our own syntactic constraints,
@@
 
-\bibliographystyle{splncs}
-\bibliography{../biblio}
+\begin{thebibliography}{10}
+
+\bibitem{de2001towards}
+De~Groote, P.:
+\newblock Towards abstract categorial grammars.
+\newblock In: Proceedings of the 39th Annual Meeting on Association for
+  Computational Linguistics, Association for Computational Linguistics (2001)
+  252--259
+
+\bibitem{de2006towards}
+De~Groote, P.:
+\newblock Towards a montagovian account of dynamics.
+\newblock In: Proceedings of SALT. Volume~16. (2006)
+
+\bibitem{qian2011event}
+Qian, S., Amblard, M.:
+\newblock Event in compositional dynamic semantics.
+\newblock In: Logical Aspects of Computational Linguistics.
+\newblock (2011)
+
+\bibitem{asher2011sdrt}
+Asher, N., Pogodalla, S.:
+\newblock Sdrt and continuation semantics.
+\newblock In: New Frontiers in Artificial Intelligence. Volume 6797 of Lecture
+  Notes in Computer Science.
+\newblock (2011)
+
+\bibitem{asher2011montagovian}
+Asher, N., Pogodalla, S.:
+\newblock A montagovian treatment of modal subordination.
+\newblock In: Proceedings of SALT. Volume~20. (2011)  387--405
+
+\bibitem{pogodalla2012controlling}
+Pogodalla, S., Pompigne, F.:
+\newblock Controlling extraction in abstract categorial grammars.
+\newblock In: Formal Grammar. (2012)
+
+\bibitem{degroote02}
+de~Groote, P.:
+\newblock Tree-adjoining grammars as abstract categorial grammars.
+\newblock In: TAG+6, Proceedings of the sixth International Workshop on Tree
+  Adjoining Grammars and Related Frameworks, Universit\`a di Venezia (2002)
+  145--150
+
+\bibitem{pogodalla2007generalizing}
+Pogodalla, S.:
+\newblock Generalizing a proof-theoretic account of scope ambiguity.
+\newblock In: 7th International Workshop on Computational Semantics. (2007)
+
+\bibitem{montague1973proper}
+Montague, R.:
+\newblock The proper treatment of quantification in ordinary english
+
+\bibitem{perrier2007french}
+Perrier, G.:
+\newblock A french interaction grammar.
+\newblock In: International Conference on Recent Advances in Natural Language
+  Processing. (2007)
+
+\bibitem{de2007type}
+de~Groote, P., Maarek, S.:
+\newblock Type-theoretic extensions of abstract categorial grammars.
+\newblock ESSLLI 2007 (2007)
+
+\end{thebibliography}
 
